### 2013-06-14 LATEX/outline.tex
First commit

\documentclass[12pt]{article}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{chicago}

\setlength{\paperwidth}{8.5in} \setlength{\paperheight}{11.0in}
\setlength{\topmargin}{0.0in} \setlength{\headheight}{0.4in}
\setlength{\headsep}{0.0in} \setlength{\textwidth}{7.2in}
\setlength{\textheight}{8.5in} \setlength{\oddsidemargin}{0.0in}
\setlength{\oddsidemargin}{-0.4in}
\setlength{\evensidemargin}{-0.4in}

\renewcommand{\baselinestretch}{1.5}
\renewcommand{\textfraction}{0.33}
\begin{document}

\bigskip

\bigskip

\clearpage

\renewcommand{\baselinestretch}{1.0}

\begin{center}
{\normalsize \thispagestyle{empty} }

{\normalsize \medskip }

{\normalsize {\Large Forcasting Net Interest Margins in a Zero Interest Rate Environment  $^{*}$%
} \medskip }

{\normalsize \bigskip \bigskip }

{\normalsize Valentin Bolotnyy, Rochelle Edge, and Luca Guerrieri }

{\normalsize \bigskip \bigskip }

{\normalsize April 29, 2013 }

{\normalsize \bigskip }
\end{center}

{\normalsize \bigskip }

\abstract{We forecast net interest margins (NIMs) using a simple vector autoregression model that includes a short interest rate and a long interest rate,
and using a term structure modeled after Diebold, et al.. Comparing NIM forecasts produced by each of these models, we find that .... Modeling the zero lower bound as a regime change produces .... Additionally, we find that NIM forecasts aggregated for the top 25 U.S. banks, by assets, are ... compared to firm-specific forecasts which are .... Exploring the effects of shocks
at different points in the term structure, we find ....}

{\normalsize \vspace{3.0cm} }

{\normalsize \noindent \textbf{Keywords}:  \vspace{1cm} }

{\normalsize \noindent \textbf{JEL Classification}:  }

{\normalsize \vspace{2cm} }

\renewcommand{\baselinestretch}{1} \footnotesize \noindent

\textbf{\ Affiliation and contact information}: Valentin Bolotnyy,
Federal Reserve Board, telephone (202) 452-6428, email
valentin.bolotnyy@frb.gov; Rochelle Edge, Federal Reserve Board, telephone (202) 452-2339, email
rochelle.m.edge@frb.gov; Luca Guerrieri, Federal Reserve Board,
telephone (202) 452-2550, email luca.guerrieri@frb.gov.

\vspace{2cm}

{\footnotesize \noindent $^{*}$ The views expressed in this paper are solely
the responsibility of the authors and should not be interpreted as
reflecting the views of the Board of Governors of the Federal Reserve System
or of any other person associated with the Federal Reserve System. We want to thank Michiel De Pooter, Canlin Li, Marcel Priebsch, John Sears, Emre Yoldas... }

\clearpage \renewcommand{\baselinestretch}{1.5} \normalsize

\section{Outline}
1. Introduction: Motivate desire/need to forecast NIMs. Describe (chart) historical NIM behavior for top 25 U.S. banks. Summarize our approaches and findings.
\newline
2. Lit Review: Discuss approaches previously taken in the literature to forecasting NIMs.
\newline
3. Methodology:

    a) Describe VAR model with 3 month and 10 year yields and how we employ it to forecast NIMs

    b) Describe Diebold, et al. approach to modeling yield curve and how we employ it to forecast NIMs

    c) Discuss why we chose to focus on these two models

    d) Discuss how why we chose to deal with zero lower bound as a regime shift

    e) Describe how we test forecasts (Diebold-Mariano and Clark-West)
\newline
4. Results: Discuss forecasts produced by different models and how well they do. Do firm-specific forecasts reveal more than aggregate (top 25) forecasts?
\newline
5. Interest Rate Path Scenarios: Discuss how shocks at different points in the term structure would affect NIMs (at aggregate and individual firm levels?)
\newline
6. Appendix: Discuss results of our effort to forecast NIMs by asset class
\newline
Potential Robustness Checks:

    a) Instead of 3-month and 6-month Svensson model interest rates, use implied rates posted by NY Fed.

\end{document}

### 2013-06-14 MATLAB/estimation_utilities/anneal.m

@@ -1,0 +1,210 @@
+function [minimum,fval] = anneal(loss, parent, options)
+% ANNEAL  Minimizes a function with the method of simulated annealing
+% (Kirkpatrick et al., 1983)
+%
+%  ANNEAL takes three input parameters, in this order:
+%
+%  LOSS is a function handle (anonymous function or inline) with a loss
+%  function, which may be of any type, and needn't be continuous. It does,
+%  however, need to return a single value.
+%
+%  PARENT is a vector with initial guess parameters. You must input an
+%  initial guess.
+%
+%  OPTIONS is a structure with settings for the simulated annealing. If no
+%  OPTIONS structure is provided, ANNEAL uses a default structure. OPTIONS
+%  can contain any or all of the following fields (missing fields are
+%  filled with default values):
+%
+%       Verbosity: Controls output to the screen.
+%                  0 suppresses all output
+%                  1 gives final report only [default]
+%                  2 gives temperature changes and final report 
+%       Generator: Generates a new solution from an old one.
+%                  Any function handle that takes a solution as input and
+%                  gives a valid solution (i.e. some point in the solution
+%                  space) as output.
+%                  The default function generates a row vector which
+%                  slightly differs from the input vector in one element:
+%                  @(x) (x+(randperm(length(x))==length(x))*randn/100)
+%                  Other examples of possible solution generators:
+%                  @(x) (rand(3,1)): Picks a random point in the unit cube
+%                  @(x) (ceil([9 5].*rand(2,1))): Picks a point in a 9-by-5
+%                                                 discrete grid
+%                  Note that if you use the default generator, ANNEAL only
+%                  works on row vectors. For loss functions that operate on
+%                  column vectors, use this generator instead of the
+%                  default:
+%                  @(x) (x(:)'+(randperm(length(x))==length(x))*randn/100)'
+%        InitTemp: The initial temperature, can be any positive number.
+%                  Default is 1.
+%        StopTemp: Temperature at which to stop, can be any positive number
+%                  smaller than InitTemp. 
+%                  Default is 1e-8.
[…]

### 2013-06-14 MATLAB/estimation_utilities/companion.m

@@ -1,0 +1,15 @@
+function coefcompanion=companion(coefb,const)
+
+nvars = size(coefb,1);
+varlag = size(coefb,3);
+
+coefcompanion = [coefb(:,:,1)];
+
+for i = 2:varlag
+coefcompanion = [coefcompanion coefb(:,:,i)];
+end
+
+coefcompanion = [coefcompanion; [eye(nvars*(varlag-1)),zeros(nvars*(varlag-1),nvars) ] ];
+
+% append constant at the end of the state space (initialize constant to 1 when hooking up to the kalman filter)
+coefcompanion = [[coefcompanion [const;zeros(nvars*(varlag-1),1)]]; [zeros(1,nvars*varlag) 1] ];

### 2013-06-14 MATLAB/estimation_utilities/confint.m

@@ -1,0 +1,69 @@
+
+function [confint_lb confint_ub] = confint(confint_perctl,cofb,const,errmat,y,varlag,nreps,isconstant,iscontemp,yw,h,y_data,xi10,P10,q,ntrain)
+
+T = size(errmat,1);
+neqs = size(errmat,2);
+
+nperiods = size(y_data,1);
+nim_mat = zeros(nreps,nperiods);
+
+
+bad_draw=0;
+for confindx=1:nreps
+    issingular=-1;
+    
+    %stay in while loop until artificial data is generated that leads a nonsingular VAR    
+    while (issingular<0) % keep drawing a new monte carlo series until one is found
+        % that produces a covariance stationary VAR
+        
+        %%% Sample with replacement from the fitted residuals
+        errpos = round((T-1)*rand(T,1)+ones(T,1));
+%         for posindx=1:T
+%             errmonte(posindx,:)=errmat(errpos(posindx),:);
+%         end
+        errmonte=errmat(errpos,:);
+        
+        %%% Using VAR structure and bootstrapped residuals, compute new y data
+        %ymonte = mkymonteg(y,coefs,errmonte);
+        ymonte = mkymonte(y,cofb,const,errmonte);
+        if (yw == 0) 
+            %[coefsmonte,coverr,errmatmonte]=estimateg(varlag,ymonte);
+            varlagmat = varlag*ones(neqs);
+            [coefsmonte,coverr,errmatmonte]=estimate(varlagmat,zeros(neqs),ones(neqs,1),ymonte);
+            
+
+            [cofbmonte,const_bmonte]=ols2ar(coefsmonte(1:end,:),isconstant,iscontemp);
+            cofamonte = iscontemp*coefsmonte(1+isconstant:neqs+isconstant,:);
+        else    
+            [Amonte, a0monte, coverr] = ywestimate(ymonte,varlag);
+            [cofbmonte,const_bmonte]=yw2ar(Amonte,a0monte);
+            cofamonte = zeros(neqs);
+        end
+        
+        cofainvmonte = (eye(neqs)-cofamonte)^(-1);
+        cofcmonte=isconstant*cofainvmonte*const_bmonte; 
+        
+        
+        [issingular] = singulartest(cofbmonte);
+        if issingular < 0
+            bad_draw = bad_draw + 1;
+        end
+        
+    end  %end of while loop
[…]

### 2013-06-14 MATLAB/estimation_utilities/confint_partial.m

@@ -1,0 +1,65 @@
+
+function [replication_forecast] = confint_partial(cofb,const,errmat,y,varlag,nreps,isconstant,iscontemp,yw,h,y_data,xi10,P10,q,ntrain,errpos)
+
+T = size(errmat,1);
+neqs = size(errmat,2);
+
+nperiods = size(y_data,1);
+nim_mat = zeros(nreps,nperiods);
+
+
+bad_draw=0;
+
+    issingular=-1;
+    
+    %stay in while loop until artificial data is generated that leads a nonsingular VAR    
+    while (issingular<0) % keep drawing a new monte carlo series until one is found
+        % that produces a covariance stationary VAR
+        
+        %%% Sample with replacement from the fitted residuals
+        %errpos = round((T-1)*rand(T,1)+ones(T,1));
+%         for posindx=1:T
+%             errmonte(posindx,:)=errmat(errpos(posindx),:);
+%         end
+        errmonte=errmat(errpos,:);
+        
+        %%% Using VAR structure and bootstrapped residuals, compute new y data
+        %ymonte = mkymonteg(y,coefs,errmonte);
+        ymonte = mkymonte(y,cofb,const,errmonte);
+        if (yw == 0) 
+            %[coefsmonte,coverr,errmatmonte]=estimateg(varlag,ymonte);
+            varlagmat = varlag*ones(neqs);
+            [coefsmonte,coverr,errmatmonte]=estimate(varlagmat,zeros(neqs),ones(neqs,1),ymonte);
+            
+
+            [cofbmonte,const_bmonte]=ols2ar(coefsmonte(1:end,:),isconstant,iscontemp);
+            cofamonte = iscontemp*coefsmonte(1+isconstant:neqs+isconstant,:);
+        else    
+            [Amonte, a0monte, coverr] = ywestimate(ymonte,varlag);
+            [cofbmonte,const_bmonte]=yw2ar(Amonte,a0monte);
+            cofamonte = zeros(neqs);
+        end
+        
+        cofainvmonte = (eye(neqs)-cofamonte)^(-1);
+        cofcmonte=isconstant*cofainvmonte*const_bmonte; 
+        
+        
+        [issingular] = singulartest(cofbmonte);
+        if issingular < 0
+            bad_draw = bad_draw + 1;
+        end
+        
+    end  %end of while loop
+     cofcompanion = companion(cofbmonte,const_bmonte);
+    
[…]

### 2013-06-14 MATLAB/estimation_utilities/confint_projection.m

@@ -1,0 +1,81 @@
+
+function [irfs1, irfs2] = confint(coefs,errmat,y,varlag,nreps,nperiods,isconstant,iscontemp,yw,)
+
+T = size(errmat,1);
+neqs = size(errmat,2);
+
+irfs1 = zeros(nreps,neqs,nperiods);
+irfs2 = irfs1;
+
+bad_draw=0;
+for confindx=1:nreps
+    confindx
+    issingular=-1;
+    
+    %stay in while loop until artificial data is generated that leads a nonsingular VAR    
+    while (issingular<0) % keep drawing a new monte carlo series until one is found
+        % that produces a covariance stationary VAR
+        
+        %%% Sample with replacement from the fitted residuals
+        errpos = round((T-1)*rand(T,1)+ones(T,1));
+%         for posindx=1:T
+%             errmonte(posindx,:)=errmat(errpos(posindx),:);
+%         end
+        errmonte=errmat(errpos,:);
+        
+        %%% Using VAR structure and bootstrapped residuals, compute new y data
+        %ymonte = mkymonteg(y,coefs,errmonte);
+        ymonte = mkymonte(y,coefs,errmonte);
+        if (yw == 0) 
+            %[coefsmonte,coverr,errmatmonte]=estimateg(varlag,ymonte);
+            varlagmat = varlag*ones(neqs);
+            startpos =1;
+            endpos = size(ymonte,1);
+            [coefsmonte,coverr,errmatmonte]=estimate(varlagmat,zeros(neqs),ones(neqs,1),ymonte,startpos,endpos);
+
+            [cofbmonte,const_bmonte]=ols2ar(coefsmonte(1:end,:),isconstant,iscontemp);
+            cofamonte = iscontemp*coefsmonte(1+isconstant:neqs+isconstant,:);
+        else    
+            [Amonte, a0monte, coverr] = ywestimate(ymonte,varlag);
+            [cofbmonte,const_bmonte]=yw2ar(Amonte,a0monte);
+            cofamonte = zeros(neqs);
+        end
+        
+        cofainvmonte = (eye(neqs)-cofamonte)^(-1);
+        cofcmonte=isconstant*cofainvmonte*const_bmonte; 
+        
+        
+        [issingular] = singulartest(cofbmonte);
+        if issingular < 0
+            bad_draw = bad_draw + 1;
+        end
[…]

### 2013-06-14 MATLAB/estimation_utilities/confint_symmetric.m

@@ -1,0 +1,70 @@
+
+function [confint_lb confint_ub] = confint_symmetric(cofb,const,errmat,y,varlag,nreps,isconstant,iscontemp,yw,h,y_data,xi10,P10,q,ntrain,forecast_path)
+
+T = size(errmat,1);
+neqs = size(errmat,2);
+
+nperiods = size(y_data,1);
+nim_mat = zeros(nreps,nperiods);
+
+
+bad_draw=0;
+for confindx=1:nreps
+    issingular=-1;
+    
+    %stay in while loop until artificial data is generated that leads a nonsingular VAR    
+    while (issingular<0) % keep drawing a new monte carlo series until one is found
+        % that produces a covariance stationary VAR
+        
+        %%% Sample with replacement from the fitted residuals
+        errpos = round((T-1)*rand(T,1)+ones(T,1));
+%         for posindx=1:T
+%             errmonte(posindx,:)=errmat(errpos(posindx),:);
+%         end
+        errmonte=errmat(errpos,:);
+        
+        %%% Using VAR structure and bootstrapped residuals, compute new y data
+        %ymonte = mkymonteg(y,coefs,errmonte);
+        ymonte = mkymonte(y,cofb,const,errmonte);
+        if (yw == 0) 
+            %[coefsmonte,coverr,errmatmonte]=estimateg(varlag,ymonte);
+            varlagmat = varlag*ones(neqs);
+            [coefsmonte,coverr,errmatmonte]=estimate(varlagmat,zeros(neqs),ones(neqs,1),ymonte);
+            
+
+            [cofbmonte,const_bmonte]=ols2ar(coefsmonte(1:end,:),isconstant,iscontemp);
+            cofamonte = iscontemp*coefsmonte(1+isconstant:neqs+isconstant,:);
+        else    
+            [Amonte, a0monte, coverr] = ywestimate(ymonte,varlag);
+            [cofbmonte,const_bmonte]=yw2ar(Amonte,a0monte);
+            cofamonte = zeros(neqs);
+        end
+        
+        cofainvmonte = (eye(neqs)-cofamonte)^(-1);
+        cofcmonte=isconstant*cofainvmonte*const_bmonte; 
+        
+        
+        [issingular] = singulartest(cofbmonte);
+        if issingular < 0
+            bad_draw = bad_draw + 1;
+        end
+        
+    end  %end of while loop
[…]

### 2013-06-14 MATLAB/estimation_utilities/eggmodfun.m

@@ -1,0 +1,166 @@
+function  [cofb,endog_,scof,eqtype_]=eggmodfun(modnam,parnam,defmodnam)
+% modnam is the name of the model file
+% the model file resides in the model directory whose path is hardwired
+% below
+
+% parnam is the name of the parameter file
+% the parameter file sits in the calling directory, that is recorded under
+% currentdir below
+
+% Based on /mq/home/aim/utility/solve.m
+% Edited for use with PC AIM - CM 8/98
+
+% Solve a linear rational expectations model with AIM.  Check the 
+% accuracy of the solution, display the roots of the system, and
+% compute the observable structure.  Uses Matlab version of parser.
+
+currentdir = cd;
+
+directory = [currentdir,'\'];
+%eval(['chdir ',directory])
+%olddirnam = dirnam;
+
+%  Parse model
+
+parseflag = 1;
+
+% clean up old model files
+if parseflag
+    eval(['!del ',modnam,'_*']);
+end
+
+%parnam = 'parmmod1';
+parmflag = 1;
+
+
+%  Request printflag
+printflag = 1;
+
+%  Modelez syntax parser:
+pcparsem;
+
+%  Define the parameter vector (always execute definitions file)
+if( length(param_) )
+   if( parmflag )
+       %capture the parameter file from the launching directory
+       %then revert to the model directory
+       eval(['chdir ',currentdir]);
+	   eval(parnam);
+       eval(['chdir ',directory])
+   end
+   if (nargin==3)
+	eval(defmodnam)
+   end 
+   [npar,ncols] = size(param_);
+end
+
+
+
+% Numerical tolerances for aim
+epsi   = 2.2e-16;
+condn  = 1.e-8;
+uprbnd = 1 + 1.e-6;
+
+if (printflag)
+   space;
+	disp('Numerical Tolerances:');
+	space
+	disp([' epsi       = ',num2str(epsi)]);
+	disp([' condn      = ',num2str(condn)]);
+	disp([' uprbnd - 1 = ',num2str(uprbnd-1)]);
+   space;
+end
+
+if(printflag & length(param_) )
+   space(2)
+   disp('Parameter values:')
+   space
+   disp('Name          Value')
+   tabulator(param_);
+   space(2)
+end
+
+% ---------------------------------------------------------------------
+% Construct structural coefficient matrix.
[…]

### 2013-06-14 MATLAB/estimation_utilities/eggmodfunestim.m

@@ -1,0 +1,138 @@
+function  [cofb,scof,endog_,eqtype_,eqname_,aimcode,err]=eggmodfunestim(modnam,parnam1,parnam2,parseflag,params,paramlabels)
+% modnam is the name of the model file
+% it sits in the calling directory
+
+% parnam is the name of the parameter file
+% the parameter file sits in the calling directory
+
+% Based on /mq/home/aim/utility/solve.m
+% Edited for use with PC AIM - CM 8/98
+
+% Solve a linear rational expectations model with AIM.  Check the 
+% accuracy of the solution, display the roots of the system, and
+% compute the observable structure.  Uses Matlab version of parser.
+
+currentdir = cd;
+
+directory = [currentdir,'\'];
+
+%  Parse model
+if nargin<5
+parseflag = 1;
+end
+
+% clean up old model files
+if parseflag
+    eval(['!del ',modnam,'_*']);
+end
+
+parmflag = 1;
+
+
+%  Request printflag
+% if nargin>4
+%     printflag = 0;
+% else
+%     printflag = 1;
+% end
+printflag=0;
+
+%  Modelez syntax parser:
+pcparsem;
+
+%  Define the parameter vector (always execute definitions file)
+if( length(param_) )
+   if( parmflag )
+	   eval(parnam1);
+       if nargin>4
+          nestims = max(size(params));
+          for paramindx = 1:nestims
+              eval([paramlabels(paramindx,:),'=params(paramindx);'])
+          end
+       end
+       eval(parnam2);
+   end
+   [npar,ncols] = size(param_);
+end
+
+
+
+% Numerical tolerances for aim
+epsi   = 2.2e-16;
+condn  = 1.e-8;
+uprbnd = 1 + 1.e-6;
+
+if (printflag)
+   space;
+	disp('Numerical Tolerances:');
+	space
+	disp([' epsi       = ',num2str(epsi)]);
+	disp([' condn      = ',num2str(condn)]);
+	disp([' uprbnd - 1 = ',num2str(uprbnd-1)]);
+   space;
+end
+
+if(printflag & length(param_) )
+   space(2)
+   disp('Parameter values:')
+   space
+   disp('Name          Value')
+   tabulator(param_);
+   space(2)
+end
+
+% ---------------------------------------------------------------------
+% Construct structural coefficient matrix.
+% ---------------------------------------------------------------------
+
+%  When using pcparsem run compute_aim_matrices directly as a script.
+eval([modnam,'_aim_matrices']);
+
[…]

### 2013-06-14 MATLAB/estimation_utilities/estimate.m

@@ -1,0 +1,120 @@
+function [coefs,errcov,errmat]=estimate(varlags,contemp,constant,y)
+
+% coefs will store the regression coefficients
+% each equation is on a different column
+% if there is a constant, the first entry is the intercept term
+% then variables entering contemporaneously,
+% then the first lags, and so on.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%% Check structure of the VAR
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+neqs = size(varlags,1);
+nvars = size(varlags,2);
+
+constant = ones(neqs,1);  % if the nth entry of constant is 1 
+                         % a constant is included in the nth equation of the VAR
+if max(constant)==1
+    isconstant=1;
+else
+    isconstant=0;
+end
+
+
+                      
+% check that none of the
+% diagonal elements of contemp is 1
+% if so, one would be regressing a variable on itself.
+contempcheck=diag(contemp);
+if max(contempcheck)>=1 
+    error('The matrix contemp cannot have ones along its diagonal')
+end
+                      
+% Count number of variables entering contemporaneously in the VAR                                 
+countcontemp=contemp'*ones(neqs,1);
+ncontemp=0;    
+for indxi=1:neqs
+    if countcontemp(indxi)>0
+       ncontemp=ncontemp+1;
+    end
+end
+
+% set iscontemp to 1 if there are any variables entering contemporaneously
+if ncontemp>0
+   iscontemp=1;
+else
+   iscontemp=0;
+end
+                
+varmaxlag=max(max(varlags));
+
+if (neqs~=nvars) 
+    error('nlags needs to be a square matrix')
+end 
+
+if (size(y,2)~=neqs)
+   error('y needs to have as many columns as varlags')
+end
+
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%% Estimate the VAR    
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% coefs will store the regression coefficients
[…]

### 2013-06-14 MATLAB/estimation_utilities/estimate_ols.m

@@ -1,0 +1,3 @@
+function beta = estimate_ols(yreg,xreg)
+
+beta = mldivide((xreg*xreg'),xreg'*yreg);

### 2013-06-14 MATLAB/estimation_utilities/estimfun.m

@@ -1,0 +1,93 @@
+function [optparams, exitflag, fval, ...
+          fitted, smoothedhistory, ...
+          errhistory b0]=estimfun(paramlabels,params,upperbound,lowerbound,...
+                                  observedlabels,fittedlabels,modnam,parnam1,parnam2,ntrain)
+      
+global zdata slist errlist slistfull errlistfull
+
+
+eval(parnam1); 
+eval(parnam2);
+
+ 
+
+%filename = 'U:\TQS\luca\SECOND_DERIVATIVE\oilmod\NODLL\zdatanew_010510.txt';
+%exportmat(zdata',filename);
+%save estimdata zdata
+
+listdrop = makelistdrop(strvcat(slist,errlist),strvcat(slistfull,errlistfull));
+
+%[likel,q,history,amatsmall,bmatsmall,endogsmall smoothedhistory] = updatemod(params,paramlabels,modnam,parnam1,parnam2,observedlabels,upperbound,lowerbound,listdrop);
+
+display('Starting optimization')
+
+%options = optimset('HessUpdate','bfgs','display','iter','MaxIter',1e10,'MaxFunEvals',1e10,'TolFun',10e-3,'TolX',1e-4);
+
+%code for tomlab optimizer
+%options = optimset;
+%Prob=ProbDef;
+%Prob.Solver.Tomlab='snopt'
+% options = optimset('HessUpdate','bfgs','display','iter','MaxIter',5e2,'MaxFunEvals',1e10,'TolFun',10e-3,'TolX',1e-4);
+
+
+options = optimset('HessUpdate','bfgs','display','iter','MaxIter',1000,'MaxFunEvals',1e10,'TolFun',10e-3,'TolX',1e-4);
+
+
+
+% fminunc
+%
+%[likel,optparams,firstd,secondd] = csminwel('updatemod',params,0.01*eye(max(size(params))),[],10e-4,10000,paramlabels,modnam,parnam1,parnam2,observedlabels,upperbound,lowerbound,listdrop)
+
+
+optparams = params;
+%annealing
+for j = 1:10
+% loss = @(x0)updatemod(x0,paramlabels,modnam,parnam1,parnam2,observedlabels,upperbound,lowerbound,listdrop,ntrain);
+% option = anneal();
+% option.Verbosity = 2;
+% [optparams, fval] = anneal(loss,params);
+% exitflag = 0;
+
+
+%fminsearch
+for i = 1:3
[…]

### 2013-06-14 MATLAB/estimation_utilities/get_h_prime.m

@@ -1,0 +1,10 @@
+function h_prime = get_h_prime(tau, lambda)
+
+if size(tau,2)>size(tau,1)
+    tau=transpose(tau);
+end
+h1 = ones(length(tau),1);
+h2 = ((1-exp(-tau*lambda))./(tau*lambda));
+h3 = ((1-exp(-tau*lambda))./(tau*lambda)) - exp(-tau*lambda);
+
+h_prime = [h1 h2 h3];

### 2013-06-14 MATLAB/estimation_utilities/get_simple_ols.m

@@ -1,0 +1,7 @@
+function [beta, yreg, xreg] = get_simple_ols(y,x)
+
+% run ols of y on lag of y and lag of x and a constant
+
+yreg = y(2:end);
+xreg = [y(1:end) x(1:end-1,:) ones(length(y)-1,1)];
+beta = estimate_ols(yreg,xreg);

### 2013-06-14 MATLAB/estimation_utilities/kalmanFilterMatlab.m

@@ -1,0 +1,85 @@
+% !the inputs follow the notation in Hamilton, section 13.2
+% !
+% !f is the transition matrix for the hidden state
+% !h relates the observable state to the hidden state
+% !y holds the observed data
+% !x10 is the initial prior on the initial state vector
+% !q is the variance covariance matrix on error term in the hidden state
+% !
+% !likel is the log likelihood
+% !x_t_t-1 is the history of priors
+% !x_t_t is the history of posteriors
+
+% ! full(T)         => f
+% ! full(Z')        => h
+% ! full(zdata')    => y
+% ! full(b0)        => xi10
+% ! full(Q)         => q
+% ! ntrains         => ntrain
+
+% Note:  Not Checking Input
+function [logLikel,errcode]= kalmanFilterMatlab(f, h, y, a, x, xi10, p10, q, r, ntrain)
+
+tol0 = 1e-5;
+errcode = 0;
+likel = 0;
+logLikel = 0;
+
+n = size(h,2);
+t = size(y,2);
+
+xi10History(:,1) = xi10(:,1);
+
+xi11History(:,1) = xi10History(:,1);
+
+likel = zeros(t,1);
+% see formula in 13.4
+for i=1:t
+    if min(diag(p10))<0+tol0
+        logLikel = -10e300; errcode=-1;
+    else
+        
+        det_hpp10hr = det(h'*p10*h+r);
+        if cond(h'*p10*h+r)<1e10
+            %(det_hpp10hr>10e-300)
+            hP_p10_hr = (h'*p10*h+r)^-1;
+            part1(:,1) = y(:, i) - a'*x - h'*xi10History(:,1);
+            
+            likel(i) = 0;
+            if i > ntrain
+                part2 = -0.5 * part1'*hP_p10_hr*part1;
+                
+                if det(h'*p10*h+r)<0
+                likel(i) = - 10e300;
+                errcode = -1;
+                else
+                likel(i) = (-n/2)*log(2*pi)-0.5*log(det(h'*p10*h+r))+part2;
+                end
+                if logLikel+likel > 10e300
+                    errcode = -1;
+                    logLikel = -10e300;
+                else
+                    logLikel=logLikel + likel(i);
+                end
+            end
+            
+            % 13.2.15
+            xi11History(:,1) = xi10History(:,1) + p10*h*hP_p10_hr*part1(:,1);
+            
+            % 13.2.16
+            p11 = p10 - p10*h*hP_p10_hr*h'*p10;
+            
[…]

### 2013-06-14 MATLAB/estimation_utilities/kalmanFilterMatlab_singleob.m

@@ -1,0 +1,68 @@
+% !the inputs follow the notation in Hamilton, section 13.2
+% !
+% !f is the transition matrix for the hidden state
+% !h relates the observable state to the hidden state
+% !y holds the observed data
+% !x10 is the initial prior on the initial state vector
+% !q is the variance covariance matrix on error term in the hidden state
+% !
+% !likel is the log likelihood
+% !x_t_t-1 is the history of priors
+% !x_t_t is the history of posteriors
+
+% ! full(T)         => f
+% ! full(Z')        => h
+% ! full(zdata')    => y
+% ! full(b0)        => xi10
+% ! full(Q)         => q
+% ! ntrains         => ntrain
+
+% Note:  Not Checking Input
+function [logLikel,errcode]= kalmanFilterMatlab_singleob(f, h, y, xi10,p10,q,r,ntrain)
+
+
+errcode = 0;
+likel = 0;
+logLikel = 0;
+
+n = size(h,2);
+t = size(y,2);
+
+xi10History(:,1) = xi10(:,1);
+
+xi11History(:,1) = xi10History(:,1);
+
+% see formula in 13.4
+for i=1:t
+    hP_p10_h = (h'*p10*h)^-1;
+    part1(:,1) = y(:, i) - h'*xi10History(:,1);
+    
+    if i > ntrain
+        part2 = -0.5 * part1'*hP_p10_h*part1;
+        det_hpp10h = det(h'*p10*h);
+        if (det_hpp10h>10e-300)
+            likel = (-n/2)*log(2*pi)-0.5*log(det(h'*p10*h))+part2;
+            if logLikel+likel > 10e300
+                errcode = -1;
+                logLikel = 10e300;
+            else
+                logLikel=logLikel + likel;
+            end
+        else 
+            errcode = -1;
+        end
+    end
+    
+    % 13.2.15
+    xi11History(:,1) = xi10History(:,1) + p10*h*hP_p10_h*part1(:,1);
+
+    % 13.2.16
+    p11 = p10 - p10*h*hP_p10_h*h'*p10;
+
+    % 13.2.17
+    xi10History(:,1) = f*xi11History(:,1);
+
+    % 13.2.21
+    p10 = f*p11*f' + q;
+end
+logLikel = -logLikel;

### 2013-06-14 MATLAB/estimation_utilities/kalmanFilterSmoother.m

@@ -1,0 +1,87 @@
+% !the inputs follow the notation in Hamilton, section 13.2
+% !
+% !f is the transition matrix for the hidden state
+% !h relates the observable state to the hidden state
+% !y holds the observed data
+% !x10 is the initial prior on the initial state vector
+% !q is the variance covariance matrix on error term in the hidden state
+% !
+% !likel is the log likelihood
+% !x_t_t-1 is the history of priors
+% !x_t_t is the history of posteriors
+
+% ! full(T)         => f
+% ! full(Z')        => h
+% ! full(zdata')    => y
+% ! full(b0)        => xi10
+% ! full(Q)         => q
+% ! ntrains         => ntrain
+
+% Note:  Not Checking Input
+function [logLikel,errcode,xi10History,xi1tHistory,errHistory]= kalmanFilterSmoother(f, h, y, xi10, q, bmatsmall, ntrain)
+errcode = 0;
+likel = 0;
+logLikel = 0;
+
+n = size(h,2);
+t = size(y,2);
+nstates = size(xi10,1);
+xi10History =zeros(nstates,t+1);
+xi10History(:,1) = xi10(:,1);
+
+xi11History =zeros(nstates,t);
+errhistory = zeros(size(bmatsmall,2),t);
+
+% see 13.2.21
+p10 = q;
+p10History = zeros(nstates,nstates,t+1);
+p11History = zeros(nstates,nstates,t);
+p10History(:,:,1) = p10;
+
+for i=1:t
+    %(H'*(Pt|t-1)*H)^-1
+    hP_p10_h = (h'*p10*h)^-1;
+    part1(:,1) = y(:, i) - h'*xi10History(:,i);
+    
+    if i > ntrain
+        part2 = -0.5 * part1'*hP_p10_h*part1;
+        likel = (2*pi)^(-n/2)*det(h'*p10*h)^(-0.5)*exp(part2);
+        
+        if likel > exp(-300*log(10))
+            logLikel=logLikel + log(likel);
+        else
+            errcode = -1;
+        end
+    end
+    
+    % 13.6.5
+    xi11History(:,i) = xi10History(:,i) + p10*h*hP_p10_h*part1(:,1);
+
+    % 13.6.7
+    p11 = p10 - p10*h*hP_p10_h*h'*p10;
+    p11History(:,:,i) = p11;
[…]

### 2013-06-14 MATLAB/estimation_utilities/kalmanFilterSmoother_v2.m

@@ -1,0 +1,128 @@
+% !the inputs follow the notation in Hamilton, section 13.2
+% !
+% !f is the transition matrix for the hidden state
+% !h relates the observable state to the hidden state
+% !y holds the observed data
+% !x10 is the initial prior on the initial state vector
+% !q is the variance covariance matrix on error term in the hidden state
+% !
+% !likel is the log likelihood
+% !x_t_t-1 is the history of priors
+% !x_t_t is the history of posteriors
+
+% ! full(T)         => f
+% ! full(Z')        => h
+% ! full(zdata')    => y
+% ! full(b0)        => xi10
+% ! full(Q)         => q
+% ! ntrains         => ntrain
+
+% Note:  Not Checking Input
+function [logLikel,errcode,xi1tHistory,xi10History,xi11History]= ...
+    kalmanFilterSmoother_v2(f, h, y, a, x, xi10, p10, q, r, ntrain)
+
+tol0 = 1e-5;
+errcode = 0;
+likel = 0;
+logLikel = 0;
+
+n = size(h,2);
+t = size(y,2);
+
+nstates = size(xi10,1);
+xi10History =zeros(nstates,t+1);
+xi10History(:,1) = xi10(:,1);
+
+xi11History =zeros(nstates,t);
+%errHistory = zeros(size(q,2),t);
+
+% see 13.2.21
+p10History = zeros(nstates,nstates,t+1);
+p11History = zeros(nstates,nstates,t);
+p10History(:,:,1) = p10;
+
+likel = zeros(t,1);
+
+for i=1:t
+    if min(diag(p10))<0+tol0
+        logLikel = -10e300; errcode=-1;
+    else
+        det_hpp10hr = det(h'*p10*h+r);
+        if cond(h'*p10*h+r)<1e10
+            %(det_hpp10hr>10e-300)
+            hP_p10_hr_inv = (h'*p10*h+r)^-1;
+            part1(:,1) = y(:, i) - a'*x - h'*xi10History(:,i);
+            
+            likel(i) = 0;
+            if i > ntrain
+                part2 = -0.5 * part1'*hP_p10_hr_inv*part1;
+                
+                likel(i) = (-n/2)*log(2*pi)-0.5*log(det(h'*p10*h+r))+part2;
+                if logLikel+likel > 10e300
[…]

### 2013-06-14 MATLAB/estimation_utilities/makechart.m

@@ -1,0 +1,71 @@
+function makechart(titlelist,legendlist,figlabel,zdata1,zdata2,zdata3)
+
+
+figure
+
+titlelist = char(strrep(cellstr(titlelist),'_','.'));
+
+ndsets=3;       % default, changed below as applicable
+if nargin==4
+    zdata2=nan*zdata1;
+    zdata3=nan*zdata1;
+    ndsets =1;
+elseif nargin == 5
+    zdata3 =nan*zdata1;
+    ndsets=2;
+elseif ((nargin>6) | (nargin <=3))
+    error ('makechart takes 4 to 6 arguments')
+end
+
+nobs = size(zdata1,1);
+xvalues = (1:nobs)';
+
+nvars = size(titlelist,1);
+if nvars==1 
+    nrows=1;
+    ncols = 1;
+elseif nvars==2
+    nrows =2;
+    ncols = 1;
+elseif (nvars == 3 | nvars ==4)
+    nrows = 2;
+    ncols =2;
+elseif (nvars==5 |nvars ==6)
+    nrows = 3;
+    ncols = 2;
+elseif (nvars==7 | nvars==8)
+    nrows = 4;
+    ncols = 2;
+else 
+    error('too many variables (makechart)')
+end
+
+for i = 1:nvars
+    subplot(nrows,ncols,i)
+    h1=plot(xvalues,zdata1(:,i),'k-',xvalues,zdata2(:,i),'r--',xvalues,zdata3(:,i),'b:');
+    [x0 x1 y10 y11] = pickaxes(xvalues,zdata1(:,i));
+    [x0 x1 y20 y21] = pickaxes(xvalues,zdata2(:,i));
+    [x0 x1 y30 y31] = pickaxes(xvalues,zdata3(:,i));
+    y0 = min([y10,y20,y30]);
+    y1 = max([y11,y21,y31]);
+    if y0==y1
+        y1=y0+1;
+    end
+    
+    axis([x0 x1 y0 y1])
+    set(h1,'linewidth',2);
+    if i==1
+        legend(legendlist)
+        text('String',figlabel,'Units','normalized','Position',[1.2 1.24],...
+       'FontSize',14,'FontWeight','bold','HorizontalAlignment','center');
+    end
+ 
+    %set(gca,'XTick',xtick)
+    %set(gca,'XTickLabel',xticklabel)
+    
+    title(titlelist(i,:));
+    
+end
+
+% sets printing preferences
+% printpref

### 2013-06-14 MATLAB/estimation_utilities/makelistdrop.m

@@ -1,0 +1,15 @@
+function [listdrop]=makelistdrop(listkeep,listfull)
+
+nfull = size(listfull,1);
+rowdrop = [];
+
+% create list of variables in the state space to be dropped
+rowindx = 0;
+for indxi = 1:nfull
+    if (isempty(strmatch(deblank(listfull(indxi,:)),listkeep,'exact')))
+       rowindx = rowindx+1;
+       rowdrop(rowindx)=indxi;
+    end
+end
+
+listdrop = listfull(rowdrop,:);

### 2013-06-14 MATLAB/estimation_utilities/mkdata.m

@@ -1,0 +1,52 @@
+function [zdata]=mkdata(nperiods,cofb,endog_,wishlist,errlist,irfshock,scalefactormod)
+
+%[nsim, ksim, ysim, isim, csim] = mkdata(nperiods,cofb,endog_)
+
+% given decision rule 
+neqs = size(endog_,1);
+
+if nargin>5
+if isempty(scalefactormod);
+    scalefactormod=1;
+end
+end
+
+history = zeros(neqs,nperiods+1);
+
+if nargin>5
+    irfshockpos = strmatch(irfshock,endog_,'exact');
+end
+
+% find position of shocks in the vector of endogenous variables
+nerrs = size(errlist,1);
+errpos = zeros(size(errlist,1),1);
+for errindx = 1:nerrs
+    errpos(errindx) = strmatch(errlist(errindx,:),endog_,'exact');
+end
+
+% generate data
+% history will contain data, the state vector at each period in time will
+% be stored columnwise.
+history = zeros(neqs,nperiods+1);
+if nargin>5
+    history(irfshockpos,1) = 1/scalefactormod;
+end
+for i = 2:nperiods+1
+    % set shocks
+    if nargin<=5 
+        history(errpos,i-1) = randn(nerrs,1);
+    end
+    % update endogenous variables
+    history(:,i) = cofb * history(:,i-1);
+end
+
+% list of variables we wish to extract from data matrix
+%wishlist = char('eff','ukp','ugdp','uinv','ucc','kp');
+nwish=size(wishlist,1);
+wishpos = zeros(nwish,1);
+
+history=history';
+for i=1:nwish
+    wishpos(i) = strmatch(wishlist(i,:),endog_,'exact');
+end
+zdata = history(2:end,wishpos);

### 2013-06-14 MATLAB/estimation_utilities/mkymonte.m

@@ -1,0 +1,17 @@
+function montey = mkymonte(y,cofb,const,errmonte)
+
+errmonte=errmonte';
+
+nvars = size(y,2);
+nobs = size(y,1);
+montey = y';
+
+nlags = size(cofb,3);
+for obs=nlags+1:nobs
+        montey(:,obs) = errmonte(:,obs-nlags)+const;
+        for k=1:nlags
+            montey(:,obs) = montey(:,obs)+cofb(:,:,k)*montey(:,obs-k);
+        end 
+end
+
+montey=montey';

### 2013-06-14 MATLAB/estimation_utilities/pccheckaim_luca.m

@@ -1,0 +1,45 @@
+function err = pccheckaim_luca(neq, nlag, nlead, h, b)
+
+%   Substitute the reduced form, b, into the structural 
+%   equations, h, and display the maximum absolute error.
+
+% Append negative identity matrix to b
+
+b = [b, -eye(neq)];
+
+%  Define indexes into the lagged part (minus) and the current and
+%  lead part (plus) of h 
+
+minus =          1:           neq*nlag;
+plus  = neq*nlag+1: neq*(nlag+1+nlead);
+
+% Initialize q
+
+q = zeros(neq*(nlead+1), neq*(nlag+1+nlead));
+
+% Stuff b into the upper left-hand block of q
+
+[rb,cb] = size(b);
+q(1:rb,1:cb) = b(1:rb,1:cb);
+
+%  Copy b into the (nlead) row blocks of q, shifting right by neq once
+%  more for each block: this produces a coefficient matrix that 
+%  solves for (x(t),..., x(t+nlead))' in terms of (x(t-nlag),..., x(t-1))'.
+
+for i = 1:nlead
+   rows = i*neq + (1:neq);
+   q(rows,:) = shiftright( q(rows-neq,:), neq );
+end
+
+%  Premultiply the left block of q by the negative inverse of the right
+%  block of q
+
+q(:,minus) =  -q(:,plus)\q(:,minus);
+
+% Define the solution error
+
+error = h(:,minus) + h(:,plus)*q(:,minus);
+err = max(max(abs( error )));
+
+
+return

### 2013-06-14 MATLAB/estimation_utilities/runvar_general.m

@@ -1,0 +1,10 @@
+function [coefb,const,coverr] = runvar_general(var_data,varlag)
+
+%var_data is in column format
+
+nvars = size(var_data,2);
+varlagmat = varlag*ones(nvars);
+[coefs,coverr,errmat]=estimate(varlagmat,zeros(nvars),ones(nvars,1),var_data(1:end-14,:));
+isconstant = 1;
+iscontemp = 0;
+[coefb,const]=ols2ar(coefs(1:end,:),isconstant,iscontemp);

### 2013-06-14 MATLAB/estimation_utilities/shrinkspacesmall.m

@@ -1,0 +1,30 @@
+function [amatsmall, bmatsmall, endogsmall]=shrinkspacesmall(amat,bmat,endog,observed)
+% This function splits the state space into endogenous and control
+% variables (which are discarded)
+%
+% The function looks for columns in the law of motion whose entries are all zeros
+% those will correspond to the positions  of the control variables
+
+nendog = size(endog,1);
+tolzero = 10e-10;
+
+countc = 0;
+counte = 0;
+for i = 1:nendog
+     if ( max(abs(amat(:,i)))<tolzero & isempty( strmatch(deblank(endog(i,:)),observed,'exact') ) )
+         countc=countc+1;
+         controlselect(countc) = i;
+     else
+         counte = counte +1;
+         endogselect(counte) = i;
+     end
+ end
+
+ 
+ amatsmall = amat(endogselect,endogselect);
+ bmatsmall = bmat(endogselect,:);
+ 
+ 
+ endogsmall=endog(endogselect,:);
+
+ 

### 2013-06-14 MATLAB/estimation_utilities/splitmat.m

@@ -1,0 +1,74 @@
+function [amat,bmat,endogs,newerrlist]=splitmat(cofb,scof,endog,errlist,droplist)
+% given law of motion of the form
+% [x_t]' = cofb [x_{t-1}]' 
+% this function produces
+% amat and bmat such that
+% y_t = amat y_{t-1} + bmat e_t
+% 
+% the names of the variables in x_t are listed in endog
+% the names of the variables in e_t are listed in errlist
+% if 'one' is a variable name the column and the row corresponding to the
+% variable are excluded from cofb
+% if 'one' is not a variable name the program assumes that the model
+% conforms to the certainty-equivalent representation
+
+
+% find position of variables in errlist within endog
+nerr = size(errlist,1);
+nendog = size(endog,1);
+nendogs =  nendog - nerr;
+
+% find positions of variables to drop
+rowkeep = [];
+keepindx = 0;
+for indxi=1:nendog
+    if (isempty(strmatch(deblank(endog(indxi,:)),droplist,'exact')))
+        keepindx = keepindx+1;
+        rowkeep(keepindx)=indxi;
+    end
+end
+
+endog = endog(rowkeep,:);
+cofb = cofb(rowkeep,rowkeep);
+
+nendog = size(endog,1);
+nendogs =  nendog - nerr;
+
+% make a list of positions within endog of endogenous variables and
+% exogenous innovations
+endogposlist = zeros(nendogs,1);
+errposlist = zeros(nerr,1);
+endogscounter = 0;
+errcounter = 0;
+for i=1:nendog 
+   % case when variable in endog is not in errlist
+   if isempty(strmatch(deblank(endog(i,:)),errlist,'exact'))
+       endogscounter = endogscounter+1;
+       endogposlist(endogscounter) = i;
+   % case when variable is in errlist
+   else
+       errcounter = errcounter+1;
+       errposlist(errcounter) = i;
+   end
+end
+
+% update list of endogenous variables
+endogs = endog(endogposlist,:);
+newerrlist = endog(errposlist,:);
+
+% eliminate the column and the row of the 'one' variable (if it exists)   
+onepos= strmatch('one',endogs,'exact');
+if (~isempty(onepos))
+    selectvec = [1:onepos-1,onepos+1:length(endogposlist)];
+    endogposlist=endogposlist(selectvec);
+    endogs = endog(endogposlist,:);
+end    
+
[…]

### 2013-06-14 MATLAB/estimation_utilities/updatemod.m

@@ -1,0 +1,86 @@
+% the q was added to the output by Phillip, 7/13
+function [likel q history amatsmall bmatsmall endogsmall smoothedhistory] = updatemod(params,paramlabels,modnam,parnam1,parnam2,observedlabels,upperbound,lowerbound,listdrop,ntrain)
+                                                           
+
+% implements penalty function to ensure function evaluation is within
+% desired bounds
+
+% declare data matrix as global
+global zdata endog_ errlist slist 
+
+aimtol = 1e-4;
+
+% Define parameters and matrices for state and observable transition
+% equations.
+% 
+% The state evolution equation takes the form:
+% b_k = T b_{k-1} + q w_k.
+% The observable equation takes the form:
+% y_k = Z b_{k} + r e_k.
+% where w_k and e_k are drawn from a  multinormal distribution with variance sigma^2
+% 
+
+
+% check inputs and modify them if outside boundaries (impose penalty later)
+origparams = params;
+params = min(max(params,lowerbound),upperbound);
+
+% obtain decision rule
+printflag = 0;
+parseflag = 0;
+[cofb,scof,endog_,xxxx,xxxxx,aimcode,aimerr]=eggmodfunestim(modnam,parnam1,parnam2,parseflag,params,paramlabels);
+
+if (aimerr<aimtol & aimcode==1) 
+[amat,bmat,endogs]=splitmat(cofb,scof,endog_,errlist,listdrop);
+[amatsmall, bmatsmall, endogsmall]=shrinkspacesmall(amat,bmat,endogs,strvcat(observedlabels,slist));
+
+
+penalty =0;
+nparams = max(size(params));
+ 
+    T = amatsmall;
+
+    q = bmatsmall;
+     
+    Q= q*q';
+    
+    
+    nobserved = size(observedlabels,1);
+    nendogsmall = size(endogsmall,1);
+    b0 = zeros(nendogsmall,1);
+    Z = zeros(nobserved,nendogsmall);
+    for observedindx=1:nobserved
+        observedpos(observedindx)=strmatch(observedlabels(observedindx,:),endogsmall,'exact');
+        Z(observedindx,observedpos(observedindx))=1;
+        b0(observedpos(observedindx)) = zdata(1,observedindx);
+    end
+        
+  
+    r = zeros(nobserved,1);
+    R = r*r';
+
+    % Find unconditional variance of state vector
[…]

### 2013-06-14 MATLAB/estimation_utilities/updatemod_full.m

@@ -1,0 +1,75 @@
+% the q was added to the output by Phillip, 7/13
+function [likel q amatsmall bmatsmall endogsmall history smoothedhistory errhistory b0] = updatemod_full(params,paramlabels,modnam,parnam1,parnam2,observedlabels,upperbound,lowerbound,listdrop,ntrain)
+
+% implements penalty function to ensure function evaluation is within
+% desired bounds
+
+% declare data matrix as global
+global zdata endog_ errlist 
+
+% Define parameters and matrices for state and observable transition
+% equations.
+% 
+% The state evolution equation takes the form:
+% b_k = T b_{k-1} + q w_k.
+% The observable equation takes the form:
+% y_k = Z b_{k} + r e_k.
+% where w_k and e_k are drawn from a  multinormal distribution with variance sigma^2
+% 
+
+
+% check inputs and modify them if outside boundaries (impose penalty later)
+origparams = params;
+params = min(max(params,lowerbound),upperbound);
+
+% obtain decision rule
+printflag = 0;
+parseflag = 0;
+[cofb,scof,endog_,xxxx,xxxxx,aimcode]=eggmodfunestim(modnam,parnam1,parnam2,parseflag,params,paramlabels);
+
+[amat,bmat,endogs]=splitmat(cofb,scof,endog_,errlist,listdrop);
+
+%[amatsmall, bmatsmall, endogsmall]=shrinkspacesmall(amat,bmat,endogs,strvcat(observedlabels,slist));
+amatsmall = amat; bmatsmall = bmat; endogsmall = endogs;
+
+penalty =0;
+nparams = max(size(params));
+ 
+    T = amatsmall;
+
+    q = bmatsmall;
+     
+    Q= q*q';
+    
+    nobserved = size(observedlabels,1);
+    nendogsmall = size(endogsmall,1);
+    b0 = zeros(nendogsmall,1);
+    Z = zeros(nobserved,nendogsmall);
+    for observedindx=1:nobserved
+        observedpos(observedindx)=strmatch(observedlabels(observedindx,:),endogsmall,'exact');
+        Z(observedindx,observedpos(observedindx))=1;
+        b0(observedpos(observedindx)) = zdata(1,observedindx);
+    end
+        
+  
+    r = zeros(nobserved,1);
+    R = r*r';
+
+    % Find unconditional variance of state vector
+    % covb = varsigma(T,Q);
+    
[…]

### 2013-06-14 MATLAB/estimation_utilities/var_out_of_sample_forecasts.m

@@ -1,0 +1,18 @@
+function [xi] = var_out_of_sample_forecasts(f, var_data, varlag, nfactors, nothers, forecast_horizon)
+
+nstates = nfactors+nothers;
+
+xi = zeros((nstates*varlag)+1,forecast_horizon+1);
+
+for i_index = 1:varlag
+xi(1+(i_index-1)*nstates:i_index*nstates,1) = var_data(end-(i_index-1),:)';
+end
+
+xi(end,1) = 1;
+
+for i_index = 1:forecast_horizon
+    xi(:,i_index+1) = f*xi(:,i_index);
+end
+xi = xi(:,2:end);
+
+

### 2013-06-14 MATLAB/nim_lsc_model_1step/alt_data.m

@@ -1,0 +1,223 @@
+RTB = [8.16875
+7.476
+7.106363636
+7.165454545
+6.89953125
+6.138923077
+5.527878788
+5.353484848
+5.535
+5.656615385
+6.038030303
+5.864545455
+5.722615385
+6.214615385
+7.015
+7.732461538
+8.544615385
+8.402923077
+7.845076923
+7.653692308
+7.758769231
+7.746461538
+7.479846154
+6.996515152
+6.03078125
+5.558153846
+5.382121212
+4.546515152
+3.893538462
+3.680769231
+3.084545455
+3.071363636
+2.96015625
+2.966769231
+3.003333333
+3.060606061
+3.25125
+3.991384615
+4.478939394
+5.285076923
+5.735538462
+5.596153846
+5.367692308
+5.261846154
+4.932153846
+5.017846154
+5.097575758
+4.976060606
+5.05984375
+5.048461538
+5.04530303
+5.087727273
+5.0521875
+4.976307692
+4.824242424
+4.253030303
+4.4071875
+4.452615385
+4.65
+5.048484848
+5.525230769
+5.716153846
+6.018923077
+6.020461538
+4.819846154
+3.658769231
+3.189230769
+1.913787879
+1.7209375
+1.717384615
+1.644090909
+1.338636364
+1.156875
+1.041846154
+0.929242424
+0.916060606
+0.916615385
+1.077538462
+1.486666667
+2.013181818
+2.54296875
+2.864615385
+3.363333333
+3.828461538
+4.394
+4.706769231
+4.908307692
+4.904153846
+4.981846154
+4.736461538
+4.314461538
+3.403636364
+2.065076923
+1.623076923
+1.491969697
+0.301515152
+0.210625
+0.173230769
+0.156818182
+0.056666667
+0.10796875
+0.146461538
+0.156666667
+0.136818182
+0.12578125
+0.046461538
+0.023030303
+0.013230769
+0.198008611];
+
+
+RG10 = [11.624548
+10.916302
+10.482159
+9.9532108
+8.7948952
+7.9190407
+7.7363768
+7.612916
+7.4169621
+8.4737139
+8.9588184
+9.2136927
+8.5891619
+9.0370951
+9.2116394
+9.0436926
+9.2795486
+8.8600769
+8.2061853
+8.0263157
+8.5235233
[…]

### 2013-06-14 MATLAB/nim_lsc_model_1step/biclag.m

@@ -1,0 +1,99 @@
+function [optbiclag criterion] = biclag(vardata,startlag,endlag) 
+%%% Searches for the lag length that minimizes the BIC criterion
+
+% Estimate the VAR with a varying laglength
+% as determined by lagindx.
+% Then determine which laglength minimizes the BIC criterion
+
+criterion = zeros(endlag-endlag+1,1);
+if (startlag<1)
+    error('second argument to biclag must be a integer greater than zero');
+end
+
+for lagindx = startlag:endlag
+    
+    y = vardata;
+    startdt = 1;
+    enddt = size(y,1);
+    
+    varlags = lagindx*ones(size(y,2));
+    
+    neqs = size(varlags,1);
+    contemp = zeros(neqs);          % no contemporaneous variables present
+    
+    %%%%%% Check structure of the VAR
+    nvars = size(varlags,2);
+    
+    constant = ones(neqs,1);  % if the nth entry of constant is 1 
+    % a constant is included in the nth equation of the VAR
+    if max(constant)==1
+        isconstant=1;
+    else
+        isconstant=0;
+    end
+    
+    % check that none of the
+    % diagonal elements of contemp is 1
+    % if so, one would be regressing a variable on itself.
+    contempcheck=diag(contemp);
+    if max(contempcheck)>=1 
+        error('The matrix contemp cannot have ones along its diagonal')
+    end
+    
+    % Count number of variables entering contemporaneously in the VAR                                 
+    countcontemp=contemp'*ones(neqs,1);
+    ncontemp=0;    
+    for indxi=1:neqs
+        if countcontemp(indxi)>0
+            ncontemp=ncontemp+1;
+        end
+    end
+    
+    % set iscontemp to 1 if there are any variables entering contemporaneously
+    if ncontemp>0
+        iscontemp=1;
+    else
+        iscontemp=0;
+    end
+    
+    varmaxlag=max(max(varlags));
+    
+    if (neqs~=nvars) 
+        error('nlags needs to be a square matrix')
+    end 
+    
+    if (size(y,2)~=neqs)
+        error('y needs to have as many columns as varlags')
+    end
+    
+    %%%%% Estimate the VAR
+    
+    % coefs will store the regression coefficients
+    % each equation is on a different column
+    % if there is a constant, the first entry is the intercept term
[…]

### 2013-06-14 MATLAB/nim_lsc_model_1step/calc_rmse.m

@@ -1,0 +1,104 @@
+function [rmse_mlmat rmse_varmat rmse_nochangemat] = calc_rmse(opt_param_mat, out_of_sample_start_pos, end_sample_pos, yobs, factors, xi10, p10, ntrain, tau, nfactors, nothers, forecast_horizon)
+
+%%%Maximum-Likelihood pseudo-out of sample forecasts%%%
+nstates = nfactors+nothers;
+num_periods = end_sample_pos - out_of_sample_start_pos;
+errors_mat = zeros(num_periods, forecast_horizon);
+rmse_mat = zeros(num_periods, 1);
+
+for pos_index = out_of_sample_start_pos:end_sample_pos
+    
+    param_vec = opt_param_mat(:,pos_index-out_of_sample_start_pos+1);
+    
+    [f, q, r, x, a, lambda, h, xi_means, error] = kalmanFilterSetup(param_vec,tau,nfactors,nothers);
+    
+    [logLikel,errcode,xi1tHistory,xi10History,xi11History]= ...
+        kalmanFilterSmoother_v2(f, h, yobs, a, x, xi10, p10, q, r, ntrain);
+    
+    xi = zeros(nstates,forecast_horizon+1);
+    
+    xi(:,1) = xi11History(:,pos_index-1);
+    
+    for i_index = 1:forecast_horizon
+        xi(:,i_index+1) = f*xi(:,i_index);
+    end
+    xi = xi(:,2:end);
+    
+    forecasts = h'*(xi + kron(ones(1,forecast_horizon),xi_means));
+    forecast_errors = zeros(1,forecast_horizon);
+    
+    if pos_index <= end_sample_pos - forecast_horizon
+        forecast_errors = yobs(end,pos_index:pos_index+forecast_horizon-1) - forecasts(end,:);
+        errors_mat(pos_index-out_of_sample_start_pos+1,:) = forecast_errors;
+    else
+        forecast_errors(1:end_sample_pos-pos_index) = yobs(end,pos_index:end_sample_pos-1) - forecasts(end,1:end_sample_pos-pos_index);
+        forecast_errors(forecast_horizon - (end_sample_pos - pos_index)+1:end) = NaN;
+        errors_mat(pos_index-out_of_sample_start_pos+1,:) = forecast_errors;
+    end
+    
+end
+
+rmse_mlmat = (nansum((errors_mat.^2))./(size(errors_mat,1)-sum(isnan(errors_mat)))).^0.5;
+
+
+%%
[…]

### 2013-06-14 MATLAB/nim_lsc_model_1step/calc_rmse_conditional.m

@@ -1,0 +1,170 @@
+function [rmse_mlmat rmse_varmat rmse_nochangemat ml_forecasts nochange_forecasts] = calc_rmse(opt_param_mat, out_of_sample_start_pos, end_sample_pos, yobs, factors, xi10, p10, ntrain, tau, nfactors, nothers, forecast_horizon, dates)
+
+plotbool =1
+
+%%%Maximum-Likelihood pseudo-out of sample forecasts%%%
+nstates = nfactors+nothers;
+num_periods = end_sample_pos - out_of_sample_start_pos;
+errors_mat = zeros(num_periods, forecast_horizon);
+rmse_mat = zeros(num_periods, 1);
+
+ntau = length(tau);
+
+ml_forecasts = nan*zeros(forecast_horizon,end_sample_pos - out_of_sample_start_pos+1);
+for pos_index = out_of_sample_start_pos:end_sample_pos
+    
+    param_vec = opt_param_mat(:,pos_index-out_of_sample_start_pos+1);
+    
+    [f, q, r, x, a, lambda, h, xi_means, error] = kalmanFilterSetup(param_vec,tau,nfactors,nothers);
+    
+    [logLikel,errcode,xi1tHistory,xi10History,xi11History,p10History]= ...
+        kalmanFilterSmoother_v2(f, h, yobs , a, x, xi10, p10, q, r, ntrain);
+    
+    
+    xi10new = xi10History(:,pos_index);
+    p10new = p10History(:,:,pos_index);
+    if pos_index <= end_sample_pos - forecast_horizon
+        [logLikel,errcode,xi1tHistory,xi10History,xi11History]= ...
+            kalmanFilterSmoother_v2(f, h(:,1:ntau), yobs(1:ntau,pos_index:pos_index+forecast_horizon-1), a(1:ntau,1:ntau), x(1:ntau,:), xi10new, p10new, q, r(1:ntau,1:ntau), ntrain);
+    else
+        [logLikel,errcode,xi1tHistory,xi10History,xi11History]= ...
+        kalmanFilterSmoother_v2(f, h(:,1:ntau), yobs(1:ntau,pos_index:end_sample_pos), a(1:ntau,1:ntau), x(1:ntau,:), xi10new, p10new, q, r(1:ntau,1:ntau), ntrain);
+    end
+    
+   
+    forecast_errors = zeros(1,forecast_horizon);
+    
+    
+    
+    if pos_index <= end_sample_pos - forecast_horizon
[…]

### 2013-06-14 MATLAB/nim_lsc_model_1step/getestimate.m

@@ -1,0 +1,117 @@
+%% first step regression
+% data starts in 1974:1 and ends in 2005:6
+load monthdata.txt
+
+%varlag = [12 12 12];
+%[coefs,coverr,errmat,icode]=estimateg(varlag,monthdata);
+
+varlag = 12;
+varlagmat = varlag*ones(3);
+startpos =1;
+endpos = size(monthdata,1);
+[coefs,coverr,errmat]=estimate(varlagmat,zeros(3),[1 1 1]',monthdata,startpos,endpos);
+
+
+coverr = errmat'*errmat/(size(errmat,1)-sum(sum(varlagmat)));
+a0inv = transpose(chol(coverr));
+
+% separate out the standard deviation from the
+% diagonal of the cholesky decomposition
+sd = diag(a0inv);
+a0inv = a0inv*diag(sd.^-1);
+
+
+% get the IRFs
+isconstant = 1;
+iscontemp = 0;
+nperiods = 24;
+yw=0;     % switch for the confidence intervals (0 = OLS, 1=Yule-Walker estimation)
+nreps = 20;  % repetitions for Monte Carlo confidence intervals 
+[cofb,const_b]=ols2ar(coefs(1:end,:),isconstant,iscontemp);
+
+shock = [sd(1) 0 0]';
+history=mkirf(a0inv*shock,cofb,nperiods);
+IRF_1 = history(:,max(varlag+1):nperiods+max(varlag))';
+rand('seed',125);       % set the seed for generating Monte Carlo samples
+shock = [1 0 0]';
+[upperbound_1,lowerbound_1] = ...
+         confint(coefs,errmat,monthdata,varlag,nreps,nperiods,isconstant,iscontemp,yw,shock,IRF_1);
+
+
+shock = [0 sd(2) 0]';
+history=mkirf(a0inv*shock,cofb,nperiods);
+IRF_2 = history(:,max(varlag+1):nperiods+max(varlag))';
+rand('seed',125);       % set the seed for generating Monte Carlo samples
+shock = [0 1 0]';
+[upperbound_2,lowerbound_2] = ...
+         confint(coefs,errmat,monthdata,varlag,nreps,nperiods,isconstant,iscontemp,yw,shock,IRF_2);
+
+
+
+shock = [0 0 sd(3)]';
[…]

### 2013-06-14 MATLAB/nim_lsc_model_1step/kalmanFilterSetup.m

@@ -1,0 +1,78 @@
+function [f, q, r, x, a, lambda, h, xi_means, error] = kalmanFilterSetup(param_vec,tau,nfactors,nothers)
+
+ntaus = length(tau);
+nstates = nfactors + nothers;
+error = 0;
+
+tol0 = 1e-8;
+f = reshape(param_vec(1:(nfactors)^2),nfactors,nfactors);
+f_end = (nfactors)^2;
+f = [[f zeros(nfactors,nothers)]; reshape(param_vec(f_end+1:(f_end+nothers*(nfactors+nothers))), nothers, nfactors+nothers)];
+
+if max(abs(eig(f)))>1+tol0
+    error = -1;
+end
+
+q_start = f_end+nothers*(nfactors+nothers)+1;
+q_end =  f_end+nothers*(nfactors+nothers) + (nstates*(nstates+1))/2;
+q_all = reshape(param_vec(q_start:q_end), (nstates*(nstates+1)/2), 1);
+
+q = zeros(nstates);
+
+qdiag_start_counter = 1;
+qdiag_end_counter = 1;
+for diag_index = (nstates)-1:-1:1
+    q=q+diag(q_all(qdiag_start_counter:qdiag_end_counter),diag_index);
+    qdiag_start_counter=qdiag_end_counter+1;
+    qdiag_end_counter = qdiag_end_counter+1+(nstates)-diag_index;
+end
+q = (q + q') + diag(q_all(qdiag_start_counter:qdiag_end_counter));
+
+%    q = (diag(q_all(1), 2) + diag(q_all(2:3), 1));
+%    q = (q + q') + diag(q_all(4:6), 0);
+
+% check that q is positive semi-definite == det > 0    
+if  min(eig(q))< 0
+    error = -1;
+end
+
+r_start = q_end+1;
+r_end = q_end + ntaus+nothers;
+
+% check that r is positive semi-definite == det > 0    
+if min(param_vec(r_start:r_end))< 0
+    error = -1;
+end
+
+r = diag(param_vec(r_start:r_end));
+
+Mu_L = param_vec(r_end+1, 1);
+Mu_S = param_vec(r_end+2, 1);
+Mu_C = param_vec(r_end+3, 1);
+Mu_Nims = param_vec(r_end+4, 1);
+
+% check that lambda is positive
+lambda = param_vec(size(param_vec,1));
[…]

### 2013-06-14 MATLAB/nim_lsc_model_1step/load_data.m

@@ -1,0 +1,553 @@
+% The VAR in this version does not include the core deposit share.
+% With that exclusion, the VAR estimated by OLS is stationary and this
+% version sticks with OLS.
+clear
+
+setpath
+% from NIM_VAR_TINY_DATA.xls
+% 1987q1 to 2012q1
+% STFBNI_MA_N.Q	STFBAI_MA_N.Q	STFBLDT_XEOP_XDO_MA_N.Q	STFBLDNS_XEOP_XDO_MA_N.Q	STFBLD_XEOP_MA_N.Q	STFBEIDT_XDO_MA_N.Q	STFBEIDNS_XDO_MA_N.Q	STFBLIDT_XQA_XDO_MA_N.Q	STFBLIDNS_XQA_XDO_MA_N.Q
+
+rochelle_varlist = char('assets_securities_notrading','assets_securities_trading','assets_loans_realestate','assets_loans_commercial','assets_loans_consumer','assets_loans_other',...
+'revenue_securities_notrading','revenue_securities_trading','revenue_loans_realestate','revenue_loans_commercial','revenue_loans_consumer','revenue_loans_other','expenses_total');
+
+rochelle_data = xlsread('interestbearing_assets_revenues_expenses.xls');
+
+rochelle_dates = 1985:.25:2012;
+
+tiny_varlist = char('STFBNI','STFBAI','STFBLDT','STFBLDNS','STFBLD','STFBEIDT','STFBEIDNS','STFBLIDT','STFBLIDNS');
+
+tiny_dates = 1987:.25:2012;
+
+tiny_data=[24018	2556935	611906	545584	2231638	1992	6920	198587	539568
+24615	2554670	621709	539989	2254553	2069	6989	178516	542699
+25213	2582481	602755	533973	2256433	2120	7174	177503	538322
+25972	2622064	641163	531548	2324992	2176	7329	182212	532748
+25385	2660765	604212	542228	2308097	2181	7167	186825	535070
+25874	2691847	628041	546747	2345436	2237	7274	189739	542270
+26740	2718968	617443	537174	2363747	2299	7641	191343	543655
+29827	2741210	658068	537314	2419906	2359	7851	194406	539975
+27960	2766355	612309	524001	2416492	2346	7866	197214	527063
+28027	2807581	619060	509158	2439362	2326	8036	192787	510744
+27664	2841387	606650	519846	2454172	2323	8036	191176	515349
[…]

### 2013-06-14 MATLAB/nim_lsc_model_1step/load_data_Rochelle.m

@@ -1,0 +1,599 @@
+% The VAR in this version does not include the core deposit share.
+% With that exclusion, the VAR estimated by OLS is stationary and this
+% version sticks with OLS.
+clear
+
+setpath
+% from NIM_VAR_TINY_DATA.xls
+% 1987q1 to 2012q1
+% STFBNI_MA_N.Q	STFBAI_MA_N.Q	STFBLDT_XEOP_XDO_MA_N.Q	STFBLDNS_XEOP_XDO_MA_N.Q	STFBLD_XEOP_MA_N.Q	STFBEIDT_XDO_MA_N.Q	STFBEIDNS_XDO_MA_N.Q	STFBLIDT_XQA_XDO_MA_N.Q	STFBLIDNS_XQA_XDO_MA_N.Q
+
+rochelle_varlist = char('assets_securities_notrading','assets_securities_trading','assets_loans_realestate',...
+                        'assets_loans_commercial','assets_loans_consumer','assets_loans_other','assets_other',...
+                        'revenue_securities_notrading','revenue_securities_trading','revenue_loans_realestate',...
+                        'revenue_loans_commercial','revenue_loans_consumer','revenue_loans_other','revenue_other',...
+                        'expenses_total');
+
+nassets = 7;                    
+rochelle_data = xlsread('interestbearing_assets_revenues_expenses_allcategories.xls');
+
+rochelle_dates = 1985:.25:2012.25;
+
+% ensure that rochelle_data has the same start as all other data needed in
+% the VAR
+startpos = find(1987 == rochelle_dates);
+endpos = find(2012 == rochelle_dates);
+% luca changed this -- check with Val
+rochelle_data = rochelle_data(startpos:endpos,:);
+
+nvars_rochelle = size(rochelle_varlist,1);
+for thisvar = 1:nvars_rochelle
+    eval([rochelle_varlist(thisvar,:),'=rochelle_data(:,thisvar);'])
+end
+
+tiny_varlist = char('STFBNI','STFBAI','STFBLDT','STFBLDNS','STFBLD','STFBEIDT','STFBEIDNS','STFBLIDT','STFBLIDNS');
+
+tiny_dates = 1987:.25:2012;
+
+tiny_data=[24018	2556935	611906	545584	2231638	1992	6920	198587	539568
+24615	2554670	621709	539989	2254553	2069	6989	178516	542699
+25213	2582481	602755	533973	2256433	2120	7174	177503	538322
[…]

### 2013-06-14 MATLAB/nim_lsc_model_1step/load_data_Rochelle_loop.m

@@ -1,0 +1,632 @@
+% The VAR in this version does not include the core deposit share.
+% With that exclusion, the VAR estimated by OLS is stationary and this
+% version sticks with OLS.
+clear
+
+setpath
+% from NIM_VAR_TINY_DATA.xls
+% 1987q1 to 2012q1
+% STFBNI_MA_N.Q	STFBAI_MA_N.Q	STFBLDT_XEOP_XDO_MA_N.Q	STFBLDNS_XEOP_XDO_MA_N.Q	STFBLD_XEOP_MA_N.Q	STFBEIDT_XDO_MA_N.Q	STFBEIDNS_XDO_MA_N.Q	STFBLIDT_XQA_XDO_MA_N.Q	STFBLIDNS_XQA_XDO_MA_N.Q
+
+rochelle_varlist = char('assets_securities_notrading','assets_securities_trading','assets_loans_realestate',...
+                        'assets_loans_commercial','assets_loans_consumer','assets_loans_other','assets_other',...
+                        'revenue_securities_notrading','revenue_securities_trading','revenue_loans_realestate',...
+                        'revenue_loans_commercial','revenue_loans_consumer','revenue_loans_other','revenue_other',...
+                        'expenses_total');
+
+nassets = 7;                    
+rochelle_data = xlsread('interestbearing_assets_revenues_expenses_allcategories.xls');
+
+
+rochelle_dates = 1985:.25:2012.25;
+
+% ensure that rochelle_data has the same start as all other data needed in
+% the VAR
+startpos = find(1987 == rochelle_dates);
+endpos = find(2012 == rochelle_dates);
+% luca changed this -- check with Val
+rochelle_data = rochelle_data(startpos:endpos,:);
+
+nvars_rochelle = size(rochelle_varlist,1);
+for thisvar = 1:nvars_rochelle
+    eval([rochelle_varlist(thisvar,:),'=rochelle_data(:,thisvar);'])
+end
+
+tiny_varlist = char('STFBNI','STFBAI','STFBLDT','STFBLDNS','STFBLD','STFBEIDT','STFBEIDNS','STFBLIDT','STFBLIDNS');
+
+tiny_dates = 1987:.25:2012;
+
+tiny_data=[24018	2556935	611906	545584	2231638	1992	6920	198587	539568
+    24615	2554670	621709	539989	2254553	2069	6989	178516	542699
+    25213	2582481	602755	533973	2256433	2120	7174	177503	538322
[…]

### 2013-06-14 MATLAB/nim_lsc_model_1step/load_data_cost_loop.m

@@ -1,0 +1,961 @@
+
+% to do:
+% update path for interest rates 
+
+
+clear
+
+move_deposit_share_down = 1;
+
+setpath
+% from NIM_VAR_TINY_DATA.xls
+% 1987q1 to 2012q1
+% STFBNI_MA_N.Q	STFBAI_MA_N.Q	STFBLDT_XEOP_XDO_MA_N.Q	STFBLDNS_XEOP_XDO_MA_N.Q	STFBLD_XEOP_MA_N.Q	STFBEIDT_XDO_MA_N.Q	STFBEIDNS_XDO_MA_N.Q	STFBLIDT_XQA_XDO_MA_N.Q	STFBLIDNS_XQA_XDO_MA_N.Q
+
+rochelle_varlist = char('assets_securities_notrading','assets_securities_trading','assets_loans_realestate',...
+                        'assets_loans_commercial','assets_loans_consumer','assets_loans_other','assets_other',...
+                        'revenue_securities_notrading','revenue_securities_trading','revenue_loans_realestate',...
+                        'revenue_loans_commercial','revenue_loans_consumer','revenue_loans_other','revenue_other',...
+                        'expenses_total');
+
+nassets = 7;                    
+rochelle_data = xlsread('interestbearing_assets_revenues_expenses_allcategories_updated.xls');
+
+
+rochelle_dates = 1985:.25:2012.5;
+
+% ensure that rochelle_data has the same start as all other data needed in
+% the VAR
+startpos = find(1987 == rochelle_dates);
+endpos = find(2012.5 == rochelle_dates);
+% luca changed this -- check with Val
+rochelle_data = rochelle_data(startpos:endpos,:);
+
+nvars_rochelle = size(rochelle_varlist,1);
+for thisvar = 1:nvars_rochelle
+    eval([rochelle_varlist(thisvar,:),'=rochelle_data(:,thisvar);'])
+end
+
+tiny_varlist = char('STFBNI','STFBAI','STFBLDT','STFBLDNS','STFBLD','STFBEIDT','STFBEIDNS','STFBLIDT','STFBLIDNS');
+
+tiny_dates = 1987:.25:2012.5;
+
+tiny_data=[24018	2556935	611906	545584	2231638	1992	6920	198587	539568
+24615	2554670	621709	539989	2254553	2069	6989	178516	542699
+25213	2582481	602755	533973	2256433	2120	7174	177503	538322
+25972	2622064	641163	531548	2324992	2176	7329	182212	532748
[…]

### 2013-06-14 MATLAB/nim_lsc_model_1step/load_data_cost_loop2.m

@@ -1,0 +1,1017 @@
+
+% to do:
+% update path for interest rates 
+
+
+clear
+
+move_deposit_share_down = 1;
+
+setpath
+% from NIM_VAR_TINY_DATA.xls
+% 1987q1 to 2012q1
+% STFBNI_MA_N.Q	STFBAI_MA_N.Q	STFBLDT_XEOP_XDO_MA_N.Q	STFBLDNS_XEOP_XDO_MA_N.Q	STFBLD_XEOP_MA_N.Q	STFBEIDT_XDO_MA_N.Q	STFBEIDNS_XDO_MA_N.Q	STFBLIDT_XQA_XDO_MA_N.Q	STFBLIDNS_XQA_XDO_MA_N.Q
+
+rochelle_varlist = char('assets_securities_notrading','assets_securities_trading','assets_loans_realestate',...
+                        'assets_loans_commercial','assets_loans_consumer','assets_loans_other','assets_other',...
+                        'revenue_securities_notrading','revenue_securities_trading','revenue_loans_realestate',...
+                        'revenue_loans_commercial','revenue_loans_consumer','revenue_loans_other','revenue_other',...
+                        'expenses_total');
+
+nassets = 7;                    
+rochelle_data = xlsread('interestbearing_assets_revenues_expenses_allcategories_updated.xls');
+
+
+rochelle_dates = 1985:.25:2012.5;
+
+% ensure that rochelle_data has the same start as all other data needed in
+% the VAR
+startpos = find(1987 == rochelle_dates);
+endpos = find(2012.5 == rochelle_dates);
+% luca changed this -- check with Val
+rochelle_data = rochelle_data(startpos:endpos,:);
+
+nvars_rochelle = size(rochelle_varlist,1);
+for thisvar = 1:nvars_rochelle
+    eval([rochelle_varlist(thisvar,:),'=rochelle_data(:,thisvar);'])
+end
+
+tiny_varlist = char('STFBNI','STFBAI','STFBLDT','STFBLDNS','STFBLD','STFBEIDT','STFBEIDNS','STFBLIDT','STFBLIDNS');
+
+tiny_dates = 1987:.25:2012.5;
+
+tiny_data=[24018	2556935	611906	545584	2231638	1992	6920	198587	539568
+24615	2554670	621709	539989	2254553	2069	6989	178516	542699
+25213	2582481	602755	533973	2256433	2120	7174	177503	538322
+25972	2622064	641163	531548	2324992	2176	7329	182212	532748
[…]

### 2013-06-14 MATLAB/nim_lsc_model_1step/load_data_ml.m

@@ -1,0 +1,68 @@
+function [yobs, dates, yields, nims, nfactors, nothers, tau, factors, shadow_bank_share_assets] = load_data_ml 
+
+start_yobs = 1985.75;
+end_yobs = 2008.25;
+dates = start_yobs:.25:end_yobs;
+%SVENY yields can be found in the 'yields' FAME db under the mnemonics in
+%yields_varlist.
+if isunix
+yields = csvread('/ofs/research2/Guerrieri/valentin/nim_lsc_model/data/yields_sveny_only_1985quarterly.csv', 1, 2);
+else
+yields = csvread('.\yields_sveny_only_1985quarterly.csv', 1, 2); 
+end
+dates_yields = 1985.75:.25:2013.0;
+start_pos = find(dates==start_yobs);
+end_pos = find(dates==end_yobs);
+yields = yields(start_pos:end_pos,:)';
+yields_varlist = char('SVENY0025','SVENY0050','SVENY0075','SVENY0100','SVENY0200','SVENY0300','SVENY0500', 'SVENY0700', 'SVENY1000', 'SVENY1500', 'SVENY2000', 'SVENY3000');
+
+%Bring in Net Interest Margins from Call Report (TINY); calculated as nim_top25 =
+%100*netintinc/interest earning assets (where each component in the ratio
+%is calculated at each quarter based on the top 25 firms, by assets
+
+if isunix
+nims = csvread('/ofs/research2/Guerrieri/valentin/data/tiny_nims/tiny_nims.csv', 1, 1);
+else
+nims = csvread('.\tiny_nims.csv', 1, 1);
+end    
+
+dates_nim = 1985.0:.25:2013.0;
+start_pos_nim = find(dates_nim==start_yobs);
+end_pos_nim = find(dates_nim==end_yobs);
+nims = 4*nims(start_pos_nim:end_pos_nim,1)'; %Multiply by 4 to annualize
+nims_varlist = char('nim_top25');
+
+%Bring in shadow banking share of financial sector assets. We use it as 
+%a measure of competitiveness in the banking sector.
+%Shadow banking share calculated as:
+%1-((FL704090005.Q+ FL734090005.Q)/((FL413065005.Q+FL674090005.Q+FL614090005.Q+FL664090005.Q+FL504090005.Q)+(FL704090005.Q+ FL734090005.Q)))
+%Data comes from Flow of Funds FAME db fof.
+
+if isunix
[…]

### 2013-06-14 MATLAB/nim_lsc_model_1step/ols1lag.m

@@ -1,0 +1,6 @@
+function [coefs] = ols1lag(nims, factors)
+
+yreg = transpose(nims(:,2:end));
+xreg = [transpose(factors(:,1:end-1)) transpose(nims(:,1:end-1))]; 
+
+coefs = (xreg'*xreg)\xreg'*yreg;

### 2013-06-14 MATLAB/nim_lsc_model_1step/optimize_likelihood.m

@@ -1,0 +1,135 @@
+
+clear;
+
+setpath;
+
+load_paramvec_from_disk = 1;  % set to 1 to load saved param_vec from disk
+                             % set to 0 to use first guess as defined below
+
+[yobs, dates, yields, nims, nfactors, nothers, tau, factors] = load_data_ml; 
+
+nstates = nfactors + nothers;
+ntaus = length(tau);
+
+ntrain = 15; %We have 153 observations per time t; we're setting ntrain to ~15% of observation pool.
+
+factor_means = mean(factors,2);
+nims_mean = mean(nims);
+
+[f_factors] = runvar(factors');
+
+% If OLS is preferred, uncomment the following section of code.
+%factors_demeaned = factors-kron(ones(1,size(factors,2)),factor_means);
+% CHANGE THIS 
+%xi10_demeaned = xi_demeaned(:,1); % Initial guess of unobserved states
+%var_data = factors_demeaned';
+
+%nvars = size(var_data,2);
+%varlag = 1;
+%varlagmat = varlag*ones(nvars);
+%[coefs,coverr,errmat]=estimate(varlagmat,zeros(nvars),ones(nvars,1),var_data(1:end-14,:));
+%isconstant = 1;
+%iscontemp = 0;
+%[f_factors,const]=ols2ar(coefs(1:end,:),isconstant,iscontemp);
+
+[f_nims] = transpose(ols1lag(nims, factors));
+
+f = [[f_factors zeros(nfactors,nothers)]; f_nims];
+
+[f_junk,coverr] = runvar([factors' nims']);
+
+q = coverr;
+p10 = coverr;
+
+%f = diag(0.5*ones(nstates,1));
+%q = diag(1*ones(nstates,1));
+r = 0.1*ones(ntaus+nothers,1);
+
+q_n = length(diag(q));
+r_n = length(diag(r));
+
+q_diag_vec = [];
+for this_n=1:q_n
+    q_diag_vec = [q_diag_vec 
+                  diag(q,q_n-this_n)];
+end
+
+%r_diag_vec = [];
+%for this_n=1:r_n
+%    r_diag_vec = [r_diag_vec 
+%                  diag(r,r_n-this_n)];
+%end
+
+lambda = .18;
+
+xi = [factors; nims];
+xi_means = mean(xi,2);
[…]

### 2013-06-14 MATLAB/nim_lsc_model_1step/out_of_sample_forecasts.m

@@ -1,0 +1,22 @@
+function [forecasts] = out_of_sample_forecasts(param_vec, yobs, xi10, p10, ntrain, tau, nfactors, nothers, forecast_horizon)
+
+%%%Maximum-Likelihood out of sample forecasts%%%
+
+[f, q, r, x, a, lambda, h, xi_means, error] = kalmanFilterSetup(param_vec,tau,nfactors,nothers);
+
+[logLikel,errcode,xi1tHistory,xi10History,xi11History]= ...
+    kalmanFilterSmoother_v2(f, h, yobs, a, x, xi10, p10, q, r, ntrain);
+
+nstates = nfactors+nothers;
+
+xi = zeros(nstates,forecast_horizon+1);
+
+xi(:,1) = xi11History(:,end);
+
+for i_index = 1:forecast_horizon
+    xi(:,i_index+1) = f*xi(:,i_index);
+end
+xi = xi(:,2:end);
+
+forecasts = h'*(xi + kron(ones(1,forecast_horizon),xi_means));
+

### 2013-06-14 MATLAB/nim_lsc_model_1step/run_calc_rmse.m

@@ -1,0 +1,58 @@
+clear;
+
+setpath;
+
+load_paramvec_from_disk = 1;  % set to 1 to load saved param_vec from disk
+% set to 0 to use first guess as defined below
+
+ntrain = 15; %We have 153 observations per time t; we're setting ntrain to ~15% of observation pool.
+
+[yobs, dates, yields, nims, nfactors, nothers, tau, factors] = load_data_ml; 
+
+out_of_sample_start_pos = find(dates==2000.0);
+end_sample_pos = length(dates);
+
+nstates = nfactors + nothers;
+ntaus = length(tau);
+
+load this_iter; %includes the full sample param_vec, xi10_demeaned, p10, ntrain
+
+param_vec_full_sample = param_vec;
+opt_param_mat = zeros(length(param_vec), end_sample_pos-out_of_sample_start_pos+1);
+options = optimset('HessUpdate','bfgs','display','iter','MaxIter',10000,'MaxFunEvals',1e10,'TolFun',10e-3,'TolX',1e-4);
+
+compute_param_mat = 0;
+
+if compute_param_mat
+for pos_index = out_of_sample_start_pos:end_sample_pos
+    [yobs, dates, yields, nims, nfactors, nothers, tau] = load_data_ml;
+    
+    display(['working on observation for date ', num2str(dates(pos_index))]);
+    
+    yobs = yobs(:,1:pos_index-1);
+    dates = dates(:,1:pos_index-1);
+    yields = yields(:,1:pos_index-1);
+    nims = nims(:,1:pos_index-1);
+    
+    
+    for fminsearch_rep = 1:1
+        loss = @(x0)mylikelihood(x0,yobs,xi10_demeaned,tau,nfactors,nothers,ntrain,p10);
+        option = anneal();
+        option.Verbosity = 2;
+        [opt_param_mat(:,pos_index-(out_of_sample_start_pos-1)), fval] = anneal(loss,param_vec_full_sample);
+                
+        [opt_param_mat(:,pos_index-(out_of_sample_start_pos-1)),fval,exitflag] = ...
+            fminsearch('mylikelihood',opt_param_mat(:,pos_index-(out_of_sample_start_pos-1)),options,yobs,xi10_demeaned,tau,nfactors,nothers,ntrain,p10);
+        
+        save this_out_of_sample opt_param_mat xi10_demeaned p10 ntrain
+        
+    end
+    
+end
+end
+%%
[…]

### 2013-06-14 MATLAB/nim_lsc_model_1step/run_calc_rmse_conditional.m

@@ -1,0 +1,90 @@
+clear;
+
+setpath;
+
+load_paramvec_from_disk = 1;  % set to 1 to load saved param_vec from disk
+% set to 0 to use first guess as defined below
+
+ntrain = 15; %We have 153 observations per time t; we're setting ntrain to ~15% of observation pool.
+
+[yobs, dates, yields, nims, nfactors, nothers, tau, factors] = load_data_ml; 
+
+out_of_sample_start_pos = find(dates==2000.0);
+end_sample_pos = length(dates);
+
+nstates = nfactors + nothers;
+ntaus = length(tau);
+
+load this_iter; %includes the full sample param_vec, xi10_demeaned, p10, ntrain
+
+param_vec_full_sample = param_vec;
+opt_param_mat = zeros(length(param_vec), end_sample_pos-out_of_sample_start_pos+1);
+options = optimset('HessUpdate','bfgs','display','iter','MaxIter',10000,'MaxFunEvals',1e10,'TolFun',10e-3,'TolX',1e-4);
+
+compute_param_mat = 0;
+
+if compute_param_mat
+for pos_index = out_of_sample_start_pos:end_sample_pos
+    [yobs, dates, yields, nims, nfactors, nothers, tau] = load_data_ml;
+    
+    display(['working on observation for date ', num2str(dates(pos_index))]);
+    
+    yobs = yobs(:,1:pos_index-1);
+    dates = dates(:,1:pos_index-1);
+    yields = yields(:,1:pos_index-1);
+    nims = nims(:,1:pos_index-1);
+    
+    
+    for fminsearch_rep = 1:1
+        loss = @(x0)mylikelihood(x0,yobs,xi10_demeaned,tau,nfactors,nothers,ntrain,p10);
+        option = anneal();
+        option.Verbosity = 2;
+        [opt_param_mat(:,pos_index-(out_of_sample_start_pos-1)), fval] = anneal(loss,param_vec_full_sample);
+                
+        [opt_param_mat(:,pos_index-(out_of_sample_start_pos-1)),fval,exitflag] = ...
+            fminsearch('mylikelihood',opt_param_mat(:,pos_index-(out_of_sample_start_pos-1)),options,yobs,xi10_demeaned,tau,nfactors,nothers,ntrain,p10);
+        
+        save this_out_of_sample opt_param_mat xi10_demeaned p10 ntrain
+        
[…]

### 2013-06-14 MATLAB/nim_lsc_model_1step/run_get_rmse.m

@@ -1,0 +1,58 @@
+clear;
+
+setpath;
+
+load_paramvec_from_disk = 1;  % set to 1 to load saved param_vec from disk
+% set to 0 to use first guess as defined below
+
+ntrain = 15; %We have 153 observations per time t; we're setting ntrain to ~15% of observation pool.
+
+[yobs, dates, yields, nims, nfactors, nothers, tau, factors] = load_data_ml; 
+
+out_of_sample_start_pos = find(dates==2000.0);
+end_sample_pos = length(dates);
+
+nstates = nfactors + nothers;
+ntaus = length(tau);
+
+load this_iter; %includes the full sample param_vec, xi10_demeaned, p10, ntrain
+
+param_vec_full_sample = param_vec;
+opt_param_mat = zeros(length(param_vec), end_sample_pos-out_of_sample_start_pos+1);
+options = optimset('HessUpdate','bfgs','display','iter','MaxIter',10000,'MaxFunEvals',1e10,'TolFun',10e-3,'TolX',1e-4);
+
+compute_param_mat = 1;
+
+if compute_param_mat
+for pos_index = out_of_sample_start_pos:end_sample_pos
+    [yobs, dates, yields, nims, nfactors, nothers, tau] = load_data_ml;
+    
+    display(['working on observation for date ', num2str(dates(pos_index))]);
+    
+    yobs = yobs(:,1:pos_index-1);
+    dates = dates(:,1:pos_index-1);
+    yields = yields(:,1:pos_index-1);
+    nims = nims(:,1:pos_index-1);
+    
+    
+    for fminsearch_rep = 1:1
+        loss = @(x0)mylikelihood(x0,yobs,xi10_demeaned,tau,nfactors,nothers,ntrain,p10);
+        option = anneal();
+        option.Verbosity = 2;
+        [opt_param_mat(:,pos_index-(out_of_sample_start_pos-1)), fval] = anneal(loss,param_vec_full_sample);
+                
+        [opt_param_mat(:,pos_index-(out_of_sample_start_pos-1)),fval,exitflag] = ...
+            fminsearch('mylikelihood',opt_param_mat(:,pos_index-(out_of_sample_start_pos-1)),options,yobs,xi10_demeaned,tau,nfactors,nothers,ntrain,p10);
+        
+        save this_out_of_sample opt_param_mat xi10_demeaned p10 ntrain
+        
+    end
+    
+end
+end
+%%
[…]

### 2013-06-14 MATLAB/nim_lsc_model_1step/run_out_of_sample_forecasts.m

@@ -1,0 +1,31 @@
+
+[yobs, dates, yields, nims, nfactors, nothers, tau, factors] = load_data_ml; 
+
+load this_iter; 
+
+forecast_horizon = 10;
+nobs_history = 90;
+
+[forecast] = out_of_sample_forecasts(param_vec, yobs, xi10_demeaned, p10, ntrain, tau, nfactors, nothers, forecast_horizon);
+
+figure
+plot(dates(end-nobs_history:end), yobs(end,end-nobs_history:end), 'k-');
+hold on
+plot(dates(end)+.25:.25:dates(end)+forecast_horizon/4, forecast(end,:), 'k--');
+
+
+%% Now redo for VAR
+
+var_data= [factors' nims'];
+
+varlag = 4;
+
+[coefb,const,coverr] = runvar_general(var_data,varlag);
+
+f_var=companion(coefb,const);
+
+[forecast_var] = var_out_of_sample_forecasts(f_var, var_data, varlag, nfactors, nothers, forecast_horizon);
+
+plot(dates(end)+.25:.25:dates(end)+forecast_horizon/4, forecast_var(nfactors+1,:), 'r--');
+
+vline(dates(end),'b');

### 2013-06-14 MATLAB/nim_lsc_model_1step/runvar.m

@@ -1,0 +1,13 @@
+function [coefb,coverr] = runvar(var_data)
+
+%var_data is in column format
+var_data = var_data - kron(ones(size(var_data,1),1),mean(var_data));
+
+
+nvars = size(var_data,2);
+varlag = 1;
+varlagmat = varlag*ones(nvars);
+[coefs,coverr,errmat]=estimate(varlagmat,zeros(nvars),ones(nvars,1),var_data(1:end-14,:));
+isconstant = 1;
+iscontemp = 0;
+[coefb,const]=ols2ar(coefs(1:end,:),isconstant,iscontemp);

### 2013-06-14 MATLAB/nim_lsc_model_1step/setpath.m

@@ -1,0 +1,11 @@
+
+restoredefaultpath
+
+path1 = './VARutilities';
+
+path (path,path1);
+
+
+dir2 = './plot_support';
+
+path(path,dir2);

### 2013-06-14 MATLAB/nim_lsc_model_1step/yield_calc.m

@@ -1,0 +1,17 @@
+function [y] = yield_calc(tau, lambda, xi_means, xi1t)
+
+if size(tau,2)>size(tau,1)
+tau = tau';
+end
+
+h_prime = get_h_prime(tau, lambda);
+
+%h = eye(3); %should be 3x3
+%h_prime = h';
+
+%Calculate the yield at given maturity using the measurement equation.
+%There are three instances of xi1tHistory because we have three state
+%variables.
+%We are adding in the constants (the Mu_L, Mu_S, and Mu_C) because
+%estimates of the state variables are de-meaned in xi1tHistory.
+y = h_prime*(xi1t + kron(ones(1,size(xi1t,2)), xi_means));

### 2013-06-14 MATLAB/nim_lsc_model_2steps/get_yield.m

@@ -1,0 +1,80 @@
+clear;
+
+setpath;
+
+load_paramvec_from_disk = 0;  % set to 1 to load saved param_vec from disk
+                             % set to 0 to use first guess as defined below
+
+%SVENY yields can be found in the 'yields' FAME db under the mnemonics in
+%yields_varlist.
+yields = csvread('/ofs/research2/Guerrieri/valentin/nim_lsc_model/data/yields_sveny_only_1985quarterly.csv', 1, 2);
+dates = 1985.75:.25:2013.0;
+end_pos = find(dates==2008.25);
+yields = yields(1:end_pos,:)';
+yields_varlist = char('SVENY0025','SVENY0050','SVENY0075','SVENY0100','SVENY0200','SVENY0300','SVENY0500', 'SVENY0700', 'SVENY1000', 'SVENY1500', 'SVENY2000', 'SVENY3000');
+
+nstates = 3;
+tau = [3 6 9 12 24 36 60 84 120 180 240 360];
+tau = tau([1,5,9]);
+yields = yields([1,5,9],:);
+ntaus = length(tau);
+lambda = .18;
+ntrain = 15; %We have 153 observations per time t; we're setting ntrain to ~15% of observation pool.
+
+factor1 = (yields(find(tau == 3),:)... 
+           +  yields(find(tau == 24),:)...
+           +  yields(find(tau == 120),:) )/3;
+       
+%factor1 = (yields(find(tau == 60),:));       
+       
+factor2 = (yields(find(tau == 3),:) - yields(find(tau == 120),:));
+
+factor3 = ( 2*yields(find(tau == 24),:) ...
+             - yields(find(tau == 120),:) ...
+             - yields(find(tau == 3),:)  )*5; 
+
+xi = [factor1; factor2; factor3];
+xi_means = mean(xi,2);
+
+% If OLS is preferred, uncomment the following section of code.
+xi_demeaned = xi-kron(ones(1,size(xi,2)),xi_means);
+xi10_demeaned = xi_demeaned(:,1); % Initial guess of unobserved states
+var_data = xi_demeaned';
+
+nvars = size(var_data,2);
+varlag = 1;
+varlagmat = varlag*ones(nvars);
+[coefs,coverr,errmat]=estimate(varlagmat,zeros(nvars),ones(nvars,1),var_data(1:end-14,:));
[…]

### 2013-06-14 MATLAB/nim_lsc_model_2steps/kalmanFilterSetup.m

@@ -1,0 +1,91 @@
+function [f, q, r, x, a, lambda, h, xi_means, error] = kalmanFilterSetup(param_vec,tau,nfactors,nothers)
+
+ntaus = length(tau);
+nstates = nfactors + nothers;
+error = 0;
+
+tol0 = 1e-8;
+f = reshape(param_vec(1:(nfactors)^2),nfactors,nfactors);
+f_end = (nfactors)^2;
+%f = [[f zeros(nfactors,nothers)]; reshape(param_vec(f_end+1:(f_end+nothers*(nfactors+nothers))), nothers, nfactors+nothers)];
+
+if max(abs(eig(f)))>1+tol0
+    error = -1;
+end
+
+q_start = f_end+nothers*(nfactors+nothers)+1;
+q_end =  f_end+nothers*(nfactors+nothers) + (nstates*(nstates+1))/2;
+q_all = reshape(param_vec(q_start:q_end), (nstates*(nstates+1)/2), 1);
+
+q = zeros(nstates);
+
+qdiag_start_counter = 1;
+qdiag_end_counter = 1;
+for diag_index = (nstates)-1:-1:1
+    q=q+diag(q_all(qdiag_start_counter:qdiag_end_counter),diag_index);
+    qdiag_start_counter=qdiag_end_counter+1;
+    qdiag_end_counter = qdiag_end_counter+1+(nstates)-diag_index;
+end
+q = (q + q') + diag(q_all(qdiag_start_counter:qdiag_end_counter));
+
+%    q = (diag(q_all(1), 2) + diag(q_all(2:3), 1));
+%    q = (q + q') + diag(q_all(4:6), 0);
+
+% check that q is positive semi-definite == det > 0    
+if  min(eig(q))< 0
+    error = -1;
+end
+
+r_start = q_end+1;
+r_end = q_end + ((ntaus+nothers)*((ntaus+nothers)+1))/2;
+r_all = reshape(param_vec(r_start:r_end), ((ntaus+nothers)*((ntaus+nothers)+1)/2), 1);
+
+%write for loop that automates this based on ntaus
+
+r = zeros(ntaus+nothers);
+
+rdiag_start_counter = 1;
+rdiag_end_counter = 1;
+for diag_index = (ntaus+nothers)-1:-1:1
+    r=r+diag(r_all(rdiag_start_counter:rdiag_end_counter),diag_index);
[…]

### 2013-06-14 MATLAB/nim_lsc_model_2steps/load_data_ml.m

@@ -1,0 +1,57 @@
+function [yobs, dates, yields, nims, nfactors, nothers, tau, factors, shadow_bank_share_assets] = load_data_ml 
+
+start_yobs = 1985.75;
+end_yobs = 2008.25;
+dates = start_yobs:.25:end_yobs;
+%SVENY yields can be found in the 'yields' FAME db under the mnemonics in
+%yields_varlist.
+yields = csvread('.\data\yields_sveny_only_1985quarterly.csv', 1, 2);
+dates_yields = 1985.75:.25:2013.0;
+start_pos = find(dates==start_yobs);
+end_pos = find(dates==end_yobs);
+yields = yields(start_pos:end_pos,:)';
+yields_varlist = char('SVENY0025','SVENY0050','SVENY0075','SVENY0100','SVENY0200','SVENY0300','SVENY0500', 'SVENY0700', 'SVENY1000', 'SVENY1500', 'SVENY2000', 'SVENY3000');
+
+%Bring in Net Interest Margins from Call Report (TINY); calculated as nim_top25 =
+%100*netintinc/interest earning assets (where each component in the ratio
+%is calculated at each quarter based on the top 25 firms, by assets).
+nims = csvread('.\data\tiny_nims.csv', 1, 1);
+dates_nim = 1985.0:.25:2013.0;
+start_pos_nim = find(dates_nim==start_yobs);
+end_pos_nim = find(dates_nim==end_yobs);
+nims = 4*nims(start_pos_nim:end_pos_nim,1)'; %Multiply by 4 to annualize
+nims_varlist = char('nim_top25');
+
+%Bring in shadow banking share of financial sector assets. We use it as 
+%a measure of competitiveness in the banking sector.
+%Shadow banking share calculated as:
+%1-((FL704090005.Q+ FL734090005.Q)/((FL413065005.Q+FL674090005.Q+FL614090005.Q+FL664090005.Q+FL504090005.Q)+(FL704090005.Q+ FL734090005.Q)))
+%Data comes from Flow of Funds FAME db fof.
+
+if isunix
+shadow_bank_share_assets_file = csvread('/ofs/research2/Guerrieri/valentin/data/measures_of_competitiveness/shadow_banking_share_assets.csv', 1, 1);
+else
+shadow_bank_share_assets_file = csvread('.\data\shadow_banking_share_assets.csv', 1, 1);
+end
+
+%This data has same dates as yields
[…]

### 2013-06-14 MATLAB/nim_lsc_model_2steps/mylikelihood.m

@@ -1,0 +1,23 @@
+function [logLikel] = mylikelihood(param_vec,yobs,xi10,tau,nfactors,nothers,ntrain,p10)
+
+[f, q, r, x, a, lambda, h, xi_means, error] = kalmanFilterSetup(param_vec,tau,nfactors,nothers);
+    
+if error == 0    
+%Call to Kalman Filter
+%[logLikel1,errcode] = kalmanFilterMatlab(f, h, yields, a, x, xi10, p10, q, r, ntrain);
+
+[logLikel,errcode,xi1tHistory,xi10History,xi11History]= ...
+    kalmanFilterSmoother_v2(f, h, yobs, a, x, xi10, p10, q, r, ntrain);
+
+%[logLikel,errcode]= kalmanFilterMatlab(f, h, yields, a, x, xi10, p10, q, r, ntrain);
+
+% check that that yield curve does not go too far down at any one point in
+% time
+[y] = yield_calc(tau, lambda, xi_means(1:end-nothers), xi1tHistory(1:end-nothers,:));
+if min(min(y))<0.01
+    logLikel = 1e300;
+end
+
+else
+    logLikel = 1e300;
+end

### 2013-06-14 MATLAB/nim_lsc_model_2steps/run_calc_rmse.m

@@ -1,0 +1,64 @@
+%Update last few lines. calc_rmse.m still needs to be updated (it's in
+%estimation utilities). This program was used to get the subsample
+%estimates for the yield curve model.
+
+clear;
+
+setpath;
+
+load_paramvec_from_disk = 1;  % set to 1 to load saved param_vec from disk
+% set to 0 to use first guess as defined below
+
+ntrain = 15; %We have 153 observations per time t; we're setting ntrain to ~15% of observation pool.
+
+[yobs, dates, yields, nims, nfactors, nothers, tau, factors] = load_data_ml; 
+
+out_of_sample_start_pos = find(dates==2000.0);
+end_sample_pos = length(dates);
+
+nstates = nfactors + nothers;
+ntaus = length(tau);
+
+if load_paramvec_from_disk == 1
+    load this_iter %includes the full sample param_vec, xi10_demeaned, p10, ntrain
+end
+
+param_vec_full_sample = param_vec;
+opt_param_mat = zeros(length(param_vec), end_sample_pos-out_of_sample_start_pos+1);
+options = optimset('HessUpdate','bfgs','display','iter','MaxIter',10000,'MaxFunEvals',1e10,'TolFun',10e-3,'TolX',1e-4);
+
+compute_param_mat = 1;
+
+if compute_param_mat
+for pos_index = out_of_sample_start_pos:end_sample_pos
+    [yobs, dates, yields, nims, nfactors, nothers, tau] = load_data_ml;
+    
+    display(['working on observation for date ', num2str(dates(pos_index))]);
+    
+    yobs = yobs(:,1:pos_index-1); %%Just the yields, no NIMs
+    dates = dates(:,1:pos_index-1);
+    yields = yields(:,1:pos_index-1);
+    %nims = nims(:,1:pos_index-1);
+    
+    
+    for fminsearch_rep = 1:1
+        loss = @(x0)mylikelihood(x0,yobs,xi10_demeaned,tau,nfactors,nothers,ntrain,p10);
+        option = anneal();
+        option.Verbosity = 2;
+        [opt_param_mat(:,pos_index-(out_of_sample_start_pos-1)), fval] = anneal(loss,param_vec_full_sample);
+                
+        [opt_param_mat(:,pos_index-(out_of_sample_start_pos-1)),fval,exitflag] = ...
[…]

### 2013-06-14 MATLAB/nim_lsc_model_2steps/run_calc_rmse_conditional.m

@@ -1,0 +1,83 @@
+% This is an old copy that needs to be updated. -- Copy again from 1step
+% directory and update for 2step procedure.
+
+
+
+clear;
+
+setpath;
+
+load_paramvec_from_disk = 1;  % set to 1 to load saved param_vec from disk
+% set to 0 to use first guess as defined below
+
+ntrain = 15; %We have 153 observations per time t; we're setting ntrain to ~15% of observation pool.
+
+[yobs, dates, yields, nims, nfactors, nothers, tau, factors] = load_data_ml; 
+
+out_of_sample_start_pos = find(dates==2000.0);
+end_sample_pos = length(dates);
+
+nstates = nfactors + nothers;
+ntaus = length(tau);
+
+load this_iter; %includes the full sample param_vec, xi10_demeaned, p10, ntrain
+
+param_vec_full_sample = param_vec;
+opt_param_mat = zeros(length(param_vec), end_sample_pos-out_of_sample_start_pos+1);
+options = optimset('HessUpdate','bfgs','display','iter','MaxIter',10000,'MaxFunEvals',1e10,'TolFun',10e-3,'TolX',1e-4);
+
+compute_param_mat = 0;
+
+if compute_param_mat
+for pos_index = out_of_sample_start_pos:end_sample_pos
+    [yobs, dates, yields, nims, nfactors, nothers, tau] = load_data_ml;
+    
+    display(['working on observation for date ', num2str(dates(pos_index))]);
+    
+    yobs = yobs(:,1:pos_index-1);
+    dates = dates(:,1:pos_index-1);
+    yields = yields(:,1:pos_index-1);
+    nims = nims(:,1:pos_index-1);
+    
+    
+    for fminsearch_rep = 1:1
+        loss = @(x0)mylikelihood(x0,yobs,xi10_demeaned,tau,nfactors,nothers,ntrain,p10);
+        option = anneal();
+        option.Verbosity = 2;
+        [opt_param_mat(:,pos_index-(out_of_sample_start_pos-1)), fval] = anneal(loss,param_vec_full_sample);
+                
+        [opt_param_mat(:,pos_index-(out_of_sample_start_pos-1)),fval,exitflag] = ...
[…]

### 2013-06-14 MATLAB/nim_lsc_model_2steps/run_optimize_likelihood.m

@@ -1,0 +1,136 @@
+
+clear;
+
+setpath;
+nreps_minimizer = 0;           % set to 0 to avoid re-optimizing
+
+load_paramvec_from_disk = 1;  % set to 1 to load saved param_vec from disk
+                             % set to 0 to use first guess as defined below
+
+[yobs, dates, yields, nims, nfactors, nothers, tau, factors] = load_data_ml; 
+
+nstates = nfactors + nothers;
+ntaus = length(tau);
+
+ntrain = 15; %We have 153 observations per time t; we're setting ntrain to ~15% of observation pool.
+
+factor_means = mean(factors,2);
+%nims_mean = mean(nims);
+
+[f_factors] = runvar(factors');
+
+% If OLS is preferred, uncomment the following section of code.
+%factors_demeaned = factors-kron(ones(1,size(factors,2)),factor_means);
+% CHANGE THIS 
+%xi10_demeaned = xi_demeaned(:,1); % Initial guess of unobserved states
+%var_data = factors_demeaned';
+
+%nvars = size(var_data,2);
+%varlag = 1;
+%varlagmat = varlag*ones(nvars);
+%[coefs,coverr,errmat]=estimate(varlagmat,zeros(nvars),ones(nvars,1),var_data(1:end-14,:));
+%isconstant = 1;
+%iscontemp = 0;
+%[f_factors,const]=ols2ar(coefs(1:end,:),isconstant,iscontemp);
+
+%[f_nims] = transpose(ols1lag(nims, factors));
+
+f = [f_factors zeros(nfactors,nothers)]; %Before also had f_nims
+
+[f_junk,coverr] = runvar(factors'); %Before also had nims'
+
+q = coverr;
+p10 = coverr;
+
+%f = diag(0.5*ones(nstates,1));
+%q = diag(1*ones(nstates,1));
+r = diag(0.1*ones(ntaus+nothers,1));
+
+q_n = length(diag(q));
+r_n = length(diag(r));
+
+q_diag_vec = [];
+for this_n=1:q_n
+    q_diag_vec = [q_diag_vec 
+                  diag(q,q_n-this_n)];
+end
+
+r_diag_vec = [];
+for this_n=1:r_n
+    r_diag_vec = [r_diag_vec 
+                  diag(r,r_n-this_n)];
+end
+
[…]

### 2013-06-14 MATLAB/nim_lsc_model_2steps/run_out_of_sample_forecasts.m

@@ -1,0 +1,33 @@
+%% To be updated.
+
+
+[yobs, dates, yields, nims, nfactors, nothers, tau, factors] = load_data_ml; 
+
+load this_iter; 
+
+forecast_horizon = 10;
+nobs_history = 90;
+
+[forecast] = out_of_sample_forecasts(param_vec, yobs, xi10_demeaned, p10, ntrain, tau, nfactors, nothers, forecast_horizon);
+
+figure
+plot(dates(end-nobs_history:end), yobs(end,end-nobs_history:end), 'k-');
+hold on
+plot(dates(end)+.25:.25:dates(end)+forecast_horizon/4, forecast(end,:), 'k--');
+
+
+%% Now redo for VAR
+
+var_data= [factors' nims'];
+
+varlag = 4;
+
+[coefb,const,coverr] = runvar_general(var_data,varlag);
+
+f_var=companion(coefb,const);
+
+[forecast_var] = var_out_of_sample_forecasts(f_var, var_data, varlag, nfactors, nothers, forecast_horizon);
+
+plot(dates(end)+.25:.25:dates(end)+forecast_horizon/4, forecast_var(nfactors+1,:), 'r--');
+
+vline(dates(end),'b');

### 2013-06-14 MATLAB/nim_lsc_model_2steps/setpath.m

@@ -1,0 +1,11 @@
+
+restoredefaultpath
+
+path1 = './estimation_utilities';
+
+path (path,path1);
+
+
+dir2 = './plot_support';
+
+path(path,dir2);

### 2013-06-14 MATLAB/nim_lsc_model_2steps/utilities_2steps/calc_rmse.m

@@ -1,0 +1,101 @@
+function [rmse_mlmat rmse_varmat rmse_nochangemat] = calc_rmse(opt_param_mat, out_of_sample_start_pos, end_sample_pos, yobs, factors, xi10, p10, ntrain, tau, nfactors, nothers, forecast_horizon)
+
+%%%Maximum-Likelihood pseudo-out of sample forecasts%%%
+nstates = nfactors+nothers;
+num_periods = end_sample_pos - out_of_sample_start_pos;
+errors_mat = zeros(num_periods, forecast_horizon);
+rmse_mat = zeros(num_periods, 1);
+
+for pos_index = out_of_sample_start_pos:end_sample_pos
+    
+    param_vec = opt_param_mat(:,pos_index-out_of_sample_start_pos+1);
+    
+    [f, q, r, x, a, lambda, h, xi_means, error] = kalmanFilterSetup(param_vec,tau,nfactors,nothers);
+    
+    [logLikel,errcode,xi1tHistory,xi10History,xi11History]= ...
+        kalmanFilterSmoother_v2(f, h, yobs, a, x, xi10, p10, q, r, ntrain);
+    
+    xi = zeros(nstates,forecast_horizon+1);
+    
+    xi(:,1) = xi11History(:,pos_index-1);
+    
+    for i_index = 1:forecast_horizon
+        xi(:,i_index+1) = f*xi(:,i_index);
+    end
+    xi = xi(:,2:end);
+    
+    forecasts = h'*(xi + kron(ones(1,forecast_horizon),xi_means));
+    forecast_errors = zeros(1,forecast_horizon);
+    
+    if pos_index <= end_sample_pos - forecast_horizon
+        forecast_errors = yobs(end,pos_index:pos_index+forecast_horizon-1) - forecasts(end,:);
+        errors_mat(pos_index-out_of_sample_start_pos+1,:) = forecast_errors;
+    else
+        forecast_errors(1:end_sample_pos-pos_index) = yobs(end,pos_index:end_sample_pos-1) - forecasts(end,1:end_sample_pos-pos_index);
+        forecast_errors(forecast_horizon - (end_sample_pos - pos_index)+1:end) = NaN;
+        errors_mat(pos_index-out_of_sample_start_pos+1,:) = forecast_errors;
+    end
+    
+end
+
+rmse_mlmat = (nansum((errors_mat.^2))./(size(errors_mat,1)-sum(isnan(errors_mat)))).^0.5;
+
+
+%%
[…]

### 2013-06-14 MATLAB/nim_lsc_model_2steps/utilities_2steps/calc_rmse_ml_conditional.m

@@ -1,0 +1,63 @@
+function [rmse_mlmat, forecast_mat] = calc_rmse_ml_conditional(opt_param_mat, out_of_sample_start_pos, end_sample_pos, yobs, xi10, p10, ntrain, tau, nfactors, nothers, forecast_horizon, dates,plotbool)
+
+%%%Maximum-Likelihood pseudo-out of sample forecasts%%%
+nstates = nfactors+nothers;
+num_periods = end_sample_pos - out_of_sample_start_pos +1;
+forecast_errors_mat = nan*ones(num_periods, forecast_horizon);
+forecast_mat = nan*ones(num_periods, forecast_horizon);
+
+
+ntau = length(tau);
+
+
+mat_counter = 0;
+for pos_index = out_of_sample_start_pos:end_sample_pos
+    mat_counter = mat_counter+1;
+    param_vec = opt_param_mat(:,pos_index-out_of_sample_start_pos+1);
+    
+    [f, q, r, x, a, lambda, h, xi_means, error] = kalmanFilterSetup(param_vec,tau,nfactors,nothers);
+    
+    [logLikel,errcode,xi1tHistory,xi10History,xi11History,p10History]= ...
+        kalmanFilterSmoother_v2(f, h, yobs , a, x, xi10, p10, q, r, ntrain);
+    
+    
+    xi10new = xi10History(:,pos_index);
+    p10new = p10History(:,:,pos_index);
+    if pos_index <= end_sample_pos - forecast_horizon
+        [logLikel,errcode,xi1tHistory,xi10History,xi11History]= ...
+            kalmanFilterSmoother_v2(f, h(:,1:ntau), yobs(1:ntau,pos_index:pos_index+forecast_horizon-1), a(1:ntau,1:ntau), x(1:ntau,:), xi10new, p10new, q, r(1:ntau,1:ntau), ntrain);
+    else
+        [logLikel,errcode,xi1tHistory,xi10History,xi11History]= ...
+            kalmanFilterSmoother_v2(f, h(:,1:ntau), yobs(1:ntau,pos_index:end_sample_pos), a(1:ntau,1:ntau), x(1:ntau,:), xi10new, p10new, q, r(1:ntau,1:ntau), ntrain);
+    end
+    
+    
+    forecast_errors = zeros(1,forecast_horizon);
+    
+    
+    
+    if  end_sample_pos - pos_index + 1 >= forecast_horizon
+        forecasts = xi11History + kron(ones(1,forecast_horizon),xi_means);
+        if plotbool
[…]

### 2013-06-14 MATLAB/nim_lsc_model_2steps/utilities_2steps/calc_rmse_nochange.m

@@ -1,0 +1,39 @@
+function [rmse_nochangemat, forecast_mat] = calc_rmse_nochange(yobs, out_of_sample_start_pos, end_sample_pos, forecast_horizon)
+%This function computes rmse for a nochange forecast and also returns
+%forecast history over a pseudo-out of sample portion of the sample.
+%Inputs: 
+%yobs is a row vector, a series spanning entire available sample.
+%out_of_sample_start_pos is the position of the first pseudo-out of sample
+%observation.
+%end_sample_pos is the position of the last pseudo-out of sample
+%observation.
+%Outputs:
+% rmse_nochangemat is a row vector containing RMSEs for each forecast step.
+% forecast_mat is a matrix containing a forecast for each period in the
+% pseudo-out of sample. Each row of forecast_mat contains a forecast for
+% the horizon desired.
+
+%% No Change Forecast
+num_periods = end_sample_pos - out_of_sample_start_pos + 1;
+forecast_errors_mat = nan*ones(num_periods, forecast_horizon);
+forecast_mat = nan*(ones(num_periods, forecast_horizon));
+
+mat_counter = 0;
+for pos_index = out_of_sample_start_pos:end_sample_pos
+    mat_counter = mat_counter+1;
+
+    forecast = yobs(pos_index-1)*ones(1,forecast_horizon);
+ 
+    if  end_sample_pos - pos_index + 1 >= forecast_horizon 
+          
+        forecast_errors = yobs(pos_index:pos_index+forecast_horizon-1) - forecast(end,:);
+        forecast_errors_mat(pos_index-out_of_sample_start_pos+1,:) = forecast_errors;
+        forecast_mat(mat_counter,1:length(forecast)) = forecast;
+    else
+        forecast_errors = yobs(pos_index:end_sample_pos) - forecast(1:end_sample_pos-pos_index+1);
+        forecast_errors_mat(mat_counter,1:end_sample_pos-pos_index+1) = forecast_errors;
+        forecast_mat(mat_counter,1:end_sample_pos-pos_index+1) = forecast(1:end_sample_pos-pos_index+1);
+    end
+
+end
+rmse_nochangemat = (nansum((forecast_errors_mat.^2))./(size(forecast_errors_mat,1)-sum(isnan(forecast_errors_mat)))).^0.5;

### 2013-06-14 MATLAB/nim_lsc_model_2steps/utilities_2steps/calc_rmse_var_conditional.m

@@ -1,0 +1,105 @@
+function [rmse_varmat, forecast_mat] = calc_rmse_var_conditional(yobs, other_var_obs, out_of_sample_start_pos, end_sample_pos, forecast_horizon, varlag)
+%This function computes rmse for a VAR forecast and also returns
+%forecast history over a pseudo-out of sample portion of the sample.
+%
+%Inputs: 
+%yobs is a row vector, a series spanning entire available sample.
+%other_var_obs is a matrix, each row is an additional series entering the
+%VAR and spanning the entire available sample. NB: The rmse is going to be
+%conditional on everything in other_var_obs. Needs to have the same number
+%of columns as yobs.
+%out_of_sample_start_pos is the position of the first pseudo-out of sample
+%observation.
+%end_sample_pos is the position of the last pseudo-out of sample
+%observation.
+%varlag is the number of lags used in the VAR.
+%
+%Outputs:
+% rmse_varmat is a row vector containing RMSEs for each forecast step.
+% var_forecasts is a matrix containing a forecast for each period in the
+% pseudo-out of sample. Each row of forecast_mat contains a forecast for
+% the horizon desired.
+
+%Now get RMSEs using the VAR model
+n_other_var_obs = size(other_var_obs, 1);
+
+num_periods = end_sample_pos - out_of_sample_start_pos + 1;
+forecast_errors_mat = nan*ones(num_periods, forecast_horizon);
+forecast_mat = nan*ones(num_periods, forecast_horizon);
+
+for pos_index = out_of_sample_start_pos:end_sample_pos
+    
+    if pos_index <= end_sample_pos - forecast_horizon
+    other_var_obs_out_of_sample = other_var_obs(:,pos_index:pos_index+forecast_horizon-1);
+    else
+      other_var_obs_out_of_sample = other_var_obs(:,pos_index:end);  
+    end
+    
+    this_yobs = yobs(:,1:pos_index-1);
+    this_other_var_obs = other_var_obs(:,1:pos_index-1); %factors
+    
+    var_data= [this_other_var_obs' this_yobs'];
+    
+    % prepare matrices needed for call to Kalman smoother
+    
+    [coefb,const,coverr] = runvar_general(var_data,varlag);
+    
+    f_var=companion(coefb,const);
+    
+    
[…]

### 2013-06-14 MATLAB/plot_support/plot_yield.m

@@ -1,0 +1,46 @@
+function plot_yield(param_vec,tau,nfactors,nothers,yobs,ntrain,xi10_demeaned,p10,dates,factor1,factor2,factor3)
+
+[f, q, r, x, a, lambda, h, xi_means, error] = kalmanFilterSetup(param_vec,tau,nfactors,nothers);
+
+[logLikel,errcode,xi1tHistory]= ...
+    kalmanFilterSmoother_v2(f, h, yobs, a, x, xi10_demeaned, p10, q, r, ntrain);  
+
+%t = find(dates==1990.25);
+
+yields = yobs(1:end-nothers,:);
+
+figure 
+tau_grid = 3:360;
+for tindx = 1:8
+    t = 15+tindx*6; 
+  subplot(4,2,tindx)
+plot(tau_grid, yield_calc(tau_grid, lambda, xi_means(1:end-nothers), xi1tHistory(1:end-nothers,t))); %xi1tHistory(:,t)
+hold on
+plot(tau, yields(:,t), 'o')
+title(['Yield Curve in ',num2str(dates(t))])
+  
+end
+
+
+
+figure
+subplot(3,1,1)
+plot(dates, factor1, 'k')
+hold on
+plot(dates, xi1tHistory(1,:)+xi_means(1), 'r--')
+title('Level Factor')
+legend('Diebold-Li Estimate of the Factor','Smoothed Estimate of the Factor')
+xlim([dates(16) dates(end)])
+subplot(3,1,2)
+plot(dates, factor2, 'k')
+title('Slope Factor')
+hold on
+plot(dates, xi1tHistory(2,:)+xi_means(2), 'r--')
+xlim([dates(16) dates(end)])
+
+subplot(3,1,3)
+plot(dates, factor3, 'k')
+hold on
+plot(dates, xi1tHistory(3,:)+xi_means(3), 'r--')
+title('Curvature Factor')
+xlim([dates(16) dates(end)])

### 2013-06-17 LATEX/paper.tex
draft of paper

\documentclass[12pt]{article}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{chicago}
\usepackage{rotating}

\setlength{\paperwidth}{8.5in} \setlength{\paperheight}{11.0in}
\setlength{\topmargin}{0.0in} \setlength{\headheight}{0.4in}
\setlength{\headsep}{0.0in} \setlength{\textwidth}{7.2in}
\setlength{\textheight}{8.5in} \setlength{\oddsidemargin}{0.0in}
\setlength{\oddsidemargin}{-0.4in}
\setlength{\evensidemargin}{-0.4in}

\newtheorem{theorem}{Theorem}
\newtheorem{acknowledgement}[theorem]{Acknowledgement}
\newtheorem{algorithm}[theorem]{Algorithm}
\newtheorem{axiom}[theorem]{Axiom}
\newtheorem{case}[theorem]{Case}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{conclusion}[theorem]{Conclusion}
\newtheorem{condition}[theorem]{Condition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{criterion}[theorem]{Criterion}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{solution}[theorem]{Solution}
\newtheorem{summary}[theorem]{Summary}
\newenvironment{proof}[1][Proof]{\noindent\textbf{#1.} }{\ \rule{0.5em}{0.5em}}
\input{tcilatex}

\renewcommand{\baselinestretch}{1.5}
\renewcommand{\textfraction}{0.33}
\begin{document}

\bigskip

\bigskip

\clearpage

\renewcommand{\baselinestretch}{1.0}

\begin{center}
{\normalsize \thispagestyle{empty} }

{\normalsize \medskip }

{\normalsize {\Large Forecasting Net Interest Margins of Banks$^{*}$%
} \medskip }

{\normalsize \bigskip \bigskip }

{\normalsize Valentin Bolotnyy, Rochelle Edge, and Luca Guerrieri }

{\normalsize \bigskip \bigskip }

{\normalsize April 29, 2013 }

{\normalsize \bigskip }
\end{center}

{\normalsize \bigskip }

\abstract{}

{\normalsize \vspace{3.0cm} }

{\normalsize \noindent \textbf{Keywords}:  \vspace{1cm} }

{\normalsize \noindent \textbf{JEL Classification}:  }

{\normalsize \vspace{2cm} }

\renewcommand{\baselinestretch}{1} \footnotesize \noindent

\textbf{\ Affiliation and contact information}: Valentin Bolotnyy,
Federal Reserve Board, telephone (202) 452-6428, email
valentin.bolotnyy@frb.gov; Rochelle Edge, Federal Reserve Board, telephone (202) 452-2339, email
rochelle.m.edge@frb.gov; Luca Guerrieri, Federal Reserve Board,
telephone (202) 452-2550, email luca.guerrieri@frb.gov.

\vspace{2cm}

{\footnotesize \noindent $^{*}$ The views expressed in this paper are solely
the responsibility of the authors and should not be interpreted as
reflecting the views of the Board of Governors of the Federal Reserve System
or of any other person associated with the Federal Reserve System.}

\clearpage \renewcommand{\baselinestretch}{1.5} \normalsize

\section{Introduction}

Regular bank stress testing (and the capital planning that it implies) is one of the three complementary reforms that has been made to the capital regulatory regime made in response to the shortcomings of the pre-crisis system.   While stress testing can take many forms - such as, being based on a small number of scenarios, being based on a large number of scenarios, or even being undertaken in reverse form to achieve a certain adverse outcome - the form of stress testing mandated under the Dodd-Frank Act for U.S. bank regulators and the form being used by the Federal Reserve for its annual Comprehensive Capital Analysis and Review (CCAR) is one based around small number of scenarios.  This approach involves specifying a small set of macroeconomic and financial scenarios that represents stressful conditions over the time horizon of the stress test and then projecting, for each BHC in the stress test, its losses, income, and path of pro-forma capital based on the scenario.  Clearly, therefore, there are at least two important parts to conducting bank stress tests:  Formulating the macroeconomic and financial scenarios that a priori would seem stressful to banks and developing models that are able to translate such stressful conditions into losses, income, and pro-forma capital.  This paper considers an aspect of the latter of these two issues.
	In developing models to translate stressful macroeconomic and financial conditions into bank income- and balance-sheet variables it is useful to consider the types of developments, and for which particular variables, that are likely to be stressful to banks.  This is because adverse developments in these variables are very likely to be featured in one of the specified stressful scenarios.  Moreover, it is also desirable that the models being used for stress testing are able to translate developments in these variables well into the key income- or balance-sheet variable that the model is designed to project.  In this paper, we examine this property for models of bank net interest margins (NIMs) defined as the ratio of net interest income (NII) to interest earning assets, where NII is the difference between a bank's interest income and interest expenses.
	Our motivation for focusing on models of NIMs is based on a number of considerations.  First,  bank losses that arise from banks' having to pay out more in interest of their liabilities than they are earning from their assets should not be overlooked as an important source of risk to the banking and broader financial sector.  To be sure, this form of losses was much less prominent in the most recent crisis relative to the form of bank losses that stem from loan defaults.  But, there are ample examples of other earlier banking crises in which adverse developments in NIMs have played a significant role.  The most familiar of these from the U.S. perspective is the Savings and Loans (S\&L) crisis of the late 1980s and early 1990s in which - as a result of the Volcker disinflation - short-term interest rates rose above long-term interest rates, interest expenses rose above interest income, NII and NIMs turned negative (in the thrift sector), and sizable losses and a substantial number of bank failures resulted.  Likewise, the three Nordic banking crises of the late 1980s and early 1990s (for Finland, Sweden, and Norway) - which represent three out of Reinhart's and Rogoff's  (2009) "big five" banking crises - were associated with high short-term interest rates that turned NII and NIMs negative and lead to sizable bank losses and failures.   Additionally, the U.K.'s secondary banking crisis of the early 1970s was associated with a sudden increase in short-term interest rates that abruptly drove up banks' costs of funds so too leading to losses and failures.  To be sure, large increases in short-term rates and declines in net interest income were not the only reason for any of these crises but the role of short-term interest rates - and, in particular, their elevated level relative to long-term interest rates was material and in most cases equivalent in importance to the other causes of the crisis.
	Another motivation for our focus on models of NIM stems from the recent lackluster pace of bank profit growth.   Bank net income growth in recent years has been largely supported by unsustainable cost cutting and reductions in provisioning, rather than from traditional sources of bank revenue-generation like NII.  NII growth, for example, (in both nominal and real terms) has been well below average in recent years while NIMs (NII relative to interest earning assets) have also been low.   Given their already low level, any adverse development that puts further downward pressure on NIMs could be very stressful for banks and, as such, it is important that the economic developments (likely involving interest rates) that would drive these outcomes are well-modeled.
Figure 1 - Net interest income growth (LHS) and net interest margins (RHS)

	Our final motivation for examining NIMs is that despite the importance of NIMs for the financial condition of banks, the modeling of NIMs has received much less attention in the literature relative to modeling of bank financial-statement variables related to credit losses.  To be sure, the recent increase in the importance of stress testing has led to the development of a number of models that link NIMs to macroecnomic variables - such as, short-term Treasury rates and the slope of the Treasury yield curve - where Kovner et al. (2011) and Covas et al. (2012) are notable examples.   However, this research has given relatively little attention to the conditional forecasting properties of these models, which is important if these models are ultimately to be used to project bank losses, income, and pro-forma given macroeconomic scenarios.
	In this paper we consider a range of different models of NIMs, developed with varying degrees of modeling complexity, and examine how well these models explain NIMs.  An important emphasis in examining how well these models explain NIMs is pseudo-out-of-sample forecast performance.  We adopt this approach for the simple reason that the evaluation of models based on their in-sample forecast performance, frequently leads to the econometrician over-fitting the model within the sample so as to maximize in-sample forecast performance.  And this leads to the model performing very poorly once it starts to be used beyond the estimation sample.  Model evaluation based on pseudo-out-of-sample forecast performance indemnifies the econometrician somewhat from this tendency.  We also try to limit the tendency to fine-tune the model to obtain better performing (now pseudo-out-of-sample) forecasts by remaining fairly restrained in terms of choosing additional variables to include in the model.  For the most part we use only the most obvious variables in our models; specifically, those that we can make a fairly strong theoretical justification for and those that appear frequently in other models of NIMs.
	We have already noted a couple of papers that attempt to model NIMs, but there are several additional ones.  OF RECENT PAPERS SKANDER AND LISA.  NOTE TWO PARTS TO THE LITERATURE OLD AND NEW AND BANKING AND MACRO.

\section{Models}
\indent 1. Forecast Combination:

2. Dynamic Factor Model, with restrictions:

Transition equation, which governs the dynamics of the state vector:

$\begin{pmatrix} {L_t - \mu_L} \\ {S_t - \mu_S} \\ {C_t - \mu_C} \end{pmatrix}$ =
$\begin{pmatrix} a_{11} & a_{12} & a_{13} \\ a_{21} & a_{22} & a_{23} \\ a_{31} & a_{32} & a_{33} \end{pmatrix}$
$\begin{pmatrix} {L_t - \mu_L} \\ {S_t - \mu_S} \\ {C_t - \mu_C} \end{pmatrix}$ +
$\begin{pmatrix} {\eta_t(L)} \\ {\eta_t(S)} \\ {\eta_t(C)} \end{pmatrix}$

Measurement equation, which relates a set of $N$ yields (in our case, 12), to the three unobservable factors and NIMs:

%$\begin{pmatrix} y_t(\tau_1) \\ y_t(\tau_2) \\ \vdots \\ y_t(\tau_$N$) \end{pmatrix}$

3. Dynamic Factor Model with 2-Step Regression:

	a. Multivariate (3 Factors added all together to NIM equation)

	b. Forecast Combination (3 Factors added one at a time and combined)

	c. With Poor-Man's Factors

4. VAR:

	a. Poor-Man's Factors:

	b. Smoothed Factors:

5. No-Change Forecast:

This forecast uses $y_t$ as the forecast for $y_{t+1}$.

\section{Comparing Forecast Performance}

\clearpage

\begin{sidewaystable}
\center
\caption{RMSE for shortened sample}
\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|}
\hline
&Step 1 &Step 2 &Step 3 &Step 4 &Step 5 &Step 6 &Step 7 &Step 8 &Step 9 &Step 10\\
\hline
Forecast Combination of Yields        &0.107&0.133&0.148&0.168&0.187&0.202&0.213&0.209&0.216&0.223\\
DFM                                   &0.122&0.148&0.169&0.212&0.247&0.278&0.316&0.340&0.373&0.404\\
DFM + 2nd Step Regression             &0.118&0.167&0.209&0.260&0.308&0.356&0.398&0.433&0.474&0.511\\
DFM + Forecast Combination            &0.112&0.159&0.199&0.247&0.300&0.351&0.399&0.438&0.489&0.535\\
Forecast Combination of Simple Factors&0.118&0.169&0.216&0.273&0.336&0.399&0.456&0.509&0.567&0.622\\
VAR on DF                             &0.173&0.234&0.314&0.379&0.412&0.436&0.436&0.476&0.523&0.571\\
VAR on Simple Factors                 &0.166&0.209&0.263&0.302&0.342&0.375&0.380&0.430&0.451&0.499\\
No-Change Forecast                    &0.107&0.138&0.154&0.181&0.206&0.228&0.244&0.248&0.271&0.293\\
\hline
\end{tabular}
\end{sidewaystable}

\clearpage
\begin{sidewaystable}
\caption{RMSEs over full sample}
\center
\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|}
\hline
&Step 1 &Step 2 &Step 3 &Step 4 &Step 5 &Step 6 &Step 7 &Step 8 &Step 9 &Step 10\\
\hline
Forecast Combination of Yields        &0.105&0.140&0.164&0.191&0.218&0.244&0.265&0.281&0.305&0.327\\
DFM + 2nd Step Regression             &0.128&0.190&0.244&0.299&0.350&0.397&0.438&0.475&0.514&0.549\\
DFM + Forecast Combination            &0.112&0.158&0.197&0.240&0.290&0.337&0.373&0.407&0.446&0.480\\
Forecast Combination of Simple Factors&0.116&0.174&0.227&0.284&0.347&0.409&0.459&0.508&0.554&0.595\\
VAR on DF                             &0.156&0.220&0.245&0.218&0.187&0.223&0.260&0.312&0.382&0.431\\
VAR on Simple Factors                 &0.148&0.196&0.243&0.275&0.283&0.296&0.319&0.377&0.456&0.530\\
No-Change Forecast                    &0.107&0.138&0.154&0.181&0.206&0.228&0.244&0.248&0.271&0.293\\
\hline
\end{tabular}
\end{sidewaystable}

\clearpage
\begin{sidewaystable}
\caption{RMSEs for models that include the share of asset of the shadow banking sector}
\center
\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|}
\hline
&Step 1 &Step 2 &Step 3 &Step 4 &Step 5 &Step 6 &Step 7 &Step 8 &Step 9 &Step 10\\
\hline
Forecast Combination of Yields        &0.107&0.138&0.157&0.182&0.207&0.232&0.247&0.253&0.266&0.278\\
DFM + 2nd Step Regression             &0.121&0.170&0.212&0.262&0.312&0.360&0.403&0.436&0.479&0.516\\
DFM + Forecast Combination            &0.103&0.135&0.156&0.183&0.207&0.232&0.247&0.253&0.266&0.278\\
Forecast Combination of Simple Factors&0.106&0.136&0.156&0.183&0.207&0.232&0.247&0.253&0.266&0.278\\
VAR on DF                             &0.357&0.425&0.469&0.498&0.569&0.645&0.731&0.781&0.821&0.847\\
VAR on Simple Factors                 &0.338&0.355&0.395&0.356&0.372&0.393&0.459&0.491&0.545&0.572\\
No-Change Forecast                    &0.107&0.138&0.154&0.181&0.206&0.228&0.244&0.248&0.271&0.293\\
\hline
\end{tabular}
\end{sidewaystable}

\end{document}

### 2013-06-17 MATLAB/plot_support/tablelatex.m
added a few semicolons to suppress output

@@ -8,7 +8,7 @@ ncols = size(resultmat,2);
 header = char('\begin{table}','\center');
 header = cellstr(header);
 
-tabular = ['\begin{tabular}{|l|']
+tabular = ['\begin{tabular}{|l|'];
 
 for i=1:ncols
     tabular = [tabular,'c|'];
@@ -37,8 +37,8 @@ for i=1:nrows
     line = [line,'\\'];
     table = [table;cellstr(line)];
 end
-table = [table;cellstr('\hline')]
-footer = char('\end{tabular}','\end{table}')
+table = [table;cellstr('\hline')];
+footer = char('\end{tabular}','\end{table}');
 
 table = [table;cellstr(footer)];
 

### 2013-06-17 MATLAB/nim_lsc_model_2steps/run_calc_rmse_conditional.m
cleaned up the tables

@@ -106,29 +106,29 @@ rowlabels = char('Forecast Combination of Yields',...
                  'VAR on DF',...
                  'VAR on Simple Factors',...
                  'No-Change Forecast');
-table1_tex = tablelatex(table1,columnlabels,rowlabels)
+table1_tex = tablelatex(table1,columnlabels,rowlabels);
+char(table1_tex)
 
 
-
-for i = 1:forecast_horizon
-   figure
-   plot(dates(15:end),nims(end,15:end),'k','lineWidth',2)
-   hold on
-   plot(dates(out_of_sample_start_pos+i-1:end),forecast_combination_mat1(1:end-i+1,i)','b--','lineWidth',2)
-   plot(dates(out_of_sample_start_pos+i-1:end),forecast_nochange_mat1(1:end-i+1,i)','r:','lineWidth',2)
-   
-   for j = out_of_sample_start_pos+i-1:end_sample_pos
-       if abs(forecast_combination_mat1(j-out_of_sample_start_pos-i+2,i)'-yobs(end,j))<=abs(forecast_nochange_mat1(j-out_of_sample_start_pos-i+2,i)'-nims(j))
-           plot(dates(j),nims(j),'bo','lineWidth',2)
-       else
-           plot(dates(j),nims(j),'rx','lineWidth',2)
-       end
-   end
-   legend('Data','Forecast Combination','No-change forecast')
-   title(['Assessing the ',num2str(i),'-step-ahead forecast'])
-   
-   xlim([dates(15) dates(end)])
-end
+% for i = [2, 4, 10]
+%    figure
+%    plot(dates(15:end),nims(end,15:end),'k','lineWidth',2)
+%    hold on
+%    plot(dates(out_of_sample_start_pos+i-1:end),forecast_combination_mat1(1:end-i+1,i)','b--','lineWidth',2)
+%    plot(dates(out_of_sample_start_pos+i-1:end),forecast_nochange_mat1(1:end-i+1,i)','r:','lineWidth',2)
+%    
+%    for j = out_of_sample_start_pos+i-1:end_sample_pos
+%        if abs(forecast_combination_mat1(j-out_of_sample_start_pos-i+2,i)'-yobs(end,j))<=abs(forecast_nochange_mat1(j-out_of_sample_start_pos-i+2,i)'-nims(j))
+%            plot(dates(j),nims(j),'bo','lineWidth',2)
+%        else
+%            plot(dates(j),nims(j),'rx','lineWidth',2)
+%        end
+%    end
[…]

### 2013-06-17 MATLAB/nim_lsc_model_1step/run_calc_rmse_conditional.m
cleaned up the table

@@ -68,14 +68,15 @@ lag = 1;
 
 
 [rmse_nochangemat, forecast_nochange_mat1] = calc_rmse_nochange(nims, out_of_sample_start_pos, end_sample_pos, forecast_horizon);
-columnlabels = char('Step 1','Step 2','Step 3','Step 4','Step 5','Step 6','Step 7','Step 8','Step 9','Step 10');
-rowlabels = char('DFM')
-table1_tex = tablelatex(rmse_nochangemat,columnlabels,rowlabels)
 
 
 plotbool = 0;                                   
 [rmse_mlmat, forecast_ml_mat1] = calc_rmse_ml_conditional(opt_param_mat, out_of_sample_start_pos, end_sample_pos, yobs, xi10_demeaned, p10, ntrain, tau, nfactors, nothers, forecast_horizon, dates,plotbool);
 
+columnlabels = char('Step 1','Step 2','Step 3','Step 4','Step 5','Step 6','Step 7','Step 8','Step 9','Step 10');
+rowlabels = char('DFM');
+table1_tex = tablelatex(rmse_mlmat,columnlabels,rowlabels);
+char(table1_tex)
 
 
 for i = 1:forecast_horizon

### 2013-06-17 LATEX/data_description.tex
Includes description of NIM and yields data used in models.

\documentclass[12pt]{article}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{chicago}
\usepackage{hyperref}
\usepackage{sectsty}

\sectionfont{\fontsize{15}{15}\selectfont}

\setlength{\paperwidth}{8.5in} \setlength{\paperheight}{11.0in}
\setlength{\topmargin}{0.0in} \setlength{\headheight}{0.4in}
\setlength{\headsep}{0.0in} \setlength{\textwidth}{7.2in}
\setlength{\textheight}{8.5in} \setlength{\oddsidemargin}{0.0in}
\setlength{\oddsidemargin}{-0.4in}
\setlength{\evensidemargin}{-0.4in}

\begin{document}

\section{Data}
The data on bank net interest margins (NIMs) that we use is from the quarterly Consolidated Reports of Condition and Income (Call Report) that every national, state member, and insured nonmember bank is required to file on the last day of each quarter by the Federal Financial Institutions Examination Council (FFIEC). The Federal Deposit Insurance Corporation is tasked as the overseer, collecting and reviewing all submissions. Call Report data used in this analysis are cleaned and adjusted for bank mergers and acquisitions, using structure data from the National Information Clearinghouse (NIC) on mergers and acquisitions. Foreign entities are excluded and domestic subsidiaries are aggregated up to the parent, bank-holding-company (BHC), level. \footnote{NIM data are adjusted for mergers between commercial banks by comparing balance sheet values of interest income, interest expenses, and interest-earning assets at the end of the quarter with those at the beginning of the quarter, accounting for amounts acquired or lost during the period because of mergers. For information on the merger-adjustment procedure for income, see the appendix in English and Nelson (1998).} \newline \newline
NIMs in our models are expressed as the annualized percentage of net interest income over interest-earning assets. To get an aggregate banking sector measure of NIMs, we aggregate NIMs for the top 25 BHCs, as ranked by total assets, which is assessed quarterly. \newline \newline
The yields data we use are derived using a smoothing technique employed in Gurkaynak, et al. (2007), based on Nelson and Siegel (1987) and Svensson (1994), which allows for daily measures of an off-the-run Treasury yield curve. Although Gurkaynak, et al. produce and publish daily rates for the entire maturity range of outstanding Treasury securities, we only use rates for twelve maturities in our models: 3 month, 6 month, 9 month, 1 year, 2 year, 3 year, 5 year, 7 year, 10 year, 15 year, 20 year, and 30 year. \footnote{Daily yields are published with a two-day lag under the base mnemonic SVENY here: \url{www.federalreserve.gov/econresdata/researchdata/feds200628_1.html}.} \newline \newline
The data on NIMs and yields start in 1985Q4 and end in 2008Q2, a period that stops short of the interest rate regime shift to the zero lower bound.

\section{References}

English, W. B., and W. R. Nelson (1998), ``Profits and Balance Sheet Developments at U.S. Commercial Banks in 1997,'' \textit{Federal Reserve Bulletin} 84 (6), pp. 391-419. \newline  
Gurkaynak, R., B. Sack and J. Wright (2007), ``The U.S. Treasury Yield Curve: 1961 to the Present,'' \textit{Journal of Monetary Economics} 54(8), pp. 2291-2304. \newline
Nelson, C. R. and A. F. Siegel (1987), ``Parsimonious Modeling of Yield Curves,'' \textit{Journal of Business} 60, pp. 473-489. \newline
Svensson, Lars E. O. (1994), ``Estimating and Interpreting Forward Rates: Sweden 1992-4,'' \textit{National Bureau of Economic Research}, Working Paper 4871.

\end{document}

### 2013-06-18 LATEX/paper.tex
paper now includes a description of the forecasting models

@@
 \documentclass[12pt]{article}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \usepackage{amssymb}
@@
 \usepackage{rotating}
+
+\setcounter{MaxMatrixCols}{10}
+%TCIDATA{OutputFilter=Latex.dll}
+%TCIDATA{Version=5.50.0.2960}
+%TCIDATA{<META NAME="SaveForMode" CONTENT="1">}
+%TCIDATA{BibliographyScheme=Manual}
+%TCIDATA{LastRevised=Tuesday, June 18, 2013 11:50:19}
+%TCIDATA{<META NAME="GraphicsSave" CONTENT="32">}
 
@@
 \setlength{\evensidemargin}{-0.4in}
-
 \newtheorem{theorem}{Theorem}
@@
 \input{tcilatex}
-
 \renewcommand{\baselinestretch}{1.5}
 \renewcommand{\textfraction}{0.33}
+
 \begin{document}
@@
 
-{\normalsize {\Large Forecasting Net Interest Margins of Banks$^{*}$%
-} \medskip }
+{\normalsize {\Large Forecasting Net Interest Margins of Banks$^{*}$}
+\medskip }
 
@@
 
-\renewcommand{\baselinestretch}{1} \footnotesize \noindent
+\renewcommand{\baselinestretch}{1} {\footnotesize \noindent }
 
-\textbf{\ Affiliation and contact information}: Valentin Bolotnyy,
-Federal Reserve Board, telephone (202) 452-6428, email
-valentin.bolotnyy@frb.gov; Rochelle Edge, Federal Reserve Board, telephone (202) 452-2339, email
-rochelle.m.edge@frb.gov; Luca Guerrieri, Federal Reserve Board,
-telephone (202) 452-2550, email luca.guerrieri@frb.gov.
+{\footnotesize \textbf{\ Affiliation and contact information}: Valentin
+Bolotnyy, Federal Reserve Board, telephone (202) 452-6428, email
+valentin.bolotnyy@frb.gov; Rochelle Edge, Federal Reserve Board, telephone
+(202) 452-2339, email rochelle.m.edge@frb.gov; Luca Guerrieri, Federal
+Reserve Board, telephone (202) 452-2550, email luca.guerrieri@frb.gov. }
 
-\vspace{2cm}
+{\footnotesize \vspace{2cm} }
 
@@
 reflecting the views of the Board of Governors of the Federal Reserve System
-or of any other person associated with the Federal Reserve System.}
+or of any other person associated with the Federal Reserve System. }
 
-\clearpage \renewcommand{\baselinestretch}{1.5} \normalsize
+{\footnotesize \clearpage \renewcommand{\baselinestretch}{1.5} }
 
-\section{Introduction}
+\section{\protect\normalsize Introduction}
 
-Regular bank stress testing (and the capital planning that it implies) is one of the three complementary reforms that has been made to the capital regulatory regime made in response to the shortcomings of the pre-crisis system.
-While stress testing can take many forms - such as, being based on a small number of scenarios, being based on a large number of scenarios, or even being undertaken in reverse form to achieve a certain adverse outcome - the form of stress testing mandated under the Dodd-Frank Act for U.S.
-bank regulators and the form being used by the Federal Reserve for its annual Comprehensive Capital Analysis and Review (CCAR) is one based around small number of scenarios.
-This approach involves specifying a small set of macroeconomic and financial scenarios that represents stressful conditions over the time horizon of the stress test and then projecting, for each BHC in the stress test, its losses, income, and path of pro-forma capital based on the scenario.
-Clearly, therefore, there are at least two important parts to conducting bank stress tests:  Formulating the macroeconomic and financial scenarios that a priori would seem stressful to banks and developing models that are able to translate such stressful conditions into losses, income, and pro-forma capital.
-This paper considers an aspect of the latter of these two issues.
-	In developing models to translate stressful macroeconomic and financial conditions into bank income- and balance-sheet variables it is useful to consider the types of developments, and for which particular variables, that are likely to be stressful to banks.
-This is because adverse developments in these variables are very likely to be featured in one of the specified stressful scenarios.
-Moreover, it is also desirable that the models being used for stress testing are able to translate developments in these variables well into the key income- or balance-sheet variable that the model is designed to project.
-In this paper, we examine this property for models of bank net interest margins (NIMs) defined as the ratio of net interest income (NII) to interest earning assets, where NII is the difference between a bank's interest income and interest expenses.
-	Our motivation for focusing on models of NIMs is based on a number of considerations.
-First,  bank losses that arise from banks' having to pay out more in interest of their liabilities than they are earning from their assets should not be overlooked as an important source of risk to the banking and broader financial sector.
-To be sure, this form of losses was much less prominent in the most recent crisis relative to the form of bank losses that stem from loan defaults.
-But, there are ample examples of other earlier banking crises in which adverse developments in NIMs have played a significant role.
-The most familiar of these from the U.S.
-perspective is the Savings and Loans (S\&L) crisis of the late 1980s and early 1990s in which - as a result of the Volcker disinflation - short-term interest rates rose above long-term interest rates, interest expenses rose above interest income, NII and NIMs turned negative (in the thrift sector), and sizable losses and a substantial number of bank failures resulted.
-Likewise, the three Nordic banking crises of the late 1980s and early 1990s (for Finland, Sweden, and Norway) - which represent three out of Reinhart's and Rogoff's  (2009) "big five" banking crises - were associated with high short-term interest rates that turned NII and NIMs negative and lead to sizable bank losses and failures.
-Additionally, the U.K.'s secondary banking crisis of the early 1970s was associated with a sudden increase in short-term interest rates that abruptly drove up banks' costs of funds so too leading to losses and failures.
-To be sure, large increases in short-term rates and declines in net interest income were not the only reason for any of these crises but the role of short-term interest rates - and, in particular, their elevated level relative to long-term interest rates was material and in most cases equivalent in importance to the other causes of the crisis.
-	Another motivation for our focus on models of NIM stems from the recent lackluster pace of bank profit growth.
-Bank net income growth in recent years has been largely supported by unsustainable cost cutting and reductions in provisioning, rather than from traditional sources of bank revenue-generation like NII.
-NII growth, for example, (in both nominal and real terms) has been well below average in recent years while NIMs (NII relative to interest earning assets) have also been low.
-Given their already low level, any adverse development that puts further downward pressure on NIMs could be very stressful for banks and, as such, it is important that the economic developments (likely involving interest rates) that would drive these outcomes are well-modeled.
-Figure 1 - Net interest income growth (LHS) and net interest margins (RHS)
+{\normalsize Regular bank stress testing (and the capital planning that it
+implies) is one of the three complementary reforms that has been made to the
+capital regulatory regime made in response to the shortcomings of the
+pre-crisis system. While stress testing can take many forms - such as, being
+based on a small number of scenarios, being based on a large number of
+scenarios, or even being undertaken in reverse form to achieve a certain
+adverse outcome - the form of stress testing mandated under the Dodd-Frank
+Act for U.S. bank regulators and the form being used by the Federal Reserve
+for its annual Comprehensive Capital Analysis and Review (CCAR) is one based
+around small number of scenarios. This approach involves specifying a small
+set of macroeconomic and financial scenarios that represents stressful
+conditions over the time horizon of the stress test and then projecting, for
+each BHC in the stress test, its losses, income, and path of pro-forma
+capital based on the scenario. Clearly, therefore, there are at least two
+important parts to conducting bank stress tests: Formulating the
+macroeconomic and financial scenarios that a priori would seem stressful to
+banks and developing models that are able to translate such stressful
+conditions into losses, income, and pro-forma capital. This paper considers
+an aspect of the latter of these two issues. In developing models to
+translate stressful macroeconomic and financial conditions into bank income-
+and balance-sheet variables it is useful to consider the types of
+developments, and for which particular variables, that are likely to be
+stressful to banks. This is because adverse developments in these variables
+are very likely to be featured in one of the specified stressful scenarios.
+Moreover, it is also desirable that the models being used for stress testing
+are able to translate developments in these variables well into the key
+income- or balance-sheet variable that the model is designed to project. In
+this paper, we examine this property for models of bank net interest margins
+(NIMs) defined as the ratio of net interest income (NII) to interest earning
+assets, where NII is the difference between a bank's interest income and
+interest expenses. Our motivation for focusing on models of NIMs is based on
+a number of considerations. First, bank losses that arise from banks' having
+to pay out more in interest of their liabilities than they are earning from
+their assets should not be overlooked as an important source of risk to the
+banking and broader financial sector. To be sure, this form of losses was
+much less prominent in the most recent crisis relative to the form of bank
+losses that stem from loan defaults. But, there are ample examples of other
+earlier banking crises in which adverse developments in NIMs have played a
+significant role. The most familiar of these from the U.S. perspective is
+the Savings and Loans (S\&L) crisis of the late 1980s and early 1990s in
+which - as a result of the Volcker disinflation - short-term interest rates
+rose above long-term interest rates, interest expenses rose above interest
+income, NII and NIMs turned negative (in the thrift sector), and sizable
+losses and a substantial number of bank failures resulted. Likewise, the
+three Nordic banking crises of the late 1980s and early 1990s (for Finland,
+Sweden, and Norway) - which represent three out of Reinhart's and Rogoff's
+(2009) "big five" banking crises - were associated with high short-term
+interest rates that turned NII and NIMs negative and lead to sizable bank
+losses and failures. Additionally, the U.K.'s secondary banking crisis of
+the early 1970s was associated with a sudden increase in short-term interest
+rates that abruptly drove up banks' costs of funds so too leading to losses
+and failures. To be sure, large increases in short-term rates and declines
+in net interest income were not the only reason for any of these crises but
+the role of short-term interest rates - and, in particular, their elevated
+level relative to long-term interest rates was material and in most cases
+equivalent in importance to the other causes of the crisis. Another
+motivation for our focus on models of NIM stems from the recent lackluster
+pace of bank profit growth. Bank net income growth in recent years has been
+largely supported by unsustainable cost cutting and reductions in
+provisioning, rather than from traditional sources of bank
+revenue-generation like NII. NII growth, for example, (in both nominal and
+real terms) has been well below average in recent years while NIMs (NII
+relative to interest earning assets) have also been low. Given their already
+low level, any adverse development that puts further downward pressure on
+NIMs could be very stressful for banks and, as such, it is important that
+the economic developments (likely involving interest rates) that would drive
+these outcomes are well-modeled. Figure 1 - Net interest income growth (LHS)
+and net interest margins (RHS) }
 
-	Our final motivation for examining NIMs is that despite the importance of NIMs for the financial condition of banks, the modeling of NIMs has received much less attention in the literature relative to modeling of bank financial-statement variables related to credit losses.
-To be sure, the recent increase in the importance of stress testing has led to the development of a number of models that link NIMs to macroecnomic variables - such as, short-term Treasury rates and the slope of the Treasury yield curve - where Kovner et al.
-(2011) and Covas et al.
-(2012) are notable examples.
-However, this research has given relatively little attention to the conditional forecasting properties of these models, which is important if these models are ultimately to be used to project bank losses, income, and pro-forma given macroeconomic scenarios.
-	In this paper we consider a range of different models of NIMs, developed with varying degrees of modeling complexity, and examine how well these models explain NIMs.
-An important emphasis in examining how well these models explain NIMs is pseudo-out-of-sample forecast performance.
-We adopt this approach for the simple reason that the evaluation of models based on their in-sample forecast performance, frequently leads to the econometrician over-fitting the model within the sample so as to maximize in-sample forecast performance.
-And this leads to the model performing very poorly once it starts to be used beyond the estimation sample.
-Model evaluation based on pseudo-out-of-sample forecast performance indemnifies the econometrician somewhat from this tendency.
-We also try to limit the tendency to fine-tune the model to obtain better performing (now pseudo-out-of-sample) forecasts by remaining fairly restrained in terms of choosing additional variables to include in the model.
-For the most part we use only the most obvious variables in our models; specifically, those that we can make a fairly strong theoretical justification for and those that appear frequently in other models of NIMs.
-	We have already noted a couple of papers that attempt to model NIMs, but there are several additional ones.
-OF RECENT PAPERS SKANDER AND LISA.
-NOTE TWO PARTS TO THE LITERATURE OLD AND NEW AND BANKING AND MACRO.
+{\normalsize Our final motivation for examining NIMs is that despite the
+importance of NIMs for the financial condition of banks, the modeling of
+NIMs has received much less attention in the literature relative to modeling
+of bank financial-statement variables related to credit losses. To be sure,
+the recent increase in the importance of stress testing has led to the
+development of a number of models that link NIMs to macroecnomic variables -
+such as, short-term Treasury rates and the slope of the Treasury yield curve
+- where Kovner et al. (2011) and Covas et al. (2012) are notable examples.
+However, this research has given relatively little attention to the
+conditional forecasting properties of these models, which is important if
+these models are ultimately to be used to project bank losses, income, and
+pro-forma given macroeconomic scenarios. In this paper we consider a range
+of different models of NIMs, developed with varying degrees of modeling
+complexity, and examine how well these models explain NIMs. An important
+emphasis in examining how well these models explain NIMs is
+pseudo-out-of-sample forecast performance. We adopt this approach for the
+simple reason that the evaluation of models based on their in-sample
+forecast performance, frequently leads to the econometrician over-fitting
+the model within the sample so as to maximize in-sample forecast
+performance. And this leads to the model performing very poorly once it
+starts to be used beyond the estimation sample. Model evaluation based on
+pseudo-out-of-sample forecast performance indemnifies the econometrician
+somewhat from this tendency. We also try to limit the tendency to fine-tune
+the model to obtain better performing (now pseudo-out-of-sample) forecasts
+by remaining fairly restrained in terms of choosing additional variables to
+include in the model. For the most part we use only the most obvious
+variables in our models; specifically, those that we can make a fairly
+strong theoretical justification for and those that appear frequently in
+other models of NIMs. We have already noted a couple of papers that attempt
+to model NIMs, but there are several additional ones. OF RECENT PAPERS
+SKANDER AND LISA. NOTE TWO PARTS TO THE LITERATURE OLD AND NEW AND BANKING
+AND MACRO. }
 
-\section{Models}
-\indent 1. Forecast Combination:
+\section{\protect\normalsize Models}
 
-2. Dynamic Factor Model, with restrictions:
+1. \textbf{Simple Forecast Combination}. \ We regress NIMs on each yield $r(\tau )_{t-1}$ separately: \
+\begin{equation*}
+NIM_{t}=c_{\tau }+\rho _{\tau }NIM_{t-1}+\gamma _{\tau }r(\tau )_{t-1}+{\eta
+_{\tau ,t}}
+\end{equation*}
 
-Transition equation, which governs the dynamics of the state vector:
+Each regression is use to produce an $s$ step-ahead forecast of NIMs
+conditional on yields on treasuries with maturity $\tau $ observed through
+period $t+s-1.$ This conditional forecast is denoted by $NIM_{\tau ,t+s/t}.$%
+The simple forecast comnination is then given by:%
+\begin{equation*}
+NIM_{t+s/t}=\sum_{\tau }\frac{NIM_{\tau ,t+s/t}}{N},
+\end{equation*}%
+where $N$ is the number of maturities considered.
 
-$\begin{pmatrix} {L_t - \mu_L} \\ {S_t - \mu_S} \\ {C_t - \mu_C} \end{pmatrix}$ =
-$\begin{pmatrix} a_{11} & a_{12} & a_{13} \\ a_{21} & a_{22} & a_{23} \\ a_{31} & a_{32} & a_{33} \end{pmatrix}$
-$\begin{pmatrix} {L_t - \mu_L} \\ {S_t - \mu_S} \\ {C_t - \mu_C} \end{pmatrix}$ +
-$\begin{pmatrix} {\eta_t(L)} \\ {\eta_t(S)} \\ {\eta_t(C)} \end{pmatrix}$
+2. \textbf{A Dynamic Factor Model for NIMs.} \ Using the state space
+representation, the state equations of our restricted dynamic factor model
+is given by
 
-Measurement equation, which relates a set of $N$ yields (in our case, 12), to the three unobservable factors and NIMs:
+\begin{equation*}
+\left(
+\begin{array}{c}
+{L_{t+1}-\mu _{L}} \\
+{S_{t+1}-\mu _{S}} \\
+{C_{t+1}-\mu _{C}} \\
+\widehat{NIM}_{t+1}-\mu _{\widehat{NIM}}%
+\end{array}%
+\right) {\normalsize \ =\ }\left(
+\begin{array}{cccc}
+a_{11} & a_{12} & a_{13} & 0 \\
+a_{21} & a_{22} & a_{23} & 0 \\
+a_{31} & a_{32} & a_{33} & 0 \\
+a_{41} & a_{42} & a_{43} & a_{44}%
+\end{array}%
+\right) {\normalsize \ }\left(
+\begin{array}{c}
+{L_{t}-\mu _{L}} \\
+{S_{t}-\mu _{S}} \\
+{C_{t}-\mu _{C}} \\
+\widehat{NIM}_{t}-\mu _{\widehat{NIM}}%
+\end{array}%
+\right) {\normalsize +}\left(
+\begin{array}{c}
+{\eta _{Lt}} \\
+{\eta _{St}} \\
+{\eta _{Ct}} \\
+{\eta _{NIMt}}%
+\end{array}%
+{\normalsize \ }\right) ,
+\end{equation*}
 
-%$\begin{pmatrix} y_t(\tau_1) \\ y_t(\tau_2) \\ \vdots \\ y_t(\tau_$N$) \end{pmatrix}$
+where $L_{t},S_{t},$ and $C_{t}$ are level slope and curvature factors for
+the yield curve, respectively, and $\widehat{NIM}_{t}$ represent a measure
+of NIMs cleaned of observation error. The term $\mu _{x}$ denotes the mean
+of $x$. The terms ${\eta _{Lt}}$, ${\eta _{St}}$, ${\eta _{Ct}}$, ${\eta
+_{NIMt}}$ denote independently distributed innovations with joint covariance
+matrix $\Omega $.
 
-3. Dynamic Factor Model with 2-Step Regression:
+\bigskip
 
-	a. Multivariate (3 Factors added all together to NIM equation)
+{\normalsize The observation equations, which relate a set of $N$ yields (in
+our case, 12) and observed NIMs can be split into two groups. \ In the first
+group, equations for yields on treasuries of maturity }$\tau $ take the form:
 
-	b. Forecast Combination (3 Factors added one at a time and combined)
+\begin{equation*}
+r(\tau )=\left(
+\begin{array}{cccc}
+1 & \frac{1-e^{-\lambda \tau }}{\lambda \tau } & \frac{1-e^{-\lambda \tau }}{%
+\lambda \tau }-e^{-\lambda \tau } & 0%
+\end{array}%
+\right) \left(
+\begin{array}{c}
+{L_{t}-\mu _{L}} \\
+{S_{t}-\mu _{S}} \\
+{C_{t}-\mu _{C}} \\
+\widehat{NIM}_{t}-\mu _{\widehat{NIM}}%
+\end{array}%
+\right) +e(\tau )_{t},
+\end{equation*}%
+where $e(\tau )_{t}$ is i.i.d. measurement error assumed to be uncorrelated
+across observations. \bigskip In the second group, the equation for observed
+NIMs takes the form%
+\begin{equation*}
+NIM_{t}=\left(
+\begin{array}{cccc}
+0 & 0 & 0 & 1%
+\end{array}%
+\right) \left(
+\begin{array}{c}
+{L_{t}-\mu _{L}} \\
+{S_{t}-\mu _{S}} \\
+{C_{t}-\mu _{C}} \\
+\widehat{NIM}_{t}-\mu _{\widehat{NIM}}%
+\end{array}%
+\right) +e(NIM)_{t},
+\end{equation*}%
+where $e(NIM)_{t}$ is i.i.d. measurement error assumed to be uncorrelated
+across observations. \bigskip
 
-	c. With Poor-Man's Factors
+Out-of-sample conditional forecasts for NIMs from this model, are obtained
+using the Kalman filter.
 
-4. VAR:
+3. \textbf{Dynamic Factor Model with Second-Step Regression.} \ The dynamic
+factor model we consider takes the form:
 
-	a. Poor-Man's Factors:
+\bigskip
+\begin{equation*}
+\left(
+\begin{array}{c}
+{L_{t+1}-\mu _{L}} \\
+{S_{t+1}-\mu _{S}} \\
+{C_{t+1}-\mu _{C}}%
+\end{array}%
+\right) {\normalsize \ =\ }\left(
+\begin{array}{ccc}
+a_{11} & a_{12} & a_{13} \\
+a_{21} & a_{22} & a_{23} \\
+a_{31} & a_{32} & a_{33}%
+\end{array}%
+\right) {\normalsize \ }\left(
+\begin{array}{c}
+{L_{t}-\mu _{L}} \\
+{S_{t}-\mu _{S}} \\
+{C_{t}-\mu _{C}}%
+\end{array}%
+\right) {\normalsize +}\left(
+\begin{array}{c}
+{\eta _{Lt}} \\
+{\eta _{St}} \\
+{\eta _{Ct}}%
+\end{array}%
+\right)
+\end{equation*}
 
-	b. Smoothed Factors:
+\begin{equation*}
+r(\tau )=\left(
+\begin{array}{ccc}
+1 & \frac{1-e^{-\lambda \tau }}{\lambda \tau } & \frac{1-e^{-\lambda \tau }}{%
+\lambda \tau }-e^{-\lambda \tau }%
+\end{array}%
+\right) \left(
+\begin{array}{c}
+{L_{t}-\mu _{L}} \\
+{S_{t}-\mu _{S}} \\
+{C_{t}-\mu _{C}} \\
+\widehat{NIM}_{t}-\mu _{\widehat{NIM}}%
+\end{array}%
+\right) +e(\tau )_{t},
+\end{equation*}
 
-5. No-Change Forecast:
+We obtain smoothed estimates of the factors $L_{t},S_{t},$ and $C_{t}$
+conditional on information up until period $T.$We estimate two alternative
+second-step forecasting equations:
 
-This forecast uses $y_t$ as the forecast for $y_{t+1}$.
+\bigskip Alternative a. {\normalsize Multivariate (all yield curve factors
+included in one NIM equation) }
 
-\section{Comparing Forecast Performance}
+\begin{equation*}
+NIM_{t}=c+\rho NIM_{t-1}+\gamma _{L}L_{t-1}+\gamma _{S}S_{t-1}+\gamma
+_{C}C_{t-1}
+\end{equation*}
 
-\clearpage
+Alternative b. {\normalsize Forecast Combination (each factor included in a
+separate equation and forecasts from each equation combined) }
 
+\begin{equation*}
+NIM_{t}=c_{i}+\rho _{i}NIM_{t-1}+\gamma _{i}F_{i,t-1}+{\eta _{i,t},}
+\end{equation*}%
+where $F_{i}\in \left\{ L,S,C\right\} .$ Forecasts $NIM_{i,t+s/t}$ from each
+separate regression are then aggregated as:
+
+\begin{equation*}
+NIM_{t+s/t}=\sum_{i}\frac{NIM_{i,t+s/t}}{3},
+\end{equation*}
+
+4. \textbf{Forecast Combination with Observed Factors. \ }Define $\left\{
+L,S,C\right\} ,$ the observed factors as in 4. above.
+\begin{equation*}
+NIM_{t}=c_{i}+\rho _{i}NIM_{t-1}+\gamma _{i}O_{i,t-1}+{\eta _{i,t},}
+\end{equation*}%
+where $O_{i}\in \left\{ L,S,C\right\} .$ Forecasts  from each separate
+regression, $NIM_{i,t+s/t},$are then aggregated as:
+
+5. \textbf{VAR with Observed Factors. \ }{\normalsize Define level factor to
+coincide with the 10-year yield; define the slope factor as the difference
+between the 10-year and the 3-month yields; define the curvatures factor to
+be 2 times the 2-year yield minus the the sum of the 3-month and 10-year
+yields. }The VAR of order four includes the three observed factors defined
+as above and NIMs. \ Forecasts conditional on the factors are obtained using
+the Kalman filter.
+\begin{equation*}
+NIM_{t+s/t}=\sum_{i}\frac{NIM_{i,t+s/t}}{3},
+\end{equation*}
+
+6. \textbf{No-Change Forecast. \ }{\normalsize According to this model }$%
+NIM_{t+s/t}=NIM_{t}${\normalsize . }
+
+\section{\protect\normalsize Comparing Forecast Performance}
+
+{\normalsize \clearpage
+}
+
+{\normalsize
 \begin{sidewaystable}
 \center
-\caption{RMSE for shortened sample}
+\caption{RMSE for shortened sample: 1989Q2-2008Q2}
 \begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|}
@@
 \hline
-Forecast Combination of Yields        &0.107&0.133&0.148&0.168&0.187&0.202&0.213&0.209&0.216&0.223\\
-DFM                                   &0.122&0.148&0.169&0.212&0.247&0.278&0.316&0.340&0.373&0.404\\
-DFM + 2nd Step Regression             &0.118&0.167&0.209&0.260&0.308&0.356&0.398&0.433&0.474&0.511\\
-DFM + Forecast Combination            &0.112&0.159&0.199&0.247&0.300&0.351&0.399&0.438&0.489&0.535\\
-Forecast Combination of Simple Factors&0.118&0.169&0.216&0.273&0.336&0.399&0.456&0.509&0.567&0.622\\
-VAR on DF                             &0.173&0.234&0.314&0.379&0.412&0.436&0.436&0.476&0.523&0.571\\
-VAR on Simple Factors                 &0.166&0.209&0.263&0.302&0.342&0.375&0.380&0.430&0.451&0.499\\
-No-Change Forecast                    &0.107&0.138&0.154&0.181&0.206&0.228&0.244&0.248&0.271&0.293\\
+1. Simple Forecast Combination             &0.107&0.133&0.148&0.168&0.187&0.202&0.213&0.209&0.216&0.223\\
+2. DFM                                     &0.122&0.148&0.169&0.212&0.247&0.278&0.316&0.340&0.373&0.404\\
+3a. DFM with 2nd Step Regression           &0.118&0.167&0.209&0.260&0.308&0.356&0.398&0.433&0.474&0.511\\
+3b. DFM with Forecast Combination          &0.112&0.159&0.199&0.247&0.300&0.351&0.399&0.438&0.489&0.535\\
+4. Forecast Combination of Observed Factors&0.118&0.169&0.216&0.273&0.336&0.399&0.456&0.509&0.567&0.622\\
+5. VAR on Observed Factors                 &0.166&0.209&0.263&0.302&0.342&0.375&0.380&0.430&0.451&0.499\\
+6. No-Change Forecast                      &0.107&0.138&0.154&0.181&0.206&0.228&0.244&0.248&0.271&0.293\\
 \hline
@@
 \end{sidewaystable}
+}
 
-\clearpage
+{\normalsize \clearpage
 \begin{sidewaystable}
-\caption{RMSEs over full sample}
+\caption{RMSEs over full sample: 1985Q4-2008Q2}
 \center
@@
 \hline
-Forecast Combination of Yields        &0.105&0.140&0.164&0.191&0.218&0.244&0.265&0.281&0.305&0.327\\
-DFM + 2nd Step Regression             &0.128&0.190&0.244&0.299&0.350&0.397&0.438&0.475&0.514&0.549\\
-DFM + Forecast Combination            &0.112&0.158&0.197&0.240&0.290&0.337&0.373&0.407&0.446&0.480\\
-Forecast Combination of Simple Factors&0.116&0.174&0.227&0.284&0.347&0.409&0.459&0.508&0.554&0.595\\
-VAR on DF                             &0.156&0.220&0.245&0.218&0.187&0.223&0.260&0.312&0.382&0.431\\
-VAR on Simple Factors                 &0.148&0.196&0.243&0.275&0.283&0.296&0.319&0.377&0.456&0.530\\
-No-Change Forecast                    &0.107&0.138&0.154&0.181&0.206&0.228&0.244&0.248&0.271&0.293\\
+1. Simple Forecast Combination             &0.105&0.140&0.164&0.191&0.218&0.244&0.265&0.281&0.305&0.327\\
+3a. DFM with 2nd Step Regression           &0.128&0.190&0.244&0.299&0.350&0.397&0.438&0.475&0.514&0.549\\
+3b. DFM with Forecast Combination          &0.112&0.158&0.197&0.240&0.290&0.337&0.373&0.407&0.446&0.480\\
+4. Forecast Combination of Observed Factors&0.116&0.174&0.227&0.284&0.347&0.409&0.459&0.508&0.554&0.595\\
+5. VAR on Observed Factors                 &0.148&0.196&0.243&0.275&0.283&0.296&0.319&0.377&0.456&0.530\\
+6. No-Change Forecast                      &0.107&0.138&0.154&0.181&0.206&0.228&0.244&0.248&0.271&0.293\\
 \hline
@@
 \end{sidewaystable}
+}
 
-\clearpage
+{\normalsize \clearpage
 \begin{sidewaystable}
-\caption{RMSEs for models that include the share of asset of the shadow banking sector}
+\caption{RMSEs for models that include the share of asset of the shadow banking sector. Estimation sample: 1989Q2-2008Q2}
 \center
@@
 \hline
-Forecast Combination of Yields        &0.107&0.138&0.157&0.182&0.207&0.232&0.247&0.253&0.266&0.278\\
-DFM + 2nd Step Regression             &0.121&0.170&0.212&0.262&0.312&0.360&0.403&0.436&0.479&0.516\\
-DFM + Forecast Combination            &0.103&0.135&0.156&0.183&0.207&0.232&0.247&0.253&0.266&0.278\\
-Forecast Combination of Simple Factors&0.106&0.136&0.156&0.183&0.207&0.232&0.247&0.253&0.266&0.278\\
-VAR on DF                             &0.357&0.425&0.469&0.498&0.569&0.645&0.731&0.781&0.821&0.847\\
-VAR on Simple Factors                 &0.338&0.355&0.395&0.356&0.372&0.393&0.459&0.491&0.545&0.572\\
-No-Change Forecast                    &0.107&0.138&0.154&0.181&0.206&0.228&0.244&0.248&0.271&0.293\\
+1. Simple Forecast Combination             &0.110&0.143&0.165&0.192&0.222&0.248&0.267&0.274&0.287&0.300\\
+3a. DFM with 2nd Step Regression           &0.121&0.170&0.212&0.262&0.312&0.360&0.403&0.436&0.478&0.515\\
+3b. DFM with Forecast Combination          &0.106&0.140&0.165&0.193&0.222&0.248&0.267&0.274&0.287&0.300\\
+4. Forecast Combination of Observed Factors&0.109&0.141&0.165&0.193&0.222&0.249&0.267&0.274&0.288&0.300\\
+5. VAR on Observed Factors                 &0.344&0.396&0.476&0.494&0.521&0.569&0.631&0.670&0.699&0.718\\
+6. No-Change Forecast                      &0.107&0.138&0.154&0.181&0.206&0.228&0.244&0.248&0.271&0.293\\
 \hline
@@
 \end{sidewaystable}
+}
 

### 2013-06-18 LATEX/paper.tex
latest version

@@
 \usepackage{rotating}
+\usepackage{hyperref}
+\usepackage{sectsty}
 
@@
 
+\section{Data}
+The data on bank net interest margins (NIMs) that we use is from the quarterly Consolidated Reports of Condition and Income (Call Report) that every national, state member, and insured nonmember bank is required to file on the last day of each quarter by the Federal Financial Institutions Examination Council (FFIEC).
+The Federal Deposit Insurance Corporation is tasked as the overseer, collecting and reviewing all submissions.
+Call Report data used in this analysis are cleaned and adjusted for bank mergers and acquisitions, using structure data from the National Information Clearinghouse (NIC) on mergers and acquisitions.
+Foreign entities are excluded and domestic subsidiaries are aggregated up to the parent, bank-holding-company (BHC), level.
+\footnote{NIM data are adjusted for mergers between commercial banks by comparing balance sheet values of interest income, interest expenses, and interest-earning assets at the end of the quarter with those at the beginning of the quarter, accounting for amounts acquired or lost during the period because of mergers.
+For information on the merger-adjustment procedure for income, see the appendix in English and Nelson (1998).} \newline \newline
+NIMs in our models are expressed as the annualized percentage of net interest income over interest-earning assets.
+To get an aggregate banking sector measure of NIMs, we aggregate NIMs for the top 25 BHCs, as ranked by total assets, which is assessed quarterly.
+\newline \newline
+The yields data are derived using a smoothing technique from Gurkaynak, et al.
+(2007), based on Nelson and Siegel (1987) and Svensson (1994), which allows for daily measures of an off-the-run Treasury yield curve.
+We use yields for twelve maturities in our models: 3 month, 6 month, 9 month, 1 year, 2 year, 3 year, 5 year, 7 year, 10 year, 15 year, 20 year, and 30 year.\footnote{Daily yields are published with a two-day lag under the base mnemonic SVENY here: \url{www.federalreserve.gov/econresdata/researchdata/feds200628_1.html}.} The yields included are quarterly averages of daily yields.
+\newline \newline
+The data on NIMs and yields start in 1985Q4 and end in 2008Q2, a period that stops short of the interest rate regime shift to the zero lower bound.
+
 \section{\protect\normalsize Comparing Forecast Performance}

### 2013-06-18 MATLAB/nim_lsc_model_2steps/run_calc_rmse_conditional.m
added charts

@@ -93,19 +93,17 @@ rmse_forecast_combination_mat1
 rmse_multivariate_mat1
 rmse_forecast_combination_mat2
 rmse_forecast_combination_mat3
-rmse_varmat1
 rmse_varmat2
 rmse_nochangemat1];
 
 
 columnlabels = char('Step 1','Step 2','Step 3','Step 4','Step 5','Step 6','Step 7','Step 8','Step 9','Step 10');
-rowlabels = char('Forecast Combination of Yields',...
-                 'DFM + 2nd Step Regression',...
-                 'DFM + Forecast Combination',...
-                 'Forecast Combination of Simple Factors',...
-                 'VAR on DF',...
-                 'VAR on Simple Factors',...
-                 'No-Change Forecast');
+rowlabels = char('1. Simple Forecast Combination',...
+                 '3a. DFM with 2nd Step Regression',...
+                 '3b. DFM with Forecast Combination',...
+                 '4. Forecast Combination of Observed Factors',...
+                 '5. VAR on Observed Factors',...
+                 '6. No-Change Forecast');
 table1_tex = tablelatex(table1,columnlabels,rowlabels);
 char(table1_tex)
 
@@ -151,20 +149,11 @@ rmse_forecast_combination_mat1
 rmse_multivariate_mat1
 rmse_forecast_combination_mat2
 rmse_forecast_combination_mat3
-rmse_varmat1
 rmse_varmat2
 rmse_nochangemat1];
 
 
 
-columnlabels = char('Step 1','Step 2','Step 3','Step 4','Step 5','Step 6','Step 7','Step 8','Step 9','Step 10');
-rowlabels = char('Forecast Combination of Yields',...
-                 'DFM + 2nd Step Regression',...
-                 'DFM + Forecast Combination',...
-                 'Forecast Combination of Simple Factors',...
-                 'VAR on DF',...
-                 'VAR on Simple Factors',...
-                 'No-Change Forecast');
 table2_tex = tablelatex(table2,columnlabels,rowlabels);
 char(table2_tex)
 
@@ -190,24 +179,26 @@ rmse_forecast_combination_mat1
 rmse_multivariate_mat1
 rmse_forecast_combination_mat2
 rmse_forecast_combination_mat3
-rmse_varmat1
 rmse_varmat2
 rmse_nochangemat1];
 
 
-
[…]

### 2013-06-18 MATLAB/nim_lsc_model_2steps/doubleplot.m
used to plot on double axes

@@ -1,0 +1,22 @@
+function doubleplot(data1,data2,dates,thistitle)
+
+
+hl1 = line(dates,data1,'Color','k');
+ax1 = gca;
+set(ax1,'XColor','k','YColor','k')
+
+
+
+%Next, create another axes at the same location as the first, placing the x-axis on top and the y-axis on the right. Set the axes Color to none to allow the first axes to be visible and color code the x- and y-axis to match the data.
+ax2 = axes('Position',get(ax1,'Position'),...
+           'XAxisLocation','top',...
+           'YAxisLocation','right',...
+           'Color','none',...
+           'XColor','w','YColor','r','LineStyle','--','XTickLabel',{});
+
+title(thistitle)
+
+%Draw the second set of data in the same color as the x- and y-axis.
+
+
+hl2 = line(dates,data2,'Color','r','lineStyle','--','Parent',ax2);

### 2013-06-25 MATLAB/nim_lsc_model_2steps/utilities_2steps/load_data_ml.m
Included components of total interest earning asset and capability to use NIMs constructed in numerous different ways

@@ -15,12 +15,12 @@ yields_varlist = char('SVENY0025','SVENY0050','SVENY0075','SVENY0100','SVENY0200
 %Bring in Net Interest Margins from Call Report (TINY); calculated as nim_top25 =
 %100*netintinc/interest earning assets (where each component in the ratio
 %is calculated at each quarter based on the top 25 firms, by assets).
-nims = csvread('..\data\tiny_nims.csv', 1, 1);
-dates_nim = 1985.0:.25:2013.0;
-start_pos_nim = find(dates_nim==start_yobs);
-end_pos_nim = find(dates_nim==end_yobs);
-nims = 4*nims(start_pos_nim:end_pos_nim,1)'; %Multiply by 4 to annualize
-nims_varlist = char('nim_top25');
+%nims = csvread('..\data\tiny_nims.csv', 1, 1);
+%dates_nim = 1985.0:.25:2013.0;
+%start_pos_nim = find(dates_nim==start_yobs);
+%end_pos_nim = find(dates_nim==end_yobs);
+%nims = 4*nims(start_pos_nim:end_pos_nim,1)'; %Multiply by 4 to annualize
+%nims_varlist = char('nim_top25');
 
 %Bring in shadow banking share of financial sector assets. We use it as 
 %a measure of competitiveness in the banking sector.
@@ -36,6 +36,64 @@ end
 
 %This data has same dates as yields
 shadow_bank_share_assets = shadow_bank_share_assets_file(start_pos:end_pos,1)';
+
+%Bring in NIMs and subcategories of interest-earning assets. Eight (8) variants
+%of NIM are possible:
+% 1. With trading assets and based on end-of-period total assets rankings
+% 2. Without trading assets and based on end-of-period total assets rankings
+% 3. With trading assets and based on quarterly average of total assets
+% rankings (quarterly average constructed from daily data on assets)
+% 4. Without trading assets and based on quarterly average of total assets
+% rankings (quarterly average constructed from daily data on assets)
+% 5-8. Same as above, but based on interest-earning (ie) asset rankings (rather
+% than total asset rankings).
+
+if isunix
+nim_option = ['/ofs/work_mpa/m1vxb00/nim_project/MATLAB/data/data_ranked_by_total_assets/nims_assets_by_endperiod_total_assets.csv';
[…]

### 2013-06-25 LATEX/data_description.tex
Added a line about yields
No major changes made.

@@
 The data on bank net interest margins (NIMs) that we use is from the quarterly Consolidated Reports of Condition and Income (Call Report) that every national, state member, and insured nonmember bank is required to file on the last day of each quarter by the Federal Financial Institutions Examination Council (FFIEC).
-The Federal Deposit Insurance Corporation is tasked as the overseer, collecting and reviewing all submissions.
+We consider a slight variant that excludes interest and fee income on loans from foreign offices, Edge and Agreement subsidiaries, and IBFs.
+We exclude this category The Federal Deposit Insurance Corporation is tasked as the overseer, collecting and reviewing all submissions.
 Call Report data used in this analysis are cleaned and adjusted for bank mergers and acquisitions, using structure data from the National Information Clearinghouse (NIC) on mergers and acquisitions.
@@
 produce and publish daily rates for the entire maturity range of outstanding Treasury securities, we only use rates for twelve maturities in our models: 3 month, 6 month, 9 month, 1 year, 2 year, 3 year, 5 year, 7 year, 10 year, 15 year, 20 year, and 30 year.
+
+***Data are continuously compounded zero-coupon Treasury yields as estimated by the Svensson Term Structure Model used in Gurkaynak, et al. (2007).***
+
+We've verified that if we use Y9-C data to construct a corresponding/equivalent/analogous measure of NIM, the correlation is ...
+
  \footnote{Daily yields are published with a two-day lag under the base mnemonic SVENY here: \url{www.federalreserve.gov/econresdata/researchdata/feds200628_1.html}.} \newline \newline

### 2013-06-28 LATEX/paper.tex
additional charts
Added introductory charts

@@
 \right) +e(NIM)_{t},
-\end{equation*}%
+\end{equation*}
 where $e(NIM)_{t}$ is i.i.d. measurement error assumed to be uncorrelated
@@
 factor model we consider takes the form:
-
 \bigskip
@@
 We obtain smoothed estimates of the factors $L_{t},S_{t},$ and $C_{t}$
-conditional on information up until period $T.$We estimate two alternative
+conditional on information up until period $T.$ We estimate two alternative
 second-step forecasting equations:
@@
 
+\begin{figure}[tbp]
+\caption{NIMs and Treasury Yields} \label{figure_nims_rates}
+\center
+\includegraphics[scale=0.85]{figure_nims_rates.ps}
+\end{figure}
+
+\begin{figure}[tbp]
+\caption{NIMs and Observed Yield Factors} \label{figure_nims_factors}
+\center
+\includegraphics[scale=0.85]{figure_nims_factors.ps}
+\end{figure}
+
+\begin{figure}[tbp]
+\caption{NIMs and Measures of Competitions} \label{figure_nims_competition}
+\center
+\includegraphics[scale=0.85]{figure_nims_competition.ps}
+\end{figure}
+
+\begin{figure}[tbp]
+\caption{Interest Income, Interest Expenses and Short-term Treasury Rates} \label{figure_nims_components}
+\center
+\includegraphics[scale=0.85]{figure_nims_components.ps}
+\end{figure}
+
 \end{document}

### 2013-06-28 LATEX/paper.tex
added table

@@
 
+\begin{table}
+\caption{RMSEs for variants of the simple forecast combination model}
+\center
+\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|}
+\hline
+&Step 1 &Step 2 &Step 3 &Step 4 &Step 5 &Step 6 &Step 7 &Step 8 &Step 9 &Step 10\\
+\hline
+All yields              &0.107&0.133&0.148&0.168&0.187&0.202&0.213&0.209&0.216&0.223\\
+Observed Factors        &0.118&0.169&0.216&0.273&0.336&0.399&0.456&0.509&0.567&0.622\\
+Smoothed Factors        &0.112&0.159&0.199&0.247&0.300&0.351&0.399&0.438&0.489&0.535\\
+3-month, 2-year, 10-year&0.105&0.130&0.141&0.160&0.178&0.197&0.202&0.198&0.209&0.217\\
+3-month, 10-year        &0.105&0.129&0.144&0.159&0.179&0.196&0.200&0.200&0.209&0.217\\
+\hline
+\end{tabular}
+\end{table}
+
 \begin{figure}[tbp]

### 2013-06-28 MATLAB/nim_lsc_model_2steps/run_calc_rmse_conditional.m
added table

@@ -192,6 +192,45 @@ table3_tex = tablelatex(table3,columnlabels,rowlabels);
 char(table3_tex)
 
 
+
+% Determine which forecast combination is best in RMSE
+[rmse_forecast_combination_mat0, forecast_combination_mat0] = calc_rmse_forecast_combination_conditional(nims(:,15:end), yields(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, 1,4);
+[rmse_forecast_combination_mat1, forecast_combination_mat1] = calc_rmse_forecast_combination_conditional(nims(:,15:end), factors(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, 1,4);
+[rmse_forecast_combination_mat2, forecast_combination_mat1] = calc_rmse_forecast_combination_conditional(nims(:,15:end), smoothed_factors(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, 1,4);
+[rmse_forecast_combination_mat3, forecast_combination_mat1] = calc_rmse_forecast_combination_conditional(nims(:,15:end),yields([find(tau==3),find(tau==24),find(tau==120)],15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, 1,4);
+[rmse_forecast_combination_mat4, forecast_combination_mat1] = calc_rmse_forecast_combination_conditional(nims(:,15:end),yields([find(tau==3),find(tau==120)],15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, 1,4);
+
+table4 = [rmse_forecast_combination_mat0
+          rmse_forecast_combination_mat1
+          rmse_forecast_combination_mat2
+          rmse_forecast_combination_mat3
+          rmse_forecast_combination_mat4
+          ];
+
+rowlabels = char('All yields',...
+                 'Observed Factors',...
+                 'Smoothed Factors',...
+                 '3-month, 2-year, 10-year',...
+                 '3-month, 10-year');
+      
+table4_tex = tablelatex(table4,columnlabels,rowlabels);
+char(table4_tex)
+
+
+% additional tests
+lag = 1;
[…]

### 2013-07-08 MATLAB/nim_lsc_model_2steps/utilities_2steps/load_data_ml.m
fixed compatibility issues with mac

@@ -67,13 +67,13 @@ yields_varlist = char('SVENY0025','SVENY0050','SVENY0075','SVENY0100',...
 %1-((FL704090005.Q+ FL734090005.Q)/((FL413065005.Q+FL674090005.Q+FL614090005.Q+FL664090005.Q+FL504090005.Q)+(FL704090005.Q+ FL734090005.Q)))
 %Data comes from Flow of Funds FAME db fof.
 
+competition_path = '..\data\shadow_banking_share_assets.csv';
 if isunix
-    shadow_bank_share_assets_file = ...
-        csvread('../data/measures_of_competitiveness/shadow_banking_share_assets.csv', 1, 1);
-else
-    shadow_bank_share_assets_file = ...
-        csvread('..\data\shadow_banking_share_assets.csv', 1, 1);
+    competition_path = strrep(competition_path,'\','/');
 end
+shadow_bank_share_assets_file = ...
+        csvread(competition_path, 1, 1);
+
 
 %This data has same dates as yields
 shadow_bank_share_assets = shadow_bank_share_assets_file(start_pos:end_pos,1)';
@@ -114,10 +114,9 @@ nims = nim_with_trading;
 
 if dataset_option == 1
     subcomponent_path = '..\data\data_ranked_by_total_assets\nims_subcomponents_by_endperiod_total_assets.csv';
-    if ~isunix
-        subcomponent_path = strrep(subcomponent_path,'/','\');
+    if isunix
+        subcomponent_path = strrep(subcomponent_path,'\','/');
     end
-
     additional_data  = csvread(subcomponent_path, 1, 9);
     
     total_ie_assets = additional_data(:,1);

### 2013-07-08 MATLAB/nim_lsc_model_2steps/run_calc_rmse_conditional.m
added PCR and PLS

@@ -82,7 +82,16 @@ lag = 1;
 [rmse_forecast_combination_mat1, forecast_combination_mat1] = calc_rmse_forecast_combination_conditional(nims(:,15:end),yields(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, 1,4);
 
 [rmse_multivariate_mat1, forecast_multivariate_mat1] = calc_rmse_multivariate_conditional(nims(:,15:end), smoothed_factors(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, lag);
+
+
 [rmse_forecast_combination_mat2, forecast_combination_mat2] = calc_rmse_forecast_combination_conditional(nims(:,15:end),smoothed_factors(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, 1,4);
+
+npc = 3;
+[rmse_pc_mat1, forecast_pc_mat1] = calc_rmse_pc(nims(:,15:end), yields(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, npc);
+
+npls = 5;
+[rmse_pls_mat1, forecast_pls_mat1] = calc_rmse_pls_conditional(nims(:,15:end), yields(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, npls);
+
 
 [rmse_forecast_combination_mat3, forecast_combination_mat3] = calc_rmse_forecast_combination_conditional(nims(:,15:end),factors(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, 1,4);
 
@@ -97,18 +106,24 @@ table1 = [
 rmse_forecast_combination_mat1
 rmse_multivariate_mat1
 rmse_forecast_combination_mat2
+rmse_pc_mat1
+rmse_pls_mat1
 rmse_forecast_combination_mat3
 rmse_varmat2
-rmse_nochangemat1];
+rmse_nochangemat1]
+
+rmse_pls_mat1
 
 
 columnlabels = char('Step 1','Step 2','Step 3','Step 4','Step 5','Step 6','Step 7','Step 8','Step 9','Step 10');
-rowlabels = char('1. Simple Forecast Combination',...
-                 '3a. DFM with 2nd Step Regression',...
-                 '3b. DFM with Forecast Combination',...
-                 '4. Forecast Combination of Observed Factors',...
[…]

### 2013-07-08 MATLAB/nim_lsc_model_2steps/utilities_2steps/calc_rmse_pc.m
utilities for PCR and PLS

@@ -1,0 +1,83 @@
+function [rmse_mat, forecast_mat] = calc_rmse_pc(yobs, other_obs, out_of_sample_start_pos, end_sample_pos, forecast_horizon, npc )
+
+% FOR NOW -- reglag does not affect anything. The regression is run with
+% only one lag of other_obs
+
+%This function computes rmse for a VAR forecast and also returns
+%forecast history over a pseudo-out of sample portion of the sample.
+%
+%Inputs: 
+%yobs is a row vector, a series spanning entire available sample.
+%other_obs is a matrix, each row is an additional series entering the
+%seond step regression and spanning the entire available sample. NB: The rmse is going to be
+%conditional on everything in other_var_obs. Needs to have the same number
+%of columns as yobs.
+%out_of_sample_start_pos is the position of the first pseudo-out of sample
+%observation.
+%end_sample_pos is the position of the last pseudo-out of sample
+%observation.
+%NPC is the number of the largest principal components used in the regression.
+%
+%Outputs:
+% rmse_mat is a row vector containing RMSEs for each forecast step.
+% var_forecasts is a matrix containing a forecast for each period in the
+% pseudo-out of sample. Each row of forecast_mat contains a forecast for
+% the horizon desired.
+
+%Now get RMSEs using the VAR model
+n_other_obs = size(other_obs, 1);
+
+num_periods = end_sample_pos - out_of_sample_start_pos + 1;
+forecast_errors_mat = nan*ones(num_periods, forecast_horizon);
+forecast_mat = nan*ones(num_periods, forecast_horizon);
+
+for pos_index = out_of_sample_start_pos:end_sample_pos
+    
+    
+    % strategy keep this part unchanged -- do not lag here. 
+    if  end_sample_pos - pos_index + 1  >= forecast_horizon 
+        
+        % notice that since everything is lagged in the regression, it was
+        % convenient to move these back by 1 period.
+    
+        other_obs_out_of_sample = other_obs(:,pos_index-1:pos_index+forecast_horizon-1-1);
+    else
+        other_obs_out_of_sample = other_obs(:,pos_index-1:end);  
+    end
+    
+    this_yobs = yobs(1:pos_index-1);
+    this_other_obs = other_obs(:,1:pos_index-1); %factors
+    
[…]

### 2013-07-08 MATLAB/nim_lsc_model_2steps/utilities_2steps/calc_rmse_pls_conditional.m

@@ -1,0 +1,87 @@
+function [rmse_mat, forecast_mat] = calc_rmse_pls_conditional(yobs, other_obs, out_of_sample_start_pos, end_sample_pos, forecast_horizon, npls )
+
+% FOR NOW -- reglag does not affect anything. The regression is run with
+% only one lag of other_obs
+
+%This function computes rmse for a VAR forecast and also returns
+%forecast history over a pseudo-out of sample portion of the sample.
+%
+%Inputs: 
+%yobs is a row vector, a series spanning entire available sample.
+%other_obs is a matrix, each row is an additional series entering the
+%seond step regression and spanning the entire available sample. NB: The rmse is going to be
+%conditional on everything in other_var_obs. Needs to have the same number
+%of columns as yobs.
+%out_of_sample_start_pos is the position of the first pseudo-out of sample
+%observation.
+%end_sample_pos is the position of the last pseudo-out of sample
+%observation.
+%NPC is the number of the largest principal components used in the regression.
+%
+%Outputs:
+% rmse_mat is a row vector containing RMSEs for each forecast step.
+% var_forecasts is a matrix containing a forecast for each period in the
+% pseudo-out of sample. Each row of forecast_mat contains a forecast for
+% the horizon desired.
+
+%Now get RMSEs using the VAR model
+n_other_obs = size(other_obs, 1);
+
+num_periods = end_sample_pos - out_of_sample_start_pos + 1;
+forecast_errors_mat = nan*ones(num_periods, forecast_horizon);
+forecast_mat = nan*ones(num_periods, forecast_horizon);
+
+for pos_index = out_of_sample_start_pos:end_sample_pos
+    
+    
+    % strategy keep this part unchanged -- do not lag here. 
+    if  end_sample_pos - pos_index + 1  >= forecast_horizon 
+        
+        % notice that since everything is lagged in the regression, it was
+        % convenient to move these back by 1 period.
+    
+        other_obs_out_of_sample = other_obs(:,pos_index-1:pos_index+forecast_horizon-1-1);
+    else
+        other_obs_out_of_sample = other_obs(:,pos_index-1:end);  
+    end
+    
+    this_yobs = yobs(1:pos_index-1);
+    this_other_obs = other_obs(:,1:pos_index-1); %factors
+    
[…]

### 2013-07-08 LATEX/slides/nim.slides.tex
slides

%\documentclass[handout,trans]{beamer}
%\documentclass{beamer}
\documentclass[handout]{beamer}
\mode<presentation>
\usepackage{beamerthemesplit}
\usepackage{hyperref}
\usepackage{color}
\usepackage{colortbl}
\usepackage{dcolumn}
\usepackage{verbatim} 
\newcolumntype{.}{D{.}{.}{3}}
%\usetheme{Berlin}
  \setbeamercovered{transparent}

\newcommand{\colorr}{ \color{red} }
\newcommand{\colorb}{ \color{blue} }
\newcommand{\colorg}{ \color{green} }
\newcommand{\colory}{ \color{yellow} }
\newcommand{\colorm}{ \color{magenta} }
\newcommand{\colorc}{ \color{cyan} }

\newcommand{\red}{\color{red}}
\newcommand{\black}{\color{black}}
\newcommand{\blue}{\color{blue}}

\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathpazo}
\usepackage{hyperref}
\usepackage{multimedia}

\setcounter{MaxMatrixCols}{10}

\title[Stressing Bank Profitability for Interest Rate Risk]
{Stressing Bank Profitability for Interest Rate Risk}
\author[Bolotnyy, Edge, \& Guerrieri]{Valentin Bolotnyy, Federal Reserve Board, \\  
                           \vspace{0.05in}Rochelle M. Edge, Federal Reserve Board, \\ 
                           \vspace{0.05in} and \\ 
                           \vspace{0.05in}Luca Guerrieri, Federal Reserve Board}
\date[8/2013]{July 8, 2013}

\begin{document}
%\AtBeginSubsection

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
\frametitle{Motivation} 
\begin{itemize}
\item \vspace{0.075in} Regular bank stress testing is one of the three complementary reforms made to U.S. capital regulation since the crisis.
\item \vspace{0.25in} There are two important parts of stress testing.
\begin{itemize}
\item \vspace{0.075in} Formulating scenarios that are stressful to banks
\item \vspace{0.075in} Developing models that are able to translate scenarios into income, losses, and \emph{pro-forma} capital.
\end{itemize}
\item \vspace{0.25in} It is important that stress tests are able to translate accurately scenarios into bank financial statement outcomes.
\begin{itemize}
\item \vspace{0.075in} We examine this question for models of bank net interest margins (NIMs). 
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Net interest margins (for the top 25 bank holding companies, assessed quarterly)}

\vspace{-0.15in}
\begin{eqnarray}\textsf{Net interest margins}\!\!\!\!&=&\!\!\!\!\frac{\textsf{Net interest income (NII)}}{\textsf{Interest bearing assets}} \nonumber \\
                \!\!\!\!&=&\!\!\!\!\frac{\textsf{Total interest income}-\textsf{Total interest expenses}}{\textsf{Interest bearing assets}}  \nonumber 
\end{eqnarray} 
\begin{center}
\includegraphics[height=5.4cm,width=7.2cm]{nims_crop.pdf}
\end{center}
%\begin{itemize}
%\item\vspace{0.0in} Net interest margins are our main variable of interest.
%\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Why study NIMs} 
\begin{itemize}
\item \vspace{0.0in} Losses from banks' earning depressed or negative NII is an important source of risk to banks and the financial sector.
\item \vspace{0.1in} The U.S. savings and loans crisis was associated with NII and NIMs turning negative in the sector.
\end{itemize}
\begin{center}
\includegraphics[height=4.4cm,width=5.9cm]{NIMs_Thrifts_CROP.pdf}
\end{center}
\vspace{-0.2in} \hspace{1.0in} {\tiny{Source: FR Bulletin, February 1996.}}\normalsize
\end{frame}

\begin{frame}
\frametitle{Why study NIMs, continued} 
\begin{itemize}
\item \vspace{0.0in} A few other banking crises have also been associated with depressed or negative NIMs; including,       
\begin{itemize}
\item \vspace{0.125in} The Nordic banking crises (in Finland, Sweden, and Norway) of the late 1980s/early 1990s; and, 
\item \vspace{0.125in} The U.K. secondary banking of 1973-75. 
\end{itemize}
\item \vspace{0.25in} NII is an important part of bank profitability.
\begin{itemize}
\item \vspace{0.125in} Bank equity analysis direct considerable attention to NII and NIMs in evaluating future bank profitability.
\item \vspace{0.125in} This reflects the fact that NII and NIMs are less volatile than other components of bank revenue.
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Objective of this paper}
\begin{itemize}
\item \vspace{0.0in} To develop a model that can be used to forecast bank net interest margins (NIMs). 
\end{itemize}
\vspace{-0.05in} 
\hspace{0.35in}\[\textsf{Net interest margins (NIMs)}=\frac{\textsf{Net interest income (NIIs)}}{\textsf{Interest bearing assets}} \] 
%\item \vspace{0.075in} $\textsf{Net interest margins (NIMs)} = 
%\frac{\textsf{Net interest income (NII)}}{\textsf{Interest bearing assets}}$ 
\begin{itemize}
\item \vspace{0.05in} We focus on conditional forecast performance.  
\begin{itemize}
\item \vspace{0.075in} Our interest is not in forecasting the macroeconomy (and especially not interest rates). 
\item \vspace{0.075in} Our interest is in forecasting bank profitability given the macroeconomy. 
\item \vspace{0.075in} Stress tests -- an important use of our model -- 
                       are conducted conditional on a set of macroeconomic scenarios.
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Key variables for modeling NIMs}
\begin{itemize}
\item \vspace{0.0in} The slope of the Treasury yield curve. 
\begin{itemize}
\item \vspace{0.05in} Reflects banks' return on maturity-transformation serivces -- one of the key services provided by banks.  
\end{itemize}
\item \vspace{0.1in} The level of short-term interest rates. 
\begin{itemize}
\item \vspace{0.05in} Indirectly reflects banks' return on transactions services -- another key service provided by banks.
\item \vspace{0.05in} The level of the short rate puts an upper limit on how much banks can earn from transactions services.  
\end{itemize}
\item \vspace{0.1in} The use of these variables is common in the macro-banking literature modeling NIMs.
\begin{itemize}
\item \vspace{0.05in} Kovner and Vickery (2012)
\item \vspace{0.05in} Covas, Rump, and Zakrajsek (2012)
\item \vspace{0.05in} English (2002)
\item \vspace{0.05in} English, Van den Heuvel, and Zakrajsek (2012)
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Key variables for modeling NIMs, continued}

\vspace{0.1in}\hspace{0.1in}
\includegraphics[height=3.6cm,width=4.8cm]{nims_slope_crop.pdf}  \hspace{0.25in}
\includegraphics[height=3.6cm,width=4.8cm]{nims_level_crop.pdf}
\begin{itemize}
\item\vspace{0.1in} NIMs increase when the yield-curve steepens.
\begin{itemize}
\item \vspace{0.025in} Reflects the increased return to maturity transformation.
\end{itemize}
\item \vspace{0.075in} Changes in the short rate drive changes in the slope of the yield curve. 
\item \vspace{0.075in} There is an evident downward trend in the time series. 
\begin{itemize}
\item \vspace{0.025in} This may be tied to the lower level of interest rates.
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Key variables for modeling NIMs, continued}

%\begin{columns}[4cm]
%\column{7.2cm}
\vspace{0.1in}\hspace{0.65in}\includegraphics[height=5.4cm,width=7.2cm]{nims_components_crop.pdf}

%\column{4.8cm}
\begin{itemize}
\item\vspace{0.1in} The influence of the level of interest rates is evident in the two components of NIMs:
\begin{itemize}
\item \vspace{0.025in} Total interest income/Interest bearing assets; and,
\item \vspace{0.025in} Total interest expense/Interest bearing assets.
\end{itemize}
\end{itemize}
%\end{columns}
\end{frame}

\begin{frame}
\frametitle{Other key variables for modeling NIMs}
\begin{itemize}
\item \vspace{0.0in} The micro-banking literature on modeling NIMs emphasizes some different variables; such as,
\begin{itemize}
\item \vspace{0.05in} The degree of competition faced by banks in loan and deposit markets; and,
\item \vspace{0.05in} The volatility of interest rates.
\end{itemize}
\item\vspace{0.1in} The greater the degree of competition faced by banks in loan and deposit markets, 
\begin{itemize}
\item \vspace{0.05in} The lower the level of NIMs that banks can set. 
\end{itemize}
\item\vspace{0.1in} If banks are risk averse, the greater the volatility of interest rates,
\begin{itemize}
\item \vspace{0.05in} The more compensation for risk that banks will require to take deposits and make loans, and,
\item \vspace{0.05in} The higher the level of NIMs that banks will set. 
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Other key variables for modeling NIMs, continued}
\begin{itemize}
\item \vspace{0.15in} Relevant papers in the empirical micro-banking literature on modeling NIMs include:
\begin{itemize}
\item \vspace{0.15in} Ho and Saunders (1981) 
\item \vspace{0.15in} Angbazo (1997) 
\item \vspace{0.15in} Saunders and Schumacher (2000)
\end{itemize}

\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Other key variables for modeling NIMs, continued}

\begin{columns}[4cm]

\column{7.2cm}
\includegraphics[height=5.4cm,width=7.2cm]{nims_competition_crop.pdf}

\column{4.8cm}
\begin{itemize}
\item\vspace{0.0in} The banking system's share of credit inter- mediation has shrunk.
\begin{itemize}
\item \vspace{0.05in} From 82\% in 1985 to 47\% in 2007.
\item \vspace{0.05in} It is now at 55\%.
\end{itemize}
\item \vspace{0.125in} Competition from the shadow banking sector has put pressure on deposit and loan rates.
\item \vspace{0.125in} One other variable we have not yet considered is interest rate volatility.
\end{itemize}
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Net interest margins:  A few quirks in the data}

\begin{columns}[4cm]

\column{7.2cm}
\includegraphics[height=5.4cm,width=7.2cm]{nims_bugs_crop.pdf}

\column{4.8cm}
\begin{itemize}
\item\vspace{-0.15in} The spike in 1988:Q4 reflects a jump then drop in ``other'' interest income.
\begin{itemize}
\item \vspace{0.075in} This reflects interest from Brazil overdue since early 1987.
\end{itemize}
\item \vspace{0.15in} The jump in 2010:Q1 reflects FAS 167 going into effect. 
\begin{itemize}
\item \vspace{0.075in} We will adjust for this jump.
\end{itemize}
\end{itemize}
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Forecasting models}

We are interesting in comparing the pseudo-out-of-sample forecasting performance of different models:
\begin{enumerate}
\item Forecast combination based on yields
\item Dynamic factor model
\item Dynamic factor model with second-step regression
\item Principal Component Regression based on yields
\item Partial Least Squares based on yields
\item Forecast combination based on observed yield-curve factors
\item A VAR with observed yield curve factors
\item A random walk without a drift
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{What's next}
\begin{itemize}

\item Next -- We describe each of the models in our horse race

\item We focus on the baseline specification for each model

\item Each model can be extended to include additional variables (sensitivity analysis)

\end{itemize}
\end{frame}

\begin{frame}

\frametitle{1. Forecast Combination Based on Yields }
We regress NIMs on each yield $r(\tau )_{t-1}$ separately: \
\begin{equation*}
NIM_{t}=c_{\tau }+\rho _{\tau }NIM_{t-1}+\gamma _{\tau }r(\tau )_{t-1}+{\eta
_{\tau ,t}}
\end{equation*}

Each regression is used to produce a recursive $s$ step-ahead forecast of NIMs
conditional on yields on treasuries with maturity $\tau $ observed through
period $t+s-1.$ This conditional forecast is denoted by $NIM_{\tau ,t+s/t}.$%
The simple forecast combination is then given by:%
\begin{equation*}
NIM_{t+s/t}=\sum_{\tau }\frac{NIM_{\tau ,t+s/t}}{N},
\end{equation*}%
where $N$ is the number of maturities considered.

\end{frame}

\begin{frame}
\frametitle{2. A Dynamic Factor Model}
Using the state space
representation, the state equations of our dynamic factor model
is given by
\begin{equation*}
\left(
\begin{array}{c}
{L_{t+1}-\mu _{L}} \\
{S_{t+1}-\mu _{S}} \\
{C_{t+1}-\mu _{C}} \\
\widehat{NIM}_{t+1}-\mu _{\widehat{NIM}}%
\end{array}%
\right) {\normalsize \ =\ }\left(
\begin{array}{cccc}
a_{11} & a_{12} & a_{13} & 0 \\
a_{21} & a_{22} & a_{23} & 0 \\
a_{31} & a_{32} & a_{33} & 0 \\
a_{41} & a_{42} & a_{43} & a_{44}%
\end{array}%
\right) {\normalsize \ }\left(
\begin{array}{c}
{L_{t}-\mu _{L}} \\
{S_{t}-\mu _{S}} \\
{C_{t}-\mu _{C}} \\
\widehat{NIM}_{t}-\mu _{\widehat{NIM}}%
\end{array}%
\right) {\normalsize +}\left(
\begin{array}{c}
{\eta _{Lt}} \\
{\eta _{St}} \\
{\eta _{Ct}} \\
{\eta _{NIMt}}%
\end{array}%
{\normalsize \ }\right) ,
\end{equation*}

$L_{t},S_{t},$ and $C_{t}$ are level slope and curvature factors for
the yield curve, respectively, and $\widehat{NIM}_{t}$ represent a measure
of NIMs cleaned of observation error.

The term $\mu _{x}$ denotes the mean
of $x$. The terms ${\eta _{Lt}}$, ${\eta _{St}}$, ${\eta _{Ct}}$, ${\eta
_{NIMt}}$ denote independently distributed innovations with joint covariance
matrix $\Omega $.

\end{frame}

\begin{frame}
\frametitle{2. A Dynamic Factor Model -- Observation Equation}

The observation equations for yields on treasuries of maturity $\tau $ take the form:
\begin{equation*}
r(\tau )=\left(
\begin{array}{cccc}
1 & \frac{1-e^{-\lambda \tau }}{\lambda \tau } & \frac{1-e^{-\lambda \tau }}{%
\lambda \tau }-e^{-\lambda \tau } & 0%
\end{array}%
\right) \left(
\begin{array}{c}
{L_{t}-\mu _{L}} \\
{S_{t}-\mu _{S}} \\
{C_{t}-\mu _{C}} \\
\widehat{NIM}_{t}-\mu _{\widehat{NIM}}%
\end{array}%
\right) +e(\tau )_{t},
\end{equation*}%
The observation equation for NIMs takes the form:
\begin{equation*}
NIM_{t}=\left(
\begin{array}{cccc}
0 & 0 & 0 & 1%
\end{array}%
\right) \left(
\begin{array}{c}
{L_{t}-\mu _{L}} \\
{S_{t}-\mu _{S}} \\
{C_{t}-\mu _{C}} \\
\widehat{NIM}_{t}-\mu _{\widehat{NIM}}%
\end{array}%
\right) +e(NIM)_{t},
\end{equation*}

Out-of-sample conditional forecasts for NIMs from this model, are obtained
using the Kalman filter.

\end{frame}

\begin{frame}

\frametitle{3. Dynamic Factor Model with Second-Step Regression}

First, we estimate the level, slope, and curvature factors in isolation:
\begin{equation*}
\left(
\begin{array}{c}
{L_{t+1}-\mu _{L}} \\
{S_{t+1}-\mu _{S}} \\
{C_{t+1}-\mu _{C}}%
\end{array}%
\right) {\normalsize \ =\ }\left(
\begin{array}{ccc}
a_{11} & a_{12} & a_{13} \\
a_{21} & a_{22} & a_{23} \\
a_{31} & a_{32} & a_{33}%
\end{array}%
\right) {\normalsize \ }\left(
\begin{array}{c}
{L_{t}-\mu _{L}} \\
{S_{t}-\mu _{S}} \\
{C_{t}-\mu _{C}}%
\end{array}%
\right) {\normalsize +}\left(
\begin{array}{c}
{\eta _{Lt}} \\
{\eta _{St}} \\
{\eta _{Ct}}%
\end{array}%
\right)
\end{equation*}

\begin{equation*}
r(\tau )=\left(
\begin{array}{ccc}
1 & \frac{1-e^{-\lambda \tau }}{\lambda \tau } & \frac{1-e^{-\lambda \tau }}{%
\lambda \tau }-e^{-\lambda \tau }%
\end{array}%
\right) \left(
\begin{array}{c}
{L_{t}-\mu _{L}} \\
{S_{t}-\mu _{S}} \\
{C_{t}-\mu _{C}} 
\end{array}%
\right) +e(\tau )_{t},
\end{equation*}

We obtain smoothed estimates of the factors $L_{t},S_{t},$ and $C_{t}$
conditional on information up until period $T.$ 
\end{frame}

\begin{frame}

\frametitle{3. Dynamic Factor Model with Second-Step Regression -- Continued }

{\blue Alternative a.}  A multivariate regression
\begin{equation*}
NIM_{t}=c+\rho NIM_{t-1}+\gamma _{L}L_{t-1}+\gamma _{S}S_{t-1}+\gamma
_{C}C_{t-1}
\end{equation*}

{\blue Alternative b.}  Forecast Combination (each factor included in a
separate equation and forecasts from each equation combined) 
\begin{equation*}
NIM_{t}=c_{i}+\rho _{i}NIM_{t-1}+\gamma _{i}F_{i,t-1}+{\eta _{i,t},}
\end{equation*}%
where $F_{i}\in \left\{ L,S,C\right\} .$ Forecasts $NIM_{i,t+s/t}$ from each
separate regression are then aggregated as:
\begin{equation*}
NIM_{t+s/t}=\sum_{i}\frac{NIM_{i,t+s/t}}{3}.
\end{equation*}

\end{frame}

\begin{frame}
\frametitle{4. Principal Component Regression}

\begin{itemize}
\item Let $X$ a $T \times n$ matrix of $n$ Treasury yields.  Using the spectral decomposition, one obtains $X'X = VDV'$.

\item Let $V_i$ be the eigenvector corresponding to $i^{th}$ largest eigenvalue.

\item The $i^{th}$ principal component is defined as $PC_i = XV_i$.

\item We estimate:
\begin{equation*}
NIM_t = c + \rho NIM_{t-1} + \sum_{i=1}^3 \gamma_i PC_{i,t-1}. 
\end{equation*}
\end{itemize}

\end{frame}

\begin{frame}

\frametitle{5. Partial Least Squares}

\begin{itemize}
\item Partial least squares is a compression technique analogous to principal components.
\item Project the regressors onto a lower dimensional space. 
\item Each new regressor is a linear combination of the original regressors.
\item What distinguishes PLS from PCR is how the
projection is done.
\item In particular, PCR chooses basis vectors of its low
dimensional projection to describe as much as
possible the variation in the data $X$.
\item However nothing guarantees that the principal
components, which “explain” $X$ optimally, will be
relevant for the prediction of $Y$.
\item Solution: incorporate information from $Y$ when
choosing the projection (use SVD of $X'Y$).

\end{itemize}
\end{frame}

\begin{frame}
\frametitle{5. Partial Least Squares -- continued}

Choose X to include 
\begin{itemize}
\item Lagged NIM
\item Yields 
\end{itemize}

Focus on PLS with 4 factors.

\end{frame}

\begin{frame}
\frametitle{6. Forecast Combination Based on Observed Factors }

\begin{itemize}
\item Define $\left\{
L,S,C\right\}$ the observed factors as:

$L = \frac{r(3m) + r(2yr) + r(10yr)}{3}$

$S = r(3m)-r(2yr)$

$C = \left[ r(3m)-r(2yr) \right] - \left[ r(2yr)-r(10yr) \right]$

\item Consider
\begin{equation*}
NIM_{i,t}=c_{i}+\rho _{i}NIM_{t-1}+\gamma _{i}O_{i,t-1}+{\eta _{i,t},}
\end{equation*}%
where $O_{i}\in \left\{ L,S,C\right\} .$

\item Forecasts  from each separate
regression, $NIM_{i,t+s/t}$, are aggregated as:
\begin{equation*}
NIM_{t+s/t}=\sum_{\tau }\frac{NIM_{i,t+s/t}}{N}.
\end{equation*}%

\end{itemize}
\end{frame}

\begin{frame}

\frametitle{7. VAR}

\begin{itemize}

\item The VAR(4) includes:\begin{itemize} \item Observed Level factor
\item Observed slope factor
\item Observed curvature factor
\item NIMs. 
\end{itemize}

\item Forecasts conditional on the factors are obtained using
the Kalman filter.

\end{itemize}

\end{frame}

\begin{frame}

\frametitle{8. No-Change Forecast.}

Using a random walk without a drift, the forecast for NIMs is:

\begin{equation*}
NIM_{t+s/t}=NIM_{t}.
\end{equation*}

\end{frame}

\begin{frame}

\frametitle{Regression Data}

\begin{itemize}

\item The data NIMs  is from the quarterly Consolidated Reports of Condition and Income (Call Report) -- we roll the bank-level information up to BHC level.

\item We aggregate NIMs for the top 25 BHCs, as ranked by total assets, which is assessed quarterly.

\item The yields data are derived using a smoothing technique from Gurkaynak, et al. (2007), based on Nelson and Siegel (1987) and Svensson (1994), which allows for daily measures of an off-the-run Treasury yield curve.

\item We use yields for twelve maturities in our models: 3 month, 6 month, 9 month, 1 year, 2 year, 3 year, 5 year, 7 year, 10 year, 15 year, 20 year, and 30 year.

\item The yields included are quarterly averages of daily yields.

\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Pseudo-out-of-sample Forecasts}

\begin{itemize}

\item Our data on NIMs start 1985Q4.

\item To avoid the spike in NIMS associated with late payments from the Latin American debt crisis, our estimation sample starts in 1989q2.

\item Where necessary, we still use the observations between 1985q4 and 1989a2 to train the Kalman filter.

\item We use a recursive estimation window.

\item The shortest window ends in 1999q4.

\item The evaluation window spans 2000q1 through 2008q2.

\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Baseline Results -- RMSEs} 
\footnotesize
\begin{table}                                                                                     
\center                                                                                           
\begin{tabular}{|l|c|c|c|c|c|c|c|c|}                                                          
\hline                                                                                            
&Step 1 &Step 2 &Step 3 &Step 4 &Step 5 &Step 6 &Step 7 &Step 8 \\                
\hline                                                                                            
1.           &0.107&0.133&0.148&0.168&0.187&0.202&0.213&0.209\\
2.           &0.122&0.148&0.169&0.212&0.247&0.278&0.316&0.340\\
3a.         &0.118&0.167&0.209&0.260&0.308&0.356&0.398&0.433\\
3b.         &0.112&0.159&0.199&0.247&0.300&0.351&0.399&0.438\\
4.           &0.118&0.167&0.208&0.259&0.307&0.355&0.397&0.432\\
5.           &0.127&0.182&0.228&0.279&0.325&0.368&0.406&0.434\\
6. 	     &0.118&0.169&0.216&0.273&0.336&0.399&0.456&0.509\\
7.           &0.166&0.209&0.263&0.302&0.342&0.375&0.380&0.430\\
8.           &0.107&0.138&0.154&0.181&0.206&0.228&0.244&0.248\\
\hline                                                                                            
\end{tabular}                                                                                     
\end{table}

1. F. Combination of yields;  2. DFM; 3a. DFM with 2nd step multivariate regression; 3b. DFM with 2nd step forecast combination; 4. PCR; 5. PLS. 6. F. Combination of observed yields; 7. VAR on observed yields; 8. no change forecast.

\end{frame}

\begin{frame}
\frametitle{Competition from the shadow banking sector}

ADD SOMETHING ON WHAT IS BEING CAPTURED AS A SHADOW BANK

\end{frame}

\begin{frame}
\frametitle{Results -- accounting for competition from the shadow-banking sector} 
\footnotesize
\begin{table}                                                                                     
\center                                                                                           
\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|}                                                          
\hline                                                                                            
&Step 1 &Step 2 &Step 3 &Step 4 &Step 5 &Step 6 &Step 7 &Step 8 \\                
\hline                                                                                            
1.   &0.110&0.143&0.165&0.192&0.222&0.248&0.267&0.274\\
3a. &0.121&0.170&0.212&0.262&0.312&0.360&0.403&0.436\\
3b. &0.106&0.140&0.165&0.193&0.222&0.248&0.267&0.274\\
4.   &0.117&0.160&0.194&0.229&0.263&0.299&0.333&0.360\\
5.   &0.127&0.182&0.228&0.279&0.325&0.368&0.406&0.435\\
6.   &0.109&0.141&0.165&0.193&0.222&0.249&0.267&0.274\\
7.   &0.344&0.396&0.476&0.494&0.521&0.569&0.631&0.670\\
8.   &0.107&0.138&0.154&0.181&0.206&0.228&0.244&0.248\\
\hline                                                                                            
\end{tabular}                                                                                     
\end{table}                                                                                       
1. F. Combination of yields; 3a. DFM with 2nd step multivariate regression; 3b. DFM with 2nd step forecast combination; 4. PCR; 5. PLS. 6. F. Combination of observed yields; 7. VAR on observed yields; 8. no change forecast.
\end{frame}

\begin{frame}
\frametitle{Results -- Alternative Forecast Combinations} 
\footnotesize
\begin{table}                                                                         
\center                                                                               
\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|}                                              
\hline                                                                                
&Step 1 &Step 2 &Step 3 &Step 4 &Step 5 &Step 6 &Step 7 &Step 8 \\    
\hline                                                                                
a.  &0.107&0.133&0.148&0.168&0.187&0.202&0.213&0.209\\
b.  &0.118&0.169&0.216&0.273&0.336&0.399&0.456&0.509\\
c.  &0.112&0.159&0.199&0.247&0.300&0.351&0.399&0.438\\
d.  &0.105&0.130&0.141&0.160&0.178&0.197&0.202&0.198\\
e.  &0.105&0.129&0.144&0.159&0.179&0.196&0.200&0.200\\
\hline                                                                                
\end{tabular}                                                                         
\end{table}     
a. All 12 yields; b. observed factors; c. smoothed factors; d. 3-month, 2-year, 10-year;  e. 3-month, 10-year.
\end{frame}

%\begin{frame}
%\frametitle{Scenario analysis} 
%\begin{itemize}
%\item \vspace{0.0in} SHOW RESULTS FOR CCAR-2013 SEVERELY ADVERSE AND ADVERSE SCENARIOS.
%\end{itemize}
%\end{frame}

\begin{frame}
\frametitle{Planned extensions} 
\begin{itemize}
\item \vspace{0.0in} Switching regime at the zero lower bound
\item \vspace{0.0in} Panel estimation on pro-forma data
\item \vspace{0.0in} Forecasting stock-market returns
\item \vspace{0.0in} Additional sensitivity analysis
\end{itemize}
\end{frame}

\end{document}

### 2013-07-08 LATEX/beg.draft.2013.tex
Rochelle's draft

\documentclass[11pt]{article}
\usepackage[dvips]{graphics}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{chicago}
\usepackage{rotating}
\usepackage{hyperref}
\usepackage{sectsty}

\addtolength{\topmargin}{-0.375in}
\addtolength{\oddsidemargin}{-0.35in}
\addtolength{\evensidemargin}{-0.35in}
\addtolength{\textheight}{0.5in}
\addtolength{\textwidth}{0.5in}
\addtolength{\parskip}{0.05in}
\renewcommand{\baselinestretch}{1.5}
\newcommand{\vsp}{\vspace{\baselineskip}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\gr}[1]{\ensuremath  {\frac{\dot {#1}}{#1}}}

\begin{document}
\renewcommand{\baselinestretch}{1}
\author{{\Large Valentin Bolotnyy} \\
Federal Reserve Board\thanks{E-mail: valentin.bolotnyy@frb.gov.\vspace{0.05in}} \and
{\Large Rochelle M. Edge} \\
Federal Reserve Board\thanks{E-mail: rochelle.m.edge@frb.gov.\vspace{0.05in}}\and
{\Large Luca Guerrieri} \\
\addtocounter{footnote}{5} Federal Reserve Board\thanks{%
E-mail: luca.guerrieri@frb.gov.}}
\title{\textbf{Stressing Bank Profitability for Interest Rate Risk}}
\date{\today}
\maketitle

\begin{abstract}
\vspace{-0.06in} \noindent  Regular bank stress testing (and the capital planning that it implies) is one of the three complementary reforms that has been made to U.S. capital regulation in response to the shortcomings of the pre-crisis regime.
 The form of stress testing mandated under the Dodd-Frank Act for the Federal Reserve to undertake and the form being used by the Federal Reserve for its annual Comprehensive Capital Analysis and Review (CCAR) is one based around a small number of scenarios.  This approach involves specifying macroeconomic and financial scenarios that represents stressful conditions over the time horizon of the stress test and then projecting, for each BHC in the stress test, its losses, income, and path of capital \textit{given that scenario}. In this paper, we focus on the revenue generate aspect of banks and consider the ability of empirical models to translate developments in interest rates into bank net interest margins (NIMs). We consider an extensive number of time-series models. While our best-performing models beat a no-change forecast, the confidence bands are so large that baseline and stress scenarios taken from CCAR appear indistinguishable.

\vspace{.125in}

\noindent Key Words:

\vspace{.125in}

\noindent JEL Classification: To be completed.

\end{abstract}
\renewcommand{\baselinestretch}{1.5}
\thispagestyle{empty}

\newpage
\setcounter{page}{1}

\section{Introduction}

\vspace{-0.1in}
Regular bank stress testing (and the capital planning that it implies) is one of the three complementary reforms that has been made to U.S. capital regulation in response to the shortcomings of the pre-crisis regime.\renewcommand{\baselinestretch}{1}\footnote{See Governor Tarullo’s November 2011 speech ``The Evolution of Capital Regulation'' for this characterization of stress testing. The two other reforms that he mentions are improvements to the traditional, individual firm approach to capital regulation and the introduction of a macroprudential component to capital regulation.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} While stress testing can take many forms---for example, being based on a small number of scenarios, being based on a large number of scenarios, or even being undertaken in reverse form to achieve a certain adverse outcome---the form of the stress tests mandated under the Dodd-Frank Act for the Federal Reserve to undertake and the form being used by the Federal Reserve for its annual Comprehensive Capital Analysis and Review (CCAR) is one based around a small number of scenarios.  This approach involves specifying a small number of macroeconomic and financial scenarios that represents stressful conditions over the time horizon of the stress test and then projecting, for each BHC in the stress test, its losses, income, and path of capital \textit{given that scenario}. Clearly, there are at least two important parts to conducting bank stress tests (i)~formulating macroeconomic and financial scenarios that should be stressful to banks and (ii)~developing econometric models that are able to translate such stressful conditions into losses, income, and capital.
%This paper considers the latter of these two issues; that is, translating
%stressful economic conditions into bank income.

In developing econometric models to translate stressful economic conditions into key bank financial-statement variables it is important to consider what types of developments are likely to be stressful to banks and what particular economic variables are likely to capture these stresses. It is then critical that the models being used for stress testing are able to translate developments in these economic variables into the key bank financial-statement variables that the model is designed to project.  In this paper, we examine this property for models of bank net interest margins (NIMs), defined as the ratio of net interest income (NII) to interest earning assets, where NII is the difference between a bank's interest income and interest expenses.

Our motivation for focusing on models of NIMs is based on a number of considerations. First, bank losses that arise from banks' having to pay out more in interest on their liabilities than they are earning from their assets should not be overlooked as an important source of risk to the banking and broader financial sector.  To be sure, losses that arise from this source of risk---called, interest rate risk---were much less prominent in the most recent crisis relative to bank losses that stem from loan defaults---that is, credit risk---but there are ample examples of other earlier banking crises in which adverse developments in NIMs have played a significant role.  The most familiar of these from the U.S. perspective is the Savings and Loans (S\&L) crisis that started in the early 1980s with short-term interest rates rising above long-term interest rates---in part, as a result of the Volcker disinflation.  This resulted in interest expenses rising above interest income, NII and NIMs turning negative (in the thrift sector), and sizable losses and ultimately a substantial number of bank failures.  Likewise, the Nordic banking crises of the late 1980s and early 1990s (for Finland, Sweden, and Norway)---which represent three of Reinhart's and Rogoff's~(2009) ``big five'' banking crises---were associated with elevated short-term interest rates that depressed NII and NIMs and lead to sizable bank losses and failures. In the Nordic crises, high short-term interest rates resulted from these countries attempting to maintain currency pegs with the Deutschemark in the face of the tight German monetary policy that accompanied its loose fiscal policy associated with re-unification spending. Finally, there is also the U.K.'s secondary banking crisis of the early 1970s, which was associated with a sudden increase in short-term interest rates that abruptly drove up banks' costs of funds so too leading to bank losses and failures.  To be sure, large increases in short-term rates and declines in net interest income were not the \textit{only} reason for any of these crises---indeed, there was at least one major class of asset prices that collapsed in each of these crises---but short-term interest rates---and, in particular, their elevated levels relative to long-term interest rates---were significant and in most cases equivalent in importance to the other causes of the crisis.

Another motivation for our focus on models of NIM stems from the recent lackluster pace of bank profit growth, which is becoming an increasing source of concern to market commentators and regulators alike. Bank net income growth in recent years has been largely supported by unsustainable cost cutting and reductions in provisioning, rather than from sources like NII.  NII growth, for example, (in both nominal and real terms) has been well below average in recent years while NIMs (NII relative to interest earning assets) have also been depressed.%\renewcommand{\baselinestretch}{1}\footnote{We should probably show both these  charts with adjustments for FAS 167.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5}  
Given their already low level, any adverse development that puts further downward pressure on NIMs could be very stressful for banks and, as such, it is important that the economic developments (likely involving interest rates) that would drive these outcomes are well-modeled.

Our final motivation for examining NIMs is that despite the importance of NIMs for the financial condition of banks \textbf{[and the attention that it receives from bank equity analysts]}, the modeling of NIMs has received much less attention in the risk-modeling literature relative to the modeling of bank variables related to credit risk.  To be sure, the recent increase in the importance of stress testing has led to the development of a number of models that link NIMs to macroeconomic variables---such as, short-term Treasury rates and the slope of the Treasury yield curve---with Kovner \textit{et al}.~(2011) and Covas \textit{et al}.~(2012) being two notable examples for the U.S.  However, this research places relatively less emphasis on the conditional forecasting properties of these models, which is important given that these models are ultimately to be used to project bank losses, income, and capital given macroeconomic scenarios.\renewcommand{\baselinestretch}{1}\footnote{Covas \textit{et al}. do present some forecast evaluation results in order to demonstrate the benefits of focusing on quantile projections. However, they only report results for net chargeoffs and pre-provisioning net revenue and their focus is on density forecast evaluation.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5}

In summing up our motivation on the importance of modeling NIMs and studying interest rate risk, we would also note that the Federal Reserve chose to feature interest rate risk in one of the scenarios specified for the 2013 Dodd-Frank Act (DFA) stress tests.  In particular, for the 2013 DFA stress tests, the Federal Reserve specified the adverse scenario as one in which the Treasury yield curve shifted up and flattened.  As noted by the Federal Reserve in its (proposed) Policy Statement on scenario design, the Board varies ``the approach it uses for the adverse scenario each year so that the results of the scenario provide the most value to supervisors, in light of current condition of the economy and the financial services industry.''

In this paper we examine how well a range of time-series models can explain NIMs. We emphasize the models' pseudo-out-of-sample \textit{conditional} forecast performance. We adopt this pseudo-out-of-sample approach for the simple reason that the evaluation of models based on their in-sample forecast performance, frequently leads to the econometrician over-fitting the model within the sample so as to maximize in-sample forecast performance. We focus on conditional forecasts for the reason that scenario-based stress testing of the banking system is the underlying motivation for our analysis and it is along the dimensions of \textit{conditional} forecasts that model forecast performance is important.  Focusing on conditional forecast performance also means that we can separate the task of modeling and forecasting NIMs from that of forecasting interest rates, which is itself a sizable topic (see Duffee, 2012).

In determining which variables to include in our model of NIMs we look to the variables that have been emphasized by two existing literatures on NIMs.  The first literature emphasizes the link between interest rates of different maturities and NIMs and is more in the macro-banking tradition. The second literature focuses on the optimal margin set by banks based on considerations like the competitive structure of the lending and deposit markets in which the bank participates, the degree of managerial risk aversion, and the degree of uncertainty associated with the future path of interest rates and is much more in the micro-banking tradition.

We proceed as follows. Section~two provides a review of the literature on NIMs with an eye to cataloguing variables that have been found useful to explain the evolution of NIMs. Section~three turns to the models that we develop to forecast NIMS. Section 4 concludes.

\section{Related literature}

\vspace{-0.1in}
There are primarily two empirical literatures on modeling NIMs.  The first literature emphasizes the link between interest rates of different maturities and NIMs and lies more in the macro-banking tradition.  The second literature focuses on the optimal margin set by banks and lies more in the micro-banking tradition.

The papers for the U.S. by Kovner \textit{et al}. and Covas \textit{et al}. noted earlier fall into the category of papers in the macro-banking tradition, where in both cases these models link NIMs to  a measure of the slope of the yield curve and a measure of short-term market rate.  Banks' earnings from interest income reflect two key services that banks supply---maturity transformation services (that is, borrowing short and lending long) and deposit transactions services---and the returns to these services are reflected by these two interest rate variables.  The slope of the yield curve reflects the return to banks from maturity transformation and thus is expected to enter NIM models with a positive sign. The short-term market rates in NIM models reflect the fact that bank deposit rates, while typically lower (since they provide transactions services), are constrained by the zero lower bound.  As such, the short rate places an upper limit on what banks can earn from the provision of their transactions services. As a result, the short rate is expected to enter the model, also, with a positive sign.

Both Kovner \textit{et al}. and Covas \textit{et al}. measure the slope of the yield curve in their respective NIM models with the 10-year to 3-month Treasury term spread and short-term market rates with the 3-month Treasury bill rate, as do English~(2002) and English \textit{et al}.~(2012)---two other papers in this tradition.  All of these papers estimate NIM models with slightly different purposes in mind.  The NIM models estimated by Kovner \textit{et al}. and Covas \textit{et al}. are developed for stress testing and are part of much larger aggregative models of a number of components of banks' financial statements. These models are both estimated on panel data with the most notable difference between these models being the estimation techniques they employ: Least squares regression in the case of Kovner \textit{et al}. and quantile regression in the case of Covas \textit{et al}.

Like our paper, English also only studies NIMs, although he does so for a number of countries (that is, all G-7 countries except for France, and for Australia, Norway, Sweden, and Switzerland).  Addditionally, (and as we also do later), he considers asset yields and liability yields separately---the two measures that when differenced imply NIMs.  English's main focus is understanding how well the aggregate banking sectors in the countries he studies appear to have managed the interest rate risk that they face on their earnings, as such he uses aggregate data (rather than firm-level data) and does not focus on the conditional forecasting performance of his model.

Finally, English \textit{et al} consider the relationship between NIMs and the slope of the yield curve and short-term rates for a panel of banks, however, in contrast to the other studies they allow for the specific coefficients estimated for each bank to vary as linear functions of bank balance sheet ratios that one \textit{a priori} might think would be relevant to net interest income.  In particular, these ratios include the maturity gap between bank assets and liabilities, the deposit share of liabilities, and the loan share of assets. Again, since English \textit{et al} are focussed on understanding the determinants of NIMs they do not consider conditional forecast performance.

  A separate strand of the NIM modeling literature comes from the micro-banking tradition, which focuses on the determinants of the loan rates and deposit rates that banks set as implied by the maximization of their profits. Papers in this tradition---such as, Ho and Saunders~(1981), Angbazo~(1997), and Saunders and Schumacher~(2000)---are based around what is known as the ``dealership'' model of banks; that is, banks in these models are viewed as dealers in the credit market, acting as an intermediary between demanders and suppliers of funds.  Banks are modeled as being able to raise funds from and lend funds to short-term money markets as well as being able to raise deposits from and make loans to nonfinancial agents.

In borrowing or lending in money markets banks face a market rate, which can vary over time, while in raising deposits from or extending loans to nonfinancial agents banks have the ability to set rates.  In raising deposits, banks can pay a below market rate that is fixed for the full life of the deposit and in extending loans they can earn an above market rate that is also fixed for the full maturity of the loan.  Importantly, the life of the deposit is the same as that of the loan, reflecting the fact that while core deposits can, in principal, be withdrawn on demand from a bank they are in practice extremely sticky.  Interest rate risk arises in this model from the fact that deposit-taking and lending opportunities do not arrive at the bank at precisely the same time but rather arrive at any time with some probability, where these probabilities---which are different for deposits and lending opportunities---depend on the interest rates set by the bank on deposits and loans.  Consequently, banks can find themselves holding deposits without loans, that they must then invest at variable market rates (and that could adversely fall below the deposit rate), or loans without deposits, that they must then fund at variable market rates (that could adverse rise above the lending rate).

The solution to the bank's profit-maximization problem is the spread between the loan rate and the deposit rate and this has two parts. The first part is the spread that a bank would set if it were completely risk neutral. This depends only on the degree of competition faced by the bank in the loan markets and the deposit markets in which it operates, where, in the model, this is captured by the slope of the loan demand schedule and slope of the deposit supply schedule facing the bank. The second part of the spread reflects banks' risk aversion to the situation of either having deposit supply without corresponding loan demand or having loan demand without the corresponding deposit supply. Consequently, the second part of the spread is the compensation for this risk that banks require in order for them to be willing to take deposits and lend. This depends positively on the banks' degree of risk aversion as well as the volatility of interest rates.

In Ho and Saunders the degree of competition faced by the bank in loan and deposit markets does not enter their estimated model, since they argue that this variable is quite slow moving and their analysis is undertaken only over a period of three years.  Saunders and Schumacher do consider the degree of competition in their analysis--which is cross-country in nature---although this takes the form of a country dummy variable in the regressions. In all cases the banks' degree of risk aversion does not enter the specification for estimated models of NIMs since this is difficult to measure, although the volatility of interest rates does and this is measured by the annual or quarterly standard deviation of weekly Treasury rates of different maturities (\textit{e.g.}, 3-month, 1-year, 2-year, 3-year, and 5-year rates).

Most papers in this literature also add a few more controls to the model before estimation, although in all cases this is done in a fairly \textit{ad hoc} way.  These variables include banks' payments of implicit (instead of, explicit) interest on deposits;\renewcommand{\baselinestretch}{1}\footnote{This takes place through service charge remissions and other types of depositor subsidies due to regulatory restrictions of explicit interest payments and is measured by the difference between noninterest expenses and noninterest revenue, all divided by interest earning assets.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} banks' non-interest-bearing reserve requirements;\renewcommand{\baselinestretch}{1}\footnote{This captures the influence of the economic cost of funds over and above the published interest expense and is measured by non-interest bearing reserves divided by total assets.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} banks' default risk, since this should imply an additional premium charged on the rate of the banks loans given the potential loss in bank capital;\renewcommand{\baselinestretch}{1}\footnote{This is measured by net chargeoffs divided by total loans.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} and, bank capital, since capital is costly and banks with higher capital may seek to cover this with higher lending rates.  Note that all of the empirical analysis in the ``dealership'' model literature for studying NIMs is undertaken at the firm-level and as such many of the \textit{ad hoc} controls that are added to the model are included so as to address cross-firm differences.

We would note that Angbazo does add default risk to the bank's profit-maximization problem and, as such, default risk does enter his model of NIMs in a more rigorous way.  Here he finds that for money center banks net interest margins are more sensitive to default risk than they are to interest-rate risk, while super-regional and regional banks are sensistive to interest-rate risk but less so to default risk. These findings he notes are consistent with these money center banks' greater concentration of short-term assets and greater use of off-balance sheet (OBS) hedging instruments.

There is a further strand of the micro-banking literature looking at determinants of banks' NIMs---which includes papers by Zarruk~(1989) and Wong~(1997)---that we mention but do not discuss at length.  In these models deposits and lending opportunities arrive at the same time and depend on the deposit and loan rates being set by banks but there is some stochastic element to the volume of deposits that banks will raise given the interest rate they set, thus requiring the banks in the model, who are risk average, to borrow or lend in the federal funds market.  These models also include features like administrative costs on deposits and loans as well as costs associated with deposit insurance premia and the primary use of these models is for comparative static exercises that examine how changes in these parameters---as well as considerations like uncertainty---will affect bank spreads.  Utlimately, however, these papers are not empirical and since the majority of their theoretical findings confirm those of those of papers like Ho and Saunders, Angbazo, and Saunders and Schumacher, we do not delve too deeply into them here.

In summing up our review of other empirical analyses of NIMs a few variables stand out---in addition, to short- and long-term interest rates---as relevant for our analysis.  The degree of competition faced by the bank in the loan markets and the deposit markets in which it operates stands out as an important consideration, especially given that the sample we use in our analysis is much longer than the samples considered by Ho and Saunders, Angbazo, and Saunders and Schumacher.  Interest rate volatility, also appears important, since this has likely also changed over our sample.  The other variables that previous authors in the micro-banking tradition have added to their models in an \textit{ad hoc} way we hold-off for the time being from adding to our analysis.  This is largely because our analysis at this stage is at the aggregate level and these additional variables appear to have been added to explain variations in NIMs across firms.

\section{Models for forecasting NIMs}

\vspace{-0.1in}
\noindent \textbf{1. Simple forecast combination.} We regress NIMs on each yield $r(\tau )_{t-1}$ separately: \
\[
NIM_{t}=c_{\tau }+\rho _{\tau }NIM_{t-1}+\gamma _{\tau }r(\tau )_{t-1}+{\eta
_{\tau ,t}}
\]

Each regression is use to produce an $s$ step-ahead forecast of NIMs conditional on yields on treasuries with maturity $\tau $ observed through period $t+s-1.$ This conditional forecast is denoted by $NIM_{\tau ,t+s/t}.$ The simple forecast comnination is then given by:
\[
NIM_{t+s/t}=\sum_{\tau }\frac{NIM_{\tau ,t+s/t}}{N},
\]
where $N$ is the number of maturities considered.

\noindent \textbf{2. A dynamic factor model for NIMs.} Using the state space representation, the state equations of our restricted dynamic factor model is given by
\[
\left(
\begin{array}{c}
{L_{t+1}-\mu _{L}} \\
{S_{t+1}-\mu _{S}} \\
{C_{t+1}-\mu _{C}} \\
\widehat{NIM}_{t+1}-\mu _{\widehat{NIM}}%
\end{array}%
\right) {\normalsize \ =\ }\left(
\begin{array}{cccc}
a_{11} & a_{12} & a_{13} & 0 \\
a_{21} & a_{22} & a_{23} & 0 \\
a_{31} & a_{32} & a_{33} & 0 \\
a_{41} & a_{42} & a_{43} & a_{44}%
\end{array}%
\right) {\normalsize \ }\left(
\begin{array}{c}
{L_{t}-\mu _{L}} \\
{S_{t}-\mu _{S}} \\
{C_{t}-\mu _{C}} \\
\widehat{NIM}_{t}-\mu _{\widehat{NIM}}%
\end{array}%
\right) {\normalsize +}\left(
\begin{array}{c}
{\eta _{Lt}} \\
{\eta _{St}} \\
{\eta _{Ct}} \\
{\eta _{NIMt}}%
\end{array}%
{\normalsize \ }\right) ,
\]
where $L_{t},S_{t},$ and $C_{t}$ are level slope and curvature factors for the yield curve, respectively, and $\widehat{NIM}_{t}$ represent a measure of NIMs cleaned of observation error. The term $\mu _{x}$ denotes the mean of $x$. The terms ${\eta _{Lt}}$, ${\eta _{St}}$, ${\eta _{Ct}}$, ${\eta_{NIMt}}$ denote independently distributed innovations with joint covariance matrix $\Omega $.

The observation equations, which relate a set of $N$ yields (in our case, 12) and observed NIMs can be split into two groups. In the first group, equations for yields on treasuries of maturity $\tau $ take the form:
\[
r(\tau )=\left(
\begin{array}{cccc}
1 & \frac{1-e^{-\lambda \tau }}{\lambda \tau } & \frac{1-e^{-\lambda \tau }}{%
\lambda \tau }-e^{-\lambda \tau } & 0%
\end{array}%
\right) \left(
\begin{array}{c}
{L_{t}-\mu _{L}} \\
{S_{t}-\mu _{S}} \\
{C_{t}-\mu _{C}} \\
\widehat{NIM}_{t}-\mu _{\widehat{NIM}}%
\end{array}%
\right) +e(\tau )_{t},
\]%
where $e(\tau )_{t}$ is i.i.d. measurement error assumed to be uncorrelated
across observations. \bigskip In the second group, the equation for observed
NIMs takes the form%
\[
NIM_{t}=\left(
\begin{array}{cccc}
0 & 0 & 0 & 1%
\end{array}%
\right) \left(
\begin{array}{c}
{L_{t}-\mu _{L}} \\
{S_{t}-\mu _{S}} \\
{C_{t}-\mu _{C}} \\
\widehat{NIM}_{t}-\mu _{\widehat{NIM}}%
\end{array}%
\right) +e(NIM)_{t},
\]
where $e(NIM)_{t}$ is i.i.d. measurement error assumed to be uncorrelated across observations.

Out-of-sample conditional forecasts for NIMs from this model, are obtained using the Kalman filter.

\noindent \textbf{3. Dynamic factor model with second-step regression for NIMs.} The dynamic factor model we consider takes the form
\[
\left(
\begin{array}{c}
{L_{t+1}-\mu _{L}} \\
{S_{t+1}-\mu _{S}} \\
{C_{t+1}-\mu _{C}}%
\end{array}%
\right) {\normalsize \ =\ }\left(
\begin{array}{ccc}
a_{11} & a_{12} & a_{13} \\
a_{21} & a_{22} & a_{23} \\
a_{31} & a_{32} & a_{33}%
\end{array}%
\right) {\normalsize \ }\left(
\begin{array}{c}
{L_{t}-\mu _{L}} \\
{S_{t}-\mu _{S}} \\
{C_{t}-\mu _{C}}%
\end{array}%
\right) {\normalsize +}\left(
\begin{array}{c}
{\eta _{Lt}} \\
{\eta _{St}} \\
{\eta _{Ct}}%
\end{array}%
\right)
\]

\[
r(\tau )=\left(
\begin{array}{ccc}
1 & \frac{1-e^{-\lambda \tau }}{\lambda \tau } & \frac{1-e^{-\lambda \tau }}{%
\lambda \tau }-e^{-\lambda \tau }%
\end{array}%
\right) \left(
\begin{array}{c}
{L_{t}-\mu _{L}} \\
{S_{t}-\mu _{S}} \\
{C_{t}-\mu _{C}} \\
\widehat{NIM}_{t}-\mu _{\widehat{NIM}}%
\end{array}%
\right) +e(\tau )_{t},
\]

We obtain smoothed estimates of the factors $L_{t},S_{t},$ and $C_{t}$ conditional on information up until period $T.$ We estimate two alternative second-step forecasting equations:

\noindent \textbf{Alternative a. Multivariate} (all yield curve factors included in one NIM equation)
\[
NIM_{t}=c+\rho NIM_{t-1}+\gamma _{L}L_{t-1}+\gamma _{S}S_{t-1}+\gamma
_{C}C_{t-1}
\]

\noindent \textbf{Alternative b. Forecast combination} (each factor included in a separate equation and forecasts from each equation combined)
\[
NIM_{t}=c_{i}+\rho _{i}NIM_{t-1}+\gamma _{i}F_{i,t-1}+{\eta _{i,t},}
\]%
where $F_{i}\in \left\{ L,S,C\right\} .$ Forecasts $NIM_{i,t+s/t}$ from each separate regression are then aggregated as:

\[
NIM_{t+s/t}=\sum_{i}\frac{NIM_{i,t+s/t}}{3},
\]

\noindent \textbf{4. Forecast combination with observed factors.} Define $\left\{L,S,C \right\} ,$ the observed factors as in subsection 3.1 above.
\[
NIM_{t}=c_{i}+\rho _{i}NIM_{t-1}+\gamma _{i}O_{i,t-1}+{\eta _{i,t},}
\]%
where $O_{i}\in \left\{ L,S,C\right\} .$ Forecasts  from each separate regression, $NIM_{i,t+s/t},$are then aggregated as:

\noindent \textbf{5. VAR with observed factors.} Define level factor to coincide with the 10-year yield; define the slope factor as the difference between the 10-year and the 3-month yields; define the curvatures factor to be 2 times the 2-year yield minus the the sum of the 3-month and 10-year yields. The VAR of order four includes the three observed factors defined as above and NIMs. Forecasts conditional on the factors are obtained using the Kalman filter.
\[
NIM_{t+s/t}=\sum_{i}\frac{NIM_{i,t+s/t}}{3},
\]

\noindent \textbf{6. No-change forecast.} According to this model $NIM_{t+s/t}=NIM_{t}${\normalsize . }

%\renewcommand{\baselinestretch}{1}\footnote{...\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5}

\section{Data}

\vspace{-0.1in}

\subsection{Net interest margins}

\vspace{-0.1in} The data on bank net interest margins (NIMs) that we use is from the quarterly Consolidated Reports of Condition and Income (Call Report) that every national, state member, and insured nonmember bank is required to file on the last day of each quarter by the Federal Financial Institutions Examination Council (FFIEC). The Federal Deposit Insurance Corporation is tasked as the overseer, collecting and reviewing all submissions. Call Report data used in this analysis are cleaned and adjusted for bank mergers and acquisitions, using structure data from the National Information Clearinghouse (NIC) on mergers and acquisitions. Foreign entities are excluded and domestic subsidiaries are aggregated up to the parent, bank-holding-company (BHC), level.\renewcommand{\baselinestretch}{1}\footnote{NIM data are adjusted for mergers between commercial banks by comparing balance sheet values of interest income, interest expenses, and interest-earning assets at the end of the quarter with those at the beginning of the quarter, accounting for amounts acquired or lost during the period because of mergers. For information on the merger-adjustment procedure for income, see the appendix in English and Nelson (1998).\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5}

NIMs in our models are expressed as the annualized percentage of net interest income over interest-earning assets. To get an aggregate banking sector measure of NIMs, we aggregate NIMs for the top 25 BHCs, as ranked by total assets, which is assessed quarterly.

\subsection{Treasury yields}

The yields data are derived using a smoothing technique from Gurkaynak, et al. (2007), based on Nelson and Siegel (1987) and Svensson (1994), which allows for daily measures of an off-the-run Treasury yield curve. We use yields for twelve maturities in our models: 3 month, 6 month, 9 month, 1 year, 2 year, 3 year, 5 year, 7 year, 10 year, 15 year, 20 year, and 30 year.\renewcommand{\baselinestretch}{1}\footnote{Daily yields are published with a two-day lag under the base mnemonic SVENY at  www.federalreserve.gov/econresdata/researchdata/feds200628\_1.html.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} The yields included are quarterly averages of daily yields.  The data on NIMs and yields start in 1985Q4 and end in 2008Q2, a period that stops short of the interest rate regime shift to the zero lower bound.

\subsection{Micro-banking variables}

\vspace{-0.1in} The main variable that we use to capture the degree of competition that banks face is the size of the traditional
banking industry (as measured by the assets in that industry) relative to the size of the total banking industry, that is, both
traditional and shadow banking (as measured by the combined assets of these these two industries). We use this measure for the
reason that competition from the shadow banking sector is the most relevant form of increased competition faced by the
traditional banking sector.  Importantly, competition from the shadow banking sector affects both sides of traditional banks'
balance sheets. For example, it puts upward pressure on the deposit rate that traditional banks must pay to attract deposits---since the shadow banking sector offers alternatives to deposits like money market mutual funds---and it puts downward pressure on the lending rate that traditional banks can earn---since the shadow banking sector offers alternatives to bank loans like loans from finance companies. This measure can be easily calculated from the U.S. Financial Accounts (previously known as the Flow of Funds Accounts), where the traditional banking sector consists of commercial banks, savings institutions, and credit unions and the shadow banking sector consists of broker-dealers, ABS issuers, finance companies, mortgage pools, and funding corporations.

In papers that have modeled NIMs as a function of more microeconomic variables interest rate volatility has been measured by
the quarterly variance of weekly Treasury interest rate series. (See, for example, Ho and Saunders, 1981, and Saunders
and Schumacher, 2000.)  Typically, also, it is the quarterly variance of several Treasury interest rate series that are
considered---such as, the 3-month rate, 1-year rate, 2-year rate, 3-year rate, and 5-year rate---with the best fitting series then
included in the regression.  We also consider the Merrill Lynch MOVE index, which is the bond market equivalent to the VIX.

% Suggestion to look at the MOVE index.  It looks like it goes back to at least April 1988 (http://www.businessinsider.com/the-move-index-bonds-2010-3) This (http://business.financialpost.com/2012/11/01/investors-seeking-yield-should-use-move-and-vix/) also says it was created in 1988.

\newpage
\begin{thebibliography}{99}
\vspace{-0.1in}
\bibitem{abel82} Abel, A.B. (1982).  ``Dynamic Effects of Permanent and Temporary Tax Policies in
a $q$ Model of Investment.''  \textit{Journal of Monetary Economics}, 9, 353-373.

\bibitem{auerb1} Auerbach, A.J. (1979). ``Wealth Maximization and the Cost of Capital.''
\textit{Quarterly Journal of Economics}, 93, 433-446.

\bibitem{auerb2} Auerbach, A.J. (1989). ``Tax Reform and Adjustment Costs:  The Impact on
Investment and Market Value.''  \textit{International Economic Review}, 30, 939-962.

\bibitem{auerb3} Auerbach, A.J., and L.H. Summers (1979). ``The Investment Tax Credit: An Evaluation.''
NBER Working Paper No.~404.

\end{thebibliography}

\newpage
{\normalsize \clearpage
\begin{sidewaystable}
\center
\caption{RMSE for shortened sample: 1989Q2-2008Q2}
\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|}
\hline
&Step 1 &Step 2 &Step 3 &Step 4 &Step 5 &Step 6 &Step 7 &Step 8 &Step 9 &Step 10\\
\hline
1. Simple Forecast Combination             &0.107&0.133&0.148&0.168&0.187&0.202&0.213&0.209&0.216&0.223\\
2. DFM                                     &0.122&0.148&0.169&0.212&0.247&0.278&0.316&0.340&0.373&0.404\\
3a. DFM with 2nd Step Regression           &0.118&0.167&0.209&0.260&0.308&0.356&0.398&0.433&0.474&0.511\\
3b. DFM with Forecast Combination          &0.112&0.159&0.199&0.247&0.300&0.351&0.399&0.438&0.489&0.535\\
4. Forecast Combination of Observed Factors&0.118&0.169&0.216&0.273&0.336&0.399&0.456&0.509&0.567&0.622\\
5. VAR on Observed Factors                 &0.166&0.209&0.263&0.302&0.342&0.375&0.380&0.430&0.451&0.499\\
6. No-Change Forecast                      &0.107&0.138&0.154&0.181&0.206&0.228&0.244&0.248&0.271&0.293\\
\hline
\end{tabular}
\end{sidewaystable}
}

\newpage
{\normalsize \clearpage
\begin{sidewaystable}
\caption{RMSEs over full sample: 1985Q4-2008Q2}
\center
\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|}
\hline
&Step 1 &Step 2 &Step 3 &Step 4 &Step 5 &Step 6 &Step 7 &Step 8 &Step 9 &Step 10\\
\hline
1. Simple Forecast Combination             &0.105&0.140&0.164&0.191&0.218&0.244&0.265&0.281&0.305&0.327\\
3a. DFM with 2nd Step Regression           &0.128&0.190&0.244&0.299&0.350&0.397&0.438&0.475&0.514&0.549\\
3b. DFM with Forecast Combination          &0.112&0.158&0.197&0.240&0.290&0.337&0.373&0.407&0.446&0.480\\
4. Forecast Combination of Observed Factors&0.116&0.174&0.227&0.284&0.347&0.409&0.459&0.508&0.554&0.595\\
5. VAR on Observed Factors                 &0.148&0.196&0.243&0.275&0.283&0.296&0.319&0.377&0.456&0.530\\
6. No-Change Forecast                      &0.107&0.138&0.154&0.181&0.206&0.228&0.244&0.248&0.271&0.293\\
\hline
\end{tabular}
\end{sidewaystable}
}

\newpage
{\normalsize \clearpage
\begin{sidewaystable}
\caption{RMSEs for models that include the share of asset of the shadow banking sector. Estimation sample: 1989Q2-2008Q2}
\center
\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|}
\hline
&Step 1 &Step 2 &Step 3 &Step 4 &Step 5 &Step 6 &Step 7 &Step 8 &Step 9 &Step 10\\
\hline
1. Simple Forecast Combination             &0.110&0.143&0.165&0.192&0.222&0.248&0.267&0.274&0.287&0.300\\
3a. DFM with 2nd Step Regression           &0.121&0.170&0.212&0.262&0.312&0.360&0.403&0.436&0.478&0.515\\
3b. DFM with Forecast Combination          &0.106&0.140&0.165&0.193&0.222&0.248&0.267&0.274&0.287&0.300\\
4. Forecast Combination of Observed Factors&0.109&0.141&0.165&0.193&0.222&0.249&0.267&0.274&0.288&0.300\\
5. VAR on Observed Factors                 &0.344&0.396&0.476&0.494&0.521&0.569&0.631&0.670&0.699&0.718\\
6. No-Change Forecast                      &0.107&0.138&0.154&0.181&0.206&0.228&0.244&0.248&0.271&0.293\\
\hline
\end{tabular}
\end{sidewaystable}
}

\newpage \clearpage
\begin{table}
\caption{RMSEs for variants of the simple forecast combination model}
\center
\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|}
\hline
&Step 1 &Step 2 &Step 3 &Step 4 &Step 5 &Step 6 &Step 7 &Step 8 &Step 9 &Step 10\\
\hline
All yields              &0.107&0.133&0.148&0.168&0.187&0.202&0.213&0.209&0.216&0.223\\
Observed Factors        &0.118&0.169&0.216&0.273&0.336&0.399&0.456&0.509&0.567&0.622\\
Smoothed Factors        &0.112&0.159&0.199&0.247&0.300&0.351&0.399&0.438&0.489&0.535\\
3-month, 2-year, 10-year&0.105&0.130&0.141&0.160&0.178&0.197&0.202&0.198&0.209&0.217\\
3-month, 10-year        &0.105&0.129&0.144&0.159&0.179&0.196&0.200&0.200&0.209&0.217\\
\hline
\end{tabular}
\end{table}

\newpage \clearpage
\begin{figure}
%[tbp]
\caption{NIMs and Treasury Yields} \label{figure_nims_rates}
\center
\includegraphics[scale=0.85]{figure_nims_rates.ps}
\end{figure}

\newpage \clearpage
\begin{figure}
%[tbp]
\caption{NIMs and Observed Yield Factors} \label{figure_nims_factors}
\center
\includegraphics[scale=0.85]{figure_nims_factors.ps}
\end{figure}

\newpage \clearpage
\begin{figure}
%[tbp]
\caption{NIMs and Measures of Competition} \label{figure_nims_competition}
\center
\includegraphics[scale=0.85]{figure_nims_competition.ps}
\end{figure}

\newpage \clearpage
\begin{figure}
%[tbp]
\caption{Interest Income, Interest Expenses and Short-term Treasury Rates} \label{figure_nims_components}
\center
\includegraphics[scale=0.85]{figure_nims_components.ps}
\end{figure}

\end{document}

### 2013-07-08 MATLAB/nim_lsc_model_2steps/run_scenario.m
Hook up forecast combination model to CCAR scenario

@@ -1,0 +1,217 @@
+
+clear;
+
+setpath;
+
+load_paramvec_from_disk = 1;  % set to 1 to load saved param_vec from disk
+% set to 0 to use first guess as defined below
+
+ntrain = 15; %We have 153 observations per time t; we're setting ntrain to ~15% of observation pool.
+
+dataset_option = 1;
+
+[yobs, dates, yields, nims, nfactors, nothers, tau, factors,...
+    shadow_bank_share_assets,...
+    total_interest_earning_assets,...
+    assets_depository_inst, assets_securities_notrade,...
+    assets_fedfunds, assets_all_loans, assets_trading_accnts, ... 
+    interest_income_to_ie_assets, interest_expense_to_ie_assets] = load_data_ml(dataset_option);
+ 
+out_of_sample_start_pos = find(dates==2000.0);
+end_sample_pos = length(dates);
+
+nstates = nfactors + nothers;
+ntaus = length(tau);
+
+load this_iter; %includes the full sample param_vec, xi10_demeaned, p10, ntrain
+
+param_vec_full_sample = param_vec;
+opt_param_mat = zeros(length(param_vec), end_sample_pos-out_of_sample_start_pos+1);
+options = optimset('HessUpdate','bfgs','display','iter','MaxIter',10000,'MaxFunEvals',1e10,'TolFun',10e-3,'TolX',1e-4);
+
+compute_param_mat = 0;
+
+if compute_param_mat
+for pos_index = out_of_sample_start_pos:end_sample_pos
+    [yobs, dates, yields, nims, nfactors, nothers, tau] = load_data_ml;
+    
+    display(['working on observation for date ', num2str(dates(pos_index))]);
+    
+    yobs = yobs(:,1:pos_index-1);
+    dates = dates(:,1:pos_index-1);
+    yields = yields(:,1:pos_index-1);
+    nims = nims(:,1:pos_index-1);
+    
+    
+    for fminsearch_rep = 1:1
+        loss = @(x0)mylikelihood(x0,yobs,xi10_demeaned,tau,nfactors,nothers,ntrain,p10);
+        option = anneal();
+        option.Verbosity = 2;
+        [opt_param_mat(:,pos_index-(out_of_sample_start_pos-1)), fval] = anneal(loss,param_vec_full_sample);
[…]

### 2013-07-08 MATLAB/nim_lsc_model_2steps/utilities_2steps/make_forecast_combination.m

@@ -1,0 +1,31 @@
+function forecast_mat = make_forecast_combination(this_yobs, this_other_obs, previous_forecast, other_obs_out_of_sample,firstlag,lastlag)
+
+
+    this_forecast_horizon = size(other_obs_out_of_sample,2)-lastlag;
+    n_other_obs = size(this_other_obs,1);
+    forecast = zeros(n_other_obs,this_forecast_horizon);
+    
+    for this_macro_factor = 1:n_other_obs
+        
+        yreg = this_yobs(1+max(lastlag,1):end);
+        xreg = [];
+        for this_lag=firstlag:lastlag
+            xreg = [xreg; this_other_obs(this_macro_factor,1+max(lastlag,1)-this_lag:end-this_lag)];
+        end
+        xreg = [ones(1,length(yreg)); this_yobs(1+lastlag-firstlag:end-max(firstlag,1)); xreg];
+        
+        ols_coef = estimate_ols(yreg,xreg);
+        
+  
+        for this_step = 1:this_forecast_horizon
+            % figure out dimensions
+            pos_vec =  (lastlag:-1:firstlag)-firstlag+this_step;
+            forecast(this_macro_factor,this_step) = [1, previous_forecast, other_obs_out_of_sample(this_macro_factor,pos_vec) ]*...
+                ols_coef;
+            previous_forecast = forecast(this_macro_factor,this_step);
+        end
+        
+    end
+    
+
+forecast_mat = mean(forecast);

### 2013-07-09 LATEX/beg.draft.2013.tex
updated draft

@@
 \author{{\Large Valentin Bolotnyy} \\
-Federal Reserve Board\thanks{E-mail: valentin.bolotnyy@frb.gov.\vspace{0.05in}} \and
+Federal Reserve Board \and
 {\Large Rochelle M. Edge} \\
-Federal Reserve Board\thanks{E-mail: rochelle.m.edge@frb.gov.\vspace{0.05in}}\and
-{\Large Luca Guerrieri} \\
-\addtocounter{footnote}{5} Federal Reserve Board\thanks{%
-E-mail: luca.guerrieri@frb.gov.}}
-\title{\textbf{Stressing Bank Profitability for Interest Rate Risk}}
+Federal Reserve Board \and
+{\Large Luca Guerrieri}\\ Federal Reserve Board} 
+\title{Stressing Bank Profitability for Interest Rate Risk$^*$}
+
 \date{\today}
@@
 \begin{abstract}
-\vspace{-0.06in} \noindent  Regular bank stress testing (and the capital planning that it implies) is one of the three complementary reforms that has been made to U.S.
-capital regulation in response to the shortcomings of the pre-crisis regime.
- The form of stress testing mandated under the Dodd-Frank Act for the Federal Reserve to undertake and the form being used by the Federal Reserve for its annual Comprehensive Capital Analysis and Review (CCAR) is one based around a small number of scenarios.
-This approach involves specifying macroeconomic and financial scenarios that represents stressful conditions over the time horizon of the stress test and then projecting, for each BHC in the stress test, its losses, income, and path of capital \textit{given that scenario}.
-In this paper, we focus on the revenue generate aspect of banks and consider the ability of empirical models to translate developments in interest rates into bank net interest margins (NIMs).
-We consider an extensive number of time-series models.
-While our best-performing models beat a no-change forecast, the confidence bands are so large that baseline and stress scenarios taken from CCAR appear indistinguishable.
+ Regular bank stress testing is one of the three complementary reforms that has been made to U.S. capital regulation in response to the shortcomings of the pre-crisis regime.
+ The form of stress testing mandated by the Dodd-Frank Act for the Federal Reserve to undertake is one based around a small number of scenarios.
+This approach involves specifying macroeconomic and financial scenarios that represents stressful conditions over the time horizon of the stress test and then projecting, for each bank holding company in the stress test, its losses, income, and path of capital \textit{given that scenario}.
+In this paper, we focus on the path of income.
+In particular, we consider the ability of empirical models to translate developments in interest rates into bank net interest margins (NIMs).
 
@@
 
-\noindent Key Words:
+\noindent Key Words: Net Interest Margins, Interest Rate Risk, Forecasting, Stress Tests.
 
@@
 
-\noindent JEL Classification: To be completed.
+\noindent JEL Classification: C53, E47, G21
+
+\vspace{2cm}
+\noindent $^*$ The views
+expressed in this paper are solely the responsibility of the
+authors and should not be interpreted as reflecting the views of
+the Board of Governors of the Federal Reserve System or of any
+other person associated with the Federal Reserve System.
+Contact Information: valentin.bolotnyy@frb.gov; rochelle.m.edge@frb.gov; luca.guerrieri@frb.gov.
 
@@
 Regular bank stress testing (and the capital planning that it implies) is one of the three complementary reforms that has been made to U.S.
-capital regulation in response to the shortcomings of the pre-crisis regime.\renewcommand{\baselinestretch}{1}\footnote{See Governor Tarullo’s November 2011 speech ``The Evolution of Capital Regulation'' for this characterization of stress testing.
+capital regulation in response to the shortcomings of the pre-crisis regime.\renewcommand{\baselinestretch}{1}\footnote{See Governor Tarullo's November 2011 speech ``The Evolution of Capital Regulation'' for this characterization of stress testing.
 The two other reforms that he mentions are improvements to the traditional, individual firm approach to capital regulation and the introduction of a macroprudential component to capital regulation.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} While stress testing can take many forms---for example, being based on a small number of scenarios, being based on a large number of scenarios, or even being undertaken in reverse form to achieve a certain adverse outcome---the form of the stress tests mandated under the Dodd-Frank Act for the Federal Reserve to undertake and the form being used by the Federal Reserve for its annual Comprehensive Capital Analysis and Review (CCAR) is one based around a small number of scenarios.
-This approach involves specifying a small number of macroeconomic and financial scenarios that represents stressful conditions over the time horizon of the stress test and then projecting, for each BHC in the stress test, its losses, income, and path of capital \textit{given that scenario}.
-Clearly, there are at least two important parts to conducting bank stress tests (i)~formulating macroeconomic and financial scenarios that should be stressful to banks and (ii)~developing econometric models that are able to translate such stressful conditions into losses, income, and capital.
-%This paper considers the latter of these two issues; that is, translating
-%stressful economic conditions into bank income.
+This approach involves specifying a small number of macroeconomic and financial scenarios that represent stressful conditions over the time horizon of the stress test and then projecting, for each bank holding company (BHC) in the stress test, its losses, income, and path of capital \textit{given that scenario}.
+Accordingly, there are at least two important parts to conducting bank stress tests (i)~formulating macroeconomic and financial scenarios that should be stressful to banks and (ii)~developing econometric models that are able to translate such stressful conditions into losses, income, and capital.
 
-In developing econometric models to translate stressful economic conditions into key bank financial-statement variables it is important to consider what types of developments are likely to be stressful to banks and what particular economic variables are likely to capture these stresses.
-It is then critical that the models being used for stress testing are able to translate developments in these economic variables into the key bank financial-statement variables that the model is designed to project.
+We examine how well a range of time-series models can explain NIMs.
+We emphasize the models' pseudo-out-of-sample \textit{conditional} forecast performance.
+We adopt this pseudo-out-of-sample approach for the simple reason that the evaluation of models based on their in-sample forecast performance, may lead to in-sample over-fitting and poor out-of-sample performance.
+An explicit focus on out-of-sample performance makes forecast uncertainty explicit.
+Moreover, concentrating on conditional forecasts, allows us to separate the task of modeling and forecasting NIMs from that of forecasting interest rates, which is itself a sizable topic (see Duffee, 2012).
+
+When developing econometric models to translate stressful economic conditions into key bank financial-statement variables, it is important to consider what types of developments are likely to be stressful to banks and what particular economic variables are likely to capture these stresses.
+It is then critical that the models being used for stress testing be able to translate developments in these economic variables into conditional forecasts for key bank financial-statement variables.
 In this paper, we examine this property for models of bank net interest margins (NIMs), defined as the ratio of net interest income (NII) to interest earning assets, where NII is the difference between a bank's interest income and interest expenses.
 
-Our motivation for focusing on models of NIMs is based on a number of considerations.
-First, bank losses that arise from banks' having to pay out more in interest on their liabilities than they are earning from their assets should not be overlooked as an important source of risk to the banking and broader financial sector.
+In determining which variables to include in our model of NIMs we look to the variables that have been emphasized by two existing strands of the literature on NIMs.
+The first strand emphasizes the link between interest rates of different maturities and NIMs and is more in the macro-banking tradition.
+The second literature focuses on the optimal margin set by banks based on considerations like the competitive structure of the lending and deposit markets in which the bank participates, the degree of managerial risk aversion, and the degree of uncertainty associated with the future path of interest rates and is much more in the micro-banking tradition.
+
+Our motivation for focusing on models of NIMs is fourfold.
+First, losses from interest rate risk have been a central feature of banking crises.
+Second, net interest rate margins capture income from the maturity transformation services at the center of the traditional banking model.
+This source of income has been lackluster given the current configuration of interest rates.
+Third, modelling of NIMs has received scant attention so far.
+Forth, the ability of producing reliable conditional forecasts for a key component of net income is a the core of the scenario-based stress testing of banks.
+
+Bank losses that arise from banks' having to pay out more in interest on their liabilities than they are earning from their assets should not be overlooked as an important source of risk to the banking and broader financial sector.
 To be sure, losses that arise from this source of risk---called, interest rate risk---were much less prominent in the most recent crisis relative to bank losses that stem from loan defaults---that is, credit risk---but there are ample examples of other earlier banking crises in which adverse developments in NIMs have played a significant role.
@@
 This resulted in interest expenses rising above interest income, NII and NIMs turning negative (in the thrift sector), and sizable losses and ultimately a substantial number of bank failures.
+
 Likewise, the Nordic banking crises of the late 1980s and early 1990s (for Finland, Sweden, and Norway)---which represent three of Reinhart's and Rogoff's~(2009) ``big five'' banking crises---were associated with elevated short-term interest rates that depressed NII and NIMs and lead to sizable bank losses and failures.
 In the Nordic crises, high short-term interest rates resulted from these countries attempting to maintain currency pegs with the Deutschemark in the face of the tight German monetary policy that accompanied its loose fiscal policy associated with re-unification spending.
-Finally, there is also the U.K.'s secondary banking crisis of the early 1970s, which was associated with a sudden increase in short-term interest rates that abruptly drove up banks' costs of funds so too leading to bank losses and failures.
+Depressed NIMs were also at the center of the U.K.'s secondary banking crisis of the early 1970s.
+A sudden increase in short-term interest rates abruptly drove up banks' costs of funds,  leading to extensive bank losses and failures.
 To be sure, large increases in short-term rates and declines in net interest income were not the \textit{only} reason for any of these crises---indeed, there was at least one major class of asset prices that collapsed in each of these crises---but short-term interest rates---and, in particular, their elevated levels relative to long-term interest rates---were significant and in most cases equivalent in importance to the other causes of the crisis.
@@
 
-Our final motivation for examining NIMs is that despite the importance of NIMs for the financial condition of banks \textbf{[and the attention that it receives from bank equity analysts]}, the modeling of NIMs has received much less attention in the risk-modeling literature relative to the modeling of bank variables related to credit risk.
+Despite the importance of NIMs for the financial condition of banks and despite the attention that it receives from bank equity analysts, the modeling of NIMs has received much less attention in the risk-management literature relative to the modeling of bank variables related to credit risk.
 To be sure, the recent increase in the importance of stress testing has led to the development of a number of models that link NIMs to macroeconomic variables---such as, short-term Treasury rates and the slope of the Treasury yield curve---with Kovner \textit{et al}.~(2011) and Covas \textit{et al}.~(2012) being two notable examples for the U.S.
@@
 
-In summing up our motivation on the importance of modeling NIMs and studying interest rate risk, we would also note that the Federal Reserve chose to feature interest rate risk in one of the scenarios specified for the 2013 Dodd-Frank Act (DFA) stress tests.
+Scenario based stress tests are predicated on the ability to produce reliable conditional forecasts for key measures of profitability such as net interest margins.
+This kind of stress test program is not limited to the U.S.
+banking system.
+Stress testing of banks based on scenarios was undertaken during the Scandinavian banking crisis to shore up confidence in the ability of banks to withstand the crisis.
+Most extensively, the European Banking Authority conducted stress tests based on scenario analysis throughout the European Union in 2009, 2010, and 2011.
+The results of such programs can only be effective in restoring confidence from the public in the midst of a crisis to the extent that the public can be convinced of the ability of regulators to assess the consequences of their scenarios on banks.
+If forecast methods at the frontier of efficiency have disentangling the implications of baseline and severe stress scenarios, the ability of stress test programs to avoid panic is called in question.
+
+%In summing up our motivation on the importance of modeling NIMs and studying interest rate risk, we would also note that the Federal Reserve chose to feature interest rate risk in one of the scenarios specified for the 2013 %Dodd-Frank Act (DFA) stress tests.
 In particular, for the 2013 DFA stress tests, the Federal Reserve specified the adverse scenario as one in which the Treasury yield curve shifted up and flattened.
-As noted by the Federal Reserve in its (proposed) Policy Statement on scenario design, the Board varies ``the approach it uses for the adverse scenario each year so that the results of the scenario provide the most value to supervisors, in light of current condition of the economy and the financial services industry.''
-
-In this paper we examine how well a range of time-series models can explain NIMs.
-We emphasize the models' pseudo-out-of-sample \textit{conditional} forecast performance.
-We adopt this pseudo-out-of-sample approach for the simple reason that the evaluation of models based on their in-sample forecast performance, frequently leads to the econometrician over-fitting the model within the sample so as to maximize in-sample forecast performance.
-We focus on conditional forecasts for the reason that scenario-based stress testing of the banking system is the underlying motivation for our analysis and it is along the dimensions of \textit{conditional} forecasts that model forecast performance is important.
-Focusing on conditional forecast performance also means that we can separate the task of modeling and forecasting NIMs from that of forecasting interest rates, which is itself a sizable topic (see Duffee, 2012).
-
-In determining which variables to include in our model of NIMs we look to the variables that have been emphasized by two existing literatures on NIMs.
-The first literature emphasizes the link between interest rates of different maturities and NIMs and is more in the macro-banking tradition.
-The second literature focuses on the optimal margin set by banks based on considerations like the competitive structure of the lending and deposit markets in which the bank participates, the degree of managerial risk aversion, and the degree of uncertainty associated with the future path of interest rates and is much more in the micro-banking tradition.
+As noted by the Federal %Reserve in its (proposed) Policy Statement on scenario design, the Board varies ``the approach it uses for the adverse scenario each year so that the results of the scenario provide the most value to supervisors, in light of %current condition of the economy and the financial services industry.''
 
@@
 \vspace{-0.1in}
-There are primarily two empirical literatures on modeling NIMs.
-The first literature emphasizes the link between interest rates of different maturities and NIMs and lies more in the macro-banking tradition.
-The second literature focuses on the optimal margin set by banks and lies more in the micro-banking tradition.
+There are two strands in the literature on modeling NIMs.
+The first strand emphasizes the link between interest rates of different maturities and NIMs and lies more in the macro-banking tradition.
+The second strand focuses on the optimal margin set by banks and lies more in the micro-banking tradition.
 
-The papers for the U.S.
-by Kovner \textit{et al}.
-and Covas \textit{et al}.
-noted earlier fall into the category of papers in the macro-banking tradition, where in both cases these models link NIMs to  a measure of the slope of the yield curve and a measure of short-term market rate.
+Kovner \textit{et al}.
+and Covas \textit{et al.} are in the macro-banking tradition.
+Both link NIMs to the slope of the yield curve and a short-term interest rate.
 Banks' earnings from interest income reflect two key services that banks supply---maturity transformation services (that is, borrowing short and lending long) and deposit transactions services---and the returns to these services are reflected by these two interest rate variables.
@@
 
-In summing up our review of other empirical analyses of NIMs a few variables stand out---in addition, to short- and long-term interest rates---as relevant for our analysis.
+In summing up, in addition, to short- and long-term interest rates, a few variables stand out as relevant for our analysis.
 The degree of competition faced by the bank in the loan markets and the deposit markets in which it operates stands out as an important consideration, especially given that the sample we use in our analysis is much longer than the samples considered by Ho and Saunders, Angbazo, and Saunders and Schumacher.
@@
 
+We construct forecasts from a wide array of methods taken from the forecasting literature.
+We used data with quarterly frequency and consider a forecast horizon up to ten quarters long.
+We are interested in forecasts for NIMs conditional on the variables in scenario $I$, which we denote $NIM_{t+s/I}$.
+Here are the methods considered.
+
 \vspace{-0.1in}
@@
 \[
-NIM_{t}=c_{\tau }+\rho _{\tau }NIM_{t-1}+\gamma _{\tau }r(\tau )_{t-1}+{\eta
+NIM_{t}=c_{\tau }+\rho _{\tau }NIM_{t-1}+\sum_{j=1}^4\gamma _{\tau,j}r(\tau)_{t-j}+{\eta
 _{\tau ,t}}
 \]
-
-Each regression is use to produce an $s$ step-ahead forecast of NIMs conditional on yields on treasuries with maturity $\tau $ observed through period $t+s-1.$ This conditional forecast is denoted by $NIM_{\tau ,t+s/t}.$ The simple forecast comnination is then given by:
+Each regression is used to produce an $s$ step-ahead forecast of NIMs conditional on yields on treasuries r(\tau) in scenario $I$ with maturity $\tau$ observed through period $t+s-1.$ This conditional forecast is denoted by $NIM_{\tau,t+s/I}.$ The simple forecast comnination is then given by:
 \[
-NIM_{t+s/t}=\sum_{\tau }\frac{NIM_{\tau ,t+s/t}}{N},
+NIM_{t+s/I}=\sum_{\tau }\frac{NIM_{\tau ,t+s/I}}{N},
 \]

### 2013-07-09 LATEX/beg.draft.2013.tex
updated draft

@@
 In particular, we consider the ability of empirical models to translate developments in interest rates into bank net interest margins (NIMs).
+\end{abstract}
 
@@
 
-\vspace{2cm}
+\vspace{3cm} \renewcommand{\baselinestretch}{1} \normalsize
 \noindent $^*$ The views
@@
 
-\end{abstract}
-\renewcommand{\baselinestretch}{1.5}
+\renewcommand{\baselinestretch}{1.5} \normalsize
 \thispagestyle{empty}
@@
 The results of such programs can only be effective in restoring confidence from the public in the midst of a crisis to the extent that the public can be convinced of the ability of regulators to assess the consequences of their scenarios on banks.
-If forecast methods at the frontier of efficiency have disentangling the implications of baseline and severe stress scenarios, the ability of stress test programs to avoid panic is called in question.
+If forecast methods at the frontier of efficiency have trouble disentangling the implications of baseline and severe stress scenarios, the ability of stress test programs to avoid panic is called in question.
 
@@
 
-We construct forecasts from a wide array of methods taken from the forecasting literature.
-We used data with quarterly frequency and consider a forecast horizon up to ten quarters long.
+We construct forecasts from a wide array of standard methods taken from the forecasting literature.
+We use data with quarterly frequency and consider a forecast horizon up to ten quarters long.
 We are interested in forecasts for NIMs conditional on the variables in scenario $I$, which we denote $NIM_{t+s/I}$.
-Here are the methods considered.
+Here are the methods considered:
 
 \vspace{-0.1in}
-\noindent \textbf{1. Simple forecast combination.} We regress NIMs on each yield $r(\tau )_{t-1}$ separately: \
+\noindent \textbf{1. Equal-weighted forecast combination.} We regress NIMs on each yield $r(\tau )_{t-1}$ separately: \
 \[
@@
 \]
-Each regression is used to produce an $s$ step-ahead forecast of NIMs conditional on yields on treasuries r(\tau) in scenario $I$ with maturity $\tau$ observed through period $t+s-1.$ This conditional forecast is denoted by $NIM_{\tau,t+s/I}.$ The simple forecast comnination is then given by:
+Each regression is used to produce an $s$ step-ahead forecast of NIMs conditional on yields on treasuries $r(\tau)$ in scenario $I$ with maturity $\tau$ observed through period $t+s-1.$ This conditional forecast is denoted by $NIM_{\tau,t+s/I}.$ The equal-weighted forecast combination is then given by:
 \[
@@
 \]
-where $N$ is the number of maturities considered.
+where $N$ is the number of maturities considered.\footnote{Experimentation with alternative lag structures led to no improvement in the forecast results reported below.}
 
-\noindent \textbf{2. A dynamic factor model for NIMs.} Using the state space representation, the state equations of our restricted dynamic factor model is given by
+\noindent \textbf{2.
+A dynamic factor model for NIMs.} We include NIMs in the dynamic factor model for the yield curve developed by Diebold and Li (2006).
+Using the state space representation, the state equations of our dynamic factor model is given by:
 \[
@@
 \]
-where $e(NIM)_{t}$ is i.i.d. measurement error assumed to be uncorrelated across observations.
+where $e(NIM)_{t}$ is i.i.d.
+measurement error assumed to be uncorrelated across observations.
+Out-of-sample conditional forecasts for NIMs from this model, are obtained using the Kalman filter.
+We focus on the filtered estimates of NIMs given a path for all yields included in the model.
 
-Out-of-sample conditional forecasts for NIMs from this model, are obtained using the Kalman filter.
-
-\noindent \textbf{3. Dynamic factor model with second-step regression for NIMs.} The dynamic factor model we consider takes the form
+\noindent \textbf{3.
+Dynamic factor model with second-step regression for NIMs.} This model extracts level, slope, and curvature factors for the yield curve following Diebold and Li (2006).
+The factors are used to forecast NIMs in a separate second-step regression.
+The dynamic factor model we consider takes the form
 \[
@@
 {S_{t}-\mu _{S}} \\
-{C_{t}-\mu _{C}} \\
-\widehat{NIM}_{t}-\mu _{\widehat{NIM}}%
+{C_{t}-\mu _{C}} 
 \end{array}%
@@
 \]
-
 We obtain smoothed estimates of the factors $L_{t},S_{t},$ and $C_{t}$ conditional on information up until period $T.$ We estimate two alternative second-step forecasting equations:
@@
 \]
+We use this equation to produce recursive forecasts for NIMs conditional on the the smoothed factors.
 
@@
 \]%
-where $F_{i}\in \left\{ L,S,C\right\} .$ Forecasts $NIM_{i,t+s/t}$ from each separate regression are then aggregated as:
-
+where $F_{i}\in \left\{ L,S,C\right\}.$ Recursive forecasts $NIM_{i,t+s/I}$ from each separate regression are then aggregated as:
 \[
-NIM_{t+s/t}=\sum_{i}\frac{NIM_{i,t+s/t}}{3},
+NIM_{t+s/I}=\sum_{i}\frac{NIM_{i,t+s/I}}{3},
 \]
 
-\noindent \textbf{4. Forecast combination with observed factors.} Define $\left\{L,S,C \right\} ,$ the observed factors as in subsection 3.1 above.
+\textbf{4. Principal component regression}
+Let $X$ a $T \times n$ matrix of $n$ Treasury yields.  Using the spectral decomposition, one obtains $X'X = VDV'$. Let $V_i$ be the normalized eigenvector corresponding to $i^{th}$ largest eigenvalue.
+The $i^{th}$ principal component is defined as $PC_i = XV_i$. We estimate
+\begin{equation*}
+NIM_t = c + \rho NIM_{t-1} + \sum_{i=1}^3 \gamma_i PC_{i,t-1} + e_t.
+\end{equation*}
+using ordinary least squares and we construct recursive forecasts for net interest margins conditional on the principal components.
+
+\begin{frame}
+
+\noindent \textbf{5.
+Forecast combination with observed factors.} Define level factor $L$ to coincide with an average of the 3-month, 2-year and 10-year Treasury yields; define the slope factor as the difference between the 10-year and the 3-month yields; define the curvatures factor to be 2 times the 2-year yield minus the the sum of the 3-month and 10-year yields.
+Define $\left\{L,S,C \right\} ,$ the observed factors as in subsection 3.1 above.
 \[
@@
 
-\noindent \textbf{5.
-VAR with observed factors.} Define level factor to coincide with the 10-year yield; define the slope factor as the difference between the 10-year and the 3-month yields; define the curvatures factor to be 2 times the 2-year yield minus the the sum of the 3-month and 10-year yields.
-The VAR of order four includes the three observed factors defined as above and NIMs.
+\noindent \textbf{6.
+VAR with observed factors.}  The VAR of order four includes the three observed factors defined as above and NIMs.
 Forecasts conditional on the factors are obtained using the Kalman filter.
-\[
-NIM_{t+s/t}=\sum_{i}\frac{NIM_{i,t+s/t}}{3},
-\]
 
-\noindent \textbf{6. No-change forecast.} According to this model $NIM_{t+s/t}=NIM_{t}${\normalsize . }
+\noindent \textbf{7. No-change forecast.} According to this model $NIM_{t+s/t}=NIM_{t}${\normalsize . }
 

### 2013-07-10 LATEX/beg.draft.2013.tex
I made a change to check things

@@
 \vspace{-0.1in}
-Regular bank stress testing (and the capital planning that it implies) is one of the three complementary reforms that has been made to U.S.
+AAA Regular bank stress testing (and the capital planning that it implies) is one of the three complementary reforms that has been made to U.S.
 capital regulation in response to the shortcomings of the pre-crisis regime.\renewcommand{\baselinestretch}{1}\footnote{See Governor Tarullo's November 2011 speech ``The Evolution of Capital Regulation'' for this characterization of stress testing.

### 2013-07-15 LATEX/beg.draft.2013.tex
Included discussion of results and cleaned up figures

--- 2013-07-09 LATEX/beg.draft.2013.tex
@@
 {\Large Luca Guerrieri}\\ Federal Reserve Board} 
-\title{Stressing Bank Profitability for Interest Rate Risk$^*$}
+\title{Stressing Bank Profitability for Interest Rate Risk$^*$ \\ \vspace{0.5cm} Preliminary and Incomplete \vspace{0.5cm}}
 
@@
 
-\vspace{3cm} \renewcommand{\baselinestretch}{1} \normalsize
+\vspace{1.5cm} \renewcommand{\baselinestretch}{1} \normalsize
 \noindent $^*$ The views
@@
 Third, modelling of NIMs has received scant attention so far.
-Forth, the ability of producing reliable conditional forecasts for a key component of net income is a the core of the scenario-based stress testing of banks.
+Fourth, the ability of producing reliable conditional forecasts for a key component of net income is at the core of the scenario-based stress testing of banks.
+We elaborate on each of these four parts of our motivation below.
 
@@
 
-Another motivation for our focus on models of NIM stems from the recent lackluster pace of bank profit growth, which is becoming an increasing source of concern to market commentators and regulators alike.
-Bank net income growth in recent years has been largely supported by unsustainable cost cutting and reductions in provisioning, rather than from sources like NII.
-NII growth, for example, (in both nominal and real terms) has been well below average in recent years while NIMs (NII relative to interest earning assets) have also been depressed.%\renewcommand{\baselinestretch}{1}\footnote{We should probably show both these  charts with adjustments for FAS 167.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5}
+The recent pace of bank profit growth has been lackluster and is becoming an increasing source of concern to market commentators and regulators alike.
+In recent years, bank net income growth has been largely supported by unsustainable cost cutting and reductions in provisioning, rather than from sources like NII.
+NII growth, for example, (in both nominal and real terms) has been well below average in recent years while NIMs (NII relative to interest earning assets) have also been depressed.
 Given their already low level, any adverse development that puts further downward pressure on NIMs could be very stressful for banks and, as such, it is important that the economic developments (likely involving interest rates) that would drive these outcomes are well-modeled.
@@
 We proceed as follows.
-Section~two provides a review of the literature on NIMs with an eye to cataloguing variables that have been found useful to explain the evolution of NIMs.
-Section~three turns to the models that we develop to forecast NIMS.
-Section 4 concludes.
+Section~2 provides a review of the literature on NIMs with an eye to cataloguing variables that have been found useful to explain the evolution of NIMs.
+Section~3 turns to the models that we develop to forecast NIMS.
+Section~4 describes the estimation data and Section~5 the estimation results.
+Section~6 concludes.
 
-\section{Related literature}
+\section{Related literature and variable selection}
 
@@
 As a result, the short rate is expected to enter the model, also, with a positive sign.
+Figure \ref{figure_nims_components} highlights the covariation of NIMs with the 3-month Treasury yield.
+In the figure NIMs are split into income and expenses components, both normalized by interest earning assets.
+In this and in the other figures described in this section, we focus on average for the top 25 bank holding companies in the United States by assets, assessed quarterly.
+The evolution of both components of NIMs bracket the evolution of the three month Treasury rate and the contours are strikingly similar.
+Figure \ref{figure_nims_rates} shows the evolution of NIMs against that of the 3-month and the 10-year Treasury yields.
 
@@
 and quantile regression in the case of Covas \textit{et al}.
+Figure \ref{figure_nims_factors} shows the evolution of average NIMs against ``observed'' level, slope, and curvature factors.\footnote{Following Diebold and Li (2006), we compute the observed factors, respectively as: the average of the 3-month, 2-year, and 10-year Treasury yields; the difference between the 10-year and the 3-month yields; and 10-year yield minus two times the 2-year yield plus the 3 month yield.}.
+As made explicit by the figure, NIMs comove strongly with the slope factor over the entire sample, but the level and curvature factors also move closely with NIMs over certain subperiods.
 
-Like our paper, English also only studies NIMs, although he does so for a number of countries (that is, all G-7 countries except for France, and for Australia, Norway, Sweden, and Switzerland).
-Addditionally, (and as we also do later), he considers asset yields and liability yields separately---the two measures that when differenced imply NIMs.
+Like our paper, English focuses exclusively on NIMs, although he does so for a number of countries (that is, all G-7 countries except for France, and for Australia, Norway, Sweden, and Switzerland).
+Addditionally (and as we also do later), he considers asset yields and liability yields separately---the two measures that when differenced imply NIMs.
 English's main focus is understanding how well the aggregate banking sectors in the countries he studies appear to have managed the interest rate risk that they face on their earnings, as such he uses aggregate data (rather than firm-level data) and does not focus on the conditional forecasting performance of his model.
@@
 The second part of the spread reflects banks' risk aversion to the situation of either having deposit supply without corresponding loan demand or having loan demand without the corresponding deposit supply.
-Consequently, the second part of the spread is the compensation for this risk that banks require in order for them to be willing to take deposits and lend.
+Accordingly, the second part of the spread is the compensation for this risk that banks require in order for them to be willing to take deposits and lend.
 This depends positively on the banks' degree of risk aversion as well as the volatility of interest rates.
@@
 In all cases the banks' degree of risk aversion does not enter the specification for estimated models of NIMs since this is difficult to measure, although the volatility of interest rates does and this is measured by the annual or quarterly standard deviation of weekly Treasury rates of different maturities (\textit{e.g.}, 3-month, 1-year, 2-year, 3-year, and 5-year rates).
+Figure \ref{figure_nims_competition} shows the evolution of NIMs against the evolution of the measure of competition that we consider -- the shadow banking sector share of assets in the overall banking sector comprising both banks and shadow banks.
+For most of the sample this share increases, but the decline is arrested and partially reversed towards the end of the sample.
+[TO DO: add a panel to that figure that includes a measure of volatility].
 
-Most papers in this literature also add a few more controls to the model before estimation, although in all cases this is done in a fairly \textit{ad hoc} way.
-These variables include banks' payments of implicit (instead of, explicit) interest on deposits;\renewcommand{\baselinestretch}{1}\footnote{This takes place through service charge remissions and other types of depositor subsidies due to regulatory restrictions of explicit interest payments and is measured by the difference between noninterest expenses and noninterest revenue, all divided by interest earning assets.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} banks' non-interest-bearing reserve requirements;\renewcommand{\baselinestretch}{1}\footnote{This captures the influence of the economic cost of funds over and above the published interest expense and is measured by non-interest bearing reserves divided by total assets.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} banks' default risk, since this should imply an additional premium charged on the rate of the banks loans given the potential loss in bank capital;\renewcommand{\baselinestretch}{1}\footnote{This is measured by net chargeoffs divided by total loans.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} and, bank capital, since capital is costly and banks with higher capital may seek to cover this with higher lending rates.
+Most papers in this literature also add a few more controls to the model before estimation.
+These variables include banks' payments of implicit (instead of, explicit) interest on deposits;\renewcommand{\baselinestretch}{1}\footnote{This takes place through service charge remissions and other types of depositor subsidies due to regulatory restrictions of explicit interest payments and is measured by the difference between noninterest expenses and noninterest revenue, all divided by interest earning assets.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} banks' non-interest-bearing reserve requirements;\renewcommand{\baselinestretch}{1}\footnote{The difference between noninterest expenses and noninterest revenue captures the influence of the economic cost of funds over and above the published interest expense and is measured by non-interest bearing reserves divided by total assets.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} banks' default risk, since this should imply an additional premium charged on the rate of the banks loans given the potential loss in bank capital;\renewcommand{\baselinestretch}{1}\footnote{This is measured by net chargeoffs divided by total loans.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} and, bank capital, since capital is costly and banks with higher capital may seek to cover this with higher lending rates.
 Note that all of the empirical analysis in the ``dealership'' model literature for studying NIMs is undertaken at the firm-level and as such many of the \textit{ad hoc} controls that are added to the model are included so as to address cross-firm differences.
@@
 We would note that Angbazo does add default risk to the bank's profit-maximization problem and, as such, default risk does enter his model of NIMs in a more rigorous way.
-Here he finds that for money center banks net interest margins are more sensitive to default risk than they are to interest-rate risk, while super-regional and regional banks are sensistive to interest-rate risk but less so to default risk.
-These findings he notes are consistent with these money center banks' greater concentration of short-term assets and greater use of off-balance sheet (OBS) hedging instruments.
+Here he finds that for money center banks net interest margins are more sensitive to default risk than they are to interest-rate risk, while super-regional and regional banks are sensitive to interest-rate risk but less so to default risk.
+These findings, he notes, are consistent with the money center banks' greater concentration of short-term assets and greater use of off-balance sheet (OBS) hedging instruments.
 
@@
 These models also include features like administrative costs on deposits and loans as well as costs associated with deposit insurance premia and the primary use of these models is for comparative static exercises that examine how changes in these parameters---as well as considerations like uncertainty---will affect bank spreads.
-Utlimately, however, these papers are not empirical and since the majority of their theoretical findings confirm those of those of papers like Ho and Saunders, Angbazo, and Saunders and Schumacher, we do not delve too deeply into them here.
+Ultimately, however, these papers are not empirical and since the majority of their theoretical findings confirm those of those of papers like Ho and Saunders, Angbazo, and Saunders and Schumacher, we do not delve too deeply into them here.
 
@@
 The terms ${\eta _{Lt}}$, ${\eta _{St}}$, ${\eta _{Ct}}$, ${\eta_{NIMt}}$ denote independently distributed innovations with joint covariance matrix $\Omega $.
+Our specification allows the yield curve factors to depend on their own lagged values but excludes lagged dependence on NIMs.
+By contrast, NIMs the specification allows for more general dependence of NIMs on lagged yield curve factors and and lagged NIMs.
 
@@
 
-\begin{frame}
-
 \noindent \textbf{5.
@@
 Forecasts conditional on the factors are obtained using the Kalman filter.
+We take as conditional forecasts the filtered estimates of NIMs conditional on the pseudo-out-of-sample observations of the remaining variables in the VAR.
+This procedure efficiently considers the  covariance between reduced-form residuals in the VAR.
 
-\noindent \textbf{7. No-change forecast.} According to this model $NIM_{t+s/t}=NIM_{t}${\normalsize . }
-
-%\renewcommand{\baselinestretch}{1}\footnote{...\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5}
+\noindent \textbf{7. No-change forecast.} Taking the forecasting model to be a random walk without a drift implies a no change forecast at all horizons $s$: $NIM_{t+s/t}=NIM_{t}${\normalsize . }
 
@@
 
-\vspace{-0.1in}
+The baseline sample starts in 1989q1 and stops in 2008q2.
+Our dataset would allow us to start even earlier, in 1985q4, but we delay the start of the estimation sample for two reasons.
+First, there is a peculiar spike evident in aggregated NIMs in Figure \ref{figure_nims_rates} in 1988q4.
+This large spike stems from late payments from Brazil during the Latin American debt crisis.
+The end point of 2008q2 is dictated by the attainment of the zero lower bound on nominal interest rates.\footnote{To estimate the dynamic factor models described under 2.
+and 3., we use observations between 1985q4 and 1989q4 as a training sample.}
 
@@
 Call Report data used in this analysis are cleaned and adjusted for bank mergers and acquisitions, using structure data from the National Information Clearinghouse (NIC) on mergers and acquisitions.
-Foreign entities are excluded and domestic subsidiaries are aggregated up to the parent, bank-holding-company (BHC), level.\renewcommand{\baselinestretch}{1}\footnote{NIM data are adjusted for mergers between commercial banks by comparing balance sheet values of interest income, interest expenses, and interest-earning assets at the end of the quarter with those at the beginning of the quarter, accounting for amounts acquired or lost during the period because of mergers.
-For information on the merger-adjustment procedure for income, see the appendix in English and Nelson (1998).\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5}
+Foreign entities are excluded and domestic subsidiaries are aggregated up to the parent, bank-holding-company (BHC), level.\footnote{NIM data are adjusted for mergers between commercial banks by comparing balance sheet values of interest income, interest expenses, and interest-earning assets at the end of the quarter with those at the beginning of the quarter, accounting for amounts acquired or lost during the period because of mergers -- see the appendix in English and Nelson (1998).}
 
@@
 (2007), based on Nelson and Siegel (1987) and Svensson (1994), which allows for daily measures of an off-the-run Treasury yield curve.
-We use yields for twelve maturities in our models: 3 month, 6 month, 9 month, 1 year, 2 year, 3 year, 5 year, 7 year, 10 year, 15 year, 20 year, and 30 year.\renewcommand{\baselinestretch}{1}\footnote{Daily yields are published with a two-day lag under the base mnemonic SVENY at  www.federalreserve.gov/econresdata/researchdata/feds200628\_1.html.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} The yields included are quarterly averages of daily yields.
-The data on NIMs and yields start in 1985Q4 and end in 2008Q2, a period that stops short of the interest rate regime shift to the zero lower bound.
+We use quarterly yields for twelve maturities in our models: 3 month, 6 month, 9 month, 1 year, 2 year, 3 year, 5 year, 7 year, 10 year, 15 year, 20 year, and 30 year.\footnote{Daily yields are published with a two-day lag under the base mnemonic SVENY at  www.federalreserve.gov/econresdata/researchdata/feds200628\_1.html.} The yields included are quarterly averages of daily yields.
 
@@
 
-% Suggestion to look at the MOVE index.
-It looks like it goes back to at least April 1988 (http://www.businessinsider.com/the-move-index-bonds-2010-3) This (http://business.financialpost.com/2012/11/01/investors-seeking-yield-should-use-move-and-vix/) also says it was created in 1988.
+\section{Results}
 
-\newpage
-\begin{thebibliography}{99}
-\vspace{-0.1in}
-\bibitem{abel82} Abel, A.B. (1982).  ``Dynamic Effects of Permanent and Temporary Tax Policies in
-a $q$ Model of Investment.''  \textit{Journal of Monetary Economics}, 9, 353-373.
+To compare the performance of the models, we consider pseudo-out-of-sample forecasts.
+We use a rolling window for estimation.
+The shortest window spans 1989q4 to 1999q4.
+The assessment window spans 2000q1 to 2008q3.
+We focus on forcasts from 1 to 10 steps ahead.
+In constructing the out-of-sample forecasts, we condition on the observation of the right-hand-side variables in each model, with the exception of lagged NIMs.
 
-\bibitem{auerb1} Auerbach, A.J. (1979). ``Wealth Maximization and the Cost of Capital.''
-\textit{Quarterly Journal of Economics}, 93, 433-446.
+Table \ref{table_rmse_baseline} reports root mean square errors (RMSE) for the pseudo-out-of-sample forecasts.
+Model 1., the equal weighted forecast combination, is the only model to have lower RMSE's relative to the no-change forecasts for Model 6.
+In relative terms, the advantage of Model 1.
+over Model 6.
+increases as the forecast horizon lengthens.
+Model 2, performs uniformly worse than model 6.
+but, excluding Model 1, it performs better than any other model we considered.
+Models 3a.
+and 3b.
+do not incorporate the information in the variance/covariance matrix of the shock in the state equation of Model 2.
+The choice of variables for the forecast combination in model 4.
+is not as broad as for 1.
+Finally, the proliferation of parameters in the VAR of model 5.
+is associated with the worst performance, overall.
 
-\bibitem{auerb2} Auerbach, A.J. (1989). ``Tax Reform and Adjustment Costs:  The Impact on
-Investment and Market Value.''  \textit{International Economic Review}, 30, 939-962.
+Table \ref{table_rmse_competition} considers variants of the model that include the share of assets of the shadow banking sector.
+\footnote{In this case, we exclude Model 2.
+from the comparison.}  In no model, does that addition  improve the performance of the conditional forecasts.
 
-\bibitem{auerb3} Auerbach, A.J., and L.H. Summers (1979). ``The Investment Tax Credit: An Evaluation.''
-NBER Working Paper No.~404.
+Table \ref{table_rmse_combination} considers the performance of alternative variants of Model 1.
+Using either our ``observed factors,'' or factors derived as smoothed estimates from the yield curve model set up as part of Model 3., does not improve the forecast performance.
+By contrast, pruning out some of the yields, lowers the RMSEs for the forecasts even further.
+A forecast combination approach that includes models conditional on the 3-month Treasury yields and the 10-year treasury yields has the best out-of-sample performance across the models we considered.
 
-\end{thebibliography}
+Figure \ref{figure_scenario} uses the forecast combination model that includes the 3-month and the 10-year yields only to forecast NIMs conditional on the three CCAR 2013 supervisory scenarios.
+The severely adverse scenario implies a downward shift of the profile of the yield curve accompanied by a sharp flattening.
+The severe scenario implies an upward shift of the yield curve with little change in the slope.
+In terms of changes in the yield curve, these two scenarios capture vastly different configurations.
+The implications of the point forecasts appear sensible.
+The severely adverse scenario reduces NIMs, but it takes a few quarters for the implications of the change in yields to transmit to NIMs.
+By contrast, the upward parallel shift in the yield curve is predicted to boost NIMs.
+
+The figure, provides a visual interpretation of the RMSEs considered thus far.
+The bands surrounding the central forecast are constructed adding and subtracting to the point forecast the RMSEs for the forecast combination model.
+Strikingly, the size of the uncertainty bands dwarfs the variation implied by the scenarios for NIMs.
+The two severe scenarios are statistically indistinguishable from the baseline and from each other.
+
+\section{Conclusion}
+
+The unweighted forecast combination model for NIMs based on just two yields outperforms the other forecast models we considered.
+No other model comes close to outperforming a no-change forecast.
+This is only a small victory as the variance of the forecast error is large even when compared to the observed variation of NIMs.
+Accordingly, even scenarios that have vastly different implications for the configuration of interest rates contain relatively little information to forecast NIMs.
+
+Our findings have implications beyond the study of net interest margins.
+Scenario-based stress tests have been used across the world to rebuild confidence in the banking sector in periods of crises.
+Yet, they are predicated on the assumption that different scenarios for the evolution of the macroeconomy have sufficiently different implications for the evolution of key balance sheet items.
+Yet, an array of time series methods found by others to be near the frontier of efficiency produces projections conditional on radically different scenarios for yields that statistically indistinguishable.
+
+Supervisory authorities that perform stress tests certainly have access to a wealth of supervisory information beyond what  the publicly available data that we have used.
+Yet, if the conclusions of stress tests are mostly predicated on supervisory sources, as our study implies, communication with the public is even more challenging.
+In that case, the effects of confidence of stress tests rely squarely on the reputation of the supervising authority, one which might be tarnished exactly in a period of crisis.
 
@@
 \center
-\caption{RMSE for shortened sample: 1989Q2-2008Q2}
+\caption{RMSE for baseline specifications} \label{table_rmse_baseline}
 \begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|}
@@
 \begin{sidewaystable}
-\caption{RMSEs over full sample: 1985Q4-2008Q2}
-\center
-\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|}
-\hline
-&Step 1 &Step 2 &Step 3 &Step 4 &Step 5 &Step 6 &Step 7 &Step 8 &Step 9 &Step 10\\
-\hline
-1. Simple Forecast Combination             &0.105&0.140&0.164&0.191&0.218&0.244&0.265&0.281&0.305&0.327\\
-3a. DFM with 2nd Step Regression           &0.128&0.190&0.244&0.299&0.350&0.397&0.438&0.475&0.514&0.549\\
-3b. DFM with Forecast Combination          &0.112&0.158&0.197&0.240&0.290&0.337&0.373&0.407&0.446&0.480\\
-4. Forecast Combination of Observed Factors&0.116&0.174&0.227&0.284&0.347&0.409&0.459&0.508&0.554&0.595\\
-5. VAR on Observed Factors                 &0.148&0.196&0.243&0.275&0.283&0.296&0.319&0.377&0.456&0.530\\
-6. No-Change Forecast                      &0.107&0.138&0.154&0.181&0.206&0.228&0.244&0.248&0.271&0.293\\
-\hline
-\end{tabular}
-\end{sidewaystable}
-}
-
-\newpage
-{\normalsize \clearpage
-\begin{sidewaystable}
-\caption{RMSEs for models that include the share of asset of the shadow banking sector. Estimation sample: 1989Q2-2008Q2}
+\caption{RMSEs for models that include the share of asset of the shadow banking sector. Estimation sample: 1989Q2-2008Q2} \label{table_rmse_competition}
 \center
@@
 \newpage \clearpage
-\begin{table}
-\caption{RMSEs for variants of the simple forecast combination model}
+\begin{sidewaystable}
+\caption{RMSEs for variants of the simple forecast combination model} \label{table_rmse_combination}
 \center
@@
 \end{tabular}
-\end{table}
+\end{sidewaystable}
 
 \newpage \clearpage
+\newpage \clearpage
+\begin{figure}
+\caption{Interest Income, Interest Expenses and Short-term Treasury Rates} \label{figure_nims_components}
+\center
+\includegraphics[scale=0.85]{figure_nims_components.ps}
+\end{figure}
+
 \begin{figure}
@@
 
-\newpage \clearpage
+\clearpage
 \begin{figure}
-%[tbp]
-\caption{NIMs and Measures of Competition} \label{figure_nims_competition}
-\center
-\includegraphics[scale=0.85]{figure_nims_competition.ps}
+\flushleft
+\caption{NIMs and Competition -- The Asset Share of the Shadow Banking Sector} \label{figure_nims_competition}
+
+\hspace{-2.5cm} \includegraphics[scale=0.85]{figure_nims_competition.ps}
 \end{figure}
 
-\newpage \clearpage
+\clearpage
 \begin{figure}
-%[tbp]
-\caption{Interest Income, Interest Expenses and Short-term Treasury Rates} \label{figure_nims_components}
-\center
-\includegraphics[scale=0.85]{figure_nims_components.ps}
+\flushleft
+\caption{Forecast for NIMs Conditional on CCAR 2012 Scenarios} \label{figure_scenario}
+
+ \includegraphics[scale=0.85]{figure_scenario.ps}
 \end{figure}

### 2013-07-15 MATLAB/nim_lsc_model_2steps/run_scenario.m
cleaned out code

@@ -197,8 +197,7 @@ title('Forecast of NIMs')
 legend('Baseline','Severely Adverse','1-RMSE band','Location','SouthWest')
 
 plot(forecast_dates, forecast_severe-rmse_forecast_combination_mat4,'r-.')
-xlim([2012 2015])
-ylim([2.4 3.2])
+
 
 
 subplot(3,2,6)
@@ -210,8 +209,7 @@ plot(forecast_dates, forecast_adverse-rmse_forecast_combination_mat4,'r-.')
 
 
 title('Forecast of Nims')
-xlim([2012 2015])
-ylim([2.4 3.2])
+
 
 
 

### 2013-07-15 MATLAB/nim_lsc_model_2steps/setpath.m
made paths compatible with linux

@@ -6,6 +6,12 @@ path1 = '.\utilities_2steps';
 path2 = '..\estimation_utilities';
 path3 = '..\plot_support';
 
+if isunix
+    path1 = strrep(path1,'\','/');
+    path2 = strrep(path2,'\','/');
+    path3 = strrep(path3,'\','/');
+end
+
 
 path (path,path1);
 path (path,path2);

### 2013-07-15 MATLAB/nim_lsc_model_2steps/run_calc_rmse_conditional.m
split out tables and figures

@@ -81,6 +81,14 @@ lag = 1;
 % Table 1 -- Shortened sample:
 [rmse_forecast_combination_mat1, forecast_combination_mat1] = calc_rmse_forecast_combination_conditional(nims(:,15:end),yields(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, 1,4);
 
+nims_change = nims(:,15:end)-nims(:,14:end-1);
+nims_change_out_of_sample = nims_change(out_of_sample_start_pos-14:end);
+nims_change_pre_out_of_sample = nims_change(out_of_sample_start_pos-15);
+
+[rmse_forecast_combination_change_mat1, forecast_combination_change_mat1] = calc_rmse_forecast_combination_conditional(nims_change,yields(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, 1,4);
+[rmse_forecast_combination_level_mat1, forecast_combination_level_mat1] = rmse_change2level(forecast_combination_change_mat1,nims_change_out_of_sample,nims_change_pre_out_of_sample); 
+
+
 [rmse_multivariate_mat1, forecast_multivariate_mat1] = calc_rmse_multivariate_conditional(nims(:,15:end), smoothed_factors(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, lag);
 
 
@@ -104,6 +112,7 @@ varlag=4;
 
 table1 = [
 rmse_forecast_combination_mat1
+rmse_forecast_combination_level_mat1
 rmse_multivariate_mat1
 rmse_forecast_combination_mat2
 rmse_pc_mat1
@@ -117,6 +126,7 @@ rmse_pls_mat1
 
 columnlabels = char('Step 1','Step 2','Step 3','Step 4','Step 5','Step 6','Step 7','Step 8','Step 9','Step 10');
 rowlabels = char('1. F. Combination - Yields',...
+                 '1b. F. Combination of change - Yields',... 
                  '3a. DFM with 2nd Step Reg.',...
                  '3b. DFM with F. Combination',...
                  '4. PCR',...

### 2013-07-15 MATLAB/nim_lsc_model_2steps/run_figures_data.m
data figures

@@ -1,0 +1,48 @@
+
+clear;
+
+setpath;
+
+dataset_option = 1;
+
+[yobs, dates, yields, nims, nfactors, nothers, tau, factors,...
+    shadow_bank_share_assets,...
+    total_interest_earning_assets,...
+    assets_depository_inst, assets_securities_notrade,...
+    assets_fedfunds, assets_all_loans, assets_trading_accnts, ... 
+    interest_income_to_ie_assets, interest_expense_to_ie_assets] = load_data_ml_long(dataset_option);
+
+
+figure
+subplot(2,1,1)
+thistitle='Nims (solid) and 3-Month Yields (dashed)';
+doubleplot(nims,yields(1,:),dates,thistitle)
+
+subplot(2,1,2)
+thistitle='Nims (solid) and 10-Year Yields (dashed)';
+doubleplot(nims,yields(9,:),dates,thistitle)
+
+
+figure
+subplot(3,1,1)
+thistitle='Nims (solid) and Observed Level Factor (dashed)';
+doubleplot(nims,factors(1,:),dates,thistitle)
+
+
+subplot(3,1,2)
+thistitle='Nims (solid) and Observed Slope Factor (dashed)';
+doubleplot(nims,-factors(2,:),dates,thistitle)
+
+subplot(3,1,3)
+thistitle='Nims (solid) and Observed Curvature Factor (dashed)';
+doubleplot(nims,-factors(3,:),dates,thistitle)
+
+
+figure
+thistitle='Nims (solid) and Asset Share of Shadow Banking Sector (dashed)';
+doubleplot(nims,1-shadow_bank_share_assets,dates,thistitle)
+xlim([dates(1) dates(end)])
+
+figure
+thistitle = 'Interest Income (% of i.e. assets), Interest Expenses (% of i.e. assets) and the 3-month Treasury yield (RHS scale)'; 
+tripleplot(interest_income_to_ie_assets, interest_expense_to_ie_assets, yields(1,:),dates,thistitle)

### 2013-07-17 LATEX/beg.draft.2013.tex
Revised into, edited throughout, added bibliography

@@
 \begin{abstract}
-Regular bank stress testing is one of the three complementary reforms that has been made to U.S. capital regulation in response to the shortcomings of the pre-crisis regime.
-The form of stress testing mandated by the Dodd-Frank Act for the Federal Reserve to undertake is one based around a small number of scenarios.
-This approach involves specifying macroeconomic and financial scenarios that represents stressful conditions over the time horizon of the stress test and then projecting, for each bank holding company in the stress test, its losses, income, and path of capital \textit{given that scenario}.
-In this paper, we focus on the path of income.
-In particular, we consider the ability of empirical models to translate developments in interest rates into bank net interest margins (NIMs).
+\noindent Regular bank stress testing (and the capital planning that it implies) is one of the three complementary reforms that has been made to U.S.
+capital regulation in response to the shortcomings of the pre-crisis regime.
+The form of stress testing currently undertaken by the Federal Reserve in implementing the stress testing requirement of the Dodd-Frank Act is based around a small number of scenarios.
+This approach involves specifying macroeconomic and financial scenarios that represent stressful conditions over the time horizon of the stress test and then projecting, for each bank holding company in the stress test, its losses, income, and path of capital \textit{given that scenario}.
+In this paper, we focus on the path of income and consider the ability of empirical models to translate developments in interest rates into bank net interest margins (NIMs).
 \end{abstract}
@@
 capital regulation in response to the shortcomings of the pre-crisis regime.\renewcommand{\baselinestretch}{1}\footnote{See Governor Tarullo's November 2011 speech ``The Evolution of Capital Regulation'' for this characterization of stress testing.
-The two other reforms that he mentions are improvements to the traditional, individual firm approach to capital regulation and the introduction of a macroprudential component to capital regulation.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} While stress testing can take many forms---for example, being based on a small number of scenarios, being based on a large number of scenarios, or even being undertaken in reverse form to achieve a certain adverse outcome---the form of the stress tests mandated under the Dodd-Frank Act for the Federal Reserve to undertake and the form being used by the Federal Reserve for its annual Comprehensive Capital Analysis and Review (CCAR) is one based around a small number of scenarios.
+The two other reforms that he mentions are improvements to the traditional, individual firm approach to capital regulation and the introduction of a macroprudential component to capital regulation.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} While stress testing can take many forms---for example, being based on a small number of scenarios, being based on a large number of scenarios, or even being undertaken in reverse form to achieve a certain adverse outcome---the form of the stress tests currently being undertaken by the Federal Reserve in implementing the stress testing requirment of the Dodd-Frank Act is one based around a small number of scenarios.
 This approach involves specifying a small number of macroeconomic and financial scenarios that represent stressful conditions over the time horizon of the stress test and then projecting, for each bank holding company (BHC) in the stress test, its losses, income, and path of capital \textit{given that scenario}.
-Accordingly, there are at least two important parts to conducting bank stress tests (i)~formulating macroeconomic and financial scenarios that should be stressful to banks and (ii)~developing econometric models that are able to translate such stressful conditions into losses, income, and capital.
+Accordingly, there are at least two important parts to conducting bank stress tests (i)~formulating stressful macroeconomic and financial scenarios around which stress tests can then be performed and (ii)~developing econometric models that are able to translate reliably such stressful conditions into losses, income, and capital.
 
-We examine how well a range of time-series models can explain NIMs.
-We emphasize the models' pseudo-out-of-sample \textit{conditional} forecast performance.
+We examine how well a range of time-series models can explain net interest margins---a key source of bank income generation.
+In doing so, we emphasize the models' pseudo-out-of-sample \textit{conditional} forecast performance.
 We adopt this pseudo-out-of-sample approach for the simple reason that the evaluation of models based on their in-sample forecast performance, may lead to in-sample over-fitting and poor out-of-sample performance.
-An explicit focus on out-of-sample performance makes forecast uncertainty explicit.
-Moreover, concentrating on conditional forecasts, allows us to separate the task of modeling and forecasting NIMs from that of forecasting interest rates, which is itself a sizable topic (see Duffee, 2012).
+Moreover, an explicit focus on out-of-sample performance makes forecast uncertainty explicit.
+We concentrate on conditional forecasts for a couple of reasons.
+First, it allows us to separate the task of modeling and forecasting NIMs from that of forecasting interest rates, which is itself a sizable topic (see \citeN{duffee2012}).
+Second, macroeconomic stress tests are in effect conditional forecasts, in this case of banking variables on stressful outcomes of aggregate macroeconomic variables.
 
 When developing econometric models to translate stressful economic conditions into key bank financial-statement variables, it is important to consider what types of developments are likely to be stressful to banks and what particular economic variables are likely to capture these stresses.
-It is then critical that the models being used for stress testing be able to translate developments in these economic variables into conditional forecasts for key bank financial-statement variables.
-In this paper, we examine this property for models of bank net interest margins (NIMs), defined as the ratio of net interest income (NII) to interest earning assets, where NII is the difference between a bank's interest income and interest expenses.
-
-In determining which variables to include in our model of NIMs we look to the variables that have been emphasized by two existing strands of the literature on NIMs.
-The first strand emphasizes the link between interest rates of different maturities and NIMs and is more in the macro-banking tradition.
-The second literature focuses on the optimal margin set by banks based on considerations like the competitive structure of the lending and deposit markets in which the bank participates, the degree of managerial risk aversion, and the degree of uncertainty associated with the future path of interest rates and is much more in the micro-banking tradition.
+It is then critical that the models being used for stress testing are able to translate developments in these economic variables into conditional forecasts for key bank financial-statement variables.
+We examine this property for models of bank net interest margins (NIMs), defined as the ratio of net interest income (NII) to interest earning assets, where NII is the difference between a bank's interest income and interest expenses.
 
 Our motivation for focusing on models of NIMs is fourfold.
-First, losses from interest rate risk have been a central feature of banking crises.
-Second, net interest rate margins capture income from the maturity transformation services at the center of the traditional banking model.
-This source of income has been lackluster given the current configuration of interest rates.
-Third, modelling of NIMs has received scant attention so far.
-Fourth, the ability of producing reliable conditional forecasts for a key component of net income is at the core of the scenario-based stress testing of banks.
+First, losses from interest rate risk have been a central feature of a number of past banking crises.
+Second, NIMs are an important component of bank revenue generation---reflecting revenues from maturity transformation services and from deposit transactions services, two of banks' traditional activities.
+Given their influence on bank earnings, NIMs are a variable that receive considerable focus from bank equity analysts.\renewcommand{\baselinestretch}{1}\footnote{Indeed, one cited  rule of thumb in the industry is that 10~basis point lower NIMs means 7~percent lower earnings.
+(See "Nomura Securities on U.S.
+Banks: Turning Japanese v2 - NIMs Hunker Down," \emph{StreetInsider.com}, June 7, 2012.) \vspace{0.05in}}\renewcommand{\baselinestretch}{1.5}   Third, the modeling of NIMs has received scant attention so far by academic researchers, despite its importance in the industry.
+Fourth, the ability to produce reliable conditional forecasts for a key component of net income is at the core of the scenario-based stress testing of banks.
 We elaborate on each of these four parts of our motivation below.
@@
 
-Likewise, the Nordic banking crises of the late 1980s and early 1990s (for Finland, Sweden, and Norway)---which represent three of Reinhart's and Rogoff's~(2009) ``big five'' banking crises---were associated with elevated short-term interest rates that depressed NII and NIMs and lead to sizable bank losses and failures.
+Likewise, the Nordic banking crises of the late 1980s and early 1990s (for Finland, Sweden, and Norway)---which represent three of \citeN{reinhart} ``big five'' banking crises---were associated with elevated short-term interest rates that depressed NII and NIMs and lead to sizable bank losses and failures.
 In the Nordic crises, high short-term interest rates resulted from these countries attempting to maintain currency pegs with the Deutschemark in the face of the tight German monetary policy that accompanied its loose fiscal policy associated with re-unification spending.
@@
 A sudden increase in short-term interest rates abruptly drove up banks' costs of funds,  leading to extensive bank losses and failures.
-To be sure, large increases in short-term rates and declines in net interest income were not the \textit{only} reason for any of these crises---indeed, there was at least one major class of asset prices that collapsed in each of these crises---but short-term interest rates---and, in particular, their elevated levels relative to long-term interest rates---were significant and in most cases equivalent in importance to the other causes of the crisis.
+To be sure, large increases in short-term rates and declines in net interest income were not the \textit{only} reason for any of these crises---indeed, there was at least one major class of asset prices that collapsed in each of these crises---but short-term interest rates---and, in particular, their elevated levels relative to long-term interest rates---were significant and in many cases equivalent in importance to the other causes of the crisis.
 
-The recent pace of bank profit growth has been lackluster and is becoming an increasing source of concern to market commentators and regulators alike.
+NIMs are an important component of bank revenue generation and receive considerable focus from bank equity analysts.
+As such, recent subdued NIM levels have become an increasing source of concern to market commentators.
 In recent years, bank net income growth has been largely supported by unsustainable cost cutting and reductions in provisioning, rather than from sources like NII.
-NII growth, for example, (in both nominal and real terms) has been well below average in recent years while NIMs (NII relative to interest earning assets) have also been depressed.
+NII growth, for example, (in both nominal and real terms) has been well below average while NIMs (NII relative to interest earning assets) have also been depressed.
 Given their already low level, any adverse development that puts further downward pressure on NIMs could be very stressful for banks and, as such, it is important that the economic developments (likely involving interest rates) that would drive these outcomes are well-modeled.
@@
 Despite the importance of NIMs for the financial condition of banks and despite the attention that it receives from bank equity analysts, the modeling of NIMs has received much less attention in the risk-management literature relative to the modeling of bank variables related to credit risk.
-To be sure, the recent increase in the importance of stress testing has led to the development of a number of models that link NIMs to macroeconomic variables---such as, short-term Treasury rates and the slope of the Treasury yield curve---with Kovner \textit{et al}.~(2011) and Covas \textit{et al}.~(2012) being two notable examples for the U.S.
-However, this research places relatively less emphasis on the conditional forecasting properties of these models, which is important given that these models are ultimately to be used to project bank losses, income, and capital given macroeconomic scenarios.\renewcommand{\baselinestretch}{1}\footnote{Covas \textit{et al}.
-do present some forecast evaluation results in order to demonstrate the benefits of focusing on quantile projections.
+To be sure, the recent increase in the importance of stress testing has led to the development of a number of models that link NIMs to macroeconomic variables with \citeN{covas2012} and \citeN{hirtle2013} being two notable examples for the U.S.
+However, this research places relatively less emphasis on the conditional forecasting properties of these models, which is important given that these models are ultimately to be used to project bank losses, income, and capital given macroeconomic scenarios.\renewcommand{\baselinestretch}{1}\footnote{\citeN{covas2012} do present some forecast evaluation results in order to demonstrate the benefits of focusing on quantile projections.
 However, they only report results for net chargeoffs and pre-provisioning net revenue and their focus is on density forecast evaluation.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5}
@@
 Section~3 turns to the models that we develop to forecast NIMS.
-Section~4 describes the estimation data and Section~5 the estimation results.
-Section~6 concludes.
+Section~4 describes the data used to estimate the forecast models and Section~5 reports our forecast results.
+Section~6 uses one of the paper's models to consider the path of NIMs under the scenarios publish by the Federal Reserve as part of last year's Dodd Frank Act stress tests.
+Section~7 concludes.
 
@@
 
-Kovner \textit{et al}.
-and Covas \textit{et al.} are in the macro-banking tradition.
-Both link NIMs to the slope of the yield curve and a short-term interest rate.
+\citeN{covas2012} and \citeN{hirtle2013} are in the macro-banking tradition and both link NIMs to the slope of the yield curve and a short-term interest rate.
 Banks' earnings from interest income reflect two key services that banks supply---maturity transformation services (that is, borrowing short and lending long) and deposit transactions services---and the returns to these services are reflected by these two interest rate variables.
@@
 The evolution of both components of NIMs bracket the evolution of the three month Treasury rate and the contours are strikingly similar.
-Figure \ref{figure_nims_rates} shows the evolution of NIMs against that of the 3-month and the 10-year Treasury yields.
+Figure~\ref{figure_nims_rates} shows the evolution of NIMs against that of the 3-month and the 10-year Treasury yields.
 
-Both Kovner \textit{et al}.
-and Covas \textit{et al}.
-measure the slope of the yield curve in their respective NIM models with the 10-year to 3-month Treasury term spread and short-term market rates with the 3-month Treasury bill rate, as do English~(2002) and English \textit{et al}.~(2012)---two other papers in this tradition.
-All of these papers estimate NIM models with slightly different purposes in mind.
-The NIM models estimated by Kovner \textit{et al}.
-and Covas \textit{et al}.
-are developed for stress testing and are part of much larger aggregative models of a number of components of banks' financial statements.
-These models are both estimated on panel data with the most notable difference between these models being the estimation techniques they employ: Least squares regression in the case of Kovner \textit{et al}.
-and quantile regression in the case of Covas \textit{et al}.
-Figure \ref{figure_nims_factors} shows the evolution of average NIMs against ``observed'' level, slope, and curvature factors.\footnote{Following Diebold and Li (2006), we compute the observed factors, respectively as: the average of the 3-month, 2-year, and 10-year Treasury yields; the difference between the 10-year and the 3-month yields; and 10-year yield minus two times the 2-year yield plus the 3 month yield.}.
-As made explicit by the figure, NIMs comove strongly with the slope factor over the entire sample, but the level and curvature factors also move closely with NIMs over certain subperiods.
+Both \citeN{covas2012} and \citeN{hirtle2013} measure the slope of the yield curve in their respective NIM models with the 10-year to 3-month Treasury term spread and short-term market rates with the 3-month Treasury bill rate, as do \citeN{english2002} and \citeN{english2012}---two other papers in this tradition.
+All of these papers estimate NIM models with slightly different purposes in mind but all use essentially the same interest rate variables to summarize the yield curve.
 
-Like our paper, English focuses exclusively on NIMs, although he does so for a number of countries (that is, all G-7 countries except for France, and for Australia, Norway, Sweden, and Switzerland).
+The NIM models estimated by \citeN{covas2012} and \citeN{hirtle2013} are developed for stress testing and are part of much larger aggregative models of a number of components of banks' financial statements.
+These models are both estimated on panel data with the most notable difference between these models being the estimation techniques they employ: Least squares regression in the case of \citeN{hirtle2013} and quantile regression in the case of \citeN{covas2012}.
+
+Like our paper, \citeN{english2002} focuses exclusively on NIMs, although he does so for a number of countries (that is, all G-7 countries except for France, and for Australia, Norway, Sweden, and Switzerland).
 Addditionally (and as we also do later), he considers asset yields and liability yields separately---the two measures that when differenced imply NIMs.
-English's main focus is understanding how well the aggregate banking sectors in the countries he studies appear to have managed the interest rate risk that they face on their earnings, as such he uses aggregate data (rather than firm-level data) and does not focus on the conditional forecasting performance of his model.
+The main focus of \citeN{english2002} is understanding how well the aggregate banking sectors in the countries he studies appear to have managed the interest rate risk that they face on their earnings, as such he uses aggregate data (rather than firm-level data) and does not focus on the conditional forecasting performance of his model.
 
-Finally, English \textit{et al} consider the relationship between NIMs and the slope of the yield curve and short-term rates for a panel of banks, however, in contrast to the other studies they allow for the specific coefficients estimated for each bank to vary as linear functions of bank balance sheet ratios that one \textit{a priori} might think would be relevant to net interest income.
+Finally, \citeN{english2012} consider the relationship between NIMs and the slope of the yield curve and short-term rates for a panel of banks, however, in contrast to the other studies they allow for the specific coefficients estimated for each bank to vary as linear functions of bank balance sheet ratios that one \textit{a priori} might think would be relevant to net interest income.
 In particular, these ratios include the maturity gap between bank assets and liabilities, the deposit share of liabilities, and the loan share of assets.
-Again, since English \textit{et al} are focussed on understanding the determinants of NIMs they do not consider conditional forecast performance.
+Again, since \citeN{english2012} are focussed on understanding the determinants of NIMs they do not consider conditional forecast performance.
+
+While the macro-banking NIM literature that we have just discussed relies on measures of the level and slope of the yield curve to summarize interest rates, the literature modeling the yield curve also considers the curvature.
+Figure~\ref{figure_nims_factors} shows the evolution of average NIMs against a set of factors summarizing the yield curve---specifically, level, slope, and curvature factors---where these factors are all simple (additive) functions of a small number of interest rates of different maturities, similar to those used by other authors in the macro-banking literature.\renewcommand{\baselinestretch}{1.0}\footnote{The factors shown in figure~\ref{figure_nims_factors} follow \citeN{diebold2006}.
+In particular, we compute the observed factors of level, slope, and curvature, respectively, as: the average of the 3-month, 2-year, and 10-year Treasury yields; the difference between the 10-year and the 3-month yields; and 10-year yield minus two times the 2-year yield plus the 3 month yield.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} In the empirical analysis that follows we will refer to the yield curve factors that are based on simple functions of a few points on the yield  curve as ``observed'' factors.
 
 A separate strand of the NIM modeling literature comes from the micro-banking tradition, which focuses on the determinants of the loan rates and deposit rates that banks set as implied by the maximization of their profits.
-Papers in this tradition---such as, Ho and Saunders~(1981), Angbazo~(1997), and Saunders and Schumacher~(2000)---are based around what is known as the ``dealership'' model of banks; that is, banks in these models are viewed as dealers in the credit market, acting as an intermediary between demanders and suppliers of funds.
+Papers in this tradition---such as, \citeN{ho1981}, \citeN{angbazo1994}, and \citeN{saunders2000}---are based around what is known as the ``dealership'' model of banks; that is, banks in these models are viewed as dealers in the credit market, acting as an intermediary between demanders and suppliers of funds.
 Banks are modeled as being able to raise funds from and lend funds to short-term money markets as well as being able to raise deposits from and make loans to nonfinancial agents.
@@
 Importantly, the life of the deposit is the same as that of the loan, reflecting the fact that while core deposits can, in principal, be withdrawn on demand from a bank they are in practice extremely sticky.
-Interest rate risk arises in this model from the fact that deposit-taking and lending opportunities do not arrive at the bank at precisely the same time but rather arrive at any time with some probability, where these probabilities---which are different for deposits and lending opportunities---depend on the interest rates set by the bank on deposits and loans.
+Interest rate risk arises in this model from the fact that deposit-taking and lending opportunities do not arrive at the bank at precisely the same time but rather arrive at any time with some probability.
+These probabilities---which are different for deposits and lending opportunities---depend on the interest rates set by the bank on deposits and loans.
 Consequently, banks can find themselves holding deposits without loans, that they must then invest at variable market rates (and that could adversely fall below the deposit rate), or loans without deposits, that they must then fund at variable market rates (that could adverse rise above the lending rate).
@@
 
-In Ho and Saunders the degree of competition faced by the bank in loan and deposit markets does not enter their estimated model, since they argue that this variable is quite slow moving and their analysis is undertaken only over a period of three years.
-Saunders and Schumacher do consider the degree of competition in their analysis--which is cross-country in nature---although this takes the form of a country dummy variable in the regressions.
+In \citeN{ho1981} the degree of competition faced by the bank in loan and deposit markets does not enter their estimated model, since they argue that this variable is quite slow moving and their analysis is undertaken only over a period of three years.
+\citeN{saunders2000} do consider the degree of competition in their analysis---which is cross-country in nature---although this takes the form of a country dummy variable in the regressions.
 In all cases the banks' degree of risk aversion does not enter the specification for estimated models of NIMs since this is difficult to measure, although the volatility of interest rates does and this is measured by the annual or quarterly standard deviation of weekly Treasury rates of different maturities (\textit{e.g.}, 3-month, 1-year, 2-year, 3-year, and 5-year rates).
-Figure \ref{figure_nims_competition} shows the evolution of NIMs against the evolution of the measure of competition that we consider -- the shadow banking sector share of assets in the overall banking sector comprising both banks and shadow banks.
+Figure~\ref{figure_nims_competition} shows the evolution of NIMs against the evolution of the measure of competition that we consider -- the shadow banking sector share of assets in the overall banking sector comprising both banks and shadow banks.
 For most of the sample this share increases, but the decline is arrested and partially reversed towards the end of the sample.
@@
 Most papers in this literature also add a few more controls to the model before estimation.
-These variables include banks' payments of implicit (instead of, explicit) interest on deposits;\renewcommand{\baselinestretch}{1}\footnote{This takes place through service charge remissions and other types of depositor subsidies due to regulatory restrictions of explicit interest payments and is measured by the difference between noninterest expenses and noninterest revenue, all divided by interest earning assets.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} banks' non-interest-bearing reserve requirements;\renewcommand{\baselinestretch}{1}\footnote{The difference between noninterest expenses and noninterest revenue captures the influence of the economic cost of funds over and above the published interest expense and is measured by non-interest bearing reserves divided by total assets.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} banks' default risk, since this should imply an additional premium charged on the rate of the banks loans given the potential loss in bank capital;\renewcommand{\baselinestretch}{1}\footnote{This is measured by net chargeoffs divided by total loans.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} and, bank capital, since capital is costly and banks with higher capital may seek to cover this with higher lending rates.
-Note that all of the empirical analysis in the ``dealership'' model literature for studying NIMs is undertaken at the firm-level and as such many of the \textit{ad hoc} controls that are added to the model are included so as to address cross-firm differences.
+These variables include banks' payments of implicit (instead of, explicit) interest on deposits;\renewcommand{\baselinestretch}{1}\footnote{This takes place through service charge remissions and other types of depositor subsidies due to regulatory restrictions of explicit interest payments and is measured by the difference between noninterest expenses and noninterest revenue, all divided by interest earning assets.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} banks' non-interest-bearing reserve requirements;\renewcommand{\baselinestretch}{1}\footnote{The difference between noninterest expenses and noninterest revenue captures the influence of the economic cost of funds over and above the published interest expense and is measured by non-interest bearing reserves divided by total assets.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} the rate of default risk facing the loans of the bank (since this should imply an additional premium on the rate of the banks loans given the potential loss in bank capital);\renewcommand{\baselinestretch}{1}\footnote{This is measured by net chargeoffs divided by total loans.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} and, bank capital, since capital is costly and banks with higher capital may seek to cover this with higher lending rates.
+Note that all of the empirical analysis in the ``dealership'' model literature for studying NIMs is undertaken at the firm-level and as such many of the \textit{ad hoc} controls that are added to the model are included so as to address cross-firm differences.\renewcommand{\baselinestretch}{1}\footnote{We would note that \citeN{angbazo1994} adds default risk to the bank's profit-maximization problem and, as such, default risk does enter his model of NIMs in a more rigorous way.
+Here he finds that for money center banks NIMs are more sensitive to default risk than they are to interest-rate risk, while super-regional and regional banks are sensitive to interest-rate risk but less so to default risk.
+These findings, he notes, are consistent with the money center banks' greater concentration of short-term assets and greater use of off-balance sheet (OBS) hedging instruments.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5}
 
-We would note that Angbazo does add default risk to the bank's profit-maximization problem and, as such, default risk does enter his model of NIMs in a more rigorous way.
-Here he finds that for money center banks net interest margins are more sensitive to default risk than they are to interest-rate risk, while super-regional and regional banks are sensitive to interest-rate risk but less so to default risk.
-These findings, he notes, are consistent with the money center banks' greater concentration of short-term assets and greater use of off-balance sheet (OBS) hedging instruments.
-
-There is a further strand of the micro-banking literature looking at determinants of banks' NIMs---which includes papers by Zarruk~(1989) and Wong~(1997)---that we mention but do not discuss at length.
-In these models deposits and lending opportunities arrive at the same time and depend on the deposit and loan rates being set by banks but there is some stochastic element to the volume of deposits that banks will raise given the interest rate they set, thus requiring the banks in the model, who are risk average, to borrow or lend in the federal funds market.
+There is a further strand of the micro-banking literature looking at determinants of banks' NIMs---which includes papers by \citeN{zarruk1989} and \citeN{wong1997}---that we mention but do not discuss at length.
+In these models deposits and lending opportunities arrive at the same time and depend on the deposit and loan rates being set by banks but there is some stochastic element to the volume of deposits that banks will obtain given the interest rate they set, thus requiring the banks in the model, who are risk average, to borrow or lend in the federal funds market.
 These models also include features like administrative costs on deposits and loans as well as costs associated with deposit insurance premia and the primary use of these models is for comparative static exercises that examine how changes in these parameters---as well as considerations like uncertainty---will affect bank spreads.
-Ultimately, however, these papers are not empirical and since the majority of their theoretical findings confirm those of those of papers like Ho and Saunders, Angbazo, and Saunders and Schumacher, we do not delve too deeply into them here.
+Ultimately, however, these papers are not empirical and since the majority of their theoretical findings confirm those of those of papers like \citeN{ho1981}, \citeN{angbazo1994}, and \citeN{saunders2000}, we do not delve too deeply into them here.
 
 In summing up, in addition, to short- and long-term interest rates, a few variables stand out as relevant for our analysis.
-The degree of competition faced by the bank in the loan markets and the deposit markets in which it operates stands out as an important consideration, especially given that the sample we use in our analysis is much longer than the samples considered by Ho and Saunders, Angbazo, and Saunders and Schumacher.
+The degree of competition faced by the bank in the loan markets and the deposit markets in which it operates stands out as an important consideration, especially given that the sample we use in our analysis is much longer than the samples considered by \citeN{ho1981} and \citeN{saunders2000}.
 Interest rate volatility, also appears important, since this has likely also changed over our sample.
@@
 We are interested in forecasts for NIMs conditional on the variables in scenario $I$, which we denote $NIM_{t+s/I}$.
-Here are the methods considered:
+This section describes the methods considered.
 
-\vspace{-0.1in}
 \noindent \textbf{1. Equal-weighted forecast combination.} We regress NIMs on each yield $r(\tau )_{t-1}$ separately: \
@@
 \]
-where $N$ is the number of maturities considered.\footnote{Experimentation with alternative lag structures led to no improvement in the forecast results reported below.}
+where $N$ is the number of maturities considered.\renewcommand{\baselinestretch}{1.0}\footnote{Experimentation with alternative lag structures led to no improvement in the forecast results reported below.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5}
 
 \noindent \textbf{2.
-A dynamic factor model for NIMs.} We include NIMs in the dynamic factor model for the yield curve developed by Diebold and Li (2006).
+A dynamic factor model for NIMs.} We include NIMs in the dynamic factor model for the yield curve developed by \citeN{diebold2006}.
 Using the state space representation, the state equations of our dynamic factor model is given by:
@@
 Our specification allows the yield curve factors to depend on their own lagged values but excludes lagged dependence on NIMs.
-By contrast, NIMs the specification allows for more general dependence of NIMs on lagged yield curve factors and and lagged NIMs.
+By contrast, the specification allows for more general dependence of NIMs on lagged yield curve factors and and lagged NIMs.
 
@@
 \noindent \textbf{3.
-Dynamic factor model with second-step regression for NIMs.} This model extracts level, slope, and curvature factors for the yield curve following Diebold and Li (2006).
+Dynamic factor model with second-step regression for NIMs.} This model extracts level, slope, and curvature factors for the yield curve following \citeN{diebold2006}.
 The factors are used to forecast NIMs in a separate second-step regression.
@@
 Our dataset would allow us to start even earlier, in 1985q4, but we delay the start of the estimation sample for two reasons.
-First, there is a peculiar spike evident in aggregated NIMs in Figure \ref{figure_nims_rates} in 1988q4.
+First, there is a peculiar spike evident in aggregated NIMs in Figure~\ref{figure_nims_rates} in 1988q4.
 This large spike stems from late payments from Brazil during the Latin American debt crisis.
-The end point of 2008q2 is dictated by the attainment of the zero lower bound on nominal interest rates.\footnote{To estimate the dynamic factor models described under 2.
-and 3., we use observations between 1985q4 and 1989q4 as a training sample.}
+The end point of 2008q2 is dictated by the attainment of the zero lower bound on nominal interest rates.\renewcommand{\baselinestretch}{1.0}\footnote{To estimate the dynamic factor models described under 2.
+and 3., we use observations between 1985q4 and 1989q4 as a training sample.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5}
 
@@
 Call Report data used in this analysis are cleaned and adjusted for bank mergers and acquisitions, using structure data from the National Information Clearinghouse (NIC) on mergers and acquisitions.
-Foreign entities are excluded and domestic subsidiaries are aggregated up to the parent, bank-holding-company (BHC), level.\footnote{NIM data are adjusted for mergers between commercial banks by comparing balance sheet values of interest income, interest expenses, and interest-earning assets at the end of the quarter with those at the beginning of the quarter, accounting for amounts acquired or lost during the period because of mergers -- see the appendix in English and Nelson (1998).}
+Foreign entities are excluded and domestic subsidiaries are aggregated up to the parent, bank-holding-company (BHC), level.\renewcommand{\baselinestretch}{1.0}\footnote{NIM data are adjusted for mergers between commercial banks by comparing balance sheet values of interest income, interest expenses, and interest-earning assets at the end of the quarter with those at the beginning of the quarter, accounting for amounts acquired or lost during the period because of mergers -- see the appendix in English and Nelson (1998).\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5}
 
@@
 (2007), based on Nelson and Siegel (1987) and Svensson (1994), which allows for daily measures of an off-the-run Treasury yield curve.
-We use quarterly yields for twelve maturities in our models: 3 month, 6 month, 9 month, 1 year, 2 year, 3 year, 5 year, 7 year, 10 year, 15 year, 20 year, and 30 year.\footnote{Daily yields are published with a two-day lag under the base mnemonic SVENY at  www.federalreserve.gov/econresdata/researchdata/feds200628\_1.html.} The yields included are quarterly averages of daily yields.
+We use quarterly yields for twelve maturities in our models: 3 month, 6 month, 9 month, 1 year, 2 year, 3 year, 5 year, 7 year, 10 year, 15 year, 20 year, and 30 year.\renewcommand{\baselinestretch}{1.0}\footnote{Daily yields are published with a two-day lag under the base mnemonic SVENY at  www.federalreserve.gov/econresdata/researchdata/feds200628\_1.html.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5} The yields included are quarterly averages of daily yields.
 
@@
 
-\section{Results}
+\section{Forecasting results}
 
 To compare the performance of the models, we consider pseudo-out-of-sample forecasts.
-We use a rolling window for estimation.
+We use a recursive window for estimation.
 The shortest window spans 1989q4 to 1999q4.
@@
 increases as the forecast horizon lengthens.
-Model 2, performs uniformly worse than model 6.
+Model 2, performs uniformly worse than Model 6.
 but, excluding Model 1, it performs better than any other model we considered.
@@
 do not incorporate the information in the variance/covariance matrix of the shock in the state equation of Model 2.
-The choice of variables for the forecast combination in model 4.
+The choice of variables for the forecast combination in Model 4.
 is not as broad as for 1.
-Finally, the proliferation of parameters in the VAR of model 5.
+Finally, the proliferation of parameters in the VAR of Model 5.
 is associated with the worst performance, overall.
 
-Table \ref{table_rmse_competition} considers variants of the model that include the share of assets of the shadow banking sector.
-\footnote{In this case, we exclude Model 2.
-from the comparison.}  In no model, does that addition  improve the performance of the conditional forecasts.
+Table \ref{table_rmse_competition} considers variants of the model that include the share of assets of the shadow banking sector.\renewcommand{\baselinestretch}{1.0}\footnote{In this case, we exclude Model 2.
+from the comparison.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5}   In no model, does that addition  improve the performance of the conditional forecasts.
 
@@
 
-Figure \ref{figure_scenario} uses the forecast combination model that includes the 3-month and the 10-year yields only to forecast NIMs conditional on the three CCAR 2013 supervisory scenarios.
-The severely adverse scenario implies a downward shift of the profile of the yield curve accompanied by a sharp flattening.
-The severe scenario implies an upward shift of the yield curve with little change in the slope.
+\section{Scenario analysis}
+
+Figure~\ref{figure_scenario} uses the forecast combination model that includes the 3-month and the 10-year yields only to forecast NIMs conditional on the three Dodd-Frank Act 2013 supervisory scenarios.\renewcommand{\baselinestretch}{1.0}\footnote{See, ``2013 Supervisory Scenarios for Annual Stress Tests Required under the Dodd-Frank Act Stress Testing Rules and the Capital Plan Rule'' available at http://www.federalreserve.gov/bankinforeg/bcreg20121115a1.pdf.\vspace{0.05in}}\renewcommand{\baselinestretch}{1.5}    The severely adverse scenario---which features a severe global recession---implies a downward shift of the profile of the yield curve accompanied by a sharp flattening.
+The adverse scenario---which is motivated by a sudden jump in inflation that accompanies a moderate recession---implies an upward shift of the yield curve with a modest flattening in the slope.
 In terms of changes in the yield curve, these two scenarios capture vastly different configurations.
@@
 By contrast, the upward parallel shift in the yield curve is predicted to boost NIMs.
+The figure also shows the path of NIMs under the baseline, which trends down for all of the scenario horizon.
 
-The figure, provides a visual interpretation of the RMSEs considered thus far.
-The bands surrounding the central forecast are constructed adding and subtracting to the point forecast the RMSEs for the forecast combination model.
+The lower panels of figure~\ref{figure_scenario} also consider the uncertainty of the forecasts.
+In addition, to showing point estimates of the paths of NIMs, the figure also shows bands surrounding the central forecast, which are constructed by adding and subtracting to the point forecast the RMSEs for the forecast combination model.
 Strikingly, the size of the uncertainty bands dwarfs the variation implied by the scenarios for NIMs.
@@
 In that case, the effects of confidence of stress tests rely squarely on the reputation of the supervising authority, one which might be tarnished exactly in a period of crisis.
+
+\normalsize \clearpage
+\renewcommand{\baselinestretch}{1}
+
+\bibliographystyle{chicago}
+\bibliography{biblio}
+
+\renewcommand{\baselinestretch}{1.5}
 

### 2013-08-19 LATEX/workplan.aug.18.tex
New file with suggestions of what we should do next.

\documentclass[11pt]{article}
\addtolength{\topmargin}{-0.375in}
\addtolength{\oddsidemargin}{-0.35in}
\addtolength{\evensidemargin}{-0.35in}
\addtolength{\textheight}{0.5in}
\addtolength{\textwidth}{1.1in}
\addtolength{\parskip}{0.05in}
\renewcommand{\baselinestretch}{1.1}
\newcommand{\vsp}{\vspace{\baselineskip}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\gr}[1]{\ensuremath  {\frac{\dot {#1}}{#1}}}

\begin{document}

\noindent

\textbf{First just the time series model (i.e., not the panel models)}

\textbf{Models with yield or factors based on yield only}

\begin{enumerate}
\item No change forecast 
\begin{itemize}
\item I'm listing this first because I think the tables should only report RMSEs for the no change forecast and show all other RMSE's as relative to the no change forecast's RMSE.  
\item It may also be worth having a table reporting Barbara's full sample $A$ measures, just to set things up for later.  From an ordering perspective they should be the same as reporting relative RMSEs.  That is, while relative RMSEs are $\sqrt{\sum \epsilon^{2}_{alt.mod.}}/\sqrt{\sum \epsilon^{2}_{no.chg}}$, Barbara's $A$ measure is  $\sum \epsilon^{2}_{alt.mod.}-\sum \epsilon^{2}_{no.chg}$.
\item Additionally, it may also be work having some charts reporting rolling window relative RMSEs and rolling window versions of Barbara's $A$ measure.  
\begin{itemize}
\item It would probably make sense to do this only for a selection of models (i.e., models that are different enough).  
\item Doing the rolling window versions of Barbara's $A$ measure would set things up later for doing the hand-wavy version of Barbara's decomposition. (Recall, that we can not do a proper version of Barbara's decomposition because she does it for DAR forecasts and while we could do DAR forecasts they do not really make sense when we want to do conditional forecasts.)  
\end{itemize}
\end{itemize}
\item Models with yields
\begin{enumerate}
\item Equally-weighted average (EWA) of 12 recursive autoregression (RAR) models that individual includes all 12 yields 
\item EWA of 3 RAR models that individually include the 3-month rate, 2-year yield, and 10-year yield
\item EWA of 2 RAR models that individually include the 3-month rate and 10-year yield
\end{enumerate}
\item Models with data-based factors ($level=10.year.yield$, $slope=10.year.yield–-3.month.rate$, and $curvature=2*2.year.yield–-3.month.rate–-10.year.yield$)
\begin{enumerate}
\item RAR model that jointly includes the 3 data-based factors
\item EWA of 3 RAR models that individually include the 3 data-based factors
\end{enumerate}
\item Models with PCA-based factors (1st PC, 2nd PC, and 3rd PC)
\begin{enumerate}
\item RAR model that jointly includes the first 3 PCA-based factors
\item EWA of 3 RAR models that individually include the first 3 PCA-based factors
\end{enumerate}
\item Models with Nelson-Siegel model-based factors (level, slope, and curvature)
\begin{enumerate}
\item RAR model that jointly includes the 3 Nelson-Siegel-based factors
\item EWA of 3 RAR models that individually include the 3 Nelson-Siegel-based factors
\end{enumerate}
\item Models with PLS-based factors
\begin{enumerate}
\item RAR model that jointly includes the first 3 PLS-based factors
\item EWA of 3 RAR models that individually include the first 3 PLS-based factors
\item RAR model that includes the first PLS-based factor
\end{enumerate}
\item Should we include the system approaches that we have already done
\begin{enumerate}
\item The VAR model with NIMs and the data-based factors
\item The VAR model with NIMs and the Nelson-Siegel model-based factors
\end{enumerate}
\item Bayesian model averaging – should we pursue?
\end{enumerate}

\textbf{Models with yield or factors based on yield and banking variables}

\begin{itemize}
\item Models listed above plus a measure of banking competition (the shadow banking sector's share of intermediation)
\item Models listed above plus a measure of interest rate volatility (i.e., the MOVE index or the volatility measures from the banking literature)
\item I think that ultimately these additions to the model fail to improve forecast performance so these results are really just for completeness and in recognition of the previous banking literature on NIMs.
\end{itemize}

\textbf{Model estimation period}

\begin{itemize}
\item Expanding windows [this is what we have been doing]
\item Fixed rolling windows
\end{itemize}

\textbf{Type of forecasting model}

\begin{itemize}
\item Recursive autoregression (RAR) forecasting models [this is what we have been using]
\item Direct autoregression (DAR) forecasting models
\end{itemize}

\textbf{Forecast evaluation period}

\begin{itemize}
\item Out-of-sample [this is what we have been doing] 
\item In-sample
\end{itemize}

\textbf{Barbara's method for comparing two models}

\begin{itemize}
\item Barbara's method applies to DAR forecasts; that is, $y_{t+h}=\alpha_{0}x_{t}+\alpha_{1}x_{t-1}+\cdots+\epsilon^{\alpha}_{t}$ and $y_{t+h}=\gamma_{0}z_{t}+\gamma_{1}z_{t-1}+\cdots+\epsilon^{\gamma}_{t}$, and not forecasts generated recursively (i.e., RAR forecasts).  Most of the paper works with rolling window estimation periods, though in the last paragraph of section~3 she says that everything goes through with expanding estimation periods and the discussion is in the appendix (though it sounds like there is some adjustment needed for variances when doing the decomposition).
\item Models A and G can be estimated for different step-ahead equations and over fixed rolling windows and used to forecast out-of-sample.  This will imply for every period and every h-step ahead forecast a squared forecast error $\left(\overrightarrow{\eta}^{\alpha}_{t+h}\right)^{2}$ and
 $\left(\overrightarrow{\eta}^{\gamma}_{t+h}\right)^{2}$.  (I'm using ``over-arrows'' to denote out-of-sample forecasts.  This is my notation, not Barbara's. I realize that over-arrows typically mean vectors but I find it more intuitive to denote out-of-sample forecasts with an arrow and in-sample with a bar.)
\begin{itemize}
\item For one-step ahead we have $\left\{\left(\overrightarrow{\eta}^{\alpha}_{R+i+1|R+i}\right)^{2}\right\}_{i=0}^{T-R-1}$ and $\left\{\left(\overrightarrow{\eta}^{\gamma}_{R+i+1|R+i}\right)^{2}\right\}_{i=0}^{T-R-1}$, where in each case the rolling window estimation period \underline{and data used to generate the forecast}
ends at $R, R+1, \dots, T-1$.
\item For two-step ahead we have $\left\{\left(\overrightarrow{\eta}^{\alpha}_{R+i+2|R+i}\right)^{2}\right\}_{i=0}^{T-R-2}$ and $\left\{\left(\overrightarrow{\eta}^{\gamma}_{R+i+2||R+i}\right)^{2}\right\}_{i=0}^{T-R-2}$, where in each case the rolling window estimation period \underline{and data used to generate the forecast} ends at $R, R+1, \dots, T-2$.
\item $\cdots$
\item $\cdots$
\item For h-step ahead we have $\left\{\left(\overrightarrow{\eta}^{\alpha}_{R+i+h||R+i}\right)^{2}\right\}_{i=0}^{T-R-h}$ and $\left\{\left(\overrightarrow{\eta}^{\gamma}_{R+i+h||R+i}\right)^{2}\right\}_{i=0}^{T-R-h}$, where in each case the rolling window estimation period \underline{and data used to generate the forecast} ends at $R, R+1, \dots, T-h$.
\item Note there is a different model estimated for each one-step, two-step, ..., h-step ahead forecast and a different model estimated for each date.  
\end{itemize}
\item One can then take the difference of these squared forecast errors $\left(\overrightarrow{\eta}^{\alpha}_{t+h}\right)^{2}$ and
 $\left(\overrightarrow{\eta}^{\gamma}_{t+h}\right)^{2}$, which I'll call $\overrightarrow{\mathcal{RL}}_{t+h}$ for ``rolling loss.''
\begin{itemize}
\item For one-step ahead we have $\left\{\overrightarrow{\mathcal{RL}}_{R+i+1|R+i}\right\}_{i=0}^{T-R-1}\!\!\!\!=\left\{\left(\overrightarrow{\eta}^{\alpha}_{R+i+1|R+i}\right)^{2}\!\!-\left(\overrightarrow{\eta}^{\gamma}_{R+i+1|R+i}\right)^{2}\right\}_{i=0}^{T-R-1}$
\item For two-step ahead we have $\left\{\overrightarrow{\mathcal{RL}}_{R+i+2|R+i}\right\}_{i=0}^{T-R-2}\!\!\!\!=\left\{\left(\overrightarrow{\eta}^{\alpha}_{R+i+2|R+i}\right)^{2}\!\!-\left(\overrightarrow{\eta}^{\gamma}_{R+i+2|R+i}\right)^{2}\right\}_{i=0}^{T-R-2}$
\item $\cdots$
\item $\cdots$
\item For h-step ahead we have $\left\{\overrightarrow{\mathcal{RL}}_{R+i+h|R+i}\right\}_{i=0}^{T-R-h}\!\!\!\!=\left\{\left(\overrightarrow{\eta}^{\alpha}_{R+i+h|R+i}\right)^{2}\!\!-\left(\overrightarrow{\eta}^{\gamma}_{R+i+h|R+i}\right)^{2}\right\}_{i=0}^{T-R-h}$
\end{itemize}
\item So for any date after $R+h$ there are $h$ rolling loss $\overrightarrow{\mathcal{RL}}$ observations, corresponding to different step ahead forecasts.  For date $R+h-1$ there are $h-1$ rolling loss $\overrightarrow{\mathcal{RL}}$ observations, and so on until $R+1$ where there is just one rolling loss $\overrightarrow{\mathcal{RL}}$ observation.
\item Models A and G can be estimated over fixed rolling windows and used to forecast in-sample.  Note that whereas when one forecasts out-of-sample there is for each estimated model only 1 one-step ahead forecast, only 1 two-step ahead forecast, ..., and only 1 h-step ahead forecast, when one forecasts in-sample there is for each fixed rolling window of length $R$, $R-1$ one-step ahead forecasts, $R-2$ two-step ahead forecasts, ..., and $R-h$ h-step ahead forecasts.  When looking at in-sample forecasts errors Barbara only looks at the \emph{very last} one-step, two-step, ..., h-step ahead forecast for that model.  
\item Models A and G estimated over fixed rolling windows and used to forecast in-sample will for every period and every h-step ahead forecast imply a squared forecast error for the last period of the estimation period $\left(\overline{\eta}^{\alpha}_{t}\right)^{2}$ and
 $\left(\overline{\eta}^{\gamma}_{t}\right)^{2}$.  (I'm using ``over arrows'' to denote out-of-sample forecasts.  This is my notation, not Barbara's.)
\begin{itemize}
\item For one-step ahead we have $\left\{\left(\overline{\eta}^{\alpha}_{R+i|R+i-1}\right)^{2}\right\}_{i=0}^{T-R}$ and $\left\{\left(\overline{\eta}^{\gamma}_{R+i|R+i-1}\right)^{2}\right\}_{i=0}^{T-R}$, where in each case the rolling window estimation period
ends at $R, R+1, \dots, T$ but the data used to generate the forecast ends in $R-1, R, \dots, T-1$.
\item For two-step ahead we have $\left\{\left(\overline{\eta}^{\alpha}_{R+i|R+i-2}\right)^{2}\right\}_{i=0}^{T-R}$ and $\left\{\left(\overline{\eta}^{\gamma}_{R+i|R+i-2}\right)^{2}\right\}_{i=0}^{T-R}$, where in each case the rolling window estimation period
ends at $R, R+1, \dots, T$ but the data used to generate the forecast ends in $R-2, R-1, \dots, T-2$.
\item $\cdots$
\item $\cdots$
\item For h-step ahead we have $\left\{\left(\overline{\eta}^{\alpha}_{R+i|R+i-h}\right)^{2}\right\}_{i=0}^{T-R}$ and $\left\{\left(\overline{\eta}^{\gamma}_{R+i|R+i-h}\right)^{2}\right\}_{i=0}^{T-R}$, where in each case the rolling window estimation period
ends at $R, R+1, \dots, T$ but the data used to generate the forecast ends in $R-h, R-h+1, \dots, T-h$.
\end{itemize}
\item One can then take the difference of these squared forecast errors $\left(\overline{\eta}^{\alpha}_{t}\right)^{2}$ and
 $\left(\overline{\eta}^{\gamma}_{t}\right)^{2}$, which I'll call $\overline{\mathcal{RL}}_{t+h}$ for ``rolling loss.''
\begin{itemize}
\item For one-step ahead we have $\left\{\overline{\mathcal{RL}}_{R+i|R+i-1}\right\}_{i=0}^{T-R-1}\!\!\!\!=\left\{\left(\overline{\eta}^{\alpha}_{R+i|R+i-1}\right)^{2}\!\!-\left(\overline{\eta}^{\gamma}_{R+i|R+i-1}\right)^{2}\right\}_{i=0}^{T-R-1}$
\item For two-step ahead we have $\left\{\overline{\mathcal{RL}}_{R+i|R+i-2}\right\}_{i=0}^{T-R-2}\!\!\!\!=\left\{\left(\overline{\eta}^{\alpha}_{R+i|R+i-2}\right)^{2}\!\!-\left(\overline{\eta}^{\gamma}_{R+i|R+i-2}\right)^{2}\right\}_{i=0}^{T-R-2}$
\item $\cdots$
\item $\cdots$
\item For h-step ahead we have $\left\{\overline{\mathcal{RL}}_{R+i|R+i-h}\right\}_{i=0}^{T-R-h}\!\!\!\!=\left\{\left(\overline{\eta}^{\alpha}_{R+i|R+i-h}\right)^{2}\!\!-\left(\overline{\eta}^{\gamma}_{R+i|R+i-h}\right)^{2}\right\}_{i=0}^{T-R-h}$
\end{itemize}
\item Then one starts regressing different losses on each other.
\begin{itemize}
\item So for the one-step ahead forecast we regress $\left\{\overrightarrow{\mathcal{RL}}_{R+i+1|R+i}\right\}_{i=0}^{T-R-1}$ on $\left\{\overline{\mathcal{RL}}_{R+i|R+i-1}\right\}_{i=0}^{T-R-1}$.  The models being used to generate these losses are estimated one rolling windows that end on $R, R+1, \cdots, T$.
\item So for the two-step ahead forecast we regress $\left\{\overrightarrow{\mathcal{RL}}_{R+i+2|R+i}\right\}_{i=0}^{T-R-1}$ on $\left\{\overline{\mathcal{RL}}_{R+i|R+i-2}\right\}_{i=0}^{T-R-1}$.  The models being used to generate these losses are estimated one rolling windows that end on $R, R+1, \cdots, T$.
\item $\cdots$
\item $\cdots$
\item So for the h-step ahead forecast we regress $\left\{\overrightarrow{\mathcal{RL}}_{R+i+h|R+i}\right\}_{i=0}^{T-R-1}$ on $\left\{\overline{\mathcal{RL}}_{R+i|R+i-h}\right\}_{i=0}^{T-R-1}$.  The models being used to generate these losses are estimated one rolling windows that end on $R, R+1, \cdots, T$.
\end{itemize}
\item The regressions are:
\begin{itemize}
\item One-step: $\overrightarrow{\mathcal{RL}}_{R+i+1|R+i}=\widehat{\beta}_{1}\overline{\mathcal{RL}}_{R+i|R+i-1}+\widehat{u}_{R+i+1}$, where $i=0, 1, \cdots, T-R-1$.
\item Two-step: $\overrightarrow{\mathcal{RL}}_{R+i+2|R+i}=\widehat{\beta}_{2}\overline{\mathcal{RL}}_{R+i|R+i-2}+\widehat{u}_{R+i+2}$, where $i=0, 1, \cdots, T-R-2$.
\item $\cdots$
\item $\cdots$
\item h-step: $\overrightarrow{\mathcal{RL}}_{R+i+h|R+i}=\widehat{\beta}_{h}\overline{\mathcal{RL}}_{R+i|R+i-h}+\widehat{u}_{R+i+h}$, where $i=0, 1, \cdots, T-R-h$.
\item It may be worth noting that every out-of-sample-based loss being regressed on a corresponding in-sample-based loss uses exactly the same two versions -- where versions refer to the estimation period -- of the A and G model.  That is the models are all estimated on rolling windows ending $R, R+1, \cdots, T$ but the in-sample forecasts use this estimated model but condition on earlier data for period $R-step, R-step+1, \cdots, T-step$ forecasts while the out-of-sample forecasts use this estimated model, condition on $R, R+1, \cdots, T-step$ data and project out  $R+step, R+step+1, \cdots, T+step-step$.
\end{itemize}
\item From the regressions we can calculate:
\begin{itemize}
\item One-step: 
\begin{eqnarray}
\underbrace{\frac{1}{T\!-\!R}\!\!\sum_{i=0}^{T-R-1}\!\!\overrightarrow{\mathcal{RL}}_{R+i+1|R+i}}_{=A_{1,T\!-\!R}}
\!\!\!\!&=&\!\!\!\!\underbrace{\widehat{\beta}_{1}\cdot\frac{1}{T\!-\!R}\!\!\sum_{i=0}^{T-R-1}\!\!\overline{\mathcal{RL}}_{R+i|R+i-1}}_{=B_{1,T\!-\!R}} \nonumber \\
\!\!\!\!&+&\!\!\!\!\underbrace{\frac{1}{T\!-\!R}\!\!\sum_{i=0}^{T-R-1}\!\!\widehat{u}_{R+i+1}}_{=U_{1,T\!-\!R}}, \nonumber
\end{eqnarray}
where $i=0, 1, \cdots, T-R-1$.
\item Two-step: 
\begin{eqnarray}
\underbrace{\frac{1}{T\!-\!R\!-\!1}\!\!\sum_{i=0}^{T-R-2}\!\!\overrightarrow{\mathcal{RL}}_{R+i+2|R+i}}_{=A_{2,T\!-\!R\!-\!1}}
\!\!\!\!&=&\!\!\!\!\underbrace{\widehat{\beta}_{2}\cdot\frac{1}{T\!-\!R\!-\!1}\!\!\sum_{i=0}^{T-R-2}\!\!\overline{\mathcal{RL}}_{R+i|R+i-2}}_{=B_{2,T\!-\!R\!-\!1}} \nonumber \\
\!\!\!\!&+&\!\!\!\!\underbrace{\frac{1}{T\!-\!R\!-\!1}\!\!\sum_{i=0}^{T-R-2}\!\!\widehat{u}_{R+i+2}}_{=U_{2,T\!-\!R\!-\!1}}, \nonumber 
\end{eqnarray}
where $i=0, 1, \cdots, T-R-2$.
\item $\cdots$
\item $\cdots$
\item h-step: 
\begin{eqnarray}
\underbrace{\frac{1}{T\!-\!R\!-\!(h\!-\!1)}\!\!\sum_{i=0}^{T-R-(h-1)}\!\!\overrightarrow{\mathcal{RL}}_{R+i+h|R+i}}_{=A_{h,T\!-\!R\!-\!\!(h\!-\!1)}}
\!\!\!\!&=&\!\!\!\!\underbrace{\widehat{\beta}_{h}\cdot\frac{1}{T\!-\!R\!-\!(h\!-\!1)}\!\!\sum_{i=0}^{T-R-(h-1)}\!\!\overline{\mathcal{RL}}_{R+i|R+i-h}}_{=B_{h,T\!-\!R\!-\!(h\!-\!1)}} \nonumber \\
\!\!\!\!&+&\!\!\!\!\underbrace{\frac{1}{T\!-\!R\!-\!(h\!-\!1)}\!\!\sum_{i=0}^{T-R-(h-1)}\!\!\widehat{u}_{R+i+h}}_{=U_{h,T\!-\!R\!-\!(h\!-\!1)}}, \nonumber
\end{eqnarray}
where $i=0, 1, \cdots, T-R-h$.
\end{itemize}
\item The interpretation of the $A$, $B$, and $U$ terms are as follows.
\begin{itemize}
\item $A_{1,T\!-\!R}$, $A_{2,T\!-\!R\!-\!1}$, ..., $A_{h,T\!-\!R\!-\!\!(h\!-\!1)}$ measures the two models' relative out-of-sample forecasting ability \underline{on average} over the \underline{full} out-of-sample forecast evaluation period.
\item $B_{1,T\!-\!R}$, $B_{2,T\!-\!R\!-\!1}$, ..., $B_{h,T\!-\!R\!-\!\!(h\!-\!1)}$ are proportional to the two models' relative in-sample forecasting ability \underline{on average} over the \underline{full} in-sample forecast evaluation period.  This represents the part of the two models' out-of-sample relative forecasting ability that is reflected in in-sample forecasting ability.  When the $B$ terms have the same sign as their corresponding $A$ term it means that in-sample relative forecast performance has predictive content for out-of-sample relative performance.  If they have the opposite sign, there is predictive content but it is misleading.
\item $U_{1,T\!-\!R}$, $U_{2,T\!-\!R\!-\!1}$, ..., $U_{h,T\!-\!R\!-\!\!(h\!-\!1)}$ represents the part of the two models' out-of-sample relative forecasting ability that is \underline{not} reflected in in-sample forecasting ability. 
\end{itemize}
\item Barbara also wanted to look at how the two models' relative out-of-sample forecasting ability varies across the full out-of-sample forecast evaluation period so they thought about rolling windows for:
\begin{eqnarray}
A_{1,T\!-\!R}\!\!\!\!&=&\!\!\!\!\frac{1}{T\!-\!R}\!\!\sum_{i=0}^{T-R-1}\!\!\overrightarrow{\mathcal{RL}}_{R+i+1|R+i} \nonumber \\
A_{2,T\!-\!R\!-\!1}\!\!\!\!&=&\!\!\!\!\frac{1}{T\!-\!R\!-\!1}\!\!\sum_{i=0}^{T-R-2}\!\!\overrightarrow{\mathcal{RL}}_{R+i+2|R+i} \nonumber \\
\!\!\!\!&&\!\!\!\!\cdots  \nonumber \\
\!\!\!\!&&\!\!\!\!\cdots  \nonumber \\
A_{h,T\!-\!R\!-\!\!(h\!-\!1)}\!\!\!\!&=&\!\!\!\!\frac{1}{T\!-\!R\!-\!(h\!-\!1)}\!\!\sum_{i=0}^{T-R-(h-1)}\!\!\overrightarrow{\mathcal{RL}}_{R+i+h|R+i}. \nonumber
\end{eqnarray}
\item If the moving windows are of length $m$ then:
\begin{itemize}
\item For $A_{1,T\!-\!R}$ there are $T-R-m$ possible windows, where the windows are: 
\[\left\{\overrightarrow{\mathcal{RL}}_{R+i+1|R+i}\ \ , \ \  . \ \  . \ \  . \ \ ,\ \ \overrightarrow{\mathcal{RL}}_{R+i+1+m|R+i+m}\right\}_{i=0}^{T-R-1-m}\] 
and so the rolling window relative out-of-sample forecast performance terms would be:
\[\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+1+j|R+i+j}\ \ \ \mathrm{for} \ \ \ i=0,1,\cdots,T\!-\!R\!-\!1\!-\!m. \]
\item For $A_{2,T\!-\!R\!-\!1}$ there are $T-R-m-1$ possible windows, where the windows are: 
\[\left\{\overrightarrow{\mathcal{RL}}_{R+i+2|R+i}\ \ , \ \  . \ \  . \ \  . \ \ ,\ \ \overrightarrow{\mathcal{RL}}_{R+i+2+m|R+i+m}\right\}_{i=0}^{T-R-2-m}\] 
and so the rolling window relative out-of-sample forecast performance terms would be:
\[\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+2+j|R+i+j}\ \ \ \mathrm{for} \ \ \ i=0,1,\cdots,T\!-\!R\!-\!2\!-\!m. \]
\item And so on for all the other steps until we get to the $h$-th step
\item For $A_{h,T\!-\!R\!-\!(h-1)}$ there are $T-R-m-(h-1)$ possible windows, where the windows are: 
\[\left\{\overrightarrow{\mathcal{RL}}_{R+i+h|R+i}\ \ , \ \  . \ \  . \ \  . \ \ ,\ \ \overrightarrow{\mathcal{RL}}_{R+i+h+m|R+i+m}\right\}_{i=0}^{T-R-h-m}\] 
and so the rolling window relative out-of-sample forecast performance terms would be:
\[\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+h+j|R+i+j}\ \ \ \mathrm{for} \ \ \ i=0,1,\cdots,T\!-\!R\!-\!h\!-\!m. \]
\end{itemize}
\item So then it is the rolling window relative out-of-sample forecast performances -- that is, $\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+1+j|R+i+j}$, $\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+2+j|R+i+j}$, . . . , $\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+h+j|R+i+j}$ -- that one wants to decompose.  These are done as follows.
\begin{eqnarray}
\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+1+j|R+i+j}\!\!\!\!&=&\!\!\!\!\underbrace{\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+1+j|R+i+j}-\underbrace{\frac{1}{T\!-\!R}\!\!\sum_{i=0}^{T-R-1}\!\!\overrightarrow{\mathcal{RL}}_{R+i+1|R+i}}_{=A_{1,T\!-\!R}}}_{=A_{i,1,T-R}} \nonumber \\
\!\!\!\!&+&\!\!\!\!B_{1,T-R}+U_{1,T-R} \ \ \ \ \   \ \ \ \ \   \ \ \ \ \   \mathrm{for} \ i=0,1,\cdots,T\!-\!R\!-\!1\!-\!m. \nonumber
\end{eqnarray}
\begin{eqnarray}
\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+2+j|R+i+j}\!\!\!\!&=&\!\!\!\!\underbrace{\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+2+j|R+i+j}-\underbrace{\frac{1}{T\!-\!R\!-\!1}\!\!\sum_{i=0}^{T-R-2}\!\!\overrightarrow{\mathcal{RL}}_{R+i+2|R+i}}_{=A_{2,T\!-\!R\!-\!1}}}_{=A_{i,2,T-R-1}} \nonumber \\
\!\!\!\!&+&\!\!\!\!B_{2,T-R-1}+U_{2,T-R-1} \ \ \ \ \   \ \ \ \ \   \mathrm{for} \ i=0,1,\cdots,T\!-\!R\!-\!2\!-\!m. \nonumber
\end{eqnarray}
\[ \cdots \]
\[ \cdots \]
\begin{eqnarray}
\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+h+j|R+i+j}\!\!\!\!&=&\!\!\!\!\underbrace{\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+h+j|R+i+j}-\underbrace{\frac{1}{T\!-\!R\!-\!(h\!-\!1)}\!\!\sum_{i=0}^{T-R-(h-1)}\!\!\overrightarrow{\mathcal{RL}}_{R+i+h|R+i}}_{=A_{h,T\!-\!R-(h-1)}}}_{=A_{i,h,T-R-(h-1)}} \nonumber \\
\!\!\!\!&+&\!\!\!\!B_{h,T-R-(h-1)}+U_{h,T-R-(h-1)} \ \ \ \ \   \ \ \ \ \   \mathrm{for} \ i=0,1,\cdots,T\!-\!R\!-\!h\!-\!m. \nonumber
\end{eqnarray}
In summary we have
\begin{eqnarray}
\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+1+j|R+i+j}\!\!\!\!&=&\!\!\!\!A_{i,1,T-R}+B_{1,T-R}+U_{1,T-R} \nonumber \\
\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+2+j|R+i+j}\!\!\!\!&=&\!\!\!\!A_{i,2,T-R-1}+B_{2,T-R-1}+U_{2,T-R-1} \nonumber \\
\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+h+j|R+i+j}\!\!\!\!&=&\!\!\!\!A_{i,h,T-R-(h-1)}+B_{h,T-R-(h-1)}+U_{h,T-R-(h-1)} \nonumber
\end{eqnarray}
\item So basically one calculates the rolling window relative out-of-sample forecast performances -- that is, $\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+1+j|R+i+j}$, $\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+2+j|R+i+j}$, . . . , $\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+h+j|R+i+j}$ -- and tries to understand them in terms of the above components. 
\begin{itemize}
\item  I think what is interesting about looking at these is that the relative out-of-sample forecasting performance of the two models could change over time.  That is, over some windows one of the two models may forecast better where as over other windows the other of the two models may forecast better. 
\end{itemize}
\item When Barbara does the exchange rate forecasting example the ultimate finding is not so interesting, since basically she just reports that ``the lack of out-of-sample predictive ability is related to the lack of in-sample predictive ability.''
\item We should be able to do something like Barbara's decomposition because we can generate in-sample forecasts (and one only ends up using the in-sample forecast errors for the very last period of the sample) but it is not clear -- indeed very unlikely -- that the components that we decompose the rolling relative out-of-sample into will be independent and that all the proofs that Barbara does will go through.
\end{itemize}

\textbf{Panel models}
\begin{itemize}
\item  Possible models: 
\begin{itemize}
\item Panel model
\item Bank-specific equations with SUR
\item Swamy random coefficients model?
\end{itemize}
\item Evaluate forecasts but also compare them to the SNL analysts forecasts.
\end{itemize}
\end{document}               % End of document.

### 2013-09-05 LATEX/workplan.aug.18.tex
Workplan explaining Rossi et al (2013) and PLS paper Slides in progress for Atlanta

@@
 \begin{itemize}
-\item I'm listing this first because I think the tables should only report RMSEs for the no change forecast and show all other RMSE's as relative to the no change forecast's RMSE.  
-\item It may also be worth having a table reporting Barbara's full sample $A$ measures, just to set things up for later.
-From an ordering perspective they should be the same as reporting relative RMSEs.
-That is, while relative RMSEs are $\sqrt{\sum \epsilon^{2}_{alt.mod.}}/\sqrt{\sum \epsilon^{2}_{no.chg}}$, Barbara's $A$ measure is  $\sum \epsilon^{2}_{alt.mod.}-\sum \epsilon^{2}_{no.chg}$.
-\item Additionally, it may also be work having some charts reporting rolling window relative RMSEs and rolling window versions of Barbara's $A$ measure.  
+\item I'm listing this first because I think we should also show tables of relative RSMEs and for these tables I think the relevant comparison should be relative to the no change forecast.    
+\begin{itemize}
+\item The first table could show all the RSMEs in absolute terms (as they are now shown).  
+\item The second table could show just the RSME for the no change forecast and then relative RSMEs for all other approaches; that is, $\sqrt{\sum \epsilon^{2}_{alt.mod.}}/\sqrt{\sum \epsilon^{2}_{no.chg}}$.
+\item The third table could show just the RSME for the no change forecast and then show Barbara's (relative) full sample $A$ measures, defined as $\sum \epsilon^{2}_{alt.mod.}-\sum \epsilon^{2}_{no.chg}$, just to set things up for later.
+From an ordering perspective Barbara's full sample $A$ measures should be the same as reporting relative RMSEs.
+\end{itemize}
+\item Additionally, it may also be worth having some charts reporting rolling window relative RMSEs and rolling window versions of Barbara's $A$ measure.  
 \begin{itemize}
@@
 \item Models A and G can be estimated for different step-ahead equations and over fixed rolling windows and used to forecast out-of-sample.
-This will imply for every period and every h-step ahead forecast a squared forecast error $\left(\overrightarrow{\eta}^{\alpha}_{t+h}\right)^{2}$ and
+This will imply for every period and every $h$-step ahead forecast a squared forecast error $\left(\overrightarrow{\eta}^{\alpha}_{t+h}\right)^{2}$ and
  $\left(\overrightarrow{\eta}^{\gamma}_{t+h}\right)^{2}$.
@@
 \begin{itemize}
-\item For one-step ahead we have $\left\{\left(\overrightarrow{\eta}^{\alpha}_{R+i+1|R+i}\right)^{2}\right\}_{i=0}^{T-R-1}$ and $\left\{\left(\overrightarrow{\eta}^{\gamma}_{R+i+1|R+i}\right)^{2}\right\}_{i=0}^{T-R-1}$, where in each case the rolling window estimation period \underline{and data used to generate the forecast}
+\item For $1$-step ahead we have $\left\{\left(\overrightarrow{\eta}^{\alpha}_{R+i+1|R+i}\right)^{2}\right\}_{i=0}^{T-R-1}$ and $\left\{\left(\overrightarrow{\eta}^{\gamma}_{R+i+1|R+i}\right)^{2}\right\}_{i=0}^{T-R-1}$, where the rolling window estimation period \underline{and data used to generate the forecast}
 ends at $R, R+1, \dots, T-1$.
-\item For two-step ahead we have $\left\{\left(\overrightarrow{\eta}^{\alpha}_{R+i+2|R+i}\right)^{2}\right\}_{i=0}^{T-R-2}$ and $\left\{\left(\overrightarrow{\eta}^{\gamma}_{R+i+2||R+i}\right)^{2}\right\}_{i=0}^{T-R-2}$, where in each case the rolling window estimation period \underline{and data used to generate the forecast} ends at $R, R+1, \dots, T-2$.
+\item For $2$-step ahead we have $\left\{\left(\overrightarrow{\eta}^{\alpha}_{R+i+2|R+i}\right)^{2}\right\}_{i=0}^{T-R-2}$ and $\left\{\left(\overrightarrow{\eta}^{\gamma}_{R+i+2||R+i}\right)^{2}\right\}_{i=0}^{T-R-2}$, where the rolling window estimation period \underline{and data used to generate the forecast} ends at $R, R+1, \dots, T-2$.
 \item $\cdots$
 \item $\cdots$
-\item For h-step ahead we have $\left\{\left(\overrightarrow{\eta}^{\alpha}_{R+i+h||R+i}\right)^{2}\right\}_{i=0}^{T-R-h}$ and $\left\{\left(\overrightarrow{\eta}^{\gamma}_{R+i+h||R+i}\right)^{2}\right\}_{i=0}^{T-R-h}$, where in each case the rolling window estimation period \underline{and data used to generate the forecast} ends at $R, R+1, \dots, T-h$.
-\item Note there is a different model estimated for each one-step, two-step, ..., h-step ahead forecast and a different model estimated for each date.  
+\item For $h$-step ahead we have $\left\{\left(\overrightarrow{\eta}^{\alpha}_{R+i+h||R+i}\right)^{2}\right\}_{i=0}^{T-R-h}$ and $\left\{\left(\overrightarrow{\eta}^{\gamma}_{R+i+h||R+i}\right)^{2}\right\}_{i=0}^{T-R-h}$, where the rolling window estimation period \underline{and data used to generate the forecast} ends at $R, R+1, \dots, T-h$.
+\item Note there is a different model estimated for each $1$-step, $2$-step, ..., $h$-step ahead forecast and a different model estimated for each date.  
 \end{itemize}
@@
 \begin{itemize}
-\item For one-step ahead we have $\left\{\overrightarrow{\mathcal{RL}}_{R+i+1|R+i}\right\}_{i=0}^{T-R-1}\!\!\!\!=\left\{\left(\overrightarrow{\eta}^{\alpha}_{R+i+1|R+i}\right)^{2}\!\!-\left(\overrightarrow{\eta}^{\gamma}_{R+i+1|R+i}\right)^{2}\right\}_{i=0}^{T-R-1}$
-\item For two-step ahead we have $\left\{\overrightarrow{\mathcal{RL}}_{R+i+2|R+i}\right\}_{i=0}^{T-R-2}\!\!\!\!=\left\{\left(\overrightarrow{\eta}^{\alpha}_{R+i+2|R+i}\right)^{2}\!\!-\left(\overrightarrow{\eta}^{\gamma}_{R+i+2|R+i}\right)^{2}\right\}_{i=0}^{T-R-2}$
+\item For $1$-step ahead: $\left\{\overrightarrow{\mathcal{RL}}_{R+i+1|R+i}\right\}_{i=0}^{T-R-1}\!\!\!\!=\left\{\left(\overrightarrow{\eta}^{\alpha}_{R+i+1|R+i}\right)^{2}\!\!-\left(\overrightarrow{\eta}^{\gamma}_{R+i+1|R+i}\right)^{2}\right\}_{i=0}^{T-R-1}$
+\item For $2$-step ahead: $\left\{\overrightarrow{\mathcal{RL}}_{R+i+2|R+i}\right\}_{i=0}^{T-R-2}\!\!\!\!=\left\{\left(\overrightarrow{\eta}^{\alpha}_{R+i+2|R+i}\right)^{2}\!\!-\left(\overrightarrow{\eta}^{\gamma}_{R+i+2|R+i}\right)^{2}\right\}_{i=0}^{T-R-2}$
 \item $\cdots$
 \item $\cdots$
-\item For h-step ahead we have $\left\{\overrightarrow{\mathcal{RL}}_{R+i+h|R+i}\right\}_{i=0}^{T-R-h}\!\!\!\!=\left\{\left(\overrightarrow{\eta}^{\alpha}_{R+i+h|R+i}\right)^{2}\!\!-\left(\overrightarrow{\eta}^{\gamma}_{R+i+h|R+i}\right)^{2}\right\}_{i=0}^{T-R-h}$
+\item For $h$-step ahead: $\left\{\overrightarrow{\mathcal{RL}}_{R+i+h|R+i}\right\}_{i=0}^{T-R-h}\!\!\!\!=\left\{\left(\overrightarrow{\eta}^{\alpha}_{R+i+h|R+i}\right)^{2}\!\!-\left(\overrightarrow{\eta}^{\gamma}_{R+i+h|R+i}\right)^{2}\right\}_{i=0}^{T-R-h}$
 \end{itemize}
-\item So for any date after $R+h$ there are $h$ rolling loss $\overrightarrow{\mathcal{RL}}$ observations, corresponding to different step ahead forecasts.
-For date $R+h-1$ there are $h-1$ rolling loss $\overrightarrow{\mathcal{RL}}$ observations, and so on until $R+1$ where there is just one rolling loss $\overrightarrow{\mathcal{RL}}$ observation.
+\item So for any date after $R+h$ there are $h$ rolling loss $\overrightarrow{\mathcal{RL}}$ observations, corresponding to different step ahead forecasts, for date $R+h-1$ there are $h-1$ rolling loss $\overrightarrow{\mathcal{RL}}$ observations, for date $R+h-2$ there are $h-2$ rolling loss $\overrightarrow{\mathcal{RL}}$ observations, and so on until $R+1$ where there is just one rolling loss $\overrightarrow{\mathcal{RL}}$ observation.
 \item Models A and G can be estimated over fixed rolling windows and used to forecast in-sample.
-Note that whereas when one forecasts out-of-sample there is for each estimated model only 1 one-step ahead forecast, only 1 two-step ahead forecast, ..., and only 1 h-step ahead forecast, when one forecasts in-sample there is for each fixed rolling window of length $R$, $R-1$ one-step ahead forecasts, $R-2$ two-step ahead forecasts, ..., and $R-h$ h-step ahead forecasts.
-When looking at in-sample forecasts errors Barbara only looks at the \emph{very last} one-step, two-step, ..., h-step ahead forecast for that model.
-\item Models A and G estimated over fixed rolling windows and used to forecast in-sample will for every period and every h-step ahead forecast imply a squared forecast error for the last period of the estimation period $\left(\overline{\eta}^{\alpha}_{t}\right)^{2}$ and
- $\left(\overline{\eta}^{\gamma}_{t}\right)^{2}$.  (I'm using ``over arrows'' to denote out-of-sample forecasts.  This is my notation, not Barbara's.)
+Note that whereas when one forecasts out-of-sample there is for each estimated model only 1 $1$-step ahead forecast, only 1 $2$-step ahead forecast, ..., and only 1 $h$-step ahead forecast, when one forecasts in-sample there is for each estimated fixed rolling window of length $R$, $R-1$ $1$-step ahead forecasts (for different quarters), $R-2$ $2$-step ahead forecasts (for different quarter), ..., and $R-h$ $h$-step ahead forecasts (for different quarters).
+When looking at in-sample forecasts errors Barbara only looks at the \emph{very last} $1$-step, \emph{very last} $2$-step, ..., \emph{very last} $h$-step ahead forecast for that model.
+\item Models A and G estimated over fixed rolling windows and used to forecast in-sample will for every period and every $h$-step ahead forecast imply a squared forecast error for the last period of the estimation period $\left(\overline{\eta}^{\alpha}_{t}\right)^{2}$ and
+ $\left(\overline{\eta}^{\gamma}_{t}\right)^{2}$.  (I'm using ``over bars'' to denote in-sample forecasts.  This is my notation, not Barbara's.)
 \begin{itemize}
-\item For one-step ahead we have $\left\{\left(\overline{\eta}^{\alpha}_{R+i|R+i-1}\right)^{2}\right\}_{i=0}^{T-R}$ and $\left\{\left(\overline{\eta}^{\gamma}_{R+i|R+i-1}\right)^{2}\right\}_{i=0}^{T-R}$, where in each case the rolling window estimation period
+\item For $1$-step ahead we have $\left\{\left(\overline{\eta}^{\alpha}_{R+i|R+i-1}\right)^{2}\right\}_{i=0}^{T-R}$ and $\left\{\left(\overline{\eta}^{\gamma}_{R+i|R+i-1}\right)^{2}\right\}_{i=0}^{T-R}$, where in each case the rolling window estimation period
 ends at $R, R+1, \dots, T$ but the data used to generate the forecast ends in $R-1, R, \dots, T-1$.
-\item For two-step ahead we have $\left\{\left(\overline{\eta}^{\alpha}_{R+i|R+i-2}\right)^{2}\right\}_{i=0}^{T-R}$ and $\left\{\left(\overline{\eta}^{\gamma}_{R+i|R+i-2}\right)^{2}\right\}_{i=0}^{T-R}$, where in each case the rolling window estimation period
+\item For $2$-step ahead we have $\left\{\left(\overline{\eta}^{\alpha}_{R+i|R+i-2}\right)^{2}\right\}_{i=0}^{T-R}$ and $\left\{\left(\overline{\eta}^{\gamma}_{R+i|R+i-2}\right)^{2}\right\}_{i=0}^{T-R}$, where in each case the rolling window estimation period
 ends at $R, R+1, \dots, T$ but the data used to generate the forecast ends in $R-2, R-1, \dots, T-2$.
@@
 \item $\cdots$
-\item For h-step ahead we have $\left\{\left(\overline{\eta}^{\alpha}_{R+i|R+i-h}\right)^{2}\right\}_{i=0}^{T-R}$ and $\left\{\left(\overline{\eta}^{\gamma}_{R+i|R+i-h}\right)^{2}\right\}_{i=0}^{T-R}$, where in each case the rolling window estimation period
+\item For $h$-step ahead we have $\left\{\left(\overline{\eta}^{\alpha}_{R+i|R+i-h}\right)^{2}\right\}_{i=0}^{T-R}$ and $\left\{\left(\overline{\eta}^{\gamma}_{R+i|R+i-h}\right)^{2}\right\}_{i=0}^{T-R}$, where in each case the rolling window estimation period
 ends at $R, R+1, \dots, T$ but the data used to generate the forecast ends in $R-h, R-h+1, \dots, T-h$.
@@
 \begin{itemize}
-\item For one-step ahead we have $\left\{\overline{\mathcal{RL}}_{R+i|R+i-1}\right\}_{i=0}^{T-R-1}\!\!\!\!=\left\{\left(\overline{\eta}^{\alpha}_{R+i|R+i-1}\right)^{2}\!\!-\left(\overline{\eta}^{\gamma}_{R+i|R+i-1}\right)^{2}\right\}_{i=0}^{T-R-1}$
-\item For two-step ahead we have $\left\{\overline{\mathcal{RL}}_{R+i|R+i-2}\right\}_{i=0}^{T-R-2}\!\!\!\!=\left\{\left(\overline{\eta}^{\alpha}_{R+i|R+i-2}\right)^{2}\!\!-\left(\overline{\eta}^{\gamma}_{R+i|R+i-2}\right)^{2}\right\}_{i=0}^{T-R-2}$
+\item For $1$-step ahead: $\left\{\overline{\mathcal{RL}}_{R+i|R+i-1}\right\}_{i=0}^{T-R-1}\!\!\!\!=\left\{\left(\overline{\eta}^{\alpha}_{R+i|R+i-1}\right)^{2}\!\!-\left(\overline{\eta}^{\gamma}_{R+i|R+i-1}\right)^{2}\right\}_{i=0}^{T-R-1}$
+\item For $2$-step ahead: $\left\{\overline{\mathcal{RL}}_{R+i|R+i-2}\right\}_{i=0}^{T-R-2}\!\!\!\!=\left\{\left(\overline{\eta}^{\alpha}_{R+i|R+i-2}\right)^{2}\!\!-\left(\overline{\eta}^{\gamma}_{R+i|R+i-2}\right)^{2}\right\}_{i=0}^{T-R-2}$
 \item $\cdots$
 \item $\cdots$
-\item For h-step ahead we have $\left\{\overline{\mathcal{RL}}_{R+i|R+i-h}\right\}_{i=0}^{T-R-h}\!\!\!\!=\left\{\left(\overline{\eta}^{\alpha}_{R+i|R+i-h}\right)^{2}\!\!-\left(\overline{\eta}^{\gamma}_{R+i|R+i-h}\right)^{2}\right\}_{i=0}^{T-R-h}$
+\item For $h$-step ahead: $\left\{\overline{\mathcal{RL}}_{R+i|R+i-h}\right\}_{i=0}^{T-R-h}\!\!\!\!=\left\{\left(\overline{\eta}^{\alpha}_{R+i|R+i-h}\right)^{2}\!\!-\left(\overline{\eta}^{\gamma}_{R+i|R+i-h}\right)^{2}\right\}_{i=0}^{T-R-h}$
 \end{itemize}
@@
 \begin{itemize}
-\item So for the one-step ahead forecast we regress $\left\{\overrightarrow{\mathcal{RL}}_{R+i+1|R+i}\right\}_{i=0}^{T-R-1}$ on $\left\{\overline{\mathcal{RL}}_{R+i|R+i-1}\right\}_{i=0}^{T-R-1}$.
-The models being used to generate these losses are estimated one rolling windows that end on $R, R+1, \cdots, T$.
-\item So for the two-step ahead forecast we regress $\left\{\overrightarrow{\mathcal{RL}}_{R+i+2|R+i}\right\}_{i=0}^{T-R-1}$ on $\left\{\overline{\mathcal{RL}}_{R+i|R+i-2}\right\}_{i=0}^{T-R-1}$.
-The models being used to generate these losses are estimated one rolling windows that end on $R, R+1, \cdots, T$.
+\item For the $1$-step ahead forecast regress $\left\{\overrightarrow{\mathcal{RL}}_{R+i+1|R+i}\right\}_{i=0}^{T-R-1}$ on $\left\{\overline{\mathcal{RL}}_{R+i|R+i-1}\right\}_{i=0}^{T-R-1}$.
+The models being used to generate these losses are estimated on rolling windows that end on $R, R+1, \cdots, T$.
+\item For the $2$-step ahead forecast regress $\left\{\overrightarrow{\mathcal{RL}}_{R+i+2|R+i}\right\}_{i=0}^{T-R-2}$ on $\left\{\overline{\mathcal{RL}}_{R+i|R+i-2}\right\}_{i=0}^{T-R-2}$.
+The models being used to generate these losses are estimated on rolling windows that end on $R, R+1, \cdots, T$.
 \item $\cdots$
 \item $\cdots$
-\item So for the h-step ahead forecast we regress $\left\{\overrightarrow{\mathcal{RL}}_{R+i+h|R+i}\right\}_{i=0}^{T-R-1}$ on $\left\{\overline{\mathcal{RL}}_{R+i|R+i-h}\right\}_{i=0}^{T-R-1}$.
-The models being used to generate these losses are estimated one rolling windows that end on $R, R+1, \cdots, T$.
+\item For the $h$-step ahead forecast regress $\left\{\overrightarrow{\mathcal{RL}}_{R+i+h|R+i}\right\}_{i=0}^{T-R-h}$ on $\left\{\overline{\mathcal{RL}}_{R+i|R+i-h}\right\}_{i=0}^{T-R-h}$.
+The models being used to generate these losses are estimated on rolling windows that end on $R, R+1, \cdots, T$.
 \end{itemize}
@@
 \begin{itemize}
-\item One-step: $\overrightarrow{\mathcal{RL}}_{R+i+1|R+i}=\widehat{\beta}_{1}\overline{\mathcal{RL}}_{R+i|R+i-1}+\widehat{u}_{R+i+1}$, where $i=0, 1, \cdots, T-R-1$.
-\item Two-step: $\overrightarrow{\mathcal{RL}}_{R+i+2|R+i}=\widehat{\beta}_{2}\overline{\mathcal{RL}}_{R+i|R+i-2}+\widehat{u}_{R+i+2}$, where $i=0, 1, \cdots, T-R-2$.
+\item $1$-step: $\overrightarrow{\mathcal{RL}}_{R+i+1|R+i}=\widehat{\beta}_{1}\overline{\mathcal{RL}}_{R+i|R+i-1}+\widehat{u}_{R+i+1}$, where $i=0, 1, \cdots, T-R-1$.
+\item $2$-step: $\overrightarrow{\mathcal{RL}}_{R+i+2|R+i}=\widehat{\beta}_{2}\overline{\mathcal{RL}}_{R+i|R+i-2}+\widehat{u}_{R+i+2}$, where $i=0, 1, \cdots, T-R-2$.
 \item $\cdots$
 \item $\cdots$
-\item h-step: $\overrightarrow{\mathcal{RL}}_{R+i+h|R+i}=\widehat{\beta}_{h}\overline{\mathcal{RL}}_{R+i|R+i-h}+\widehat{u}_{R+i+h}$, where $i=0, 1, \cdots, T-R-h$.
+\item $h$-step: $\overrightarrow{\mathcal{RL}}_{R+i+h|R+i}=\widehat{\beta}_{h}\overline{\mathcal{RL}}_{R+i|R+i-h}+\widehat{u}_{R+i+h}$, where $i=0, 1, \cdots, T-R-h$.
 \item It may be worth noting that every out-of-sample-based loss being regressed on a corresponding in-sample-based loss uses exactly the same two versions -- where versions refer to the estimation period -- of the A and G model.
-That is the models are all estimated on rolling windows ending $R, R+1, \cdots, T$ but the in-sample forecasts use this estimated model but condition on earlier data for period $R-step, R-step+1, \cdots, T-step$ forecasts while the out-of-sample forecasts use this estimated model, condition on $R, R+1, \cdots, T-step$ data and project out  $R+step, R+step+1, \cdots, T+step-step$.
+That is the models are all estimated on rolling windows ending $R, R+1, \cdots, T-s$ but the in-sample forecasts use this estimated model but condition on earlier data for period $R-s$, $R+1-s$, $\cdots$, $T-s-s$ (where $s=1,2,\cdots,h$) in generating the forecasts for periods $R, R+1, \cdots, T$ while the out-of-sample forecasts use this estimated model, condition on $R$, $R+1$, $\cdots$, $T-s$ data and project out  $R+s$, $R+1+s$, $\cdots$, $T$ (where $s=1,2,\cdots,h$).
 \end{itemize}
@@
 \begin{itemize}
-\item One-step: 
+\item $1$-step: 
 \begin{eqnarray}
-\underbrace{\frac{1}{T\!-\!R}\!\!\sum_{i=0}^{T-R-1}\!\!\overrightarrow{\mathcal{RL}}_{R+i+1|R+i}}_{=A_{1,T\!-\!R}}
-\!\!\!\!&=&\!\!\!\!\underbrace{\widehat{\beta}_{1}\cdot\frac{1}{T\!-\!R}\!\!\sum_{i=0}^{T-R-1}\!\!\overline{\mathcal{RL}}_{R+i|R+i-1}}_{=B_{1,T\!-\!R}} \nonumber \\
-\!\!\!\!&+&\!\!\!\!\underbrace{\frac{1}{T\!-\!R}\!\!\sum_{i=0}^{T-R-1}\!\!\widehat{u}_{R+i+1}}_{=U_{1,T\!-\!R}}, \nonumber
+\underbrace{\frac{1}{T\!-\!R}\!\!\sum_{i=0}^{T-R-1}\!\!\overrightarrow{\mathcal{RL}}_{R+i+1|R+i}}_{=A_{1,T-R}}
+\!\!\!\!&=&\!\!\!\!\underbrace{\widehat{\beta}_{1}\cdot\frac{1}{T\!-\!R}\!\!\sum_{i=0}^{T-R-1}\!\!\overline{\mathcal{RL}}_{R+i|R+i-1}}_{=B_{1,T-R}} 
++\underbrace{\frac{1}{T\!-\!R}\!\!\sum_{i=0}^{T-R-1}\!\!\widehat{u}_{R+i+1}}_{=U_{1,T-R}}, \nonumber
 \end{eqnarray}
-where $i=0, 1, \cdots, T-R-1$.
-\item Two-step: 
+\item $2$-step: 
 \begin{eqnarray}
-\underbrace{\frac{1}{T\!-\!R\!-\!1}\!\!\sum_{i=0}^{T-R-2}\!\!\overrightarrow{\mathcal{RL}}_{R+i+2|R+i}}_{=A_{2,T\!-\!R\!-\!1}}
-\!\!\!\!&=&\!\!\!\!\underbrace{\widehat{\beta}_{2}\cdot\frac{1}{T\!-\!R\!-\!1}\!\!\sum_{i=0}^{T-R-2}\!\!\overline{\mathcal{RL}}_{R+i|R+i-2}}_{=B_{2,T\!-\!R\!-\!1}} \nonumber \\
-\!\!\!\!&+&\!\!\!\!\underbrace{\frac{1}{T\!-\!R\!-\!1}\!\!\sum_{i=0}^{T-R-2}\!\!\widehat{u}_{R+i+2}}_{=U_{2,T\!-\!R\!-\!1}}, \nonumber 
+\underbrace{\frac{1}{T\!-\!R\!-\!1}\!\!\sum_{i=0}^{T-R-2}\!\!\overrightarrow{\mathcal{RL}}_{R+i+2|R+i}}_{=A_{2,T-R-1}}
+\!\!\!\!&=&\!\!\!\!\underbrace{\widehat{\beta}_{2}\cdot\frac{1}{T\!-\!R\!-\!1}\!\!\sum_{i=0}^{T-R-2}\!\!\overline{\mathcal{RL}}_{R+i|R+i-2}}_{=B_{2,T-R-1}} 
++\underbrace{\frac{1}{T\!-\!R\!-\!1}\!\!\sum_{i=0}^{T-R-2}\!\!\widehat{u}_{R+i+2}}_{=U_{2,T-R-1}}, \nonumber 
 \end{eqnarray}
-where $i=0, 1, \cdots, T-R-2$.
 \item $\cdots$
 \item $\cdots$
-\item h-step: 
+\item $h$-step: 
 \begin{eqnarray}
-\underbrace{\frac{1}{T\!-\!R\!-\!(h\!-\!1)}\!\!\sum_{i=0}^{T-R-(h-1)}\!\!\overrightarrow{\mathcal{RL}}_{R+i+h|R+i}}_{=A_{h,T\!-\!R\!-\!\!(h\!-\!1)}}
-\!\!\!\!&=&\!\!\!\!\underbrace{\widehat{\beta}_{h}\cdot\frac{1}{T\!-\!R\!-\!(h\!-\!1)}\!\!\sum_{i=0}^{T-R-(h-1)}\!\!\overline{\mathcal{RL}}_{R+i|R+i-h}}_{=B_{h,T\!-\!R\!-\!(h\!-\!1)}} \nonumber \\
-\!\!\!\!&+&\!\!\!\!\underbrace{\frac{1}{T\!-\!R\!-\!(h\!-\!1)}\!\!\sum_{i=0}^{T-R-(h-1)}\!\!\widehat{u}_{R+i+h}}_{=U_{h,T\!-\!R\!-\!(h\!-\!1)}}, \nonumber
+\underbrace{\frac{1}{T\!-\!R\!-\!(h\!-\!1)}\!\!\!\!\sum_{i=0}^{T-R-h}\!\!\!\!\overrightarrow{\mathcal{RL}}_{R+i+h|R+i}}_{=A_{h,T-R-(h-1)}}
+\!\!\!\!&=&\!\!\!\!\underbrace{\widehat{\beta}_{h}\cdot\frac{1}{T\!-\!R\!-\!(h\!-\!1)}\!\!\!\!\sum_{i=0}^{T-R-h}\!\!\!\!\overline{\mathcal{RL}}_{R+i|R+i-h}}_{=B_{h,T-R-(h-1)}} \ \ \ \   \ \ \ \    \ \ \ \   \ \
+\nonumber \\
+\!\!\!\!&+&\!\!\!\!
+\underbrace{\frac{1}{T\!-\!R\!-\!(h\!-\!1)}\!\!\!\!\sum_{i=0}^{T-R-h}\!\!\!\!\widehat{u}_{R+i+h}}_{=U_{h,T-R-(h-1)}}, \nonumber
 \end{eqnarray}
-where $i=0, 1, \cdots, T-R-h$.
 \end{itemize}
@@
 \!\!\!\!&&\!\!\!\!\cdots  \nonumber \\
-A_{h,T\!-\!R\!-\!\!(h\!-\!1)}\!\!\!\!&=&\!\!\!\!\frac{1}{T\!-\!R\!-\!(h\!-\!1)}\!\!\sum_{i=0}^{T-R-(h-1)}\!\!\overrightarrow{\mathcal{RL}}_{R+i+h|R+i}. \nonumber
+A_{h,T\!-\!R\!-\!\!(h\!-\!1)}\!\!\!\!&=&\!\!\!\!\frac{1}{T\!-\!R\!-\!(h\!-\!1)}\!\!\sum_{i=0}^{T-R-h}\!\!\overrightarrow{\mathcal{RL}}_{R+i+h|R+i}. \nonumber
 \end{eqnarray}
@@
 \begin{itemize}
-\item For $A_{1,T\!-\!R}$ there are $T-R-m$ possible windows, where the windows are: 
-\[\left\{\overrightarrow{\mathcal{RL}}_{R+i+1|R+i}\ \ , \ \  . \ \  . \ \  . \ \ ,\ \ \overrightarrow{\mathcal{RL}}_{R+i+1+m|R+i+m}\right\}_{i=0}^{T-R-1-m}\] 
+\item For $A_{1,T\!-\!R}$ there are $T-R-m+1$ possible windows, where the windows are: 
+\[\left\{\overrightarrow{\mathcal{RL}}_{R+i+1+0|R+i+0}\ \ , \ \  . \ \  . \ \  . \ \ ,\ \ \overrightarrow{\mathcal{RL}}_{R+i+1+(m-1)|R+i+(m-1)}\right\}_{i=0}^{T-R-1+1-m}\] 
 and so the rolling window relative out-of-sample forecast performance terms would be:
-\[\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+1+j|R+i+j}\ \ \ \mathrm{for} \ \ \ i=0,1,\cdots,T\!-\!R\!-\!1\!-\!m. \]
-\item For $A_{2,T\!-\!R\!-\!1}$ there are $T-R-m-1$ possible windows, where the windows are: 
-\[\left\{\overrightarrow{\mathcal{RL}}_{R+i+2|R+i}\ \ , \ \  . \ \  . \ \  . \ \ ,\ \ \overrightarrow{\mathcal{RL}}_{R+i+2+m|R+i+m}\right\}_{i=0}^{T-R-2-m}\] 
+\[\frac{1}{m}\sum_{j=0}^{m-1}\overrightarrow{\mathcal{RL}}_{R+i+1+j|R+i+j}\ \ \ \mathrm{for} \ \ \ i=0,1,\cdots,T\!-\!R\!-\!1\!+\!1\!-\!m. \]
+\item For $A_{2,T\!-\!R\!-\!1}$ there are $T-R-m$ possible windows, where the windows are: 
+\[\left\{\overrightarrow{\mathcal{RL}}_{R+i+2+0|R+i+0}\ \ , \ \  . \ \  . \ \  . \ \ ,\ \ \overrightarrow{\mathcal{RL}}_{R+i+2+(m-1)|R+i+(m-1)}\right\}_{i=0}^{T-R-2+1-m}\] 
 and so the rolling window relative out-of-sample forecast performance terms would be:
-\[\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+2+j|R+i+j}\ \ \ \mathrm{for} \ \ \ i=0,1,\cdots,T\!-\!R\!-\!2\!-\!m. \]
+\[\frac{1}{m}\sum_{j=0}^{m-1}\overrightarrow{\mathcal{RL}}_{R+i+2+j|R+i+j}\ \ \ \mathrm{for} \ \ \ i=0,1,\cdots,T\!-\!R\!-\!2\!+\!1\!-\!m. \]
 \item And so on for all the other steps until we get to the $h$-th step
 \item For $A_{h,T\!-\!R\!-\!(h-1)}$ there are $T-R-m-(h-1)$ possible windows, where the windows are: 
-\[\left\{\overrightarrow{\mathcal{RL}}_{R+i+h|R+i}\ \ , \ \  . \ \  . \ \  . \ \ ,\ \ \overrightarrow{\mathcal{RL}}_{R+i+h+m|R+i+m}\right\}_{i=0}^{T-R-h-m}\] 
+\[\left\{\overrightarrow{\mathcal{RL}}_{R+i+h+0|R+i+0}\ \ , \ \  . \ \  . \ \  . \ \ ,\ \ \overrightarrow{\mathcal{RL}}_{R+i+h+(m-1)|R+i+(m-1)}\right\}_{i=0}^{T-R-h+1-m}\] 
 and so the rolling window relative out-of-sample forecast performance terms would be:
-\[\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+h+j|R+i+j}\ \ \ \mathrm{for} \ \ \ i=0,1,\cdots,T\!-\!R\!-\!h\!-\!m. \]
+\[\frac{1}{m}\sum_{j=0}^{m-1}\overrightarrow{\mathcal{RL}}_{R+i+h+j|R+i+j}\ \ \ \mathrm{for} \ \ \ i=0,1,\cdots,T\!-\!R\!-\!h\!+\!1\!-\!m. \]
 \end{itemize}
-\item So then it is the rolling window relative out-of-sample forecast performances -- that is, $\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+1+j|R+i+j}$, $\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+2+j|R+i+j}$, .
+\item So then it is the rolling window relative out-of-sample forecast performances -- that is, $\frac{1}{m}\sum_{j=0}^{m-1}\overrightarrow{\mathcal{RL}}_{R+i+1+j|R+i+j}$, $\frac{1}{m}\sum_{j=0}^{m-1}\overrightarrow{\mathcal{RL}}_{R+i+2+j|R+i+j}$, .
 .
 .
-, $\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+h+j|R+i+j}$ -- that one wants to decompose.
+, $\frac{1}{m}\sum_{j=0}^{m-1}\overrightarrow{\mathcal{RL}}_{R+i+h+j|R+i+j}$ -- that one wants to decompose.
 These are done as follows.
 \begin{eqnarray}
-\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+1+j|R+i+j}\!\!\!\!&=&\!\!\!\!\underbrace{\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+1+j|R+i+j}-\underbrace{\frac{1}{T\!-\!R}\!\!\sum_{i=0}^{T-R-1}\!\!\overrightarrow{\mathcal{RL}}_{R+i+1|R+i}}_{=A_{1,T\!-\!R}}}_{=A_{i,1,T-R}} \nonumber \\
-\!\!\!\!&+&\!\!\!\!B_{1,T-R}+U_{1,T-R} \ \ \ \ \   \ \ \ \ \   \ \ \ \ \   \mathrm{for} \ i=0,1,\cdots,T\!-\!R\!-\!1\!-\!m. \nonumber
+\frac{1}{m}\sum_{j=0}^{m-1}\overrightarrow{\mathcal{RL}}_{R+i+1+j|R+i+j}\!\!\!\!&=&\!\!\!\!\underbrace{\frac{1}{m}\sum_{j=0}^{m-1}\overrightarrow{\mathcal{RL}}_{R+i+1+j|R+i+j}-\underbrace{\frac{1}{T\!-\!R}\!\!\sum_{i=0}^{T-R-1}\!\!\overrightarrow{\mathcal{RL}}_{R+i+1|R+i}}_{=A_{1,T\!-\!R}}}_{=A_{i,1,T-R}} \nonumber \\
+\!\!\!\!&+&\!\!\!\!B_{1,T-R}+U_{1,T-R} \ \ \ \ \   \ \ \ \ \   \ \  \mathrm{for} \ i=0,1,\cdots,T\!-\!R\!-\!1\!+\!1\!-\!m. \nonumber
 \end{eqnarray}
 \begin{eqnarray}
-\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+2+j|R+i+j}\!\!\!\!&=&\!\!\!\!\underbrace{\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+2+j|R+i+j}-\underbrace{\frac{1}{T\!-\!R\!-\!1}\!\!\sum_{i=0}^{T-R-2}\!\!\overrightarrow{\mathcal{RL}}_{R+i+2|R+i}}_{=A_{2,T\!-\!R\!-\!1}}}_{=A_{i,2,T-R-1}} \nonumber \\
-\!\!\!\!&+&\!\!\!\!B_{2,T-R-1}+U_{2,T-R-1} \ \ \ \ \   \ \ \ \ \   \mathrm{for} \ i=0,1,\cdots,T\!-\!R\!-\!2\!-\!m. \nonumber
+\frac{1}{m}\sum_{j=0}^{m-1}\overrightarrow{\mathcal{RL}}_{R+i+2+j|R+i+j}\!\!\!\!&=&\!\!\!\!\underbrace{\frac{1}{m}\sum_{j=0}^{m-1}\overrightarrow{\mathcal{RL}}_{R+i+2+j|R+i+j}-\underbrace{\frac{1}{T\!-\!R\!-\!1}\!\!\sum_{i=0}^{T-R-2}\!\!\overrightarrow{\mathcal{RL}}_{R+i+2|R+i}}_{=A_{2,T\!-\!R\!-\!1}}}_{=A_{i,2,T-R-1}} \nonumber \\
+\!\!\!\!&+&\!\!\!\!B_{2,T-R-1}+U_{2,T-R-1} \ \ \ \ \   \ \  \mathrm{for} \ i=0,1,\cdots,T\!-\!R\!-\!2\!+\!1\!-\!m. \nonumber
 \end{eqnarray}
@@
 \begin{eqnarray}
-\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+h+j|R+i+j}\!\!\!\!&=&\!\!\!\!\underbrace{\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+h+j|R+i+j}-\underbrace{\frac{1}{T\!-\!R\!-\!(h\!-\!1)}\!\!\sum_{i=0}^{T-R-(h-1)}\!\!\overrightarrow{\mathcal{RL}}_{R+i+h|R+i}}_{=A_{h,T\!-\!R-(h-1)}}}_{=A_{i,h,T-R-(h-1)}} \nonumber \\
-\!\!\!\!&+&\!\!\!\!B_{h,T-R-(h-1)}+U_{h,T-R-(h-1)} \ \ \ \ \   \ \ \ \ \   \mathrm{for} \ i=0,1,\cdots,T\!-\!R\!-\!h\!-\!m. \nonumber
+\frac{1}{m}\sum_{j=0}^{m-1}\overrightarrow{\mathcal{RL}}_{R+i+h+j|R+i+j}\!\!\!\!&=&\!\!\!\!\underbrace{\frac{1}{m}\sum_{j=0}^{m-1}\overrightarrow{\mathcal{RL}}_{R+i+h+j|R+i+j}-\underbrace{\frac{1}{T\!-\!R\!-\!(h\!-\!1)}\!\!\sum_{i=0}^{T-R-h}\!\!\overrightarrow{\mathcal{RL}}_{R+i+h|R+i}}_{=A_{h,T\!-\!R-(h-1)}}}_{=A_{i,h,T-R-(h-1)}} \nonumber \\
+\!\!\!\!&+&\!\!\!\!B_{h,T-R-(h-1)}+U_{h,T-R-(h-1)} \ \ \ \ \   \mathrm{for} \ i=0,1,\cdots,T\!-\!R\!-\!h\!+\!1\!-\!m. \nonumber
 \end{eqnarray}
@@
 \begin{eqnarray}
-\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+1+j|R+i+j}\!\!\!\!&=&\!\!\!\!A_{i,1,T-R}+B_{1,T-R}+U_{1,T-R} \nonumber \\
-\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+2+j|R+i+j}\!\!\!\!&=&\!\!\!\!A_{i,2,T-R-1}+B_{2,T-R-1}+U_{2,T-R-1} \nonumber \\
-\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+h+j|R+i+j}\!\!\!\!&=&\!\!\!\!A_{i,h,T-R-(h-1)}+B_{h,T-R-(h-1)}+U_{h,T-R-(h-1)} \nonumber
+\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+1+j|R+i+j}\!\!\!\!&=&\!\!\!\!A_{i,1,T-R}\!+\!B_{1,T-R}\!+\!U_{1,T-R} \ \ \ \ \    \ \ \ \ \  \mathrm{for} \ i=0,1,\cdots,T\!-\!R\!-\!1\!+\!1\!-\!m\nonumber \\
+\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+2+j|R+i+j}\!\!\!\!&=&\!\!\!\!A_{i,2,T-R-1}\!+\!B_{2,T-R-1}\!+\!U_{2,T-R-1} \ \mathrm{for} \ i=0,1,\cdots,T\!-\!R\!-\!2\!+\!1\!-\!m\nonumber \\
+\cdots \nonumber \\
+\cdots \nonumber \\
+\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+h+j|R+i+j}\!\!\!\!&=&\!\!\!\!A_{i,h,T-R-(h-1)}\!+\!B_{h,T-R-(h-1)}\!+\!U_{h,T-R-(h-1)}\ \nonumber \\
+&& \ \ \ \    \ \ \ \   \ \ \ \    \ \ \ \    \ \ \ \    \ \ \ \   \ \ \ \    \ \ \ \   \ \ \ \   \ \ \ \    \ \ 
+\mathrm{for} \ i=0,1,\cdots,T\!-\!R\!-\!h\!+\!1\!-\!m \nonumber
 \end{eqnarray}
-\item So basically one calculates the rolling window relative out-of-sample forecast performances -- that is, $\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+1+j|R+i+j}$, $\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+2+j|R+i+j}$, .
+\item So basically one calculates the rolling window relative out-of-sample forecast performances -- that is, $\frac{1}{m}\sum_{j=0}^{m-1}\overrightarrow{\mathcal{RL}}_{R+i+1+j|R+i+j}$, $\frac{1}{m}\sum_{j=0}^{m-1}\overrightarrow{\mathcal{RL}}_{R+i+2+j|R+i+j}$, .
 .
 .
-, $\frac{1}{m}\sum_{j=0}^{m}\overrightarrow{\mathcal{RL}}_{R+i+h+j|R+i+j}$ -- and tries to understand them in terms of the above components.
+, $\frac{1}{m}\sum_{j=0}^{m-1}\overrightarrow{\mathcal{RL}}_{R+i+h+j|R+i+j}$ -- and tries to understand them in terms of the above components.
 \begin{itemize}
@@
 \end{itemize}
+\item The next step is to makes test statistics out of the $\left\{\left\{A_{i,s,T-R}\right\}_{i=0}^{T-R-s+1-m},B_{s,T-R},U_{s,T-R}\right\}_{s=1}^{h}$
+terms calculated above; though a few more terms are needed to calculate these. [Note that we are back to calculating things over the full forecast sample.]  Let:
+\begin{eqnarray}
+\widehat{rl}_{R+j+1}\!\!\!\!&=&\!\!\!\!
+\left(\begin{array}{c} \overrightarrow{\mathcal{RL}}_{R+j+1|R+j} \\ \overline{\mathcal{RL}}_{R+j|R+j-1}\cdot\overrightarrow{\mathcal{RL}}_{R+j+1|R+j} \end{array} \right),
+\textrm{where} \ \ j=0, 1, \cdots, T-R-1
+\nonumber \\
+\widehat{rl}_{R+j+2}\!\!\!\!&=&\!\!\!\!
+\left(\begin{array}{c} \overrightarrow{\mathcal{RL}}_{R+j+2|R+j} \\ \overline{\mathcal{RL}}_{R+j|R+j-2}\cdot\overrightarrow{\mathcal{RL}}_{R+j+2|R+j} \end{array} \right),
+\textrm{where} \ \ j=0, 1, \cdots, T-R-2
+\nonumber \\
+\cdots
+\nonumber \\
+\cdots
+\nonumber \\
+\widehat{rl}_{R+j+h}\!\!\!\!&=&\!\!\!\!
+\left(\begin{array}{c} \overrightarrow{\mathcal{RL}}_{R+j+h|R+j} \\ \overline{\mathcal{RL}}_{R+j|R+j-h}\cdot\overrightarrow{\mathcal{RL}}_{R+j+h|R+j} \end{array} \right),
+\textrm{where} \ \ j=0, 1, \cdots, T-R-h
+\nonumber
+\end{eqnarray}
+\item Barbara then defines (where I assumed ``$d$" denotes de-meaned, again over the full forecast sample):
+\begin{eqnarray}
+\underbrace{\widehat{rl}^{d}_{R+j+1}}_{(2x1)}\!\!\!\!&=&\!\!\!\!\widehat{rl}_{R+j+1}-\frac{1}{T-R-1+1}\sum_{i=0}^{T-R-1}\widehat{rl}_{R+i+1}, \ \textrm{where} \ \ j=0, 1, \cdots, T-R-1 \nonumber \\
+\underbrace{\widehat{rl}^{d}_{R+j+2}}_{(2x1)}\!\!\!\!&=&\!\!\!\!\widehat{rl}_{R+j+2}-\frac{1}{T-R-2+1}\sum_{i=0}^{T-R-2}\widehat{rl}_{R+i+2}, \ \textrm{where} \ \ j=0, 1, \cdots, T-R-2\nonumber \\
+\cdots
+\nonumber \\
+\cdots
+\nonumber \\
+\underbrace{\widehat{rl}^{d}_{R+j+h}}_{(2x1)}\!\!\!\!&=&\!\!\!\!\widehat{rl}_{R+j+h}-\frac{1}{T-R-h+1}\sum_{i=0}^{T-R-h}\widehat{rl}_{R+i+h}, \ \textrm{where} \ \ j=0, 1, \cdots, T-R-h
+\nonumber
+\end{eqnarray}
+\item *****************************************************************************
+\item There seem to be parallels with this an Newey-West (1987), as Barbara alludes to.  See section 15.4.3 (on pages 422 and 423) of Greene. 
+\item For the $1$-step ahead forecast Barbara calculates
+\[ \widehat{\Omega}_{roll,1}=\frac{1}{T-R}\sum_{i=-(q(T-R)-1)}^{+(q(T-R)-1)}\left(1-\left|\frac{i}{q(T-R)}\right|\right)\sum_{j=0}^{T-R-1}(\widehat{rl}^{d}_{R+j+1})(\widehat{rl}^{d}_{R+j+1})^{\prime}
+\]
+(So if $q=4$ then $t$ would be $\{-3,-2,-1,0,1,2,3\}$ and then the $\left(1-\left|\frac{i}{q(T-R)}\right|\right)$ term would be $\{1-\frac{3}{4},1-\frac{2}{4},1-\frac{1}{4},1-\frac{0}{4},1-\frac{1}{4},1-\frac{2}{4},1-\frac{2}{4}\}$ or $\{\frac{1}{4},\frac{2}{4},\frac{3}{4},1,\frac{3}{4},\frac{2}{4},\frac{1}{4}\}$  The thing that I don't understand is that there doesn't seem to be any interaction between the first and second summations.
+That is, $i$ doesn't enter the $\sum_{j=0}^{T-R-1}(\widehat{rl}^{d}_{R+j+1})(\widehat{rl}^{d}_{R+j+1})^{\prime}$ term.)
+\item For the $2$-step ahead forecast Barbara calculates (I think)
+\[ \widehat{\Omega}_{roll,2}=\frac{1}{T-R-1}\sum_{i=-(q(T-R-1)-1)}^{+(q(T-R-1)-1)}\left(1-\left|\frac{i}{q(T-R-1)}\right|\right)\sum_{j=0}^{T-R-2}(\widehat{rl}^{d}_{R+j+2})(\widehat{rl}^{d}_{R+j+2})^{\prime}
+\]
+\item ...
+\item ...
+\item For the $h$-step ahead forecast Barbara calculates (I think)
+\[ \widehat{\Omega}_{roll,h}=\!\!
+\frac{1}{T\!-\!R\!-\!(h\!-\!1)}\!\!\sum_{i=-(q(T-R-(h-1))-1)}^{+(q(T-R-(h-1))-1)}\!\!\left(1\!-\!\left|\frac{i}{q(T\!-\!R\!-\!(h\!-\!1))}\right|\right)\!\!\sum_{j=0}^{T-R-h}\!\!\!(\widehat{rl}^{d}_{R+j+2})(\widehat{rl}^{d}_{R+j+2})^{\prime}
+\]
+\item *****************************************************************************
+\item For the $1$-step ahead forecast one takes:
+\begin{eqnarray}&&\!\!\!\!\widehat{\sigma}^{2}_{A,1}=\widehat{\Omega}_{(1,1),roll,1},\ \widehat{\sigma}^{2}_{B,1}=\Phi^{2}_{T-R}\widehat{\Omega}_{(2,2),roll,1}\ \mathrm{and} \ \widehat{\sigma}^{2}_{U,1}=\widehat{\sigma}^{2}_{A,1}-\widehat{\sigma}^{2}_{B,1}, \nonumber \\
+&&\!\!\!\!\mathrm{where}\ \Phi^{2}_{T-R}=\left(\frac{1}{T-R}\sum_{i=0}^{T-R-1} \overline{\mathcal{RL}}_{R+i|R+i-1}\right) \left(\frac{1}{T-R}\sum_{i=0}^{T-R-1} \left(\overline{\mathcal{RL}}_{R+i|R+i-1}\right)^{2}\right)^{-1} \nonumber
+\end{eqnarray}
+\item For the $2$-step ahead forecast one takes:
+\begin{eqnarray}&&\!\!\!\!\widehat{\sigma}^{2}_{A,2}=\widehat{\Omega}_{(1,1),roll,2},\ \widehat{\sigma}^{2}_{B,2}=\Phi^{2}_{T-R}\widehat{\Omega}_{(2,2),roll,2}\ \mathrm{and} \ \widehat{\sigma}^{2}_{U,2}=\widehat{\sigma}^{2}_{A,2}-\widehat{\sigma}^{2}_{B,2}, \nonumber \\
+&&\!\!\!\!\mathrm{where}\ \Phi^{2}_{T-R-1}=\left(\frac{1}{T\!-\!R\!-\!1}\sum_{i=0}^{T-R-2} \overline{\mathcal{RL}}_{R+i|R+i-2}\right) \left(\frac{1}{T\!-\!R\!-\!1}\sum_{i=0}^{T-R-2} \left(\overline{\mathcal{RL}}_{R+i|R+i-2}\right)^{2}\right)^{-1} \nonumber
+\end{eqnarray}
+\item ...
+\item ...
+\item For the $h$-step ahead forecast one takes:
+\begin{eqnarray}&&\!\!\!\!\widehat{\sigma}^{2}_{A,h}=\widehat{\Omega}_{(1,1),roll,h},\ \widehat{\sigma}^{2}_{B,h}=\Phi^{2}_{T-R}\widehat{\Omega}_{(2,2),roll,h}\ \mathrm{and} \ \widehat{\sigma}^{2}_{U,h}=\widehat{\sigma}^{2}_{A,h}-\widehat{\sigma}^{2}_{B,h}, \nonumber \\
+&&\!\!\!\!\mathrm{where}\ \Phi^{2}_{T-R-(h-1)}=\left(\frac{1}{T\!-\!R\!-\!(h\!-\!1)}\sum_{i=0}^{T-R-h} \overline{\mathcal{RL}}_{R+i|R+i-h}\right) \nonumber \\ && \ \ \ \   \ \ \ \   \ \ \ \   \ \ \ \   \ \ \ \   \ \ \ \   \ \ \ \   \ \ \ \   \times\left(\frac{1}{T\!-\!R\!-\!(h\!-\!1)}\sum_{i=0}^{T-R-h}\left (\overline{\mathcal{RL}}_{R+i|R+i-h}\right)^{2}\right)^{-1} \ \ \ \   \ \ \ \    \ \ \ \   \nonumber
+\end{eqnarray}
+\item Then one calculates the test-statistics for seeing if $\left\{\left\{A_{i,s,T-R}\right\}_{i=0}^{T-R-s-m},B_{s,T-R},U_{s,T-R}\right\}_{s=1}^{h}$ terms are significant or not.
+Barbara just does one example of this -- that is for one $h$-step ahead forecast -- and these are her equations in section~4.
+I think one actually has to do things sligthly differently for $1$-step, $2$-step, ..., $h$-step ahead forecasts.
+\item I think the statistics for the $1$-step ahead forecast will be:
+\begin{eqnarray}
+\Gamma^{(A)}_{1,T-R}\!\!\!\!&=&\!\!\!\!\sup_{i=0,1,\cdots,T-R-1+1-m}\left|\sqrt{T-R}\cdot \widehat{\sigma}^{-1}_{A,1}\cdot A_{i,1,T-R} \right|,
+\nonumber \\
+\Gamma^{(B)}_{1,T-R}\!\!\!\!&=&\!\!\!\! \sqrt{T-R}\cdot\widehat{\sigma}^{-1}_{B,1}\cdot B_{1,T-R}, \nonumber \\
+\Gamma^{(U)}_{1,T-R}\!\!\!\!&=&\!\!\!\! \sqrt{T-R}\cdot\widehat{\sigma}^{-1}_{U,1}\cdot U_{1,T-R}, \nonumber
+\end{eqnarray}
+\item I think the statistics for the $2$-step ahead forecast will be:
+\begin{eqnarray}
+\Gamma^{(A)}_{2,T-R-1}\!\!\!\!&=&\!\!\!\!\sup_{i=0,1,\cdots,T-R-2+1-m}\left|\sqrt{T-R-1}\cdot \widehat{\sigma}^{-1}_{A,2}\cdot A_{i,2,T-R-1} \right|,
+\nonumber \\
+\Gamma^{(B)}_{2,T-R-1}\!\!\!\!&=&\!\!\!\! \sqrt{T-R-1}\cdot\widehat{\sigma}^{-1}_{B,2}\cdot B_{2,T-R-1}, \nonumber \\
+\Gamma^{(U)}_{2,T-R-1}\!\!\!\!&=&\!\!\!\! \sqrt{T-R-1}\cdot\widehat{\sigma}^{-1}_{U,2}\cdot U_{2,T-R-1}, \nonumber
+\end{eqnarray}
+\item ...
+\item ... 
+\item I think the statistics for the $h$-step ahead forecast will be:
+\begin{eqnarray}
+\Gamma^{(A)}_{h,T-R-(h-1)}\!\!\!\!&=&\!\!\!\!\sup_{i=0,1,\cdots,T-R-h+1-m}\left|\sqrt{T-R-(h-1)}\cdot \widehat{\sigma}^{-1}_{A,h}\cdot A_{i,h,T-R-(h-1)} \right|,
+\nonumber \\
+\Gamma^{(B)}_{h,T-R-(h-1)}\!\!\!\!&=&\!\!\!\! \sqrt{T-R-(h-1)}\cdot\widehat{\sigma}^{-1}_{B,h}\cdot B_{h,T-R-(h-1)}, \nonumber \\
+\Gamma^{(U)}_{h,T-R-(h-1)}\!\!\!\!&=&\!\!\!\! \sqrt{T-R-(h-1)}\cdot\widehat{\sigma}^{-1}_{U,h}\cdot U_{h,T-R-(h-1)}, \nonumber
+\end{eqnarray}
+\item The critical values for the test statistic associated with the $\Gamma^{(A)}$ terms has an important parameter, which doesn't appear with my notations.
+This parameter is $\tau$, which is given by $\tau=[\lambda\cdot P_{s}]$ (recall $P_{s}=T-R-s+1$ with $s$ denoting step), where $\lambda\in[\mu,1]$.
+The $\lambda$ parameter is actually the important one, this is what shows up in critical values for the test statistic associated with the $\Gamma^{(A)}$.
+(See Table~1.)
+\begin{itemize}
+\item Barbara writes her $1$-step ahead forecast as: 
+\[\frac{1}{m}\sum_{t=R+\tau-m}^{R+\tau-1}\widehat{L}_{t+1}=A_{\tau,P}\!+\!B_{P}\!+\!U_{P} \ \mathrm{where} \  \tau=\lambda\cdot P_{1} \ \mathrm{and} \  \lambda\in[\mu,1].
+\]
+\item I write the $1$-step ahead forecast as: 
+\[\frac{1}{m}\sum_{j=0}^{m-1}\overrightarrow{\mathcal{RL}}_{R+i+1+j|R+i+j}=A_{1,i,T-R}+B_{1,T-R}+U_{1,T-R} \ \mathrm{where}\  i=0,1,\cdots,T-R-1+1-m.
+\]
+\item The $\tau$ parameter doesn't really affect the last window since $\lambda=1$ for the last window, which means that $\tau=P$. But given this Barbara's and my last windows are the same.
+\begin{itemize}
+\item For me, the last window goes from $j=0$ to $j=m-1$ for $i=T-R+1-1-m$, so from $\overrightarrow{\mathcal{RL}}_{T-m+1|T-m}$ to $\overrightarrow{\mathcal{RL}}_{T|T-1}$.
+\item For Barbara, the $\tau$ for the last window is $\tau=P_{1}$ and it goes from $t=R+\tau-m=R+P_{1}-m$ (the 1st observation) to $t=R+\tau-1=R+P_{1}-1$ (the last observation), so from $\widehat{L}_{R+P_{1}-m+1}$ to $\widehat{L}_{R+P_{1}}$.
+\item Since $T=R+P_{1}+1-1$ these are the same across the two notations.
+\end{itemize}
+\item We look at the observations for the 1st window. 
+\begin{itemize}
+\item For me, the 1st window goes from $j=0$ to $j=m-1$ for $i=0$, so from
+$\overrightarrow{\mathcal{RL}}_{R+1|R}$ to $\overrightarrow{\mathcal{RL}}_{R+m|R+m-1}$.
+\item For Barbara, the $\tau$ for the 1st window is $\tau=\lambda\cdot P_{1}=\mu\cdot P_{1}$ and it goes from $t=R+\tau-m=R+\mu\cdot P_{1}-m$ (the 1st observation) to $t=R+\tau-1=R+\mu\cdot P_{1}-1$ (the last observation), so from $\widehat{L}_{R+\mu\cdot P_{1}-m+1}$ to $\widehat{L}_{R+\mu\cdot P_{1}}$.
+\item This means that $R+\mu\cdot P_{1}-m+1=R+1$ and $R+\mu\cdot P_{1}=R+m$ or $\underline{\mu=m/P_{1}=m/(T-R)}$.
+The lowest value that Barbara reports in Table 1 is $0.1$ so we probably don't want the window that we use for looking at instability to be less than $10$~percent of the full forecast evaluation range, but this seems OK.
+\end{itemize}
+\item We look at the observations for the 2nd window. 
+\begin{itemize}
+\item For me, the 2nd window goes from $j=0$ to $j=m-1$ for $i=1$, so from
+$\overrightarrow{\mathcal{RL}}_{R+2|R+1}$ to $\overrightarrow{\mathcal{RL}}_{R+m+1|R+m}$.
+\item For Barbara, the $\tau$ for the 2nd window is $\tau=\lambda\cdot P_{1}=\widetilde{\mu}\cdot P_{1}$ and it goes from $t=R+\tau-m=R+\widetilde{\mu}\cdot P_{1}-m$ (the 1st observation) to $t=R+\tau-1=R+\widetilde{\mu}\cdot P_{1}-1$ (the last observation), so from $\widehat{L}_{R+\widetilde{\mu}\cdot P_{1}-m+1}$ to $\widehat{L}_{R+\widetilde{\mu}\cdot P_{1}}$.
+\item This means that $R+\widetilde{\mu}\cdot P_{1}-m+1=R+2$ and $R+\widetilde{\mu}\cdot P_{1}=R+m+1$ or $\widetilde{\mu}=(m+1)/P_{1}=(m+1)/(T-R)$.
+But I don't think this $\widetilde{\mu}$ is really needed for anything.
+\end{itemize}
+\item So I think Barbara's $\mu$'s for the $1$-step ahead forecasts are $\mu\in [\frac{m+0}{T-R},\frac{m+1}{T-R},\cdots,\frac{m+T-R-m}{T-R}]$ or $\mu=\frac{m+i}{T-R}$ for $i=0,1,\cdots,T-R-m$.
+I think that only
+the lowest $\mu$ that is relevant for the critical values in Table~1.
+\end{itemize}
+\item Now consider the $2$-step ahead forecasts.
+\begin{itemize}
+\item Barbara writes her $2$-step ahead forecast as: 
+\[\frac{1}{m}\sum_{t=R+\tau-m}^{R+\tau-1}\widehat{L}_{t+2}=A_{\tau,P}\!+\!B_{P}\!+\!U_{P} \ \mathrm{where} \  \tau=\lambda\cdot P_{2}=\lambda\cdot(T-R-1) \ \mathrm{and} \  \lambda\in[\mu,1].\]
+\item I write the $2$-step ahead forecast as: 
+\begin{eqnarray}\frac{1}{m}\sum_{j=0}^{m-1}\overrightarrow{\mathcal{RL}}_{R+i+2+j|R+i+j}\!\!\!\!&=&\!\!\!\!A_{2,i,T-R-1}+B_{2,T-R-1}+U_{2,T-R-1} \nonumber \\
+&&\ \ \ \ \   \ \ \ \ \   \ \ \ \ \   \mathrm{where}\  i=0,1,\cdots,T-R-2+1-m. \nonumber
+\end{eqnarray}
+\item The $\tau$ parameter doesn't really affect the last window since $\lambda=1$ for the last window, which means that $\tau=P_{2}$. But given this Barbara's and my last windows are the same.
+\begin{itemize}
+\item For me, the last window goes from $j=0$ to $j=m-1$ for $i=T-R+2-1-m$, so from $\overrightarrow{\mathcal{RL}}_{T-m+1|T-m-1}$ to $\overrightarrow{\mathcal{RL}}_{T|T-2}$.
+\item For Barbara, the $\tau$ for the last window is $\tau=P_{2}$ and it goes from $t=R+\tau-m=R+P_{2}-m$ (the 1st observation) to $t=R+\tau-1=R+P_{2}-1$ (the last observation), so from $\widehat{L}_{R+P_{2}-m+2}$ to $\widehat{L}_{R+P_{2}+1}$.
+\item Since $T=R+P_{2}+1$ these are the same across the two notations.
+\end{itemize}
+\item We look at the observations for the 1st window.
+\begin{itemize}
+\item For me, the 1st window goes from $j=0$ to $j=m-1$ for $i=0$, so from $\overrightarrow{\mathcal{RL}}_{R+2|R}$ to $\overrightarrow{\mathcal{RL}}_{R+m+1|R+m-1}$.
+\item For Barbara, the $\tau$ for the 1st window is $\tau=\lambda\cdot P_{2}=\mu\cdot P_{2}$ and it goes from $t=R+\tau-m=R+\mu\cdot P_{2}-m$ (the 1st observation) to $t=R+\tau-1=R+\mu\cdot P_{2}-1$ (the last observation), so from $\widehat{L}_{R+\mu\cdot P_{2}-m+2}$ to $\widehat{L}_{R+\mu\cdot P_{2}+1}$.
+\item This means that $R+\mu\cdot P_{2}-m+2=R+2$ and $R+\mu\cdot P_{2}+1=R+m+1$ or $\underline{\mu=m/P_{2}=m/(T-R-1)}$.
+So this means that the $\mu$ for the $2$-step ahead forecasts will be a bit different from the $\mu$ for the $1$-step ahead forecasts ($m/(T-R-1)$ versus $m/(T-R)$).
+\end{itemize}
+\item We look at the observations for the 2nd window.
+\begin{itemize}
+\item For me, the 2nd window goes from $j=0$ to $j=m-1$ for $i=1$, so from $\overrightarrow{\mathcal{RL}}_{R+3|R+1}$ to $\overrightarrow{\mathcal{RL}}_{R+m+2|R+m}$.
+\item For Barbara, the $\tau$ for the 2nd window is $\tau=\lambda\cdot P_{2}=\widetilde{\mu}\cdot P_{2}$ and it goes from $t=R+\tau-m=R+\widetilde{\mu}\cdot P_{2}-m$ (the 1st observation) to $t=R+\tau-1=R+\widetilde{\mu}\cdot P_{2}-1$ (the last observation), so from $\widehat{L}_{R+\widetilde{\mu}\cdot P_{2}-m+2}$ to $\widehat{L}_{R+\widetilde{\mu}\cdot P_{2}+1}$.
+\item This means that $R+\widetilde{\mu}\cdot P_{2}-m+2=R+3$ and $R+\widetilde{\mu}\cdot P_{2}+1=R+m+2$ or $\widetilde{\mu}=(m+1)/P_{2}=(m+1)/(T-R-1)$.
+Again, I don't think this $\widetilde{\mu}$ is really needed for anything.
+\end{itemize}
+\item So I think Barbara's $\mu$'s for the $2$ step ahead forecasts are $\mu\in [\frac{m+0}{T-R-1},\frac{m+1}{T-R-1},\cdots,$ $\frac{m+T-R-1-m}{T-R-1}]$ or $\mu=\frac{m+i}{T-R-1}$ for $i=0,1,\cdots,T-R-m-1$.
+I think that only
+the lowest $\mu$ that is relevant for the critical values in Table~1.
+\end{itemize}
+\item Now consider the $h$-step ahead forecasts.
+\begin{itemize}
+\item Barbara writes her $h$-step ahead forecast as: 
+\[
+\frac{1}{m}\sum_{t=R+\tau-m}^{R+\tau-1}\widehat{L}_{t+h}=A_{\tau,P}\!+\!B_{P}\!+\!U_{P} \ \mathrm{where} \  \tau=\lambda\cdot P_{h}=\lambda\cdot(T-R-(h-1)) \ \mathrm{and} \  \lambda\in[\mu,1].\]
+\item I write the $h$-step ahead forecast as: 
+\begin{eqnarray}
+\frac{1}{m}\sum_{j=0}^{m-1}  \overrightarrow{\mathcal{RL}}_{R+i+h+j|R+i+j}
+\!\!\!\!&=&\!\!\!\!A_{h,i,T-R-(h-1)}+B_{2,T-R-(h-1)}+U_{2,T-R-(h-1)} \nonumber \\
+&&\ \ \ \ \   \ \ \ \ \   \ \ \ \ \   \mathrm{where}\  i=0,1,\cdots,T-R-h+1-m. \nonumber
+\end{eqnarray}
+\item The $\tau$ parameter doesn't really affect the last window since $\lambda=1$ for the last window, which means that $\tau=P_{h}$. But given this Barbara's and my last windows are the same.
+\begin{itemize}
+\item For me, the last window goes from $j=0$ to $j=m-1$ for $i=T-R+h-1-m$, so from $\overrightarrow{\mathcal{RL}}_{T-m+1|T-m-(h-1)}$ to $\overrightarrow{\mathcal{RL}}_{T|T-h}$.
+\item For Barbara, the $\tau$ for the last window is $\tau=P_{h}$ and it goes from $t=R+\tau-m=R+P_{h}-m$ (the 1st observation) to $t=R+\tau-1=R+P_{h}-1$ (the last observation), so from $\widehat{L}_{R+P_{h}-m+h}$ to $\widehat{L}_{R+P_{h}+(h-1)}$.
+\item Since $T=R+P_{h}+(h-1)$ these are the same across the two notations.
+\end{itemize}
+\item We look at the observations for the 1st window.
+\begin{itemize}
+\item For me, the 1st window goes from $j=0$ to $j=m-1$ for $i=0$, so from $\overrightarrow{\mathcal{RL}}_{R+h|R}$ to $\overrightarrow{\mathcal{RL}}_{R+m+(h-1)|R+m-1}$.
+\item For Barbara, the $\tau$ for the 1st window is $\tau=\lambda\cdot P_{h}=\mu\cdot P_{h}$ and it goes from $t=R+\tau-m=R+\mu\cdot P_{h}-m$ (the 1st observation) to $t=R+\tau-1=R+\mu\cdot P_{h}-1$ (the last observation), so from $\widehat{L}_{R+\mu\cdot P_{h}-m+h}$ to $\widehat{L}_{R+\mu\cdot P_{h}+(h-1)}$.
+\item This means that $R+\mu\cdot P_{h}-m+h=R+h$ and $R+\mu\cdot P_{h}+(h-1)=R+m+(h-1)$ or $\underline{\mu=m/P_{h}=m/(T-R-(h-1))}$.
+So this means that the $\mu$ for the $h$-step ahead forecasts will be a bit different from the $\mu$ for the $1$-step ahead forecasts and the $\mu$ for the $2$-step ahead forecast ($m/(T-R-(h-1))$ versus $m/(T-R-1)$ and $m/(T-R)$).
+\end{itemize}
+\item We look at the observations for the 2nd window.
+\begin{itemize}
+\item For me, the 2nd window goes from $j=0$ to $j=m-1$ for $i=1$, so from $\overrightarrow{\mathcal{RL}}_{R+h+1|R+1}$ to $\overrightarrow{\mathcal{RL}}_{R+m+h|R+m}$.
+\item For Barbara, the $\tau$ for the 1st window is $\tau=\lambda\cdot P_{h}=\widetilde{\mu}\cdot P_{h}$ and it goes from $t=R+\tau-m=R+\widetilde{\mu}\cdot P_{h}-m$ (the 1st observation) to $t=R+\tau-1=R+\widetilde{\mu}\cdot P_{h}-1$ (the last observation), so from $\widehat{L}_{R+\widetilde{\mu}\cdot P_{h}-m+h}$ to $\widehat{L}_{R+\widetilde{\mu}\cdot P_{h}+(h-1)}$.
+\item This means that $R+\widetilde{\mu}\cdot P_{h}-m+h=R+h+1$ and $R+\widetilde{\mu}\cdot P_{h}+(h-1)=R+m+h$ or $\widetilde{\mu}=(m+1)/P_{h}=(m+1)/(T-R-(h-1))$.
+Again, I don't think this $\widetilde{\mu}$ is really needed for anything.
+\end{itemize}
+\item So I think Barbara's $\mu$'s for the $h$ step ahead forecasts are $\mu\in [\frac{m+0}{T-R-(h-1)},\frac{m+1}{T-R-(h-1)},$ $\cdots,\frac{m+T-R-(h-1)-m}{T-R-(h-1)}]$ or $\mu=\frac{m+i}{T-R-(h-1)}$ for $i=0,1,\cdots,T-R-m-(h-1)$.
+I think that only
+the lowest $\mu$ that is relevant for the critical values in Table~1.
+\end{itemize}
 \item When Barbara does the exchange rate forecasting example the ultimate finding is not so interesting, since basically she just reports that ``the lack of out-of-sample predictive ability is related to the lack of in-sample predictive ability.''
@@
 \end{itemize}
+
+\textbf{Doing PLS with lagged dependent variables}
+\begin{itemize}
+\item My reading of Groen and Kapetanios (NY Fed Staff Report no. 327, 2009) is that the following is how we would implement PLS with lagged dependent variables.
+\item Note that the example given is Groen and Kapetanios sounds like it too is for DAR type forecasts but I would think that this would work with RAR type forecasts too.
+\item The model is:
+\[ y_{t}=\alpha+\beta^{\prime}_{\mathsf{1}}f^{pls}_{\mathsf{1},t-1}+\beta^{\prime}_{\mathsf{2}}f^{pls}_{\mathsf{2},t-1}+\beta^{\prime}_{\mathsf{3}}f^{pls}_{\mathsf{3},t-1}+\sum_{i=1}^{p}\rho_{i}y_{t-i}+\epsilon_{t}\]
+\begin{itemize}
+\item I stopped at three factors because all the other models that we have been looking at have been using three factors.
+\end{itemize}
+\item The first thing we need to do is de-mean the variable that we are interested in forecasting, $y_{t}$, and the large number of variables that we are trying to compress with PLS, that is $x^{s}_{t-1}$, where $v=1,2,\cdots S$.
+\begin{itemize}
+\item We denote the de-meaned variables by $y^{d}_{t}$ and $x^{d}_{s,t-1}$.
+\end{itemize}
+\item The way Groen and Kapetanios explain what to do with the lagged dependent variable is that ``one needs to control for the effect of $y_{t-1}$, $y_{t-2}$, $\cdots$, $y_{t-p}$ (or $y^{d}_{t-1}$, $y^{d}_{t-2}$, $\cdots$, $y^{d}_{t-p}$, though I don't think it makes a difference) on the covariances between $y^{d}_{t}$ and $x^{d}_{s,t-1}$.''
+\begin{itemize}
+\item Doing this -- which is done in the next bullet point -- is basically the Frish-Waugh~(1933) theorem (see Greene page 180).
+This says that if you have a model $f_{t}=\kappa\cdot g_{t}+\lambda\cdot h_{t}+\epsilon_{t}$ you will retrieve the same estimate of $\kappa$ from estimating this equation as you would by (i)~regressing $f_{t}$ on $h_{t}$ and saving $\widehat{\nu}^{f}_{t}=f_{t}-\widehat{\theta}^{f}\cdot h_{t}$,
+(ii)~regressing $g_{t}$ on $h_{t}$ and saving $\widehat{\nu}^{g}_{t}=g_{t}-\widehat{\theta}^{g}\cdot h_{t}$, and (iii)~then regressing $\widehat{\nu}^{f}_{t}$ on $\widehat{\nu}^{g}_{t}$.
+That is, $\widehat{\nu}^{f}_{t}=\kappa\widehat{\nu}^{g}_{t}+\omega_{t}$ would give you the same estimate of $\kappa$.
+\end{itemize}
+\item So then one regresses:
+\begin{eqnarray}
+y^{d}_{t}&=&\sum_{i=1}^{p}\gamma_{i}\cdot y_{t-i}+u_{t} \ \ \ \  \ \ \ \ \textrm{and} \  \textrm{saves} \  
+\widehat{u}_{t}=y^{d}_{t}-\sum_{i=1}^{p}\widehat{\gamma}_{i}\cdot y_{t-i} \nonumber \\
+x^{d}_{1,t-1}&=&\sum_{i=1}^{p}\delta_{1,i}\cdot y_{t-i}+v_{1,t-1} \ \ \textrm{and} \  \textrm{saves} \  
+\widehat{v}_{1,t-1}=x^{d}_{1,t-1}-\sum_{i=1}^{p}\widehat{\delta}_{1,i}\cdot y_{t-i}  \nonumber \\
+x^{d}_{2,t-1}&=&\sum_{i=1}^{p}\delta_{2,i}\cdot y_{t-i}+v_{2,t-1} \ \ \textrm{and} \  \textrm{saves} \  
+\widehat{v}_{2,t-1}=x^{d}_{2,t-1}-\sum_{i=1}^{p}\widehat{\delta}_{2,i}\cdot y_{t-i}  \nonumber \\
+&&\cdots  \nonumber \\
+&&\cdots  \nonumber \\
+x^{d}_{S,t-1}&=&\sum_{i=1}^{p}\delta_{S,i}\cdot y_{t-i}+v_{S,t-1} \ \ \textrm{and} \  \textrm{saves} \  
+\widehat{v}_{S,t-1}=x^{d}_{S,t-1}-\sum_{i=1}^{p}\widehat{\delta}_{S,i}\cdot y_{t-i} \nonumber 
+\end{eqnarray}
+\item So then what I think one wants to do the PLS on is on $\widehat{u}_{t}$ and the large number of variables $\widehat{v}_{1,t-1}$, $\widehat{v}_{2,t-1}$, $\cdots$, $\widehat{v}_{S,t-1}$ and I think one can follow the algorithm that Groen and Kapetanios describe for doing this.
+\begin{itemize}
+\item The variables we start with are $\widehat{u}_{t}$ and the large number of variables $\widehat{v}_{1,t-1}$, $\widehat{v}_{2,t-1}$, $\cdots$, $\widehat{v}_{S,t-1}$.
+\item Determine the $S\times1$ vector of indicator variables or loadings $w_{\mathsf{1}}=\left(w_{\mathsf{1},1}, w_{\mathsf{1},2},\cdots,w_{\mathsf{1},S}\right)^{\prime}$ for the \underline{first} PLS factor by computing individual covariances $w_{\mathsf{1},s}=Cov(\widehat{u}_{t},\widehat{v}_{s,t-1})$, for $s=1,2,\cdots,S$.
+\begin{itemize}
+\item Because the $\widehat{v}_{s,t-1}$ variables have unit variance $w_{\mathsf{1},s}=Cov(\widehat{u}_{t},\widehat{v}_{s,t-1})/Var(\widehat{v}_{s,t-1})$, which means that we could calculate each $w_{\mathsf{1},s}$ by regressing $\widehat{u}_{t}$ on each $\widehat{v}_{s,t-1}$ and saving the regression coefficient.
+This sort of gives this analogy to simple model averaging.
+\end{itemize}
+\item Construct the \underline{first} PLS factor as:
+\[ f_{\mathsf{1},t-1}=\sum_{s=1}^{S}w_{\mathsf{1},v}\cdot\widehat{v}_{s,t-1}=\sum_{s=1}^{S}Cov(\widehat{u}_{t},\widehat{v}_{s,t-1})\cdot\widehat{v}_{s,t-1}. \]
+\item For the next PLS factor we want to think about the part of $\widehat{u}_{t}$ and the $\widehat{v}_{s,t-1}$ terms that is orthogonal to the first factor $f_{\mathsf{1},t}$ so one regresses:
+\begin{eqnarray}
+\widehat{u}_{t}&=&\gamma_{\mathsf{1}}\cdot f_{\mathsf{1},t-1}+\widehat{\widehat{u}}_{t} \ \ \ \ \ \ \   \ \textrm{and} \  \textrm{saves} \  
+\widehat{\widehat{u}}_{t}=\widehat{u}_{t}-\widehat{\gamma}_{\mathsf{1}}\cdot f_{\mathsf{1},t-1} \nonumber \\
+\widehat{v}_{1,t-1}&=&\delta_{\mathsf{1},1}\cdot f_{\mathsf{1},t-1}+\widehat{\widehat{v}}_{1,t-1} \ \ \textrm{and} \  \textrm{saves} \  
+\widehat{\widehat{v}}_{1,t-1}=\widehat{v}_{1,t-1}-\widehat{\delta}_{\mathsf{1},1}\cdot f_{\mathsf{1},t-1}  \nonumber \\
+\widehat{v}_{2,t-1}&=&\delta_{\mathsf{1},2}\cdot f_{\mathsf{1},t-1}+\widehat{\widehat{v}}_{2,t-1} \ \ \textrm{and} \  \textrm{saves} \  
+\widehat{\widehat{v}}_{2,t-1}=\widehat{v}_{2,t-1}-\widehat{\delta}_{\mathsf{1},2}\cdot f_{\mathsf{1},t-1}  \nonumber \\
+&&\cdots  \nonumber \\
+&&\cdots  \nonumber \\
+\widehat{v}_{S,t-1}&=&\delta_{\mathsf{1},S}\cdot f_{\mathsf{1},t-1}+\widehat{\widehat{v}}_{S,t-1} \ \ \textrm{and} \  \textrm{saves} \  
+\widehat{\widehat{v}}_{S,t-1}=\widehat{v}_{S,t-1}-\widehat{\delta}_{\mathsf{1},S}\cdot f_{\mathsf{1},t-1} \nonumber 
+\end{eqnarray}
+\item Then the variables we work with are $\widehat{\widehat{u}}_{t}$ and the large number of variables $\widehat{\widehat{v}}_{1,t-1}$, $\widehat{\widehat{v}}_{2,t-1}$, $\cdots$, $\widehat{\widehat{v}}_{S,t-1}$.
+\item Determine the $S\times1$ vector of indicator variables or loadings $w_{\mathsf{2}}=\left(w_{\mathsf{2},1}, w_{\mathsf{2},2},\cdots,w_{\mathsf{2},S}\right)^{\prime}$ for the \underline{second} PLS factor by computing individual covariances $w_{\mathsf{2},s}=Cov(\widehat{\widehat{u}}_{t},\widehat{\widehat{v}}_{s,t-1})$, for $s=1,2,\cdots,S$.
+\begin{itemize}
+\item Because the $\widehat{\widehat{v}}_{s,t-1}$ variables have unit variance $w_{\mathsf{1},s}=Cov(\widehat{\widehat{u}}_{t},\widehat{\widehat{v}}_{s,t-1})/Var(\widehat{\widehat{v}}_{s,t-1})$, which means that we could calculate each $w_{\mathsf{2},s}$ by regressing $\widehat{\widehat{u}}_{t}$ on each $\widehat{\widehat{v}_{s,t-1}}$ and saving the regression coefficient.
+Simple model averaging doesn't really get to this point.
+The analogy is really just there for the first factor.
+\end{itemize}
+\item Construct the \underline{second} PLS factor as:
+\[ f_{\mathsf{2},t-1}=\sum_{s=1}^{S}w_{\mathsf{2},s}\cdot\widehat{\widehat{v}}_{s,t-1}=\sum_{s=1}^{S}Cov(\widehat{\widehat{u}}_{t},\widehat{\widehat{v}}_{s,t-1})\cdot\widehat{\widehat{v}}_{s,t-1}. \]
+\item For the next PLS factor we want to think about the part of $\widetilde{\overline{y}}^{d}_{t+1}$ and the $\widetilde{\overline{x}}^{d}_{v,t}$ terms that is orthogonal to the first factor $f_{I,t}$ so one regresses:
+\begin{eqnarray}
+\widehat{\widehat{u}}_{t}&=&\gamma_{\mathsf{2}}\cdot f_{\mathsf{2},t-1}+\widehat{\widehat{\widehat{u}}}_{t} \ \ \ \ \ \ \  \ \textrm{and} \  \textrm{saves} \  
+\widehat{\widehat{\widehat{u}}}_{t}=\widehat{\widehat{u}}_{t}-\widehat{\gamma}_{\mathsf{2}}\cdot f_{\mathsf{2},t-1} \nonumber \\
+\widehat{\widehat{v}}_{1,t-1}&=&\delta_{\mathsf{2},1}\cdot f_{\mathsf{2},t-1}+\widehat{\widehat{\widehat{v}}}_{1,t-1} \ \ \textrm{and} \  \textrm{saves} \  
+\widehat{\widehat{\widehat{v}}}_{1,t-1}=\widehat{\widehat{v}}_{1,t-1}-\widehat{\delta}_{\mathsf{2},1}\cdot f_{\mathsf{2},t-1}  \nonumber \\
+\widehat{\widehat{v}}_{2,t-1}&=&\delta_{\mathsf{2},2}\cdot f_{\mathsf{2},t-1}+\widehat{\widehat{\widehat{v}}}_{2,t-1} \ \ \textrm{and} \  \textrm{saves} \  
+\widehat{\widehat{\widehat{v}}}_{2,t-1}=\widehat{\widehat{v}}_{2,t-1}-\widehat{\delta}_{\mathsf{2},2}\cdot f_{\mathsf{2},t-1}  \nonumber \\
+&&\cdots  \nonumber \\
+&&\cdots  \nonumber \\
+\widehat{\widehat{v}}_{S,t-1}&=&\delta_{\mathsf{2},S}\cdot f_{\mathsf{2},t-1}+\widehat{\widehat{\widehat{v}}}_{S,t-1} \ \ \textrm{and} \  \textrm{saves} \  
+\widehat{\widehat{\widehat{v}}}_{S,t-1}=\widehat{\widehat{v}}_{S,t-1}-\widehat{\delta}_{\mathsf{2},S}\cdot f_{\mathsf{2},t-1} \nonumber 
+\end{eqnarray}
+\item Then the variables we work with are $\widehat{\widehat{\widehat{u}}}_{t}$ and the large number of variables $\widehat{\widehat{\widehat{v}}}_{1,t-1}$, $\widehat{\widehat{\widehat{v}}}_{2,t-1}$, $\cdots$, $\widehat{\widehat{\widehat{v}}}_{S,t-1}$.
+\item Determine the $S\times1$ vector of indicator variables or loadings $w_{\mathsf{3}}=\left(w_{\mathsf{3},1}, w_{\mathsf{3},2},\cdots,w_{\mathsf{3},S}\right)^{\prime}$ for the \underline{third} PLS factor by computing individual covariances $w_{\mathsf{3},s}=Cov(\widehat{\widehat{\widehat{u}}}_{t},\widehat{\widehat{\widehat{v}}}_{s,t-1})$, for $s=1,2,\cdots,S$.
+\begin{itemize}
+\item Because the $\widehat{\widehat{\widehat{v}}}_{s,t-1}$ variables have unit variance $w_{\mathsf{1},s}=Cov(\widehat{\widehat{\widehat{u}}}_{t},\widehat{\widehat{\widehat{v}}}_{s,t-1})/Var(\widehat{\widehat{\widehat{v}}}_{s,t-1})$, which means that we could calculate each $w_{\mathsf{2},s}$ by regressing $\widehat{\widehat{\widehat{u}}}_{t}$ on each $\widehat{\widehat{\widehat{v}}}_{s,t-1}$ and saving the regression coefficient.
+Again, simple model averaging doesn't really get to this point.
+The analogy is really just there for the first factor.
+\end{itemize}
+\item Construct the \underline{third} PLS factor as:
+\[ f_{\mathsf{3},t-1}=\sum_{s=1}^{S}w_{\mathsf{3},s}\cdot\widehat{\widehat{\widehat{v}}}_{s,t-1}=\sum_{s=1}^{S}Cov(\widehat{\widehat{\widehat{u}}}_{t},\widehat{\widehat{\widehat{v}}}_{s,t-1})\cdot\widehat{\widehat{\widehat{v}}}_{s,t-1}.
+\]
+\end{itemize}
+\item Note that the example given is Groen and Kapetanios sounds like it too is for DAR type forecasts but I would think that this would work with RAR type forecasts too.
+\item Then I think one can just go ahead and estimate the model below (which is the same as the one given at the begining of this section):
+\[ y_{t}=\alpha+\beta^{\prime}_{\mathsf{1}}f^{pls}_{\mathsf{1},t-1}+\beta^{\prime}_{\mathsf{2}}f^{pls}_{\mathsf{2},t-1}+\beta^{\prime}_{\mathsf{3}}f^{pls}_{\mathsf{3},t-1}+\sum_{i=1}^{p}\rho_{i}y_{t-i}+\epsilon_{t}\]
+\item Without lagged dependent variables the above model with just the first PLS factor would be the same as the simple model averaging approach.
+I think it would still be the same with just the first PLS factor and the lagged dependent variables.
+(Presumably we could check this.)  I think having the additional factors goes beyond the simple model averaging approach.
+\end{itemize}
+%\begin{itemize}
+%\item 
+%\end{itemize}
 

### 2013-09-05 MATLAB/nim_lsc_model_2steps/run_calc_rmse_conditional.m
Added changes to implement Barbara Rossi's tests.

@@ -79,13 +79,17 @@ smoothed_factors = xi1tHistory(1:3,:);
 lag = 1;
 
 % Table 1 -- Shortened sample:
-[rmse_forecast_combination_mat1, forecast_combination_mat1] = calc_rmse_forecast_combination_conditional(nims(:,15:end),yields(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, 1,4);
+[rmse_forecast_combination_mat1, forecast_combination_mat1, insample_forecast_combination_mat1,...
+ forecast_combination_errors_mat1 insample_forecast_combination_errors_mat1] = calc_rmse_forecast_combination_conditional(nims(:,15:end),yields(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, 1,4);
+
+
 
 nims_change = nims(:,15:end)-nims(:,14:end-1);
 nims_change_out_of_sample = nims_change(out_of_sample_start_pos-14:end);
 nims_change_pre_out_of_sample = nims_change(out_of_sample_start_pos-15);
 
 [rmse_forecast_combination_change_mat1, forecast_combination_change_mat1] = calc_rmse_forecast_combination_conditional(nims_change,yields(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, 1,4);
+
 [rmse_forecast_combination_level_mat1, forecast_combination_level_mat1] = rmse_change2level(forecast_combination_change_mat1,nims_change_out_of_sample,nims_change_pre_out_of_sample); 
 
 
@@ -106,7 +110,40 @@ npls = 5;
 varlag=4;
 [rmse_varmat1, forecast_var_mat1] = calc_rmse_var_conditional(nims(:,15:end), smoothed_factors(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, varlag);
 [rmse_varmat2, forecast_var_mat2] = calc_rmse_var_conditional(nims(:,15:end), factors(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, varlag);
-[rmse_nochangemat1, forecast_nochange_mat1] = calc_rmse_nochange(nims(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon);
[…]

### 2013-09-05 MATLAB/nim_lsc_model_2steps/utilities_2steps/calc_rmse_nochange.m

@@ -1,4 +1,4 @@
-function [rmse_nochangemat, forecast_mat, forecast_errors_mat] = calc_rmse_nochange(yobs, out_of_sample_start_pos, end_sample_pos, forecast_horizon)
+function [rmse_nochangemat, forecast_mat, forecast_errors_mat, insample_forecast_errors_mat] = calc_rmse_nochange(yobs, out_of_sample_start_pos, end_sample_pos, forecast_horizon)
 %This function computes rmse for a nochange forecast and also returns
 %forecast history over a pseudo-out of sample portion of the sample.
 %Inputs: 
@@ -17,12 +17,13 @@ function [rmse_nochangemat, forecast_mat, forecast_errors_mat] = calc_rmse_nocha
 num_periods = end_sample_pos - out_of_sample_start_pos + 1;
 forecast_errors_mat = nan*ones(num_periods, forecast_horizon);
 
-mat_counter = 0;
-
 forecast_mat = kron(yobs(out_of_sample_start_pos-1:end_sample_pos-1)',ones(1,forecast_horizon));
+insample_forecast_mat = yobs(out_of_sample_start_pos-2:end_sample_pos-2);
+
 
 [y_mat] = make_y_mat(yobs(out_of_sample_start_pos:end_sample_pos)',forecast_horizon);
 
 forecast_errors_mat = forecast_mat -y_mat;
+insample_forecast_errors_mat = insample_forecast_mat' - yobs(out_of_sample_start_pos:end)';
 
 rmse_nochangemat = (nansum((forecast_errors_mat.^2))./(size(forecast_errors_mat,1)-sum(isnan(forecast_errors_mat)))).^0.5;

### 2013-09-08 MATLAB/nim_lsc_model_2steps/utilities_2steps/calc_rmse_pls_conditional.m
changed code to parse out effect of lagged dependent variable before running PLS

@@ -51,21 +51,42 @@ for pos_index = out_of_sample_start_pos:end_sample_pos
     % if yobs has more than one row, fold all the rows after the first in
     
     ylag = 1;
-
-
-    yreg = transpose(this_yobs(2:end));
-    xreg = [transpose(this_yobs(1:end-1)) , transpose(this_other_obs(:,1:end-1))];
-
-    [XL,YL,XS,YS,beta_pls] = plsregress(xreg,yreg,npls);
     
+    mean_this_yobs = mean(this_yobs);
+    this_yobs_demeaned = this_yobs-mean_this_yobs;
+    
+    mean_this_other_obs = mean(this_other_obs,2);
+    this_other_obs_demeaned = this_other_obs-kron(mean_this_other_obs,ones(1,size(this_other_obs,2)));
+
+    beta_y = regress(this_yobs_demeaned(2:end)',this_yobs_demeaned(1:end-1)');
+    u_y = this_yobs_demeaned(2:end)' - this_yobs_demeaned(1:end-1)'*beta_y;
+    
+    %u_x = zeros(n_other_obs,size(this_yobs_demeaned,2)-1);
+    beta_x = zeros(1,n_other_obs);  
+    for x_pos =1:n_other_obs
+        beta_x(x_pos) = regress(this_other_obs_demeaned(x_pos,2:end)',this_yobs_demeaned(1:end-1)');
+    end
+    u_x = this_other_obs_demeaned(:,2:end)'-kron(this_yobs_demeaned(1:end-1)',beta_x);
+
+
+    %[XL,YL,XS,YS,beta_pls] = plsregress(u_x,u_y,npls);
+    
+    % the PLS factors are in T
+    % u_x *R = T
+    [B,C,P,T,U,R,R2X,R2Y]=plssim(u_x,u_y,npls);
+    
+    beta_pls = regress(this_yobs_demeaned(2:end)',[this_yobs_demeaned(1:end-1)',T]);
     
     this_forecast_horizon = min(end_sample_pos-pos_index+1,10);
     forecast = zeros(1,this_forecast_horizon);
     previous_forecast = this_yobs(1,end);
     for this_step = 1:this_forecast_horizon
-        % figure out dimensions
-        forecast(this_step) = [1, previous_forecast, transpose(other_obs_out_of_sample(:,this_step)) ]*...
-                              beta_pls;
+        % to finish off -- 
+        % construct u_x
+        % apply 
[…]

### 2013-09-08 LATEX/beg.draft.2013.tex
inserted table with Rossi test
added PLS and to Test Table added Figure to paper showing forecasts added table with rolling RMSEs

@@
 Our findings have implications beyond the study of net interest margins.
-Scenario-based stress tests have been used across the world to rebuild confidence in the banking sector in periods of crises.
+Scenario-based stress tests have been used across the world to boost confidence in the banking sector in periods of crises.
 Yet, they are predicated on the assumption that different scenarios for the evolution of the macroeconomy have sufficiently different implications for the evolution of key balance sheet items.
@@
 \center
-\caption{RMSE for baseline specifications} \label{table_rmse_baseline}
+\caption{RMSE for baseline specifications -- Recursive Estimation Windows} \label{table_rmse_baseline}
 \begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|}
@@
 
-\newpage
-{\normalsize \clearpage
+\begin{sidewaystable}       
+\caption{RMSE for baseline specifications -- Rolling Estimation Windows} \label{table_rmse_baseline}                                                                              
+\center                                                                                           
+\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|}                                                          
+\hline                                                                                            
+&Step 1 &Step 2 &Step 3 &Step 4 &Step 5 &Step 6 &Step 7 &Step 8 &Step 9 &Step 10\\                
+\hline                                                                                            
+1. F. Combination - Yields          &0.113&0.135&0.150&0.165&0.180&0.190&0.203&0.205&0.205&0.210\\
+3a. DFM with 2nd Step Reg.          &0.111&0.146&0.174&0.209&0.244&0.279&0.307&0.333&0.369&0.403\\
+3b. DFM with F. Combination         &0.113&0.168&0.211&0.263&0.324&0.372&0.419&0.461&0.508&0.547\\
+4. PCR                              &0.118&0.167&0.208&0.259&0.307&0.355&0.397&0.432&0.472&0.508\\
+5. PLS                              &0.111&0.147&0.173&0.195&0.215&0.234&0.251&0.268&0.300&0.333\\
+6. F. Combination - Observed Factors&0.113&0.155&0.189&0.235&0.293&0.347&0.398&0.452&0.510&0.564\\
+7. VAR on Observed Factors          &0.871&0.906&0.562&1.040&1.094&0.786&0.872&0.459&0.409&0.468\\
+8. No-Change Forecast               &0.107&0.138&0.154&0.181&0.206&0.228&0.244&0.248&0.271&0.293\\
+\hline                                                                                            
+\end{tabular}                                                                                     
+\end{sidewaystable}
+
+\normalsize \clearpage
+\renewcommand{\baselinestretch}{1} 
+\begin{table}  
+\caption{Decomposing the Relative Performance of 1-Step Ahead Forecasts}                                                        
+\center                                                                
+\begin{tabular}{|l|l|c|c|}                                             
+\hline                                                                 
+&&m = 4&m=12 \\                                                        
+\hline                                                                 
+1. F. Combination - Yields          & DMW&\multicolumn{2}{|c|}{1.170}\\
+\hline                                                                 
+& $\Gamma_P^{(A)}$&11.153*&2.477\\                                     
+& $\Gamma_P^{(B)}$&3.056*&3.056*\\                                     
+& $\Gamma_P^{(U)}$&0.490&0.490\\                                       
+\hline                                                                 
+3a. DFM with 2nd Step Reg.          & DMW&\multicolumn{2}{|c|}{0.478}\\
+\hline                                                                 
+& $\Gamma_P^{(A)}$&7.684&3.727\\                                       
+& $\Gamma_P^{(B)}$&-0.806&-0.806\\                                     
+& $\Gamma_P^{(U)}$&1.146&1.146\\                                       
+\hline                                                                 
+3b. DFM with F. Combination         & DMW&\multicolumn{2}{|c|}{0.773}\\
+\hline                                                                 
+& $\Gamma_P^{(A)}$&6.841&2.441\\                                       
+& $\Gamma_P^{(B)}$&2.923*&2.923*\\                                     
+& $\Gamma_P^{(U)}$&0.769&0.769\\                                       
+\hline                                                                 
+4. PCR                              & DMW&\multicolumn{2}{|c|}{1.040}\\
+\hline                                                                 
+& $\Gamma_P^{(A)}$&8.324&3.873\\                                       
+& $\Gamma_P^{(B)}$&-0.059&-0.059\\                                     
+& $\Gamma_P^{(U)}$&1.445&1.445\\                                       
+\hline                                                                 
+5. PLS                              & DMW&\multicolumn{2}{|c|}{0.638}\\
+\hline                                                                 
+& $\Gamma_P^{(A)}$&7.105&3.710\\                                       
+& $\Gamma_P^{(B)}$&-0.630&-0.630\\                                     
+& $\Gamma_P^{(U)}$&1.209&1.209\\                                       
+\hline                                                                 
+6. F. Combination - Observed Factors& DMW&\multicolumn{2}{|c|}{0.855}\\
+\hline                                                                 
+& $\Gamma_P^{(A)}$&6.518&2.863\\                                       
+& $\Gamma_P^{(B)}$&-2.711*&-2.711*\\                                   
+& $\Gamma_P^{(U)}$&2.025*&2.025*\\                                     
+\hline                                                                 
+7. VAR on Observed Factors          & DMW&\multicolumn{2}{|c|}{1.250}\\
+\hline                                                                 
+& $\Gamma_P^{(A)}$&9.040&2.208\\                                       
+& $\Gamma_P^{(B)}$&1.836&1.836\\                                       
+& $\Gamma_P^{(U)}$&-1.249&-1.249\\                                     
+\hline                                                                 
+\end{tabular}                                                          
+\end{table}
+
+\clearpage
+\renewcommand{\baselinestretch}{1.5} \normalsize
+
 \begin{sidewaystable}
@@
 \end{sidewaystable}
-}
 
-\newpage \clearpage
+\clearpage
 \begin{sidewaystable}
@@
 \flushleft
+\caption{Pseudo-Out-of-Sample Forecasts} \label{figure_forecasts}
+ \includegraphics[scale=0.85]{figure_forecasts.ps}
+\end{figure}
+
+\clearpage
+\begin{figure}
+\flushleft
 \caption{Forecast for NIMs Conditional on CCAR 2012 Scenarios} \label{figure_scenario}

### 2013-09-08 MATLAB/nim_lsc_model_2steps/run_calc_rmse_conditional.m

@@ -110,6 +110,12 @@ npc = 3;
 
 npls = 3;
 [rmse_pls_mat1, forecast_pls_mat1] = calc_rmse_pls_conditional(nims(:,15:end), yields(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, npls);
+[rmse_pls_rolling_mat1, forecast_pls_rolling_mat1, ...
+    pls_forecast_errors_mat, insample_pls_forecast_errors_mat] = ...
+    calc_rmse_pls_conditional_rolling(nims(:,15:end), yields(:,15:end), ...
+    out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, npls);
+
+
 
 [rmse_forecast_combination_mat3, forecast_combination_mat3] = calc_rmse_forecast_combination_conditional(nims(:,15:end),factors(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, 1,4);
 [rmse_forecast_combination_rolling_mat3, forecast_combination_rolling_mat3 forecast_combination_errors_mat3 insample_forecast_combination_errors_mat3] = calc_rmse_forecast_combination_conditional_rolling(nims(:,15:end),factors(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, 1,4);
@@ -165,7 +171,14 @@ test_results = run_rs_tests(pc_forecast_errors_mat(:,1),...
                             insample_pc_forecast_errors_mat,...
                             insample_forecast_errors_nochange_mat,...
                             qn,m);
-test_table = [test_table; test_results];  
+test_table = [test_table; test_results]; 
+
+test_results = run_rs_tests(pls_forecast_errors_mat(:,1),...
+                            forecast_errors_nochange_mat(:,1),...
+                            insample_pls_forecast_errors_mat,...
+                            insample_forecast_errors_nochange_mat,...
+                            qn,m);
+test_table = [test_table; test_results]; 
 
 test_results = run_rs_tests(forecast_combination_errors_mat3(:,1),...
                             forecast_errors_nochange_mat(:,1),...
@@ -215,6 +228,13 @@ test_results = run_rs_tests(pc_forecast_errors_mat(:,1),...
                             qn,m);
 test_table2 = [test_table2; test_results];  
 
+test_results = run_rs_tests(pls_forecast_errors_mat(:,1),...
+                            forecast_errors_nochange_mat(:,1),...
+                            insample_pls_forecast_errors_mat,...
[…]

### 2013-09-08 MATLAB/nim_lsc_model_2steps/utilities_2steps/calc_rmse_pls_conditional_rolling.m

@@ -1,0 +1,124 @@
+function [rmse_mat, forecast_mat,forecast_errors_mat,insample_forecast_errors_mat] = calc_rmse_pls_conditional_rolling(yobs, other_obs, out_of_sample_start_pos, end_sample_pos, forecast_horizon, npls )
+
+% FOR NOW -- reglag does not affect anything. The regression is run with
+% only one lag of other_obs
+
+%This function computes rmse, forecasts, and forecast errors using partial
+% least squares.
+%
+%Inputs: 
+%yobs is a row vector, a series spanning entire available sample.
+%other_obs is a matrix, each row is an additional series entering the
+%seond step regression and spanning the entire available sample. NB: The rmse is going to be
+%conditional on everything in other_var_obs. Needs to have the same number
+%of columns as yobs.
+%out_of_sample_start_pos is the position of the first pseudo-out of sample
+%observation.
+%end_sample_pos is the position of the last pseudo-out of sample
+%observation.
+%npls is the number of the pls factors to be used.
+%
+%Outputs:
+% rmse_mat is a row vector containing RMSEs for each forecast step.
+% forecast_mat is a matrix containing a forecast for each period in the
+% pseudo-out of sample. Each row of forecast_mat contains a forecast for
+% the horizon desired.
+
+
+n_other_obs = size(other_obs, 1);
+
+num_periods = end_sample_pos - out_of_sample_start_pos + 1;
+forecast_errors_mat = nan*ones(num_periods, forecast_horizon);
+forecast_mat = nan*ones(num_periods, forecast_horizon);
+insample_forecast_mat = nan*ones(num_periods, 1);
+
+sample_size = out_of_sample_start_pos-1;
+
+for pos_index = out_of_sample_start_pos:end_sample_pos
+    
+    
+    % strategy keep this part unchanged -- do not lag here. 
+    if  end_sample_pos - pos_index + 1  >= forecast_horizon 
+        
+        % notice that since everything is lagged in the regression, it was
+        % convenient to move these back by 1 period.
+    
+        other_obs_out_of_sample = other_obs(:,pos_index-1:pos_index+forecast_horizon-1-1);
+    else
[…]

### 2013-09-09 MATLAB/estimation_utilities/dmw_test.m
Added DMW test at all horizons

@@ -1,0 +1,18 @@
+function [DMW, GWp, sigma] = dmw_test(out_of_sample_errors_model1,...
+                                      out_of_sample_errors_model2,...
+                                      qn,sl)
+
+DeltaL_oos = out_of_sample_errors_model1.^2-...
+             out_of_sample_errors_model2.^2;
+
+
+sigma = sqrt(nw(DeltaL_oos,qn));
+
+
+DMW  = sqrt(size(DeltaL_oos,1))*mean( DeltaL_oos)/sigma;
+    GWp = 1-cdf('chi2',DMW^2,1);
+    if (GWp>sl)
+        GWp = 0;
+    else
+        GWp = 1;
+    end

### 2013-09-09 MATLAB/nim_lsc_model_2steps/run_calc_rmse_conditional.m

@@ -83,7 +83,7 @@ lag = 1;
  forecast_combination_errors_mat1 insample_forecast_combination_errors_mat1] = calc_rmse_forecast_combination_conditional(nims(:,15:end),yields(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, 1,4);
 
 [rmse_forecast_combination_rolling_mat1, forecast_combination_rolling_mat1,...
- forecast_combination_errors_mat1 insample_forecast_combination_errors_mat1] = calc_rmse_forecast_combination_conditional_rolling(nims(:,15:end),yields(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, 1,4);
+ forecast_combination_errors_rolling_mat1 insample_forecast_combination_errors_rolling_mat1] = calc_rmse_forecast_combination_conditional_rolling(nims(:,15:end),yields(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, 1,4);
 
 
 nims_change = nims(:,15:end)-nims(:,14:end-1);
@@ -93,44 +93,51 @@ nims_change_pre_out_of_sample = nims_change(out_of_sample_start_pos-15);
 [rmse_forecast_combination_level_mat1, forecast_combination_level_mat1] = rmse_change2level(forecast_combination_change_mat1,nims_change_out_of_sample,nims_change_pre_out_of_sample); 
 
 
-[rmse_multivariate_mat1, forecast_multivariate_mat1] = calc_rmse_multivariate_conditional(nims(:,15:end), smoothed_factors(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, lag);
+[rmse_multivariate_mat1, forecast_multivariate_mat1, ...
+ multivariate_errors_mat1] = calc_rmse_multivariate_conditional(nims(:,15:end), smoothed_factors(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, lag);
 
 [rmse_multivariate_rolling_mat1, forecast_multivariate_rolling_mat1 ...
- multivariate_errors_mat1 insample_multivariate_errors_mat1] = calc_rmse_multivariate_conditional_rolling(nims(:,15:end), smoothed_factors(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, lag);
[…]

### 2013-09-09 MATLAB/nim_lsc_model_2steps/utilities_2steps/calc_rmse_multivariate_conditional.m

@@ -1,4 +1,4 @@
-function [rmse_mat, forecast_mat] = calc_rmse_multivariate_conditional(yobs, other_obs, out_of_sample_start_pos, end_sample_pos, forecast_horizon, reglag)
+function [rmse_mat, forecast_mat, forecast_errors_mat] = calc_rmse_multivariate_conditional(yobs, other_obs, out_of_sample_start_pos, end_sample_pos, forecast_horizon, reglag)
 
 % FOR NOW -- reglag does not affect anything. The regression is run with
 % only one lag of other_obs

### 2013-09-09 MATLAB/nim_lsc_model_2steps/utilities_2steps/calc_rmse_pc.m

@@ -1,4 +1,4 @@
-function [rmse_mat, forecast_mat] = calc_rmse_pc(yobs, other_obs, out_of_sample_start_pos, end_sample_pos, forecast_horizon, npc )
+function [rmse_mat, forecast_mat, forecast_errors_mat] = calc_rmse_pc(yobs, other_obs, out_of_sample_start_pos, end_sample_pos, forecast_horizon, npc )
 
 % FOR NOW -- reglag does not affect anything. The regression is run with
 % only one lag of other_obs

### 2013-09-09 MATLAB/nim_lsc_model_2steps/utilities_2steps/calc_rmse_pls_conditional.m

@@ -1,4 +1,4 @@
-function [rmse_mat, forecast_mat] = calc_rmse_pls_conditional(yobs, other_obs, out_of_sample_start_pos, end_sample_pos, forecast_horizon, npls )
+function [rmse_mat, forecast_mat, forecast_errors_mat] = calc_rmse_pls_conditional(yobs, other_obs, out_of_sample_start_pos, end_sample_pos, forecast_horizon, npls )
 
 % FOR NOW -- reglag does not affect anything. The regression is run with
 % only one lag of other_obs

### 2013-09-09 MATLAB/nim_lsc_model_2steps/utilities_2steps/calc_rmse_var_conditional.m

@@ -1,4 +1,4 @@
-function [rmse_varmat, forecast_mat] = calc_rmse_var_conditional(yobs, other_var_obs, out_of_sample_start_pos, end_sample_pos, forecast_horizon, varlag)
+function [rmse_varmat, forecast_mat, forecast_errors_mat] = calc_rmse_var_conditional(yobs, other_var_obs, out_of_sample_start_pos, end_sample_pos, forecast_horizon, varlag)
 %This function computes rmse for a VAR forecast and also returns
 %forecast history over a pseudo-out of sample portion of the sample.
 %

### 2013-09-09 MATLAB/plot_support/table_dmw_latex.m

@@ -1,0 +1,44 @@
+function table=table_dmw_latex(resultmat,sigmat,columnlabels,rowlabels)
+
+
+nrows = size(resultmat,1);
+ncols = size(resultmat,2);
+
+
+header = char('\begin{table}','\center');
+header = cellstr(header);
+
+tabular = ['\begin{tabular}{|l|'];
+
+for i=1:ncols
+    tabular = [tabular,'c|'];
+end
+tabular = [tabular,'}'];
+
+cellarray='';
+for i=1:ncols
+    cellarray=[cellarray,'&',columnlabels(i,:)];
+end
+cellarray = [cellarray,'\\'];
+
+
+table = cellstr(header);
+table = [table;cellstr(tabular);cellstr('\hline');cellstr(cellarray);cellstr('\hline')];
+for i=1:nrows
+    line = rowlabels(i,:);
+    for j=1:ncols
+        line = [line,'&',num2str(resultmat(i,j),'%5.3f')];   
+        if sigmat(i,j)
+            line = [line,'*'];
+        end
+    end
+    line = [line,'\\'];
+    table = [table;cellstr(line)];
+end
+table = [table;cellstr('\hline')];
+footer = char('\end{tabular}','\end{table}');
+
+table = [table;cellstr(footer)];
+
+
+

### 2013-09-12 MATLAB/nim_lsc_model_2steps/run_calc_rmse_conditional_long_sample.m
completed firm-specific estimates

@@ -1,3 +1,4 @@
+
 clear;
 
 setpath;
@@ -14,7 +15,7 @@ dataset_option = 1;
     total_interest_earning_assets,...
     assets_depository_inst, assets_securities_notrade,...
     assets_fedfunds, assets_all_loans, assets_trading_accnts, ... 
-    interest_income_to_ie_assets, interest_expense_to_ie_assets] = load_data_ml(dataset_option);
+    interest_income_to_ie_assets, interest_expense_to_ie_assets] = load_data_ml(dataset_option,2013);
  
 out_of_sample_start_pos = find(dates==2000.0);
 end_sample_pos = length(dates);
@@ -80,9 +81,6 @@ lag = 1;
 % Table 1 -- Shortened sample:
 [rmse_forecast_combination_mat1, forecast_combination_mat1,...
  forecast_combination_errors_mat1 insample_forecast_combination_errors_mat1] = calc_rmse_forecast_combination_conditional(nims(:,15:end),yields(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, 1,4);
-
-
-
 
 [rmse_forecast_combination_rolling_mat1, forecast_combination_rolling_mat1,...
  forecast_combination_errors_rolling_mat1 insample_forecast_combination_errors_rolling_mat1] = calc_rmse_forecast_combination_conditional_rolling(nims(:,15:end),yields(:,15:end), out_of_sample_start_pos-14, end_sample_pos-14, forecast_horizon, 1,4);
@@ -457,11 +455,11 @@ for i = horizon_plot_list
    end
    title(['Assessing the ',num2str(i),'-step-ahead forecast'])
    
-   xlim([dates(50) dates(end)])
+   xlim([dates(15) dates(end)])
 end
 
 
-
+% 
 
 % Table 2 -- full sample
 [rmse_forecast_combination_mat1, forecast_multivariate_mat1] = calc_rmse_forecast_combination_conditional(nims,yields, out_of_sample_start_pos, end_sample_pos, forecast_horizon, 1,4);

### 2013-09-12 MATLAB/nim_lsc_model_2steps/run_calc_rmse_conditional_pro_forma.m

@@ -1,0 +1,347 @@
+
+clear;
+
+setpath;
+
+
+dataset_option = 1;
+sample_start = 1989.25; %1996;
+sample_end = 2008.25; %2013;
+    
+[yobs, dates, yields, nims, nfactors, nothers, tau, factors,...
+    shadow_bank_share_assets,...
+    total_interest_earning_assets,...
+    assets_depository_inst, assets_securities_notrade,...
+    assets_fedfunds, assets_all_loans, assets_trading_accnts, ... 
+    interest_income_to_ie_assets, interest_expense_to_ie_assets] = load_data_ml(dataset_option,sample_end,sample_start);
+ 
+
+% get list of banks
+pro_forma_data_info = load_pro_forma_data;
+
+bank_ids = pro_forma_data_info{1};
+bank_names = pro_forma_data_info{2};
+
+% overwrite nims with bank-specific data
+%bank_data= load_pro_forma_data(bank_ids(1));
+%nims = bank_data{1};
+
+
+out_of_sample_start_pos = find(dates==2000);
+end_sample_pos = length(dates);
+
+nstates = nfactors + nothers;
+ntaus = length(tau);
+
+forecast_horizon = 10;
+
+
+
+%% get RMSEs
+lag = 1;
+
+% Table 1 -- Shortened sample:
+[rmse_forecast_combination_mat1, forecast_combination_mat1,...
+ forecast_combination_errors_mat1 insample_forecast_combination_errors_mat1] = calc_rmse_forecast_combination_conditional(nims,yields, out_of_sample_start_pos, end_sample_pos, forecast_horizon, 1,4);
+
+[rmse_forecast_combination_rolling_mat1, forecast_combination_rolling_mat1,...
+ forecast_combination_errors_rolling_mat1 insample_forecast_combination_errors_rolling_mat1] = calc_rmse_forecast_combination_conditional_rolling(nims,yields, out_of_sample_start_pos, end_sample_pos, forecast_horizon, 1,4);
+
+
+npc = 3;
+[rmse_pc_mat1, forecast_pc_mat1,...
+    pc_forecast_errors_mat] = calc_rmse_pc(nims, yields, out_of_sample_start_pos, end_sample_pos, forecast_horizon, npc);
+[rmse_pc_rolling_mat1, forecast_pc_rolling_mat1,...
[…]

### 2013-09-12 MATLAB/nim_lsc_model_2steps/run_calc_rmse_conditional_pro_forma2.m

@@ -1,0 +1,350 @@
+
+clear;
+
+setpath;
+
+
+dataset_option = 1;
+sample_start = 1996; %1989.25; %1996;
+sample_end = 2013; %2008.25; %2013;
+    
+[yobs, dates, yields, nims, nfactors, nothers, tau, factors,...
+    shadow_bank_share_assets,...
+    total_interest_earning_assets,...
+    assets_depository_inst, assets_securities_notrade,...
+    assets_fedfunds, assets_all_loans, assets_trading_accnts, ... 
+    interest_income_to_ie_assets, interest_expense_to_ie_assets] = load_data_ml(dataset_option,sample_end,sample_start);
+ 
+
+out_of_sample_start_pos = find(dates==2007.75);
+end_sample_pos = length(dates);
+
+nstates = nfactors + nothers;
+ntaus = length(tau);
+
+forecast_horizon = 10;
+
+% get list of banks
+pro_forma_data_info = load_pro_forma_data;
+
+bank_ids = pro_forma_data_info{1};
+bank_names = pro_forma_data_info{2};
+
+nbanks = length(bank_ids);
+
+bank_rmse_forecast_combination_recursive = nan*ones(nbanks,forecast_horizon);
+bank_dmw_forecast_combination_recursive = nan*ones(nbanks,forecast_horizon);
+bank_dmw_sig_forecast_combination_recursive = nan*ones(nbanks,forecast_horizon);
+
+bank_rmse_forecast_combination_rolling = nan*ones(nbanks,forecast_horizon);
+bank_dmw_forecast_combination_rolling = nan*ones(nbanks,forecast_horizon);
+bank_dmw_sig_forecast_combination_rolling = nan*ones(nbanks,forecast_horizon);
+
+
+bank_rmse_pcr_recursive = nan*ones(nbanks,forecast_horizon);
+bank_dmw_pcr_recursive = nan*ones(nbanks,forecast_horizon);
+bank_dmw_sig_pcr_recursive = nan*ones(nbanks,forecast_horizon);
+
+bank_rmse_pcr_rolling = nan*ones(nbanks,forecast_horizon);
+bank_dmw_pcr_rolling = nan*ones(nbanks,forecast_horizon);
+bank_dmw_sig_pcr_rolling = nan*ones(nbanks,forecast_horizon);
+
+
+bank_rmse_pls_recursive = nan*ones(nbanks,forecast_horizon);
[…]

### 2013-09-12 MATLAB/nim_lsc_model_2steps/run_scenario.m

@@ -154,7 +154,7 @@ hold on
 plot(forecast_dates,severe_scen(1,1:10),'r--');
 title('3-Month Treasury Yields')
 legend('Baseline','Severely Adverse Scenario','Location','NorthWest')
-xlim([2012 2015])
+xlim([2012.75 2015])
 ylim([0 4])
 
 
@@ -165,7 +165,7 @@ plot(forecast_dates,adverse_scen(1,1:10),'r--');
 
 title('3-Month Treasury Yields')
 legend('Baseline','Adverse Scenario','Location','NorthWest')
-xlim([2012 2015])
+xlim([2012.75 2015])
 ylim([0 4])
 
 subplot(3,2,3)
@@ -174,7 +174,7 @@ hold on
 plot(forecast_dates,severe_scen(2,1:10),'r--');
 title('10-Year Treasury Yields')
 legend('Baseline','Severely Adverse Scenario','Location','NorthWest')
-xlim([2012 2015])
+xlim([2012.75 2015])
 ylim([0 5])
 
 
@@ -185,28 +185,28 @@ plot(forecast_dates,adverse_scen(2,1:10),'r--');
 
 title('10-Year Treasury Yields')
 legend('Baseline','Adverse Scenario','Location','NorthWest')
-xlim([2012 2015])
+xlim([2012.75 2015])
 ylim([0 5])
 
 
 subplot(3,2,5)
 plot(forecast_dates, forecast_baseline,'k'); hold on
 plot(forecast_dates, forecast_severe,'r--')
-plot(forecast_dates, forecast_severe+rmse_forecast_combination_mat4,'r-.')
+plot(forecast_dates, forecast_baseline+rmse_forecast_combination_mat4,'k-.')
 title('Forecast of NIMs')
 legend('Baseline','Severely Adverse','1-RMSE band','Location','SouthWest')
-
-plot(forecast_dates, forecast_severe-rmse_forecast_combination_mat4,'r-.')
+xlim([2012.75 2015])
+plot(forecast_dates, forecast_baseline-rmse_forecast_combination_mat4,'k-.')
 
 
 
 subplot(3,2,6)
 plot(forecast_dates, forecast_baseline,'k'); hold on
 plot(forecast_dates, forecast_adverse,'r--')
-plot(forecast_dates, forecast_adverse+rmse_forecast_combination_mat4,'r-.')
+plot(forecast_dates, forecast_baseline+rmse_forecast_combination_mat4,'k-.')
[…]

### 2013-09-12 MATLAB/nim_lsc_model_2steps/utilities_2steps/load_data_ml.m

@@ -10,10 +10,22 @@ function [yobs, dates, yields, nims, nfactors, nothers, tau, factors,...
 % 2  -- BHCs are selected by quarter-average total assets
 % 3  -- BHCs are selected by quarter-end interest earning assets
 % 4  -- BHCs are selected by quarter-average interest earning assets
-if nargin == 0
+if nargin == 0 % default if no arguments are passed
     dataset_option = 1;
-else
+    end_yobs = 2008.25;
+    start_yobs= 1985.75;
+elseif nargin == 1  % first argument is dataset option
     dataset_option = varargin{1};
+    end_yobs = 2008.25;
+    start_yobs= 1985.75;
+elseif nargin == 2  % second argument is end date of desired data
+    dataset_option = varargin{1};
+    end_yobs = varargin{2};
+    start_yobs= 1985.75;
+elseif nargin == 3  % third argument is start date of desired data
+    dataset_option = varargin{1};
+    end_yobs = varargin{2};
+    start_yobs = varargin{3};
 end
 
 nim_dataset_titles = char('nims_assets_by_endperiod_total_assets.csv',...
@@ -35,8 +47,7 @@ end
 
 
 
-start_yobs = 1985.75;
-end_yobs = 2008.25;
+%start_yobs = 1985.75;
 dates = start_yobs:.25:end_yobs;
 %SVENY yields can be found in the 'yields' FAME db under the mnemonics in
 %yields_varlist.
@@ -46,8 +57,8 @@ else
     yields = csvread('..\data\yields_sveny_only_1985quarterly.csv', 1, 2);
 end
 dates_yields = 1985.75:.25:2013.0;
-start_pos = find(dates==start_yobs);
-end_pos = find(dates==end_yobs);
+start_pos = find(dates_yields==start_yobs);
+end_pos = find(dates_yields==end_yobs);
 yields = yields(start_pos:end_pos,:)';
 yields_varlist = char('SVENY0025','SVENY0050','SVENY0075','SVENY0100',...
     'SVENY0200','SVENY0300','SVENY0500', 'SVENY0700',...

### 2013-09-12 MATLAB/nim_lsc_model_2steps/load_pro_forma_data.m
added file to load pro forma data

@@ -1,0 +1,47 @@
+function vargout=load_pro_forma_data(vargin)
+
+
+if nargin == 0
+    bank_ids = [         1027004    1037003       1039502               1068025   1068191            1069778              1070345          1073757             1074156     1078529          1111435             1119794  1120754            1131787       1132449                 1199611           1199844    1245415         1378434            1951350      2277860        3242838       3587146]';
+    bank_names = char('ZIONS BC','M\&T BK CORP','JPMORGAN CHASE \& CO','KEYCORP','HUNTINGTON BSHRS','PNC FNCL SVC GROUP','FIFTH THIRD BC','BANK OF AMER CORP','BB\&T CORP','BBVA USA BSHRS','STATE STREET CORP','U S BC','WELLS FARGO \& CO','SUNTRUST BK','RBS CITIZENS FNCL GRP','NORTHERN TR CORP','COMERICA','BMO FNCL CORP','UNIONBANCAL CORP','CITIGROUP','CAPITAL ONE FC','REGIONS FC','BANK OF NY MELLON CORP');
+    
+    if isunix
+        pro_forma = csvread('../data/pro_forma/may9_nim_emre_luca.csv', 1, 2);
+    else
+        pro_forma = csvread('..\data\pro_forma\may9_nim_emre_luca.csv', 1, 2);
+    end
+    
+    
+    bank_dates = 1990:.25:2013;
+    nobs = length(bank_dates);
+    nbanks = length(bank_ids);
+    bank_assets=pro_forma(nobs:nobs:nobs*nbanks,6);
+    [tmp pos_sort]=sort(-bank_assets);
+    
+    bank_ids=bank_ids(pos_sort);
+    bank_names = bank_names(pos_sort,:);
+    
+    vargout{1} = bank_ids;
+    vargout{2} = bank_names;
+elseif nargin == 1
+    bankid = vargin;
+    if isunix
+        pro_forma = csvread('../data/pro_forma/may9_nim_emre_luca.csv', 1, 2);
+    else
+        pro_forma = csvread('..\data\pro_forma\may9_nim_emre_luca.csv', 1, 2);
+    end
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    %entity	nim	nim_ta roa roe assets_avg
+    bank_dates = 1990:.25:2013;
+    bank_nim=pro_forma(find(pro_forma(:,1)==bankid),2);
[…]

### 2013-09-12 MATLAB/plot_support/tablelatex.m
table routine

@@ -28,10 +28,10 @@ for i=1:nrows
     line = rowlabels(i,:);
     for j=1:ncols
         if (~isempty(strmatch('s.e.',rowlabels(i,:))))
-            line = [line,'& (',num2str(resultmat(i,j),'%5.3f'),')'];   
+            line = [line,'& (',num2str(resultmat(i,j),'%5.2f'),')'];   
         
         else 
-            line = [line,'&',num2str(resultmat(i,j),'%5.3f')];   
+            line = [line,'&',num2str(resultmat(i,j),'%5.2f')];   
         end
     end
     line = [line,'\\'];

### 2013-09-12 LATEX/beg.draft.2013.tex
Added DMW test at all horizons
updated paper includes firm-specific tables of results

--- 2013-07-17 LATEX/beg.draft.2013.tex
@@
 \begin{abstract}
-\noindent Regular bank stress testing (and the capital planning that it implies) is one of the three complementary reforms that has been made to U.S.
-capital regulation in response to the shortcomings of the pre-crisis regime.
-The form of stress testing currently undertaken by the Federal Reserve in implementing the stress testing requirement of the Dodd-Frank Act is based around a small number of scenarios.
-This approach involves specifying macroeconomic and financial scenarios that represent stressful conditions over the time horizon of the stress test and then projecting, for each bank holding company in the stress test, its losses, income, and path of capital \textit{given that scenario}.
-In this paper, we focus on the path of income and consider the ability of empirical models to translate developments in interest rates into bank net interest margins (NIMs).
+\noindent Bank stress testing  is a central component of reforms to U.S.
+capital regulation promulgated by the Dodd-Frank Act in response to the shortcomings of the pre-crisis regime.
+The form of stress testing currently undertaken by the Federal Reserve is based around a small number of scenarios.
+This approach involves specifying macroeconomic and financial scenarios that represent stressful conditions and then projecting, for each bank holding company in the stress test, its losses, income, and path of capital \textit{given that scenario}.
+In this paper, we focus on the path of income and consider the ability of empirical models to translate developments in interest rates into bank net interest margins.
+Our findings have bearing on the use of stress tests to boost confidence in banks in periods of crisis.
 \end{abstract}
@@
 
-The NIM models estimated by \citeN{covas2012} and \citeN{hirtle2013} are developed for stress testing and are part of much larger aggregative models of a number of components of banks' financial statements.
+The NIM models estimated by \citeN{covas2012} and by \citeN{hirtle2013} are developed for stress testing and are part of much larger aggregative models of a number of components of banks' financial statements.
 These models are both estimated on panel data with the most notable difference between these models being the estimation techniques they employ: Least squares regression in the case of \citeN{hirtle2013} and quantile regression in the case of \citeN{covas2012}.
@@
 Like our paper, \citeN{english2002} focuses exclusively on NIMs, although he does so for a number of countries (that is, all G-7 countries except for France, and for Australia, Norway, Sweden, and Switzerland).
-Addditionally (and as we also do later), he considers asset yields and liability yields separately---the two measures that when differenced imply NIMs.
+Additionally (and as we also do later), he considers asset yields and liability yields separately---the two measures that when differenced imply NIMs.
 The main focus of \citeN{english2002} is understanding how well the aggregate banking sectors in the countries he studies appear to have managed the interest rate risk that they face on their earnings, as such he uses aggregate data (rather than firm-level data) and does not focus on the conditional forecasting performance of his model.
@@
 
-The unweighted forecast combination model for NIMs based on just two yields outperforms the other forecast models we considered.
-No other model comes close to outperforming a no-change forecast.
-This is only a small victory as the variance of the forecast error is large even when compared to the observed variation of NIMs.
-Accordingly, even scenarios that have vastly different implications for the configuration of interest rates contain relatively little information to forecast NIMs.
+The forecast combination of yields outperforms the no change model in forecasting net interest margins.
+As the forecast horizon lengthens, forecast improvement from the forecast combination method become more sizable.
+None of the other model considered outperforms a no-change forecast.
+This is only a small victory, as the variance of the forecast error is large even when compared to the observed variation of NIMs.
+Accordingly, even stress test scenarios that have vastly different implications for the configuration of interest rates contain relatively little information to forecast NIMs.
 
 Our findings have implications beyond the study of net interest margins.
-Scenario-based stress tests have been used across the world to rebuild confidence in the banking sector in periods of crises.
+Scenario-based stress tests have been used across the world to boost confidence in the banking sector in periods of crises.
 Yet, they are predicated on the assumption that different scenarios for the evolution of the macroeconomy have sufficiently different implications for the evolution of key balance sheet items.
-Yet, an array of time series methods found by others to be near the frontier of efficiency produces projections conditional on radically different scenarios for yields that statistically indistinguishable.
+An array of time series methods found by others to be near the frontier of efficiency produces projections conditional on radically different scenarios for yields that are statistically indistinguishable.
 
-Supervisory authorities that perform stress tests certainly have access to a wealth of supervisory information beyond what  the publicly available data that we have used.
-Yet, if the conclusions of stress tests are mostly predicated on supervisory sources, as our study implies, communication with the public is even more challenging.
+Supervisory authorities that perform stress tests have access to a wealth of supervisory information beyond what  the publicly available data that we have used.
+However, if the conclusions of stress tests gain precision from the use of supervisory sources, communication with the public is even more challenging.
 In that case, the effects of confidence of stress tests rely squarely on the reputation of the supervising authority, one which might be tarnished exactly in a period of crisis.
@@
 
-\renewcommand{\baselinestretch}{1.5}
+\renewcommand{\baselinestretch}{1.5} \normalsize
 
-\newpage
-{\normalsize \clearpage
+\clearpage
 \begin{sidewaystable}
 \center
-\caption{RMSE for baseline specifications} \label{table_rmse_baseline}
+\caption{Out-of-Sample RMSE for baseline specifications -- Recursive Estimation Window} \label{table_rmse_baseline}
 \begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|}
@@
 2. DFM                                                       &0.122&0.148&0.169&0.212&0.247&0.278&0.316&0.340&0.373&0.404\\
-3a. DFM with 2nd Step Regression           &0.118&0.167&0.209&0.260&0.308&0.356&0.398&0.433&0.474&0.511\\
-3b. DFM with Forecast Combination          &0.112&0.159&0.199&0.247&0.300&0.351&0.399&0.438&0.489&0.535\\
-4. Forecast Combination of Observed Factors&0.118&0.169&0.216&0.273&0.336&0.399&0.456&0.509&0.567&0.622\\
-5. VAR on Observed Factors                 &0.166&0.209&0.263&0.302&0.342&0.375&0.380&0.430&0.451&0.499\\
-6. No-Change Forecast                      &0.107&0.138&0.154&0.181&0.206&0.228&0.244&0.248&0.271&0.293\\
+3a. DFM with 2nd Step Reg.           &0.118&0.167&0.209&0.260&0.308&0.356&0.398&0.433&0.474&0.511\\
+3b. DFM with F. Combination          &0.112&0.159&0.199&0.247&0.300&0.351&0.399&0.438&0.489&0.535\\
+4. PCR                               &0.118&0.167&0.208&0.259&0.307&0.355&0.397&0.432&0.472&0.508\\
+5. PLS                               &0.115&0.158&0.192&0.234&0.278&0.323&0.365&0.400&0.443&0.482\\
+6. F. Combination - Observed Factors &0.118&0.169&0.216&0.273&0.336&0.399&0.456&0.509&0.567&0.622\\
+7. VAR on Observed Factors           &0.166&0.209&0.263&0.302&0.342&0.375&0.380&0.430&0.451&0.499\\
+8. No-Change Forecast                &0.107&0.138&0.154&0.181&0.206&0.228&0.244&0.248&0.271&0.293\\
 \hline
@@
 \end{sidewaystable}
-}
 
-\newpage
-{\normalsize \clearpage
 \begin{sidewaystable}
-\caption{RMSEs for models that include the share of asset of the shadow banking sector. Estimation sample: 1989Q2-2008Q2} \label{table_rmse_competition}
+\caption{Diebold-Mariano-West Statistics Relative to the No-Change Forecast -- Recursive Estimation Window}     
 \center                                                                                                     
@@
 \hline                                                                                                      
-1. Simple Forecast Combination             &0.110&0.143&0.165&0.192&0.222&0.248&0.267&0.274&0.287&0.300\\
-3a. DFM with 2nd Step Regression           &0.121&0.170&0.212&0.262&0.312&0.360&0.403&0.436&0.478&0.515\\
-3b. DFM with Forecast Combination          &0.106&0.140&0.165&0.193&0.222&0.248&0.267&0.274&0.287&0.300\\
-4. Forecast Combination of Observed Factors&0.109&0.141&0.165&0.193&0.222&0.249&0.267&0.274&0.288&0.300\\
-5. VAR on Observed Factors                 &0.344&0.396&0.476&0.494&0.521&0.569&0.631&0.670&0.699&0.718\\
-6. No-Change Forecast                      &0.107&0.138&0.154&0.181&0.206&0.228&0.244&0.248&0.271&0.293\\
+1. F. Combination - Yields          &0.102&-0.821&-0.788&-1.153&-1.445&-1.521&-1.433&-1.380&-1.775&-2.016*\\
+3a. DFM with 2nd Step Reg.          &1.049&1.010&1.142&1.320&1.430&1.566&1.675&1.811&1.902&1.913\\          
+3b. DFM with F. Combination         &0.879&1.434&1.660&1.603&1.723&1.858&2.009*&2.220*&2.587*&3.015*\\      
+4. PCR                              &1.056&1.010&1.123&1.281&1.384&1.512&1.622&1.754&1.847&1.862\\          
+5. PLS                              &0.983&0.922&1.042&1.170&1.329&1.534&1.728&1.903&1.964*&1.960\\         
+6. F. Combination - Observed Factors&1.256&1.211&1.365&1.440&1.610&1.799&1.955&2.263*&2.629*&3.055*\\       
+7. VAR on Observed Factors          &2.653*&3.596*&2.512*&2.321*&2.094*&2.242*&2.040*&2.145*&1.955&1.674\\  
 \hline                                                                                                      
@@
 \end{sidewaystable}
-}
 
-\newpage \clearpage
 \begin{sidewaystable}       
-\caption{RMSEs for variants of the simple forecast combination model} \label{table_rmse_combination}
+\caption{Out-of-Sample RMSE for baseline specifications -- Rolling Estimation Window} \label{table_rmse_baseline}                                                                              
+\center                                                                                           
+\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|}                                                          
+\hline                                                                                            
+&Step 1 &Step 2 &Step 3 &Step 4 &Step 5 &Step 6 &Step 7 &Step 8 &Step 9 &Step 10\\                
+\hline                                                                                            
+1. F. Combination - Yields          &0.113&0.135&0.150&0.165&0.180&0.190&0.203&0.205&0.205&0.210\\
+3a. DFM with 2nd Step Reg.          &0.111&0.146&0.174&0.209&0.244&0.279&0.307&0.333&0.369&0.403\\
+3b. DFM with F. Combination         &0.113&0.168&0.211&0.263&0.324&0.372&0.419&0.461&0.508&0.547\\
+4. PCR                              &0.118&0.167&0.208&0.259&0.307&0.355&0.397&0.432&0.472&0.508\\
+5. PLS                              &0.111&0.147&0.173&0.195&0.215&0.234&0.251&0.268&0.300&0.333\\
+6. F. Combination - Observed Factors&0.113&0.155&0.189&0.235&0.293&0.347&0.398&0.452&0.510&0.564\\
+7. VAR on Observed Factors          &0.871&0.906&0.562&1.040&1.094&0.786&0.872&0.459&0.409&0.468\\
+8. No-Change Forecast               &0.107&0.138&0.154&0.181&0.206&0.228&0.244&0.248&0.271&0.293\\
+\hline                                                                                            
+\end{tabular}                                                                                     
+\end{sidewaystable}
+
+\clearpage
+\begin{sidewaystable}   \caption{Diebold-Mariano-West Statistics Relative to the No-Change Forecast -- Rolling Estimation Window}
+\center                                                                                                     
+\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|}                                                                    
+\hline                                                                                                      
+&Step 1 &Step 2 &Step 3 &Step 4 &Step 5 &Step 6 &Step 7 &Step 8 &Step 9 &Step 10\\                          
+\hline                                                                                                      
+1. F. Combination - Yields          &1.037&-0.383&-0.408&-0.954&-1.431&-1.901&-1.727&-1.399&-1.879&-2.090*\\
+3a. DFM with 2nd Step Reg.          &0.473&0.372&0.595&0.723&0.871&1.052&1.164&1.353&1.375&1.297\\          
+3b. DFM with F. Combination         &0.796&1.901&1.848&1.611&1.723&1.684&1.728&1.895&2.123*&2.426*\\        
+4. PCR                              &1.056&1.010&1.123&1.281&1.384&1.512&1.622&1.754&1.847&1.862\\          
+5. PLS                              &0.630&0.590&0.729&0.424&0.268&0.161&0.202&0.485&0.550&0.567\\          
+6. F. Combination - Observed Factors&0.844&0.957&1.062&1.200&1.505&1.717&1.874&2.291*&2.619*&2.940*\\       
+7. VAR on Observed Factors          &1.328&1.288&1.542&1.402&1.307&1.786&1.556&2.792*&0.940&1.251\\         
+\hline                                                                                                      
+\end{tabular}
+\footnotesize A '*' denotes significance at the 95\% level.                                                                                               
+\end{sidewaystable}
+
+\normalsize \clearpage
+\renewcommand{\baselinestretch}{1} 
+\begin{table} 
+\caption{Decomposition of Forecast Performance Relative to the No-Change Forecast}                                                         
+\center                                                                
+\begin{tabular}{|l|l|c|c|}                                             
+\hline                                                                 
+&&m = 4&m=12 \\                                                        
+\hline                                                                 
+1. F. Combination - Yields          & DMW&\multicolumn{2}{|c|}{1.037}\\
+\hline                                                                 
+& $\Gamma_P^{(A)}$&9.885*&2.196\\                                      
+& $\Gamma_P^{(B)}$&2.817*&2.817*\\                                     
+& $\Gamma_P^{(U)}$&0.434&0.434\\                                       
+\hline                                                                 
+3a. DFM with 2nd Step Reg.          & DMW&\multicolumn{2}{|c|}{0.473}\\
+\hline                                                                 
+& $\Gamma_P^{(A)}$&7.600&3.686\\                                       
+& $\Gamma_P^{(B)}$&-0.845&-0.845\\                                     
+& $\Gamma_P^{(U)}$&1.104&1.104\\                                       
+\hline                                                                 
+3b. DFM with F. Combination         & DMW&\multicolumn{2}{|c|}{0.796}\\
+\hline                                                                 
+& $\Gamma_P^{(A)}$&7.052&2.516\\                                       
+& $\Gamma_P^{(B)}$&2.713*&2.713*\\                                     
+& $\Gamma_P^{(U)}$&0.792&0.792\\                                       
+\hline                                                                 
+4. PCR                              & DMW&\multicolumn{2}{|c|}{1.056}\\
+\hline                                                                 
+& $\Gamma_P^{(A)}$&8.455&3.934\\                                       
+& $\Gamma_P^{(B)}$&-0.062&-0.062\\                                     
+& $\Gamma_P^{(U)}$&1.442&1.442\\                                       
+\hline                                                                 
+5. PLS                              & DMW&\multicolumn{2}{|c|}{0.630}\\
+\hline                                                                 
+& $\Gamma_P^{(A)}$&7.007&3.659\\                                       
+& $\Gamma_P^{(B)}$&-0.681&-0.681\\                                     
+& $\Gamma_P^{(U)}$&1.147&1.147\\                                       
+\hline                                                                 
+6. F. Combination - Observed Factors& DMW&\multicolumn{2}{|c|}{0.844}\\
+\hline                                                                 
+& $\Gamma_P^{(A)}$&6.433&2.826\\                                       
+& $\Gamma_P^{(B)}$&-2.537*&-2.537*\\                                   
+& $\Gamma_P^{(U)}$&2.018*&2.018*\\                                     
+\hline                                                                 
+7. VAR on Observed Factors          & DMW&\multicolumn{2}{|c|}{1.328}\\
+\hline                                                                 
+& $\Gamma_P^{(A)}$&9.599&2.345\\                                       
+& $\Gamma_P^{(B)}$&1.309&1.309\\                                       
+& $\Gamma_P^{(U)}$&0.226&0.226\\                                       
+\hline                                                                 
+\end{tabular}
+
+\footnotesize \flushleft \noindent
+[Insert footnote on interpretation of table here]                                                        
+\end{table}
+
+\clearpage
+\renewcommand{\baselinestretch}{1.5} \normalsize
+
+\begin{sidewaystable}
+\caption{Out-of-Sample RMSEs (recursive estimation window) for models that include the share of asset of the shadow banking sector} \label{table_rmse_competition}
+\center
+\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|}
+\hline
+&Step 1 &Step 2 &Step 3 &Step 4 &Step 5 &Step 6 &Step 7 &Step 8 &Step 9 &Step 10\\
+\hline
+1. F. Combination - Yields           &0.107&0.133&0.148&0.168&0.187&0.202&0.213&0.209&0.216&0.223\\
+1b. F. Combination of change - Yields&0.163&0.173&0.186&0.217&0.251&0.294&0.327&0.346&0.396&0.434\\
+3a. DFM with 2nd Step Reg.           &0.118&0.167&0.209&0.260&0.308&0.356&0.398&0.433&0.474&0.511\\
+3b. DFM with F. Combination          &0.112&0.159&0.199&0.247&0.300&0.351&0.399&0.438&0.489&0.535\\
+4. PCR                               &0.118&0.167&0.208&0.259&0.307&0.355&0.397&0.432&0.472&0.508\\
+5. PLS                               &0.115&0.158&0.192&0.234&0.278&0.323&0.365&0.400&0.443&0.482\\
+6. F. Combination - Observed Factors &0.118&0.169&0.216&0.273&0.336&0.399&0.456&0.509&0.567&0.622\\
+7. VAR on Observed Factors           &0.166&0.209&0.263&0.302&0.342&0.375&0.380&0.430&0.451&0.499\\
+8. No-Change Forecast                &0.107&0.138&0.154&0.181&0.206&0.228&0.244&0.248&0.271&0.293\\\hline
+\end{tabular}
+\end{sidewaystable}
+
+\clearpage
+\begin{sidewaystable}
+\caption{Out-of-Sample RMSEs for variants of the simple forecast combination model (recursive estimation window)} \label{table_rmse_combination}
 \center
@@
 
-\newpage \clearpage
-\newpage \clearpage
+\begin{sidewaystable}     
+\caption{Out-of-Sample RMSE (rolling estimation window) -- Extended Assessment Sample Through 2013q1}                                                                  
+\center                                                                                            
+\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|}                                                
+\hline                                                                                  
+&Step 1 &Step 2 &Step 3 &Step 4 &Step 5 &Step 6 &Step 7 &Step 8 &Step 9 &Step 10\\      
+\hline                                                                                  
+1. F. Combination - Yields          &0.12&0.16&0.20&0.23&0.26&0.29&0.30&0.30&0.30&0.30\\
+3a. DFM with 2nd Step Reg.          &0.12&0.15&0.17&0.20&0.23&0.26&0.28&0.29&0.32&0.34\\
+3b. DFM with F. Combination         &0.12&0.16&0.20&0.25&0.30&0.34&0.37&0.41&0.44&0.47\\
+4. PCR                              &0.12&0.15&0.19&0.23&0.26&0.30&0.33&0.36&0.39&0.41\\
+5. PLS                              &0.12&0.14&0.17&0.19&0.21&0.23&0.24&0.26&0.28&0.31\\
+6. F. Combination - Observed Factors&0.12&0.16&0.19&0.23&0.27&0.32&0.35&0.39&0.43&0.46\\
+7. VAR on Observed Factors          &0.71&0.73&0.47&0.83&0.87&0.63&0.69&0.39&0.37&0.40\\
+8. No-Change Forecast               &0.11&0.15&0.17&0.20&0.22&0.25&0.28&0.28&0.30&0.32\\
+\hline                                                                                  
+\end{tabular}                                                                               
+\end{sidewaystable}
+
+\clearpage
+\renewcommand{\baselinestretch}{1} \normalsize 
+\begin{table} 
+\caption{Decomposition of Forecast Performance Relative to the No-Change Forecast --  Extended Assessment Window Through 2013q1 }
+\center                                                                
+\begin{tabular}{|l|l|c|c|}                                             
+\hline                                                                 
+&&m = 4&m=12 \\                                                        
+\hline                                                                 
+1. F. Combination - Yields          & DMW&\multicolumn{2}{|c|}{0.673}\\
+\hline                                                                 
+& $\Gamma_P^{(A)}$&15.190*&3.890\\                                     
+& $\Gamma_P^{(B)}$&3.634*&3.634*\\                                     
+& $\Gamma_P^{(U)}$&-0.458&-0.458\\                                     
+\hline                                                                 
+3a. DFM with 2nd Step Reg.          & DMW&\multicolumn{2}{|c|}{0.425}\\
+\hline                                                                 
+& $\Gamma_P^{(A)}$&9.933&5.123\\                                       
+& $\Gamma_P^{(B)}$&-1.910&-1.910\\                                     
+& $\Gamma_P^{(U)}$&1.098&1.098\\                                       
+\hline                                                                 
+3b. DFM with F. Combination         & DMW&\multicolumn{2}{|c|}{0.753}\\
+\hline                                                                 
+& $\Gamma_P^{(A)}$&11.879*&3.923\\                                     
+& $\Gamma_P^{(B)}$&-3.581*&-3.581*\\                                   
+& $\Gamma_P^{(U)}$&0.964&0.964\\                                       
+\hline                                                                 
+4. PCR                              & DMW&\multicolumn{2}{|c|}{0.465}\\
+\hline                                                                 
+& $\Gamma_P^{(A)}$&13.470*&6.846*\\                                    
+& $\Gamma_P^{(B)}$&-0.916&-0.916\\                                     
+& $\Gamma_P^{(U)}$&1.261&1.261\\                                       
+\hline                                                                 
+5. PLS                              & DMW&\multicolumn{2}{|c|}{0.598}\\
+\hline                                                                 
+& $\Gamma_P^{(A)}$&8.607&4.789\\                                       
+& $\Gamma_P^{(B)}$&-1.746&-1.746\\                                     
+& $\Gamma_P^{(U)}$&0.817&0.817\\                                       
+\hline                                                                 
+6. F. Combination - Observed Factors& DMW&\multicolumn{2}{|c|}{0.896}\\
+\hline                                                                 
+& $\Gamma_P^{(A)}$&10.675*&3.858\\                                     
+& $\Gamma_P^{(B)}$&-3.444*&-3.444*\\                                   
+& $\Gamma_P^{(U)}$&1.179&1.179\\                                       
+\hline                                                                 
+7. VAR on Observed Factors          & DMW&\multicolumn{2}{|c|}{1.343}\\
+\hline                                                                 
+& $\Gamma_P^{(A)}$&15.481*&4.311\\                                     
+& $\Gamma_P^{(B)}$&1.307&1.307\\                                       
+& $\Gamma_P^{(U)}$&0.313&0.313\\                                       
+\hline                                                                 
+\end{tabular}                                                          
+\end{table}
+
+\begin{sidewaystable}  
+\caption{Out-of-Sample RMSE for the 1-Step Ahead Forecast -- Bank-Specific Models} 
+\center                                                        
+\begin{tabular}{|l|c|c|c|c|c|c|c|c|}                                                                      
+\hline 
+&\multicolumn{3}{|c|}{Recursive}&\multicolumn{3}{|c|}{Rolling}    & & \\
+
+&1. F. Comb  &4. PCR      &5. PLS      &1. F. Comb  &4. PCR      &5. PLS      &8. No Change&SNL         \\
+\hline                                                                                                    
+JPMORGAN CHASE \& CO  &0.17&0.18&0.18&0.18&0.18&0.17&0.15&0.00\\                                          
+BANK OF AMER CORP     &0.24&0.24&0.25&0.25&0.24&0.30&0.26&0.00\\                                          
+CITIGROUP             &0.22&0.24&0.23&0.29&0.24&0.30&0.21&0.00\\                                          
+WELLS FARGO \& CO     &0.14&0.14&0.16&0.15&0.14&0.12&0.12&0.00\\                                          
+U S BC                &0.10&0.13&0.12&0.11&0.13&0.12&0.10&0.00\\                                          
+PNC FNCL SVC GROUP    &0.21&0.22&0.22&0.21&0.22&0.20&0.18&0.00\\                                          
+BANK OF NY MELLON CORP&0.32&0.31&0.32&0.33&0.31&0.29&0.37&0.00\\                                          
+STATE STREET CORP     &0.22&0.26&0.25&0.22&0.26&0.26&0.20&0.00\\                                          
+BB\&T CORP            &0.14&0.18&0.16&0.14&0.18&0.12&0.13&0.00\\                                          
+SUNTRUST BK           &0.25&0.21&0.16&0.14&0.21&0.13&0.10&0.00\\                                          
+CAPITAL ONE FC        &1.06&0.90&0.91&1.13&0.90&1.08&0.53&0.00\\                                          
+REGIONS FC            &0.16&0.19&0.14&0.16&0.19&0.14&0.12&0.00\\                                          
+RBS CITIZENS FNCL GRP &0.12&0.21&0.13&0.10&0.21&0.12&0.12&0.00\\                                          
+FIFTH THIRD BC        &0.30&0.32&0.32&0.30&0.32&0.29&0.34&0.00\\                                          
+BMO FNCL CORP         &1.05&1.04&1.02&1.04&1.04&1.06&1.17&0.00\\                                          
+UNIONBANCAL CORP      &0.36&0.36&0.33&0.35&0.36&0.35&0.35&0.00\\                                          
+NORTHERN TR CORP      &0.11&0.10&0.12&0.11&0.10&0.13&0.12&0.00\\                                          
+KEYCORP               &0.47&0.44&0.45&0.50&0.44&0.48&0.49&0.00\\                                          
+M\&T BK CORP          &0.11&0.14&0.15&0.12&0.14&0.13&0.12&0.00\\                                          
+BBVA USA BSHRS        &0.79&1.10&1.19&0.90&1.10&1.26&1.50&0.00\\                                          
+COMERICA              &0.16&0.20&0.17&0.17&0.20&0.19&0.17&0.00\\                                          
+HUNTINGTON BSHRS      &0.23&0.13&0.10&0.27&0.13&0.18&0.11&0.00\\                                          
+ZIONS BC              &0.23&0.19&0.22&0.22&0.19&0.21&0.20&0.00\\                                          
+\hline                                                                                                    
+\end{tabular}                                                        
+\end{sidewaystable}
+
+\renewcommand{\baselinestretch}{1.5} \normalsize
+
+\clearpage
 \begin{figure}
@@
 \center
-\includegraphics[scale=0.85]{figure_nims_components.ps}
+\hspace{-1.5cm}\includegraphics[scale=0.8]{figure_nims_components.ps}
 \end{figure}
@@
 \center
-\includegraphics[scale=0.85]{figure_nims_rates.ps}
+\includegraphics[scale=0.8]{figure_nims_rates.ps}
 \end{figure}
@@
 \begin{figure}
-%[tbp]
 \caption{NIMs and Observed Yield Factors} \label{figure_nims_factors}
 \center
-\includegraphics[scale=0.85]{figure_nims_factors.ps}
+\includegraphics[scale=0.8]{figure_nims_factors.ps}
 \end{figure}
@@
 \begin{figure}
-\flushleft
+
 \caption{NIMs and Competition -- The Asset Share of the Shadow Banking Sector} \label{figure_nims_competition}
-
-\hspace{-2.5cm} \includegraphics[scale=0.85]{figure_nims_competition.ps}
+\center
+\noindent
+\hspace{-1.5cm}\includegraphics[scale=0.8]{figure_nims_competition.ps}
 \end{figure}
@@
 \begin{figure} 
+\caption{Pseudo-Out-of-Sample Forecasts} \label{figure_forecasts}
+\center \includegraphics[scale=0.75]{figure_forecasts.ps}
 \flushleft 
+\renewcommand{\baselinestretch}{1} \normalsize 
+\footnotesize \noindent Observations for which the forecast combination of yields (our Model 1) beats the no change forecast (our Model 8) are marked with with an 'o;' observations for which the no change forecast beats the forecast combination of yields are marked with an 'x.'
+\end{figure}
+
+\clearpage
+\begin{figure}
+\center
 \caption{Forecast for NIMs Conditional on CCAR 2012 Scenarios} \label{figure_scenario}
 
- \includegraphics[scale=0.85]{figure_scenario.ps}
+ \includegraphics[scale=0.8]{figure_scenario.ps}
 \end{figure}
 
+\clearpage
+\appendix
+
+\section{Additional Bank-Specific Results}
+
+\begin{table}                                               
+\caption{Model 1, Forecast Combination, Out-of-Sample RMSEs (recursive window) in deviation from RMSEs for no-change forecast}
+\center                                                     
+\begin{tabular}{|l|c|c|c|c|c|c|}                            
+\hline                                                      
+&Step 1 &Step 2 &Step 4 &Step 6 &Step 8 &Step 10\\          
+\hline                                                      
+JPMORGAN CHASE \& CO  &0.02&0.04&0.04&0.05&0.02&-0.00\\     
+BANK OF AMER CORP     &-0.01&-0.04&-0.12&-0.14&-0.13&-0.12\\
+CITIGROUP             &0.01&0.02&0.05&0.04&0.05&0.01\\      
+WELLS FARGO \& CO     &0.02&0.05&0.04&0.01&-0.03&-0.07\\    
+U S BC                &-0.01&-0.02&0.00&0.02&0.04&0.02\\    
+PNC FNCL SVC GROUP    &0.03&0.07&0.10&0.10&0.09&0.11\\      
+BANK OF NY MELLON CORP&-0.05&-0.07&0.05&0.00&-0.08&-0.21\\  
+STATE STREET CORP     &0.02&0.04&0.11&0.11&0.27&0.52\\      
+BB\&T CORP            &0.01&0.01&0.03&0.01&0.06&0.05\\      
+SUNTRUST BK           &0.15&0.17&0.16&0.14&0.20&0.27\\      
+CAPITAL ONE FC        &0.53&0.57&0.39&0.30&0.17&0.13\\      
+REGIONS FC            &0.04&0.04&0.01&-0.04&-0.02&-0.06\\   
+RBS CITIZENS FNCL GRP &-0.00&-0.04&-0.10&-0.11&0.07&0.25\\  
+FIFTH THIRD BC        &-0.04&-0.07&0.02&0.02&0.10&0.12\\    
+BMO FNCL CORP         &-0.12&-0.26&-0.38&-0.47&-0.42&-0.53\\
+UNIONBANCAL CORP      &0.01&-0.01&0.00&-0.05&0.03&0.05\\    
+NORTHERN TR CORP      &-0.01&-0.01&0.02&-0.01&-0.04&-0.08\\ 
+KEYCORP               &-0.02&-0.04&-0.07&-0.10&-0.12&-0.02\\
+M\&T BK CORP          &-0.02&-0.02&0.01&-0.03&-0.01&-0.00\\ 
+BBVA USA BSHRS        &-0.71&-0.73&-0.72&-0.75&-0.70&-1.03\\
+COMERICA              &-0.00&-0.02&0.01&0.02&0.09&0.18\\    
+HUNTINGTON BSHRS      &0.12&0.08&0.09&0.11&0.18&0.24\\      
+ZIONS BC              &0.03&0.07&0.10&0.10&0.16&0.12\\      
+\hline                                                      
+\end{tabular}                                               
+\end{table}
+
+\begin{table} 
+\caption{Model 1, Forecast Combination (recursive window), Diebold-Mariano-West Statistics relative to a no-change forecast}
+
+\center                                                             
+\begin{tabular}{|l|c|c|c|c|c|c|}                                    
+\hline                                                              
+&Step 1 &Step 2 &Step 4 &Step 6 &Step 8 &Step 10\\                  
+\hline                                                              
+JPMORGAN CHASE \& CO  &1.246&1.071&0.813&0.789&0.391&-0.007\\       
+BANK OF AMER CORP     &-0.565&-0.838&-1.269&-1.472&-2.067*&-2.469*\\
+CITIGROUP             &0.403&0.451&0.676&0.629&0.484&0.300\\        
+WELLS FARGO \& CO     &0.852&1.166&1.085&0.189&-0.854&-1.724\\      
+U S BC                &-0.769&-1.162&0.397&0.425&0.457&0.245\\      
+PNC FNCL SVC GROUP    &2.021*&1.964*&1.812&3.445*&5.785*&2.819*\\   
+BANK OF NY MELLON CORP&-1.052&-0.817&1.019&0.092&-1.703&-5.676*\\   
+STATE STREET CORP     &0.778&1.058&1.349&1.837&2.224*&2.712*\\      
+BB\&T CORP            &0.594&0.354&0.940&0.866&2.505*&2.115*\\      
+SUNTRUST BK           &2.102*&1.838&1.960&2.583*&3.923*&3.554*\\    
+CAPITAL ONE FC        &3.072*&2.251*&1.917&3.005*&2.844*&1.148\\    
+REGIONS FC            &1.307&1.159&0.185&-0.929&-1.419&-1.081\\     
+RBS CITIZENS FNCL GRP &-0.059&-1.039&-2.538*&-1.799&0.584&1.226\\   
+FIFTH THIRD BC        &-0.729&-0.644&0.442&0.498&2.747*&2.152*\\    
+BMO FNCL CORP         &-0.718&-0.775&-0.985&-1.155&-1.143&-1.468\\  
+UNIONBANCAL CORP      &0.217&-0.157&0.017&-2.286*&0.633&1.116\\     
+NORTHERN TR CORP      &-0.683&-0.526&0.484&-0.180&-0.845&-4.202*\\  
+KEYCORP               &-0.558&-0.979&-1.873&-1.422&-1.078&-0.153\\  
+M\&T BK CORP          &-0.779&-0.631&0.210&-1.194&-0.255&-0.057\\   
+BBVA USA BSHRS        &-0.787&-0.756&-0.744&-0.862&-0.886&-1.064\\  
+COMERICA              &-0.178&-0.559&0.122&0.272&1.273&2.655*\\     
+HUNTINGTON BSHRS      &2.000*&1.600&2.067*&2.989*&5.065*&4.791*\\   
+ZIONS BC              &1.841&2.451*&1.914&1.747&9.565*&2.198*\\     
+\hline                                                              
+\end{tabular}
+
+\flushleft
+\footnotesize \noindent A star denotes significance at the 95\% level                                                      
+\end{table}
+
+\begin{table}
+\caption{Model 1, Forecast Combination, Out-of-Sample RMSEs (rolling window) in deviation from RMSEs for no-change forecast}                                               
+\center                                                     
+\begin{tabular}{|l|c|c|c|c|c|c|}                            
+\hline                                                      
+&Step 1 &Step 2 &Step 4 &Step 6 &Step 8 &Step 10\\          
+\hline                                                      
+JPMORGAN CHASE \& CO  &0.03&0.05&0.06&0.08&0.04&0.01\\      
+BANK OF AMER CORP     &-0.01&-0.03&-0.11&-0.12&-0.10&-0.11\\
+CITIGROUP             &0.09&0.16&0.17&0.10&0.10&0.07\\      
+WELLS FARGO \& CO     &0.03&0.07&0.07&0.02&-0.03&-0.06\\    
+U S BC                &0.00&0.01&0.07&0.09&0.09&0.08\\      
+PNC FNCL SVC GROUP    &0.03&0.08&0.12&0.13&0.12&0.15\\      
+BANK OF NY MELLON CORP&-0.04&-0.05&0.06&0.02&-0.07&-0.22\\  
+STATE STREET CORP     &0.01&0.03&0.09&0.07&0.22&0.49\\      
+BB\&T CORP            &0.01&0.00&0.03&0.01&0.05&0.04\\      
+SUNTRUST BK           &0.04&0.08&0.17&0.14&0.19&0.23\\      
+CAPITAL ONE FC        &0.60&0.62&0.38&0.27&0.10&0.04\\      
+REGIONS FC            &0.04&0.05&0.02&-0.05&0.01&-0.01\\    
+RBS CITIZENS FNCL GRP &-0.02&-0.06&-0.10&-0.10&-0.01&0.13\\ 
+FIFTH THIRD BC        &-0.04&-0.05&0.05&0.06&0.14&0.16\\    
+BMO FNCL CORP         &-0.13&-0.28&-0.38&-0.45&-0.41&-0.50\\
+UNIONBANCAL CORP      &-0.01&-0.04&-0.08&-0.12&-0.03&0.03\\ 
+NORTHERN TR CORP      &-0.01&-0.00&0.02&-0.01&-0.03&-0.09\\ 
+KEYCORP               &0.01&-0.00&0.02&-0.03&-0.04&0.05\\   
+M\&T BK CORP          &-0.01&-0.00&0.05&0.01&0.04&0.04\\    
+BBVA USA BSHRS        &-0.60&-0.57&-0.58&-0.67&-0.63&-0.96\\
+COMERICA              &-0.00&-0.02&-0.00&0.01&0.06&0.13\\   
+HUNTINGTON BSHRS      &0.15&0.10&0.07&0.09&0.18&0.17\\      
+ZIONS BC              &0.02&0.04&0.06&0.06&0.11&0.03\\      
+\hline                                                      
+\end{tabular}                                               
+\end{table}
+
+\begin{table} 
+\caption{Model 1, Forecast Combination (rolling window), Diebold-Mariano-West Statistics relative to a no-change forecast}                                                     
+\center                                                            
+\begin{tabular}{|l|c|c|c|c|c|c|}                                   
+\hline                                                             
+&Step 1 &Step 2 &Step 4 &Step 6 &Step 8 &Step 10\\                 
+\hline                                                             
+JPMORGAN CHASE \& CO  &1.270&1.084&1.059&1.082&0.681&0.152\\       
+BANK OF AMER CORP     &-0.287&-0.687&-1.187&-1.272&-1.598&-2.469*\\
+CITIGROUP             &1.297&1.408&1.254&1.076&1.029&1.834\\       
+WELLS FARGO \& CO     &1.184&1.510&1.366&0.426&-0.835&-1.721\\     
+U S BC                &0.156&0.184&1.360&1.079&1.002&0.867\\       
+PNC FNCL SVC GROUP    &1.920&1.976*&1.844&3.517*&5.693*&2.910*\\   
+BANK OF NY MELLON CORP&-0.783&-0.623&1.333&0.574&-1.660&-5.544*\\  
+STATE STREET CORP     &0.739&0.877&0.990&0.873&1.598&2.212*\\      
+BB\&T CORP            &0.957&0.131&1.335&1.215&2.626*&1.962*\\     
+SUNTRUST BK           &1.192&1.449&1.479&1.962*&2.958*&4.175*\\    
+CAPITAL ONE FC        &3.432*&2.578*&2.000*&2.828*&1.316&0.339\\   
+REGIONS FC            &1.359&1.487&0.707&-0.889&0.365&-0.113\\     
+RBS CITIZENS FNCL GRP &-1.336&-1.546&-2.309*&-2.264*&-0.148&0.994\\
+FIFTH THIRD BC        &-0.610&-0.420&1.078&1.779&3.947*&2.614*\\   
+BMO FNCL CORP         &-0.755&-0.833&-1.007&-1.141&-1.110&-1.347\\ 
+UNIONBANCAL CORP      &-0.192&-0.499&-1.114&-2.623*&-0.852&1.130\\ 
+NORTHERN TR CORP      &-0.397&-0.070&0.774&-0.183&-0.726&-4.868*\\ 
+KEYCORP               &0.387&-0.142&0.397&-1.128&-0.780&0.700\\    
+M\&T BK CORP          &-0.260&-0.007&1.257&0.362&1.777&1.449\\     
+BBVA USA BSHRS        &-0.687&-0.622&-0.623&-0.798&-0.803&-1.017\\ 
+COMERICA              &-0.033&-0.542&-0.028&0.089&0.963&2.357*\\   
+HUNTINGTON BSHRS      &1.606&1.244&1.104&1.889&3.549*&3.242*\\     
+ZIONS BC              &1.802&2.002*&1.403&0.976&5.704*&0.930\\     
+\hline                                                             
+\end{tabular}
+
+\flushleft
+\footnotesize \noindent A star denotes significance at the 95\% level                                                                                                           
+\end{table}
+
+\begin{table} 
+\caption{Model 4, Principal Component Regression, Out-of-Sample RMSEs (recursive window) in deviation from RMSEs for no-change forecast}                                              
+\center                                                     
+\begin{tabular}{|l|c|c|c|c|c|c|}                            
+\hline                                                      
+&Step 1 &Step 2 &Step 4 &Step 6 &Step 8 &Step 10\\          
+\hline                                                      
+JPMORGAN CHASE \& CO  &0.03&0.06&0.08&0.06&-0.02&-0.10\\    
+BANK OF AMER CORP     &-0.01&-0.03&-0.08&-0.09&-0.08&-0.06\\
+CITIGROUP             &0.03&0.05&0.10&0.10&0.12&0.05\\      
+WELLS FARGO \& CO     &0.02&0.01&-0.04&-0.10&-0.19&-0.23\\  
+U S BC                &0.02&0.05&0.10&0.14&0.13&0.12\\      
+PNC FNCL SVC GROUP    &0.04&0.08&0.14&0.14&0.16&0.18\\      
+BANK OF NY MELLON CORP&-0.06&-0.07&0.07&0.02&-0.05&-0.13\\  
+STATE STREET CORP     &0.06&0.13&0.22&0.30&0.30&0.24\\      
+BB\&T CORP            &0.05&0.10&0.17&0.19&0.23&0.25\\      
+SUNTRUST BK           &0.11&0.19&0.29&0.33&0.41&0.47\\      
+CAPITAL ONE FC        &0.37&0.47&0.37&0.26&0.11&0.10\\      
+REGIONS FC            &0.07&0.10&0.12&0.10&0.13&0.15\\      
+RBS CITIZENS FNCL GRP &0.09&0.17&0.29&0.32&0.41&0.48\\      
+FIFTH THIRD BC        &-0.02&-0.04&0.04&0.08&0.16&0.18\\    
+BMO FNCL CORP         &-0.13&-0.19&-0.23&-0.30&-0.33&-0.45\\
+UNIONBANCAL CORP      &0.01&0.03&0.08&0.03&0.10&0.12\\      
+NORTHERN TR CORP      &-0.02&-0.03&-0.03&-0.09&-0.18&-0.25\\
+KEYCORP               &-0.05&-0.01&0.09&0.09&0.13&0.17\\    
+M\&T BK CORP          &0.01&0.03&0.08&0.06&0.11&0.13\\      
+BBVA USA BSHRS        &-0.40&-0.64&-0.99&-1.16&-1.30&-1.51\\
+COMERICA              &0.03&0.05&0.09&0.11&0.18&0.28\\      
+HUNTINGTON BSHRS      &0.01&0.01&0.01&-0.01&0.00&0.00\\     
+ZIONS BC              &-0.01&0.01&0.03&0.01&0.03&-0.04\\    
+\hline                                                      
+\end{tabular}                                               
+\end{table}
+
+\begin{table}
+\caption{Model 4, Principal Component Regression (recursive window), Diebold-Mariano-West Statistics relative to a no-change forecast}                                                       
+\center                                                             
+\begin{tabular}{|l|c|c|c|c|c|c|}                                    
+\hline                                                              
+&Step 1 &Step 2 &Step 4 &Step 6 &Step 8 &Step 10\\                  
+\hline                                                              
+JPMORGAN CHASE \& CO  &1.585&1.431&0.940&0.484&-0.161&-0.777\\      
+BANK OF AMER CORP     &-0.324&-0.376&-0.851&-0.946&-0.900&-0.718\\  
+CITIGROUP             &1.113&0.840&0.861&0.720&1.082&0.332\\        
+WELLS FARGO \& CO     &0.548&0.292&-1.096&-2.949*&-3.555*&-2.542*\\ 
+U S BC                &0.645&0.621&0.717&0.867&0.924&0.964\\        
+PNC FNCL SVC GROUP    &1.464&1.451&1.395&1.027&0.904&1.134\\        
+BANK OF NY MELLON CORP&-1.140&-0.717&0.624&0.121&-0.370&-1.180\\    
+STATE STREET CORP     &1.469&1.510&1.190&1.042&0.897&0.767\\        
+BB\&T CORP            &1.785&1.577&1.251&1.089&1.270&1.521\\        
+SUNTRUST BK           &2.681*&2.349*&2.136*&2.030*&2.672*&3.587*\\  
+CAPITAL ONE FC        &2.284*&1.898&1.292&1.365&1.733&1.214\\       
+REGIONS FC            &2.534*&2.030*&1.124&0.661&0.796&0.818\\      
+RBS CITIZENS FNCL GRP &1.523&1.310&1.169&1.079&1.241&1.543\\        
+FIFTH THIRD BC        &-0.735&-0.487&0.535&0.788&1.630&1.922\\      
+BMO FNCL CORP         &-0.818&-0.644&-0.708&-0.927&-1.067&-1.521\\  
+UNIONBANCAL CORP      &0.238&0.706&1.434&0.587&1.784&1.429\\        
+NORTHERN TR CORP      &-1.112&-1.347&-0.685&-1.913&-3.124*&-5.452*\\
+KEYCORP               &-0.693&-0.158&0.564&0.386&0.508&0.659\\      
+M\&T BK CORP          &0.776&0.673&0.771&0.429&0.704&0.981\\        
+BBVA USA BSHRS        &-0.982&-0.983&-1.045&-1.129&-1.169&-1.230\\  
+COMERICA              &1.267&0.980&0.805&0.659&1.020&1.589\\        
+HUNTINGTON BSHRS      &0.800&0.278&0.215&-0.093&0.037&0.061\\       
+ZIONS BC              &-0.809&0.284&0.549&0.242&0.731&-2.058*\\     
+\hline                                                              
+\end{tabular}
+
+\flushleft
+\footnotesize \noindent A star denotes significance at the 95\% level
+
+\end{table}
+
+\begin{table} 
+\caption{Model 4, Principal Component Regression, Out-of-Sample RMSEs (rolling window) in deviation from RMSEs for no-change forecast}                                                  
+\center                                                     
+\begin{tabular}{|l|c|c|c|c|c|c|}                            
+\hline                                                      
+&Step 1 &Step 2 &Step 4 &Step 6 &Step 8 &Step 10\\          
+\hline                                                      
+JPMORGAN CHASE \& CO  &0.03&0.06&0.08&0.06&-0.02&-0.10\\    
+BANK OF AMER CORP     &-0.01&-0.03&-0.08&-0.09&-0.08&-0.06\\
+CITIGROUP             &0.03&0.05&0.10&0.10&0.12&0.05\\      
+WELLS FARGO \& CO     &0.02&0.01&-0.04&-0.10&-0.19&-0.23\\  
+U S BC                &0.02&0.05&0.10&0.14&0.13&0.12\\      
+PNC FNCL SVC GROUP    &0.04&0.08&0.14&0.14&0.16&0.18\\      
+BANK OF NY MELLON CORP&-0.06&-0.07&0.07&0.02&-0.05&-0.13\\  
+STATE STREET CORP     &0.06&0.13&0.22&0.30&0.30&0.24\\      
+BB\&T CORP            &0.05&0.10&0.17&0.19&0.23&0.25\\      
+SUNTRUST BK           &0.11&0.19&0.29&0.33&0.41&0.47\\      
+CAPITAL ONE FC        &0.37&0.47&0.37&0.26&0.11&0.10\\      
+REGIONS FC            &0.07&0.10&0.12&0.10&0.13&0.15\\      
+RBS CITIZENS FNCL GRP &0.09&0.17&0.29&0.32&0.41&0.48\\      
+FIFTH THIRD BC        &-0.02&-0.04&0.04&0.08&0.16&0.18\\    
+BMO FNCL CORP         &-0.13&-0.19&-0.23&-0.30&-0.33&-0.45\\
+UNIONBANCAL CORP      &0.01&0.03&0.08&0.03&0.10&0.12\\      
+NORTHERN TR CORP      &-0.02&-0.03&-0.03&-0.09&-0.18&-0.25\\
+KEYCORP               &-0.05&-0.01&0.09&0.09&0.13&0.17\\    
+M\&T BK CORP          &0.01&0.03&0.08&0.06&0.11&0.13\\      
+BBVA USA BSHRS        &-0.40&-0.64&-0.99&-1.16&-1.30&-1.51\\
+COMERICA              &0.03&0.05&0.09&0.11&0.18&0.28\\      
+HUNTINGTON BSHRS      &0.01&0.01&0.01&-0.01&0.00&0.00\\     
+ZIONS BC              &-0.01&0.01&0.03&0.01&0.03&-0.04\\    
+\hline                                                      
+\end{tabular}                                               
+\end{table}
+
+\begin{table} 
+\caption{Model 4, Principal Component Regression (rolling window), Diebold-Mariano-West Statistics relative to a no-change forecast}
+
+\center                                                             
+\begin{tabular}{|l|c|c|c|c|c|c|}                                    
+\hline                                                              
+&Step 1 &Step 2 &Step 4 &Step 6 &Step 8 &Step 10\\                  
+\hline                                                              
+JPMORGAN CHASE \& CO  &1.585&1.431&0.940&0.484&-0.161&-0.777\\      
+BANK OF AMER CORP     &-0.324&-0.376&-0.851&-0.946&-0.900&-0.718\\  
+CITIGROUP             &1.113&0.840&0.861&0.720&1.082&0.332\\        
+WELLS FARGO \& CO     &0.548&0.292&-1.096&-2.949*&-3.555*&-2.542*\\ 
+U S BC                &0.645&0.621&0.717&0.867&0.924&0.964\\        
+PNC FNCL SVC GROUP    &1.464&1.451&1.395&1.027&0.904&1.134\\        
+BANK OF NY MELLON CORP&-1.140&-0.717&0.624&0.121&-0.370&-1.180\\    
+STATE STREET CORP     &1.469&1.510&1.190&1.042&0.897&0.767\\        
+BB\&T CORP            &1.785&1.577&1.251&1.089&1.270&1.521\\        
+SUNTRUST BK           &2.681*&2.349*&2.136*&2.030*&2.672*&3.587*\\  
+CAPITAL ONE FC        &2.284*&1.898&1.292&1.365&1.733&1.214\\       
+REGIONS FC            &2.534*&2.030*&1.124&0.661&0.796&0.818\\      
+RBS CITIZENS FNCL GRP &1.523&1.310&1.169&1.079&1.241&1.543\\        
+FIFTH THIRD BC        &-0.735&-0.487&0.535&0.788&1.630&1.922\\      
+BMO FNCL CORP         &-0.818&-0.644&-0.708&-0.927&-1.067&-1.521\\  
+UNIONBANCAL CORP      &0.238&0.706&1.434&0.587&1.784&1.429\\        
+NORTHERN TR CORP      &-1.112&-1.347&-0.685&-1.913&-3.124*&-5.452*\\
+KEYCORP               &-0.693&-0.158&0.564&0.386&0.508&0.659\\      
+M\&T BK CORP          &0.776&0.673&0.771&0.429&0.704&0.981\\        
+BBVA USA BSHRS        &-0.982&-0.983&-1.045&-1.129&-1.169&-1.230\\  
+COMERICA              &1.267&0.980&0.805&0.659&1.020&1.589\\        
+HUNTINGTON BSHRS      &0.800&0.278&0.215&-0.093&0.037&0.061\\       
+ZIONS BC              &-0.809&0.284&0.549&0.242&0.731&-2.058*\\     
+\hline                                                              
+\end{tabular}
+
+\flushleft
+\footnotesize \noindent A star denotes significance at the 95\% level
+
+\end{table}
+
+\begin{table}  
+\caption{Model 5, Partial Least Squares, Out-of-Sample RMSEs (recursive window) in deviation from RMSEs for no-change forecast}                                                 
+\center                                                     
+\begin{tabular}{|l|c|c|c|c|c|c|}                            
+\hline                                                      
+&Step 1 &Step 2 &Step 4 &Step 6 &Step 8 &Step 10\\          
+\hline                                                      
+JPMORGAN CHASE \& CO  &0.03&0.06&0.08&0.05&-0.03&-0.13\\    
+BANK OF AMER CORP     &-0.01&-0.03&-0.09&-0.08&-0.06&-0.03\\
+CITIGROUP             &0.02&0.05&0.13&0.14&0.18&0.10\\      
+WELLS FARGO \& CO     &0.04&0.06&0.05&-0.01&-0.10&-0.21\\   
+U S BC                &0.02&0.03&0.05&0.07&0.02&-0.03\\     
+PNC FNCL SVC GROUP    &0.05&0.08&0.13&0.10&0.07&0.00\\      
+BANK OF NY MELLON CORP&-0.05&-0.06&0.05&0.03&-0.06&-0.22\\  
+STATE STREET CORP     &0.05&0.12&0.22&0.29&0.30&0.25\\      
+BB\&T CORP            &0.03&0.05&0.06&0.04&0.02&-0.02\\     
+SUNTRUST BK           &0.05&0.08&0.07&0.04&0.03&-0.04\\     
+CAPITAL ONE FC        &0.38&0.49&0.49&0.42&0.19&0.06\\      
+REGIONS FC            &0.02&0.01&-0.07&-0.11&-0.09&-0.02\\  
+RBS CITIZENS FNCL GRP &0.01&0.02&0.07&0.09&0.15&0.20\\      
+FIFTH THIRD BC        &-0.02&-0.03&0.02&0.01&0.02&-0.01\\   
+BMO FNCL CORP         &-0.16&-0.15&-0.11&-0.08&-0.04&-0.22\\
+UNIONBANCAL CORP      &-0.03&-0.06&-0.09&-0.16&-0.09&-0.06\\
+NORTHERN TR CORP      &-0.00&-0.01&-0.00&-0.03&-0.10&-0.15\\
+KEYCORP               &-0.03&-0.03&-0.07&-0.13&-0.16&-0.21\\
+M\&T BK CORP          &0.02&0.04&0.04&-0.02&-0.07&-0.13\\   
+BBVA USA BSHRS        &-0.31&-0.49&-0.83&-0.74&-0.95&-1.29\\
+COMERICA              &0.01&0.00&-0.07&-0.13&-0.19&-0.23\\  
+HUNTINGTON BSHRS      &-0.01&-0.03&-0.03&-0.04&-0.01&-0.01\\
+ZIONS BC              &0.02&0.09&0.14&0.21&0.31&0.36\\      
+\hline                                                      
+\end{tabular}                                               
+\end{table}
+
+\begin{table}
+\caption{Model 5, Partial Least Squares (recursive window), Diebold-Mariano-West Statistics relative to a no-change forecast}
+
+\center                                                              
+\begin{tabular}{|l|c|c|c|c|c|c|}                                     
+\hline                                                               
+&Step 1 &Step 2 &Step 4 &Step 6 &Step 8 &Step 10\\                   
+\hline                                                               
+JPMORGAN CHASE \& CO  &1.518&1.467&0.947&0.455&-0.233&-1.007\\       
+BANK OF AMER CORP     &-0.364&-0.485&-0.800&-0.789&-0.764&-0.764\\   
+CITIGROUP             &0.733&0.716&1.169&0.885&1.849&0.893\\         
+WELLS FARGO \& CO     &1.470&1.184&0.623&-0.093&-1.013&-1.813\\      
+U S BC                &1.083&0.960&0.938&1.832&0.680&-0.565\\        
+PNC FNCL SVC GROUP    &1.603&1.868&2.709*&3.598*&1.807&0.010\\       
+BANK OF NY MELLON CORP&-1.215&-1.028&1.167&1.168&-1.523&-8.018*\\    
+STATE STREET CORP     &1.357&1.604&1.393&1.114&0.921&0.854\\         
+BB\&T CORP            &2.067*&1.906&4.353*&1.461&0.537&-0.651\\      
+SUNTRUST BK           &1.802&1.820&1.183&0.479&0.304&-0.397\\        
+CAPITAL ONE FC        &2.417*&3.182*&2.805*&2.031*&1.127&0.643\\     
+REGIONS FC            &0.954&0.260&-0.826&-0.806&-0.594&-0.233\\     
+RBS CITIZENS FNCL GRP &0.273&0.286&0.543&0.514&0.804&1.239\\         
+FIFTH THIRD BC        &-0.607&-0.451&0.348&0.133&0.385&-0.177\\      
+BMO FNCL CORP         &-0.917&-0.541&-0.367&-0.257&-0.144&-0.841\\   
+UNIONBANCAL CORP      &-1.146&-1.464&-1.541&-4.153*&-4.278*&-2.349*\\
+NORTHERN TR CORP      &-0.177&-0.425&-0.026&-0.714&-2.317*&-5.474*\\ 
+KEYCORP               &-0.740&-0.526&-0.722&-0.948&-1.040&-1.539\\   
+M\&T BK CORP          &1.599&2.549*&2.572*&-0.562&-1.446&-2.908*\\   
+BBVA USA BSHRS        &-0.990&-0.997&-1.084&-1.235&-1.265&-1.290\\   
+COMERICA              &0.444&0.016&-2.252*&-2.291*&-3.623*&-5.810*\\ 
+HUNTINGTON BSHRS      &-0.550&-0.907&-0.460&-0.567&-0.196&-0.159\\   
+ZIONS BC              &0.984&2.038*&1.901&1.854&1.886&2.110*\\       
+\hline                                                               
+\end{tabular}
+
+\flushleft
+\footnotesize \noindent A star denotes significance at the 95\% level
+
+\end{table}
+
+\begin{table}  
+\caption{Model 5, Partial Least Squares, Out-of-Sample RMSEs (rolling window) in deviation from RMSEs for no-change forecast}                                                                 
+\center                                                     
+\begin{tabular}{|l|c|c|c|c|c|c|}                            
+\hline                                                      
+&Step 1 &Step 2 &Step 4 &Step 6 &Step 8 &Step 10\\          
+\hline                                                      
+JPMORGAN CHASE \& CO  &0.02&0.02&0.03&0.01&-0.06&-0.13\\    
+BANK OF AMER CORP     &0.05&0.07&0.05&0.05&0.04&0.06\\      
+CITIGROUP             &0.09&0.17&0.35&0.35&0.31&0.21\\      
+WELLS FARGO \& CO     &-0.00&-0.02&-0.06&-0.13&-0.20&-0.28\\
+U S BC                &0.02&0.02&0.02&0.04&-0.02&-0.01\\    
+PNC FNCL SVC GROUP    &0.02&0.04&0.06&0.05&-0.02&-0.08\\    
+BANK OF NY MELLON CORP&-0.08&-0.08&0.11&0.07&-0.04&-0.18\\  
+STATE STREET CORP     &0.06&0.05&0.10&0.16&0.22&0.19\\      
+BB\&T CORP            &-0.01&-0.02&-0.03&-0.08&-0.04&-0.02\\
+SUNTRUST BK           &0.03&0.07&0.08&0.07&0.15&0.20\\      
+CAPITAL ONE FC        &0.55&0.57&0.58&0.52&0.25&0.13\\      
+REGIONS FC            &0.02&0.02&-0.04&-0.07&-0.00&0.06\\   
+RBS CITIZENS FNCL GRP &-0.01&-0.02&0.01&-0.01&0.16&0.23\\   
+FIFTH THIRD BC        &-0.05&-0.04&0.09&0.15&0.23&0.26\\    
+BMO FNCL CORP         &-0.12&-0.14&-0.08&-0.16&-0.20&-0.36\\
+UNIONBANCAL CORP      &-0.01&0.00&0.04&-0.04&0.03&0.03\\    
+NORTHERN TR CORP      &0.01&0.01&0.04&0.00&-0.07&-0.14\\    
+KEYCORP               &-0.01&-0.01&-0.03&-0.07&-0.10&-0.12\\
+M\&T BK CORP          &0.01&-0.00&-0.02&-0.08&-0.13&-0.15\\ 
+BBVA USA BSHRS        &-0.23&-0.36&-0.64&-0.63&-0.78&-1.08\\
+COMERICA              &0.02&0.01&-0.07&-0.10&-0.09&-0.02\\  
+HUNTINGTON BSHRS      &0.07&0.17&0.18&-0.01&0.02&-0.01\\    
+ZIONS BC              &0.01&0.06&0.11&0.18&0.31&0.35\\      
+\hline                                                      
+\end{tabular}                                               
+\end{table}
+
+\begin{table} 
+\caption{Model 5, Partial Least Squares (rolling window), Diebold-Mariano-West Statistics relative to a no-change forecast}                                                                  
+\center                                                              
+\begin{tabular}{|l|c|c|c|c|c|c|}                                     
+\hline                                                               
+&Step 1 &Step 2 &Step 4 &Step 6 &Step 8 &Step 10\\                   
+\hline                                                               
+JPMORGAN CHASE \& CO  &1.355&0.782&0.526&0.067&-0.711&-1.291\\       
+BANK OF AMER CORP     &1.790&2.203*&1.655&0.985&0.869&1.631\\        
+CITIGROUP             &2.033*&1.957&1.769&1.437&2.155*&1.572\\       
+WELLS FARGO \& CO     &-0.387&-0.806&-1.751&-3.691*&-3.336*&-2.820*\\
+U S BC                &1.397&1.066&0.476&0.863&-0.750&-0.372\\       
+PNC FNCL SVC GROUP    &1.943&1.615&1.374&1.376&-0.400&-1.324\\       
+BANK OF NY MELLON CORP&-1.360&-0.960&2.300*&2.188*&-1.044&-7.728*\\  
+STATE STREET CORP     &1.248&0.863&0.723&0.644&0.701&0.618\\         
+BB\&T CORP            &-0.572&-1.010&-0.433&-1.130&-0.546&-0.381\\   
+SUNTRUST BK           &1.175&2.246*&1.433&1.423&2.215*&2.221*\\      
+CAPITAL ONE FC        &2.560*&3.280*&2.337*&2.148*&2.098*&1.171\\    
+REGIONS FC            &0.741&0.790&-0.535&-0.623&-0.016&0.480\\      
+RBS CITIZENS FNCL GRP &-0.204&-0.832&0.182&-0.069&0.839&1.279\\      
+FIFTH THIRD BC        &-0.703&-0.311&0.693&0.723&1.056&1.257\\       
+BMO FNCL CORP         &-0.524&-0.420&-0.275&-0.508&-0.687&-1.379\\   
+UNIONBANCAL CORP      &-0.230&0.069&0.631&-1.155&0.658&0.765\\       
+NORTHERN TR CORP      &0.726&0.760&2.672*&0.111&-1.804&-6.260*\\     
+KEYCORP               &-0.370&-0.238&-0.340&-0.573&-0.678&-0.945\\   
+M\&T BK CORP          &0.584&-0.045&-0.373&-1.438&-1.830&-2.488*\\   
+BBVA USA BSHRS        &-0.762&-0.743&-0.928&-1.070&-1.092&-1.148\\   
+COMERICA              &0.853&0.277&-3.670*&-2.282*&-1.594&-0.300\\   
+HUNTINGTON BSHRS      &1.288&1.006&0.933&-0.136&0.547&-0.364\\       
+ZIONS BC              &0.767&1.509&1.454&2.120*&2.761*&2.658*\\      
+\hline                                                               
+\end{tabular}
+
+\flushleft
+\footnotesize \noindent A star denotes significance at the 95\% level
+
+\end{table}
+
 \end{document}
